% !TEX TS-program = lualatexmk
% !TEX parameter =  --shell-escape

\documentclass[12pt]{book}

\input{commonheaders_vespers_people} %%% will need to fix spacing problem in psalms at a later date.

\begin{document}
 
 \twosided[pc] %% to allow for the left-right switching of columns needed for bilingual liturgical typesetting (like in missals); it's unclear why this goes BEFORE everything else and not when activating the environment. %% It's OK for paper booklets to not switch, but it's probably best practice to switch.
 
  \thispagestyle{empty}
 
 {\centering{\huge\capspace{{ASSUMPTION OF THE B.V.M.}}}\par}\header{assumption of the b.v.m.}
 
\smalltitle{Before the Office.}

\rubrique{These prayers are said kneeling. End when the priest stands.}

\textebilingue{orationes_ante_do}{orationes_ante_do_eng}

\smalltitle{\prayers}

\rubrique{The celebrant intones the following verse. All sing the response and doxology.}

\begin{otherlanguage}{english}
\rubrique{Make the sign of the cross at the beginning and at the Magnificat, and bow the head when the Trinity is mentioned, when mention is made of the ``Holy Name'' in some way, and at ``Orémus.''}\end{otherlanguage}

\gscore[]{℣.}{Deus_in_adjutorium_tonus_solemnior}

\noindent
\begin{otherlanguage}{english}
O God, \cc{} come to my assistance. \rr O Lord, make haste to help me. Glory be to the Father, and to the Son,~* and to the Holy Ghost. As it was in the beginning, is now,~* and ever shall be, world without end. Amen. Alleluia.\end{otherlanguage}

\smalltitle{First Antiphon. 7. a.}
\initialscore{an_assumpta_est_maria_solesmes_1961}

\rubrique{\begin{otherlanguage}{english}
Sit at the word ``Sede.'' The cantor sings odd verses; all sing even verses, and this for the hymn and Magnificat, when sung entirely in chant, as well.
\end{otherlanguage}}

\psalmus{109}{109_7a}{109_7}{109_en}

\begin{otherlanguage}{english} \noindent Ant. Mary hath been assumed into heaven; the Angels rejoice; they praise and bless the Lord. \end{otherlanguage}

\rubrique{\begin{otherlanguage}{english}
Clerics in descending order, or a cantor, intone the subsequent antiphons.\end{otherlanguage}}

 \gscore[2. Ant.]{8. G}{an_maria_virgo_assumpta_est_solesmes_1961}
 
 \begin{otherlanguage}{english} \noindent Ant. The Virgin Mary hath been assumed into the chamber on high, where the King of kings sitteth on a throne amid the stars. \end{otherlanguage}

\psalmus{112}{112_8G}{112_8}{112_en}

  \gscore[3. Ant.]{4. A*}{an_in_odorem_solesmes_1961}
  
  \begin{otherlanguage}{english} \noindent Ant. We run after thee, on the scent of thy perfumes: the virgins love thee heartily. \end{otherlanguage}

  \psalmus{121}{121_4_alt_Astar}{121_4A_A_star}{121_en}

  \gscore[4. Ant.]{7. c2}{an_benedicta_filia_solesmes_1961}

\begin{otherlanguage}{english} \noindent Ant. Blessed of the Lord art thou, O daughter, for by thee we have been given to eat of the fruit (of the tree) of Life. \end{otherlanguage}

  \psalmus{126}{126_7c2}{126_7}{126_en}

 \gscore[5. Ant.]{1. g2}{an_pulchra_es_et_decora_solesmes_1961}
 
\begin{otherlanguage}{english} \noindent Ant. Beautiful and adorned art thou, O daughter of Jerusalem, terrible as a fenced camp set in battle array. \end{otherlanguage}

\psalmus{147}{147_1g2}{147_1}{147_en}

\rubrique{\begin{otherlanguage}{english}
Having sung the last antiphon in full, stand when the celebrant does.
\end{otherlanguage}}

\smalltitle{Chapter.}
\scripture{Eccli. 24, 11 – 12.}

\textebilingue{cap_15_aug_old}{cap_15_aug_old_en}

%% Only the usual tone of the hymn is given.

\smalltitle{Hymn.}

\rubrique{The first verse is sung kneeling.}

\initialscore{hy_ave_maris_stella_mode_1_LA1949}

\begin{multicols}{2}
\raggedcolumns
\begin{otherlanguage}{english}
Hail, star of the sea,\\
loving Mother of God,\\
and also always a virgin,\\
Happy gate of heaven.

Receiving that Ave\\
from Gabriel's mouth\\
confirm us in peace,\\
Reversing Eva's name.

Break the chains of sinners,\\
Bring light to the blind,\\
Drive away our evils,\\
Ask for all good.

Show thyself to be a mother,\\
May he accept prayers through thee,\\
he who, born for us,\\
Chose to be thine.

O unique virgin,\\
Meek above all,\\
Make us, absolved from sin,\\
Gentle and chaste.

Keep life pure,
Make the journey safe,
So that, seeing Jesus,
We may always rejoice together.

Let there be praise to God the Father,\\
Glory to Christ in the highest,\\
To the Holy Spirit,\\
One honor to all three. Amen
\end{otherlanguage}
\end{multicols}

\rubrique{\begin{otherlanguage}{english}
The cantor sings the verse to which all reply.
\end{otherlanguage}}

\smallscore{versiculus_15_aug}
\begin{otherlanguage}{english} \noindent \vv The holy Mother of God hath been exalted.

\noindent \rr Over choirs of Angels, into the heavenly kingdom.
 \end{otherlanguage}
 
\rubrique{This tone is used on the feast itself, but not during the octave.}
 
\smalltitle{Antiphon at the Magnificat.}

\rubrique{Sit when the celebrant does, after he sings the first words. Rise at the end.}

 \smalltitle{At I Vespers.}
 \gscore[]{1. f}{an_virgo_prudentissima_solesmes_1961}
 
 \begin{otherlanguage}{english} \noindent Ant. Virgin most prudent, whither goest thou up, like the dawn gloriously rising? O daughter of Zion, thou art all beautiful and pleasant, fair as the moon, clear as the sun. \end{otherlanguage}
 
 \smallscore{Magnificat_1_f_Solemnis}

   \smallskip%
    \noindent%
 \begin{enumerate}[label=\arabic*.,%
			leftmargin=0pt, %
			itemindent=0mm, %
			labelsep=3pt, %
			labelwidth=0pt, %
			rightmargin=0pt, %
			parsep=0pt, %
			topsep=0pt, %
			itemsep=0pt,%
			start=3]
\input{psaumes/Magnificat_1_Solemnis}
\end{enumerate}

\myrule
 \smalltitle{At II Vespers.}
 \gscore[]{8. G*}{an_hodie_maria_virgo_caelos_solesmes_1961}
 
 \begin{otherlanguage}{english} \noindent Ant. Today the Blessed Virgin Mary ascended to heaven, rejoice, she reigns with Christ forever. \end{otherlanguage}
 
 \myrule
 
%
% \rubrique{\begin{otherlanguage}{english}
%The Magnificat may be sung entirely in chant or sung alternating between a cantor singing chant and the choir singing a polyphonic or harmonized setting. Please see the final pages for the chant.\end{otherlanguage}}

\smalltitle{Canticle of the Blessed Virgin Mary.}

\scripture{Luc. 1, 46–55.}

\begin{multicols}{2}

%\lettrine{M}agníficat~* ánima mea Dóminum.\\ Et exsultávit spíritus meus~* in Deo salutári meo.
%
% Quia respéxit humilitátem ancíllæ suæ:~* ecce enim ex hoc beátam me dicent o\-mnes generatiónes.
%
% Quia fecit mihi magna qui potens est:~* et san\-ctum nomen ejus.
%
% Et misericórdia ejus a progénie in progénies~* timéntibus eum.
%
% Fecit poténtiam in bráchio suo:~* dispérsit supérbos mente cordis sui.
%
% Depósuit poténtes de sede,~* et exaltávit húmiles.
%
% Esuriéntes implévit bonis:~* et dívites dimísit inánes.
%
% Suscépit Israel púerum suum,~* recordátus misericórdiæ suæ.
%
% Sicut locútus est ad patres no\-stros,~* Abraham, et sémini ejus in sǽcula.

% Glória Patri, et Fílio,~* et Spirítui San\-cto.
%
% Sicut erat in princípio, et nunc, et semper,~* et in sǽcula sæculórum. Amen.
 
% \switchcolumn
 \begin{otherlanguage}{english}
My soul \cc~* doth magnify the Lord.

And my spirit hath rejoiced~* in God my Saviour.

Because he hath regarded the humility of his handmaid;~* for behold from henceforth all generations shall call me blessed.

Because he that is mighty, hath done great things to me;~* and holy is his name.

And his mercy is from generation unto generations,~* to them that fear him.

He hath shewed might in his arm:~* he hath scattered the proud in the conceit of their heart.

He hath put down the mighty from their seat,~* and hath exalted the humble.

He hath filled the hungry with good things;~* and the rich he hath sent empty away.

He hath received Israel his servant,~* being mindful of his mercy:

As he spoke to our fathers,~* to Abraham and to his seed for ever.

Glory be to the Father, and to the Son,~* and to the Holy Ghost.

As it was in the beginning, is now,~* and ever shall be, world without end. Amen.
\end{otherlanguage}
 \end{multicols}

\rubrique{\begin{otherlanguage}{english}
Sit as the antiphon is repeated. Then the usual dialogue follows. All stand from now until the end of the office, unless one sits for the antiphon sung for the commemoration of Saint Joachim at II Vespers, according to custom.\end{otherlanguage}}

\textebilingue{dv}{dv_eng}

\textebilingue{oratio_15_aug_old}{oratio_15_aug_old_en}

\myrule

\rubrique{At II Vespers, a commemoration of Saint Joachim is made; a commemoration is made at both Vespers of the Sunday in occurence or concurrence.}

\gscore[Ant.]{8.}{an_laudemus_virum_solesmes_1961}

\begin{otherlanguage}{english} \noindent Ant. Let us praise a man famous in his generation, with whom the Lord didst establish the blessing of all nations, and on whose head He made his covenant to rest.\end{otherlanguage}

\textebilingue{versiculus_16_aug}{versiculus_16_aug_en}

\textebilingue{oratio_16_aug}{oratio_16_aug_en}

\myrule

\textebilingue{dv}{dv_eng}

\rubrique{\begin{otherlanguage}{english}
The cantor sings the verse to which all reply.
\end{otherlanguage}}

\gscore[]{1.}{BD_BVM}

\noindent \vv Let us bless the Lord.

\noindent \rr Thanks be to God.

\textebilingue{fidelium_animae}{fidelium_animae_en}

    %\rubrique{\begin{otherlanguage}{english}
%Then one Our Father is said, unless Compline follows, or even some pious exercises, a sermon, or Benediction of the Blessed Sacrament, in parish and other smaller churches.\end{otherlanguage}}

%\rubrique{\begin{otherlanguage}{english}
%Then the office is concluded with the following verses, the Marian antiphon, etc.
%\end{otherlanguage}}

\end{document}