% !TEX TS-program = lualatexmk
% !TEX parameter =  --shell-escape

\documentclass[12pt]{book}

\input{commonheaders_vespers_people} %%% will need to fix spacing problem in psalms at a later date.

\begin{document}
 
 \twosided[pc] %% to allow for the left-right switching of columns needed for bilingual liturgical typesetting (like in missals); it's unclear why this goes BEFORE everything else and not when activating the environment. %% It's OK for paper booklets to not switch, but it's probably best practice to switch.
 
  \thispagestyle{empty}
 
 {\centering{\huge\capspace{{CIRCUMCISION OF THE LORD.}}}\par}\header{circumcision of the lord}
 \vspace{0.5\baselineskip}
 
{\centering{\huge{At I Vespers.}}\par}
 
\smalltitle{Before the Office.}

\rubrique{These prayers are said kneeling. End when the priest stands.}

\textebilingue{orationes_ante_do}{orationes_ante_do_eng}

\smalltitle{\prayers}

\rubrique{The celebrant intones the following verse. All sing the response and doxology.}

\begin{otherlanguage}{english}
\rubrique{Make the sign of the cross at the beginning and at the Magnificat, and bow the head when the Trinity is mentioned, when mention is made of the ``Holy Name'' in some way, and at ``Orémus.''}\end{otherlanguage}

\gscore[]{℣.}{Deus_In_Adjutorium_Festivus}

\noindent
\begin{otherlanguage}{english}
O God, \cc{} come to my assistance. \rr O Lord, make haste to help me. Glory be to the Father, and to the Son,~* and to the Holy Ghost. As it was in the beginning, is now,~* and ever shall be, world without end. Amen. Alleluia.\end{otherlanguage}

\smalltitle{First Antiphon. 6. F.}
\initialscore{an_o_admirabile_commercium_solesmes_1961}

\begin{otherlanguage}{english} \noindent Ant. O wondrous interchange! the Creator of mankind, taking upon him a living body, vouchsafed to be born of a pure Virgin: and by his Humanity, which was begotten in no earthly wise, hath made us partakers of his Divinity. \end{otherlanguage}

\rubrique{\begin{otherlanguage}{english}
Sit at the word ``Sede.'' The cantor sings odd verses; all sing even verses, and this for the hymn and Magnificat, when sung entirely in chant, as well.
\end{otherlanguage}}

\psalmus{109}{109_6_alt}{109_6_alt}{109_en}

\rubrique{\begin{otherlanguage}{english}
Clerics in descending order, or a cantor, intone the subsequent antiphons.\end{otherlanguage}}

 \gscore[2. Ant.]{3. a2}{an_quando_natus_es_solesmes_1961}
 
 \begin{otherlanguage}{english} \noindent Ant. When thou wast born all ineffably of a Virgin, then were the Scriptures fulfilled; thou camest down like the dew upon the fleece of wool, to bring salvation unto all mankind; we praise thee, O our God. \end{otherlanguage}

\psalmus{112}{112_3a2}{112_3a2_g}{112_en}

  \gscore[3. Ant.]{4. E}{an_rubum_quem_solesmes_1961}
  
  \begin{otherlanguage}{english} \noindent Ant. In the bush which Moses saw unconsumed, we recognize the preservation of thy glorious virginity: holy Mother of God, intercede for us.. \end{otherlanguage}

  \psalmus{121}{121_4E}{121_4E}{121_en}

  \gscore[4. Ant.]{1. f}{an_germinavit_radix_solesmes_1961}

\begin{otherlanguage}{english} \noindent Ant. The Root of Jesse hath budded, the Star hath come out of Jacob, the Virgin hath borne the Saviour: we praise thee, O our God. \end{otherlanguage}

  \psalmus{126}{126_1f}{126_1}{126_en}

 \gscore[5. Ant.]{2. D}{an_ecce_maria_genuit_solesmes_1961}
 
\begin{otherlanguage}{english} \noindent Ant. Lo, Mary hath brought forth the Saviour of whom when John saw him, he did proclaim: Behold the Lamb of God that taketh away the sins of the world, alleluia. \end{otherlanguage}

\psalmus{147}{147_2}{147_2}{147_en}

\rubrique{\begin{otherlanguage}{english}
Having sung the last antiphon in full, stand when the celebrant does.
\end{otherlanguage}}

\smalltitle{Chapter.}
\scripture{Tit. 2, 11–12.}

\texteparacol{\lettrine{A}{p}páruit grátia Dei Salvatóris nostri ómnibus homínibus,~† erúdiens nos, ut abnegántes impietátem et sæculária desidéria,~* sóbrie, et juste, et pie vivámus in hoc sǽculo.\par
\rr Deo grátias.}{\noindent The grace of God that bringeth salvation hath appeared to all men, teaching us that, denying ungodliness and worldly lusts, we should live soberly, righteously, and godly in this present world. Thanks be to God.}

\smalltitle{Hymn. Mode I.}

\rubrique{At the words \normaltext{Jesu Red\-ém\-ptor ómnium,} the celebrant extends and then closes his hands while turning to the cross.}

\initialscore{hy_jesu_redemptor_omnium_nativitatis_solesmes_1961}

\begin{multicols}{2}
\begin{otherlanguage}{english}
Jesus, the Redeemer of all,\\
who, being the equal of the Father's glory,\\
was begotten of the Sovereign Father\\
before the beginning of light.

Thou light and splendor of the Father,\\
Thou neverfailing hope of all,\\
give ear to the prayers\\
 which thy servants throughout the world pour forth.
 
Remember, O Creator of the world,\\
that in being born\\
Thou didst once assume the form of our body\\
 from the sacred womb of a Virgin.
 
The present day recurring\\
 in the course of each year,\\
 bears witness to this, that Thou alone didst come forth\\
 from the bosom of the Father, the salvation of the world.
 
The stars, the earth, and the seas,\\
and every creature under heaven\\
doth greet him with a new canticle,\\
as the author of the new salvation.
 
We also, whom the sacred stream\\
 of thy blood hath cleansed,\\pay thee the tribute of a hymn\\
 on thy birthday.
 
Jesus, unto thee be the glory,\\
 Thou who wast born of a virgin,\\
 with the Father and the loving Spirit,\\
 in the eternal age. Amen.

\end{otherlanguage}
\end{multicols}

\rubrique{\begin{otherlanguage}{english}
The cantor sings the verse to which all reply.
\end{otherlanguage}}

%\smallscore{versiculus_i_vesperis_nativitatis_domini}
\gresetinitiallines{0}
\gabcsnippet{(c3)℣. Ver(h)bum(h) ca(h)ro(h) fa(h)ctum(h) est,(h) al(h)le(h)lú(h)ia.(g_hvGFEfgf.) (::)}

\gabcsnippet{(c3) ℟. Et(h) ha(h)bi(h)tá(h)vit(h) in(h) no(h)bis,(h) al(h)le(h)lú(h)ia.(g_hvGFEfgf.) (::)}

\gresetinitiallines{1}

\begin{otherlanguage}{english} \noindent \vv The Word was made flesh, alleluia.

\noindent \rr And dwelt among us, alleluia.
 \end{otherlanguage}
 
%\rubrique{This tone is used on the feast itself, but not during the octave.}
 
\smalltitle{Antiphon at the Magnificat.}

\rubrique{Sit when the celebrant does, after he sings the first words. Rise at the end.}

\gscore[]{8. G}{an_propter_nimiam_solesmes_1961}
 
 \begin{otherlanguage}{english} \noindent Ant. God, for his great love wherewith He loved us, sent his Own Son in the likeness of sinful flesh. Alleluia. \end{otherlanguage}
 
 \smallscore{Magnificat_8_G_Solemnis}

   \smallskip%
    \noindent%
    \begin{itemize}[%
    		label=\null, %
			leftmargin=0pt, %
			itemindent=0mm, %
			labelsep=0pt, %
			labelwidth=0pt, %
			rightmargin=0pt, %
			parsep=0pt, %
			topsep=0pt, %
			itemsep=0pt]%
    \input{psaumes/Magnificat_8_Solemnis}
    \end{itemize}

 \rubrique{\begin{otherlanguage}{english}
The Magnificat may be sung entirely in chant or sung alternating between a cantor singing chant and the choir singing a polyphonic or harmonized setting.\end{otherlanguage}}

\smalltitle{Canticle of the Blessed Virgin Mary.}

\scripture{Luc. 1, 46–55.}
\begin{multicols}{2}
%\begin{paracol}{2}
%
%\lettrine{M}agníficat~* ánima mea Dóminum.\\ Et exsultávit spíritus meus~* in Deo salutári meo.
%
% Quia respéxit humilitátem ancíllæ suæ:~* ecce enim ex hoc beátam me dicent o\-mnes generatiónes.
%
% Quia fecit mihi magna qui potens est:~* et san\-ctum nomen ejus.
%
% Et misericórdia ejus a progénie in progénies~* timéntibus eum.
%
% Fecit poténtiam in bráchio suo:~* dispérsit supérbos mente cordis sui.
%
% Depósuit poténtes de sede,~* et exaltávit húmiles.
%
% Esuriéntes implévit bonis:~* et dívites dimísit inánes.
%
% Suscépit Israel púerum suum,~* recordátus misericórdiæ suæ.
%
% Sicut locútus est ad patres no\-stros,~* Abraham, et sémini ejus in sǽcula.
%
% Glória Patri, et Fílio,~* et Spirítui San\-cto.
%
% Sicut erat in princípio, et nunc, et semper,~* et in sǽcula sæculórum. Amen.
 
% \switchcolumn
 \begin{otherlanguage}{english}
My soul \cc~* doth magnify the Lord.

And my spirit hath rejoiced~* in God my Saviour.

Because he hath regarded the humility of his handmaid;~* for behold from henceforth all generations shall call me blessed.

Because he that is mighty, hath done great things to me;~* and holy is his name.

And his mercy is from generation unto generations,~* to them that fear him.

He hath shewed might in his arm:~* he hath scattered the proud in the conceit of their heart.

He hath put down the mighty from their seat,~* and hath exalted the humble.

He hath filled the hungry with good things;~* and the rich he hath sent empty away.

He hath received Israel his servant,~* being mindful of his mercy:

As he spoke to our fathers,~* to Abraham and to his seed for ever.

Glory be to the Father, and to the Son,~* and to the Holy Ghost.

As it was in the beginning, is now,~* and ever shall be, world without end. Amen.
\end{otherlanguage}
% \end{paracol}
\end{multicols}

\rubrique{\begin{otherlanguage}{english}
Sit as the antiphon is repeated. Then the usual dialogue follows. All stand from now until the end of the office.\end{otherlanguage}}

\textebilingue{dv}{dv_eng}

%\textebilingue{oratio_15_aug_old}{oratio_15_aug_old_en}

\texteparacol{Orémus.

\lettrine{D}{e}us, qui salútis ætérnæ, beátæ Maríæ virginitáte fecúnda, humáno géneri prǽmia præstitísti:~† tríbue, quǽsumus; ut ipsam pro nobis intercédere sentiámus,~* per quam merúimus auctórem vitæ suscípere, Dóminum nostrum Jesum Christum Fílium tuum.
Qui tecum vivit et regnat.

\rr Amen.}{\noindent Let us pray. O God, who by the fruitful virginity of the Blessed Mary, hast given unto mankind the rewards of everlasting life; grant, we beseech thee, that we may continually feel the might of her intercession, through whom we have worthily received the Author of our life, our Lord Jesus Christ, thy Son: Who with thee liveth and reigneth, in the unity of the Holy Spirit, God, world without end. Amen.}


\textebilingue{dv}{dv_eng}

\rubrique{\begin{otherlanguage}{english}
The cantor sings the verse to which all reply.
\end{otherlanguage}}

\gscore[]{2.}{BD_Solemnis_I_Vesp}

\vv Let us bless the Lord.

\rr Thanks be to God.
    
%\rubrique{\begin{otherlanguage}{english}
%Then one Our Father is said, unless Compline follows, or even some pious exercises, a sermon, or Benediction of the Blessed Sacrament, in parish and other smaller churches.\end{otherlanguage}}

%\rubrique{\begin{otherlanguage}{english}
%Then the office is concluded with the following verses, the Marian antiphon, etc.
%\end{otherlanguage}}

\textebilingue{fidelium_animae}{fidelium_animae_en}

\rubrique{The \normaltext{Te Deum} is customarily sung after I Vespers of the Circumcision, where these are sung.}

\rubrique{The celebrant intones, and the cantors continue. The people may join the schola on alternating verses, which begin at double bar lines.}

\gscore[]{3.}{hy_te_deum_solemn_solesmes}

\begin{multicols}{2}
\begin{otherlanguage}{english}

We praise thee, O God; we acknowledge thee to be the Lord.\\
All the earth doth worship thee, the Father everlasting.

To thee all angels cry aloud, the heavens and all the powers therein.\\
To thee cherubin and seraphin continually do cry,\\
Holy, Holy, Holy, Lord God of Sabaoth;

Heaven and earth are full of the majesty of thy glory.\\
The glorious company of the apostles praise thee.\\
The goodly fellowship of the prophets praise thee.\\
The noble army of martyrs praise thee.

The holy Church throughout all the world doth acknowledge thee:\\
the Father of an infinite majesty;\\
thine honourable, true and only Son;\\
also the Holy Ghost the Comforter.

Thou art the King of glory, O Christ.\\
Thou art the everlasting Son of the Father.\\
When thou tookest upon thee to deliver man,\\
thou didst not abhor the Virgin’s womb.

When thou hadst overcome the sharpness of death,\\
thou didst open the kingdom of heaven to all believers.

Thou sittest at the right hand of God, in the glory of the Father.\\
We believe that thou shalt come to be our judge.\\
We therefore pray thee, help thy servants,\\
whom thou hast redeemed with thy precious blood.

Make them to be numbered with thy saints in glory everlasting.\\
O Lord, save thy people and bless thine heritage.\\
Govern them and lift them up for ever.

Day by day we magnify thee;\\
and we worship thy name, ever world without end.

Vouchsafe, O Lord, to keep us this day without sin.\\
O Lord, have mercy upon us, have mercy upon us.\\
O Lord, let thy mercy lighten upon us, as our trust is in thee.\\
O Lord, in thee have I trusted; let me never be confounded.

\end{otherlanguage}
\end{multicols}

\texteparacol{\vv Benedicámus Patrem et Fílium cum San\-cto Spíritu.

\noindent \rr Laudémus et superexaltémus eum in sǽcula.

\noindent \vv Benedí\-ctus es Dómine in firmaménto cæli.

\noindent \rr Et laudábilis, et gloriósus, et superexaltátus in sǽcula.

\noindent \vv Dómine exáudi oratiónem meam.

\noindent \rr Et clamor meus ad te véniat.}{\noindent \vv Let us bless the Father and the Son with the Holy Ghost.

\noindent \rr Let us praise and exalt him above all forever.

\noindent \vv Blessed art thou, o Lord, in the firmament of heaven.

\noindent \rr And worthy of praise, and glorious, and above all forever.

\noindent \vv O Lord hear my prayer.

\noindent \rr And let my cry come unto thee.}

\textebilingue{dv}{dv_eng}

\texteparacol{Orémus. 

\lettrine{D}{e}us, cujus misericórdiæ non est númerus, et bonitátis infinítus est thesáurus:~† piíssimæ majestáti tuæ pro collátis donis grátias ágimus, tuam semper cleméntiam exorántes;~* ut qui peténtibus postuláta concédis, eósdem non déserens, ad prǽmia futúra dispónas. Per Christum Dóminum nostrum.

\rr Amen.}{\noindent Let us pray. O God, whose mercies are numberless and whose treasurey
goodness has no end, we give thanks to
thy most gracious majesty for the gifts
Thou hast bestowed, beseeching thy
mercy, that as Thou grantest the petitions
of those who ask, so not forsaking them,
Thou wilt prepare them for rewards to
come. Through Christ our Lord. Amen.}

\end{document}