% !TEX TS-program = lualatexmk
% !TEX parameter =  --shell-escape

\documentclass[12pt]{book}

\input{commonheaders_vespers_people} %%% will need to fix spacing problem in psalms at a later date.

\begin{document}
 
 \twosided[pc] %% to allow for the left-right switching of columns needed for bilingual liturgical typesetting (like in missals); it's unclear why this goes BEFORE everything else and not when activating the environment. %% It's OK for paper booklets to not switch, but it's probably best practice to switch.
 
  \thispagestyle{empty}
 
 {\centering{\huge\capspace{{NATIVITY OF SAINT JOHN THE BAPTIST.}}}\par}\header{nativity of saint john the baptist.}
 
 \vspace{0.5\baselineskip}
 
 {\centering{\huge{At I Vespers.}}\par}
 
\smalltitle{Before the Office.}

\rubrique{These prayers are said kneeling. End when the priest stands.}

\textebilingue{orationes_ante_do}{orationes_ante_do_eng}

\smalltitle{\prayers}

\rubrique{The celebrant intones the following verse. All sing the response and doxology.}

\begin{enpars}
\rubrique{Make the sign of the cross at the beginning and at the Magnificat, and bow the head when the Trinity is mentioned, when mention is made of the ``Holy Name'' in some way, and at ``Orémus.''}\end{enpars}

\gscore[]{℣.}{Deus_in_adjutorium_tonus_solemnior}

\noindent
\begin{enpars}
O God, \cc{} come to my assistance. \rr O Lord, make haste to help me. Glory be to the Father, and to the Son, and to the Holy Ghost. As it was in the beginning, is now, and ever shall be, world without end. Amen. Alleluia.\end{enpars}

\smalltitle{First Antiphon. 7. a.}
 \initialscore{an_ipse_praeibit_ante_illum_solesmes} 
 \begin{enpars} \noindent Ant. He shall go before Him in the spirit and power of Elias, to make ready a people prepared for the Lord. \end{enpars}

\rubrique{\begin{enpars}
Sit at the word ``Sede.'' The cantor sings odd verses; all sing even verses, and this for the hymn and Magnificat, when sung entirely in chant, as well.
\end{enpars}}

\psalmus{109}{109_7a}{109_7}{109_en}
%
\rubrique{\begin{enpars}
Clerics in descending order, or a cantor, intone the subsequent antiphons.\end{enpars}}
%
 \gscore[2. Ant.]{8. G}{an_joannes_est_nomen_ejus_solesmes_1961}
% 
 \begin{enpars} \noindent Ant. His name is John. He shall drink neither wine nor strong drink, and many shall rejoice at his birth. \end{enpars}
%
 \psalmus{110}{110_8G}{110_8}{110_en}
%
\gscore[3. Ant.]{1. f}{an_ex_utero_senectutis_solesmes_1961}

  \begin{enpars} \noindent Ant. From the womb of the barren and advanced in years was born John the Forerunner of the Lord. \end{enpars}
%
 \psalmus{111}{111_1f}{111_1}{111_en}
%
 \gscore[4. Ant.]{4. A*}{an_iste_puer_magnus_solesmes_1961}
%
\begin{enpars} \noindent Ant. The child shall be great in the sight of the Lord, for His hand is with him. \end{enpars}
%
  \psalmus{112}{112_4A_A_star}{112_4A_A_star}{112_en}
%
  \gscore[5. Ant.]{5. a}{an_nazaraeus_vocabitur_solesmes_1961}
% 
\begin{enpars} \noindent Ant. The child shall be called a Nazarite. He shall drink neither wine nor strong drink, neither shall he eat any unclean thing, even from his mother's womb.\end{enpars}
%
  \psalmus{116}{116_5}{116_5}{116_en}

\rubrique{\begin{enpars}
Having sung the last antiphon in full, stand when the celebrant does.
\end{enpars}}

\chapterlateng

\scripture{Isaiæ 49, 1.}

\texteparacol{\lettrine{A}{u}udíte, ínsulæ, et atténdite pópuli de longe:~† Dóminus ab útero vocávit me,~* de ventre matris meæ recordátus est nóminis mei.

\rr Deo grátias.}{\noindent Give ear, ye islands, and hearken, ye people from afar. The Lord hath called me from the womb, from the bowels of my mother he hath been mindful of my name. Thanks be to God.}

\smalltitle{Hymn.}

\gscore[]{2.}{hy_ut_queant_laxis_solesmes_1961}

\begin{multicols}{2}
\begin{enpars}
\raggedcolumns
Uttering praises,
Raising songs of thy feats,	
Meekly thy servants\\	
Falter in our errors.	
Souls that are filthy\\	
Launder of offenses,	
Saint John the Baptist.
	
Thou wast not born yet	
When there came an angel	
To tell thy father\\	
Tidings of thy greatness.	
What he must name thee\\	
And thy way of living:	
All were predestined.	

Thy skeptic father,	
Doubting Heaven's promise,	
Promptly and fully\\	
Lost all voice and language;	
But what was taken,\\	
Ruined past all healing:	
Thy birth restored it.	

Deep in thy mother,	
From her womb thou didst sense it;	
It was the presence\\	
Of the unborn Savior.	
Thus to thy parents,\\	
Thou, before delivery,	
didst reveal the hidden.	

We praise the Father	
And the Son begotten;	
We praise the Spirit,\\	
Who has equal power.	
One God and holy\\	
For all times and seasons,	
All generations. Amen.
\end{enpars}
\end{multicols}

\textes{cantor_rubric}

\gresetinitiallines{0}
\gabcsnippet{(c3)℣. Fu(h)it(h) mis(h)sus(h) a(h) De(h)o.(g_hvGFEfgf.) (::) (Z-)

℟. Cu(h)i(h) no(h)men(h) e(h)rat(h) Jo(h)án(h)nes.(g_hvGFEfgf.) (::)}
\gresetinitiallines{1}

%\smallscore{versiculus_2_vesperis_christi_regis}
\begin{enpars} \noindent \vv There was a man sent from God.

\noindent \rr Whose name was John.
 \end{enpars}
 
\smalltitle{Antiphon at the Magnificat.}

\rubrique{Sit when the celebrant does, after he sings the first words. Rise at the end.}

 \gscore[]{8. G}{an_ingresso_zacharia_june_24_solesmes_1961}
 
 \begin{enpars} \noindent Ant. When Zacharias went into the Temple of the Lord, the Angel Gabriel appeared unto him, standing on the right side of the altar of incense. \end{enpars}

 \rubrique{\begin{enpars}
The Magnificat may be sung entirely in chant or sung alternating between a cantor singing chant and the choir singing a polyphonic or harmonized setting.\end{enpars}}


\smalltitle{Canticle of the Blessed Virgin Mary.}

\scripture{Luc. 1, 46–55.}

 \smallscore{Magnificat_8_G_Solemnis}

   \smallskip%
    \noindent%
    \begin{enumerate}[%
    		label=\arabic*., %
			leftmargin=0pt, %
			itemindent=0mm, %
			labelsep=3pt, %
			labelwidth=0pt, %
			rightmargin=0pt, %
			parsep=0pt, %
			topsep=0pt, %
			itemsep=0pt,
			start=3]%
    \input{psaumes/Magnificat_8_Solemnis}
    \end{enumerate}

\begin{multicols}{2}
%\begin{paracol}{2}

%\lettrine{M}agníficat~* ánima mea Dóminum.\\ Et exsultávit spíritus meus~* in Deo salutári meo.
%
% Quia respéxit humilitátem ancíllæ suæ:~* ecce enim ex hoc beátam me dicent o\-mnes generatiónes.
%
% Quia fecit mihi magna qui potens est:~* et san\-ctum nomen ejus.
%
% Et misericórdia ejus a progénie in progénies~* timéntibus eum.
%
% Fecit poténtiam in bráchio suo:~* dispérsit supérbos mente cordis sui.
%
% Depósuit poténtes de sede,~* et exaltávit húmiles.
%
% Esuriéntes implévit bonis:~* et dívites dimísit inánes.
%
% Suscépit Israel púerum suum,~* recordátus misericórdiæ suæ.
%
% Sicut locútus est ad patres no\-stros,~* Abraham, et sémini ejus in sǽcula.
%
% Glória Patri, et Fílio,~* et Spirítui San\-cto.
%
% Sicut erat in princípio, et nunc, et semper,~* et in sǽcula sæculórum. Amen.
% 
% \switchcolumn
 \begin{enpars}
My soul \cc~* doth magnify the Lord.

And my spirit hath rejoiced~* in God my Saviour.

Because he hath regarded the humility of his handmaid;~* for behold from henceforth all generations shall call me blessed.

Because he that is mighty, hath done great things to me;~* and holy is his name.

And his mercy is from generation unto generations,~* to them that fear him.

He hath shewed might in his arm:~* he hath scattered the proud in the conceit of their heart.

He hath put down the mighty from their seat,~* and hath exalted the humble.

He hath filled the hungry with good things;~* and the rich he hath sent empty away.

He hath received Israel his servant,~* being mindful of his mercy:

As he spoke to our fathers,~* to Abraham and to his seed for ever.

Glory be to the Father, and to the Son,~* and to the Holy Ghost.

As it was in the beginning, is now,~* and ever shall be, world without end. Amen.
\end{enpars}
% \end{paracol}
\end{multicols}

\rubrique{\begin{enpars}
Sit as the antiphon is repeated. Then the usual dialogue follows. All stand from now until the end of the office, unless one sits for the antiphon sung for the commemoration of the Sunday or of the saints, according to custom.\end{enpars}}

\textebilingue{dv}{dv_eng}

\texteparacol{Orémus.

\lettrine{D}{e}us, qui præséntem diem honorábilem nobis in beáti Joánnis nativitáte fecísti:~† da pópulis tuis spirituálium grátiam gaudiórum:~* et ómnium fidélium mentes dírige in viam salútis ætérnæ.
Per Dóminum.}{\noindent Let us pray. O God, Who hast made this day to be honourable in our eyes because of the Birth of the blessed John, graciously quicken thy people with spiritual joy, and order the minds of all the faithful in the way of everlasting salvation. Through Jesus Christ, thy Son our Lord, Who liveth and reigneth with thee, in the unity of the Holy Ghost, God, world without end. Amen.}

%\rubrique{\begin{enpars}
%The cantor sings the verse to which all reply. The tone is chosen at the choirmaster's discretion.
%\end{enpars}}

\rubrique{A commemoration is made of the Sunday when occurring or concurring with June 24.}

\textes{cantor_rubric}
%\myrule

\gscore[]{2.}{BD_Solemnis_I_Vesp}

\noindent \vv Let us bless the Lord.

\noindent \rr Thanks be to God.

%\myrule

%\gscore[]{6.}{BD_Solemnis_mode_6}
%
%\vv Let us bless the Lord.
%
%\rr Thanks be to God.
%
%\rubrique{Or the following tone may be used:}
%
%\gscore[]{5.}{BD_Solemnis_mode_5}
%
\textebilingue{fidelium_animae}{fidelium_animae_en}

%\myrule
%\smallscore{Magnificat_7a_Solemnis}
%
%   \smallskip%
%    \noindent%
%    \begin{itemize}[%
%    		label=\null, %
%			leftmargin=0pt, %
%			itemindent=0mm, %
%			labelsep=0pt, %
%			labelwidth=0pt, %
%			rightmargin=0pt, %
%			parsep=0pt, %
%			topsep=0pt, %
%			itemsep=0pt]%
%    \input{psaumes/Magnificat_7}
%    \end{itemize}

%\rubrique{\begin{enpars}
%Then one Our Father is said, unless Compline follows, or even some pious exercises, a sermon, or Benediction of the Blessed Sacrament, in parish and other smaller churches.\end{enpars}}

%\rubrique{\begin{enpars}
%Then the office is concluded with the following verses, the Marian antiphon, etc.
%\end{enpars}}

\end{document}