% !TEX TS-program = lualatexmk
% !TEX parameter =  --shell-escape

\documentclass[12pt]{book}

\input{commonheaders_vespers_people} %%% will need to fix spacing problem in psalms at a later date.

\begin{document}
 
 \twosided[pc] %% to allow for the left-right switching of columns needed for bilingual liturgical typesetting (like in missals); it's unclear why this goes BEFORE everything else and not when activating the environment. %% It's OK for paper booklets to not switch, but it's probably best practice to switch.
 
  \thispagestyle{empty}
 
 {\centering{\huge\capspace{{NATIVITY OF THE LORD.}}}\par}\header{nativity of the lord.}
 \vspace{0.5\baselineskip}
 
{\centering{\huge{At I Vespers.}}\par}
 
\smalltitle{Before the Office.}

\rubrique{These prayers are said kneeling. End when the priest stands.}

\textebilingue{orationes_ante_do}{orationes_ante_do_eng}

\smalltitle{\prayers}

\rubrique{The celebrant intones the following verse. All sing the response and doxology.}

\begin{otherlanguage}{english}
\rubrique{Make the sign of the cross at the beginning and at the Magnificat, and bow the head when the Trinity is mentioned, when mention is made of the ``Holy Name'' in some way, and at ``Orémus.''}\end{otherlanguage}

\gscore[]{℣.}{Deus_in_adjutorium_tonus_solemnior}

\noindent
\begin{otherlanguage}{english}
O God, \cc{} come to my assistance. \rr O Lord, make haste to help me. Glory be to the Father, and to the Son,~* and to the Holy Ghost. As it was in the beginning, is now,~* and ever shall be, world without end. Amen. Alleluia.\end{otherlanguage}

\smalltitle{First Antiphon. 8. G.}
\initialscore{an_rex_pacificus_solesmes_1961}

\begin{otherlanguage}{english} \noindent Ant. The King of Peace is magnified, whose face the whole earth desireth. \end{otherlanguage}

\rubrique{\begin{otherlanguage}{english}
Sit at the word ``Sede.'' The cantor sings odd verses; all sing even verses, and this for the hymn and Magnificat, when sung entirely in chant, as well.
\end{otherlanguage}}

\psalmus{109}{109_8G}{109_8G}{109_en}

\rubrique{\begin{otherlanguage}{english}
Clerics in descending order, or a cantor, intone the subsequent antiphons.\end{otherlanguage}}

 \gscore[2. Ant.]{7. a}{an_magnificatus_est_solesmes_1961}
 
 \begin{otherlanguage}{english} \noindent Ant. The Peaceful King exceeded all the kings of the whole earth. \end{otherlanguage}

\psalmus{110}{110_7a}{110_7}{110_en}

  \gscore[3. Ant.]{8. G}{an_completi_sunt_solesmes_1961}
  
  \begin{otherlanguage}{english} \noindent Ant. The days of Mary were accomplished that she should bring forth her first-born Son. \end{otherlanguage}

  \psalmus{111}{111_8G}{111_8}{111_en}

  \gscore[4. Ant.]{8. G}{an_scitote_quia_prope_solesmes_1961}

\begin{otherlanguage}{english} \noindent Ant. Know ye that the kingdom of God is at hand? Amen, I say unto you, it will not tarry. \end{otherlanguage}

  \psalmus{112}{112_8G}{112_8}{112_en}

 \gscore[5. Ant.]{1. g}{an_levate_capita_vestra_solesmes_1961}
 
\begin{otherlanguage}{english} \noindent Ant. Lift up your heads; behold, your redemption draweth nigh. \end{otherlanguage}

\psalmus{116}{116_1g}{116_1}{116_en}

\rubrique{\begin{otherlanguage}{english}
Having sung the last antiphon in full, stand when the celebrant does.
\end{otherlanguage}}

\smalltitle{Chapter.}
\scripture{Tit. 3, 4 – 5.}

\texteparacol{\lettrine{A}{p}páruit benígnitas et humánitas Salvatóris nostri Dei:~† non ex opéribus justítiæ, quæ fécimus nos,~* sed secúndum suam misericórdiam salvos nos fecit.\par
\rr Deo grátias.}{\noindent But when the goodness and kindness of God our Saviour appeared: Not by the works of justice, which we have done, but according to his mercy, he saved us. Thanks be to God.}

\smalltitle{Hymn. Mode I.}

\rubrique{At the words \normaltext{Jesu Redémptor ómnium,} the celebrant extends and then closes his hands while turning to the cross.}

\initialscore{hy_jesu_redemptor_omnium_nativitatis_solesmes_1961}

\begin{multicols}{2}
\begin{otherlanguage}{english}
Jesus, the Redeemer of all,\\
who, being the equal of the Father's glory,\\
was begotten of the Sovereign Father\\
before the beginning of light.

Thou light and splendor of the Father,\\
Thou neverfailing hope of all,\\
give ear to the prayers\\
 which thy servants throughout the world pour forth.
 
Remember, O Creator of the world,\\
that in being born\\
Thou didst once assume the form of our body\\
 from the sacred womb of a Virgin.
 
The present day recurring\\
 in the course of each year,\\
 bears witness to this, that Thou alone didst come forth\\
 from the bosom of the Father, the salvation of the world.
 
The stars, the earth, and the seas,\\
and every creature under heaven\\
doth greet Him with a new canticle,\\
as the author of the new salvation.
 
We also, whom the sacred stream\\
 of thy blood hath cleansed,\\pay thee the tribute of a hymn\\
 on Thy birthday.
 
Jesus, unto thee be the glory,\\
 Thou who wast born of a virgin,\\
 with the Father and the loving Spirit,\\
 in the eternal age. Amen.

\end{otherlanguage}
\end{multicols}

\rubrique{\begin{otherlanguage}{english}
The cantor sings the verse to which all reply.
\end{otherlanguage}}

\smallscore{versiculus_i_vesperis_nativitatis_domini}

\begin{otherlanguage}{english} \noindent \vv Tomorrow the iniquity of the earth shall be blotted out.

\noindent \rr And the Saviour of the world shall reign over us.
 \end{otherlanguage}
 
%\rubrique{This tone is used on the feast itself, but not during the octave.}
 
\smalltitle{Antiphon at the Magnificat.}

\rubrique{Sit when the celebrant does, after he sings the first words. Rise at the end.}

\gscore[]{8. G}{an_cum_ortus_fuerit_solesmes_1961}
 
 \begin{otherlanguage}{english} \noindent Ant. When the sun shall have risen from heaven, ye shall see the King of kings proceeding from the Father, as a bridegroom from his chamber. \end{otherlanguage}
 
 \smallscore{Magnificat_8_G_Solemnis}

   \smallskip%
    \noindent%
 \begin{enumerate}[label=\arabic*.,%
			leftmargin=0pt, %
			itemindent=0mm, %
			labelsep=3pt, %
			labelwidth=0pt, %
			rightmargin=0pt, %
			parsep=0pt, %
			topsep=0pt, %
			itemsep=0pt,%
			start=3]
    \input{psaumes/Magnificat_8_Solemnis}
    \end{enumerate}

 \rubrique{\begin{otherlanguage}{english}
The Magnificat may be sung entirely in chant or sung alternating between a cantor singing chant and the choir singing a polyphonic or harmonized setting.\end{otherlanguage}}

\smalltitle{Canticle of the Blessed Virgin Mary.}

\scripture{Luc. 1, 46–55.}
\begin{multicols}{2}
%\begin{paracol}{2}
%
%\lettrine{M}agníficat~* ánima mea Dóminum.\\ Et exsultávit spíritus meus~* in Deo salutári meo.
%
% Quia respéxit humilitátem ancíllæ suæ:~* ecce enim ex hoc beátam me dicent o\-mnes generatiónes.
%
% Quia fecit mihi magna qui potens est:~* et san\-ctum nomen ejus.
%
% Et misericórdia ejus a progénie in progénies~* timéntibus eum.
%
% Fecit poténtiam in bráchio suo:~* dispérsit supérbos mente cordis sui.
%
% Depósuit poténtes de sede,~* et exaltávit húmiles.
%
% Esuriéntes implévit bonis:~* et dívites dimísit inánes.
%
% Suscépit Israel púerum suum,~* recordátus misericórdiæ suæ.
%
% Sicut locútus est ad patres no\-stros,~* Abraham, et sémini ejus in sǽcula.
%
% Glória Patri, et Fílio,~* et Spirítui San\-cto.
%
% Sicut erat in princípio, et nunc, et semper,~* et in sǽcula sæculórum. Amen.
 
% \switchcolumn
 \begin{otherlanguage}{english}
My soul \cc~* doth magnify the Lord.

And my spirit hath rejoiced~* in God my Saviour.

Because he hath regarded the humility of his handmaid;~* for behold from henceforth all generations shall call me blessed.

Because he that is mighty, hath done great things to me;~* and holy is his name.

And his mercy is from generation unto generations,~* to them that fear him.

He hath shewed might in his arm:~* he hath scattered the proud in the conceit of their heart.

He hath put down the mighty from their seat,~* and hath exalted the humble.

He hath filled the hungry with good things;~* and the rich he hath sent empty away.

He hath received Israel his servant,~* being mindful of his mercy:

As he spoke to our fathers,~* to Abraham and to his seed for ever.

Glory be to the Father, and to the Son,~* and to the Holy Ghost.

As it was in the beginning, is now,~* and ever shall be, world without end. Amen.
\end{otherlanguage}
% \end{paracol}
\end{multicols}

\rubrique{\begin{otherlanguage}{english}
Sit as the antiphon is repeated. Then the usual dialogue follows. All stand from now until the end of the office.\end{otherlanguage}}

\textebilingue{dv}{dv_eng}

%\textebilingue{oratio_15_aug_old}{oratio_15_aug_old_en}

\texteparacol{Orémus.

\lettrine{C}{o}ncéde, quǽsumus, omnípotens Deus:~* ut nos Unigéniti tui nova per carnem Natívitas líberet; quos sub peccáti jugo vetústa sérvitus tenet.
Per eúmdem Dóminum.

\rr Amen.}{\noindent Let us pray. Grant, we beseech thee, almighty God, that the new birth of thy Only-begotten Son as man may set us free, who are held by the old bondage under the yoke of sin.
Through the same Jesus Christ, thy Son, Our Lord, Who liveth and reigneth with thee in the unity of the Holy Ghost, God, world without end. Amen.}

\textebilingue{dv}{dv_eng}

\rubrique{\begin{otherlanguage}{english}
The cantor sings the verse to which all reply.
\end{otherlanguage}}

\gscore[]{2.}{BD_Solemnis_I_Vesp}

\noindent \vv Let us bless the Lord.

\noindent \rr Thanks be to God.

\textebilingue{fidelium_animae}{fidelium_animae_en}
    
%\rubrique{\begin{otherlanguage}{english}
%Then one Our Father is said, unless Compline follows, or even some pious exercises, a sermon, or Benediction of the Blessed Sacrament, in parish and other smaller churches.\end{otherlanguage}}

%\rubrique{\begin{otherlanguage}{english}
%Then the office is concluded with the following verses, the Marian antiphon, etc.
%\end{otherlanguage}}

\end{document}