%% !TEX TS-program = lualatexmk
%% !TEX parameter =  --shell-escape

\documentclass[12pt]{book}

\input{commonheaders_vespers_people} %%% will need to fix spacing problem in psalms at a later date.

\begin{document}
 
 \twosided[pc] %% to allow for the left-right switching of columns needed for bilingual liturgical typesetting (like in missals); it's unclear why this goes BEFORE everything else and not when activating the environment. %% It's OK for paper booklets to not switch, but it's probably best practice to switch.
 
  \thispagestyle{empty}
 
 {\centering{\huge\capspace{{PURIFICATION OF THE B.V.M.}}}\par}\header{purification of the b.v.m.}
 \vspace{0.5\baselineskip}
 
 {\centering{\huge{At I Vespers.}}\par}
 
\smalltitle{Before the Office.}

\rubrique{These prayers are said kneeling. End when the priest stands.}

\textebilingue{orationes_ante_do}{orationes_ante_do_eng}

\smalltitle{\prayers}

\rubrique{The celebrant intones the following verse. All sing the response and doxology.}

\begin{enpars}
\rubrique{Make the sign of the cross at the beginning and at the Magnificat, and bow the head when the Trinity is mentioned, when mention is made of the ``Holy Name'' in some way, and at ``Orémus.''}\end{enpars}

\gscore[]{℣.}{Deus_In_Adjutorium_Festivus}

\noindent
\begin{enpars}
O God, \cc{} come to my assistance. \rr O Lord, make haste to help me. Glory be to the Father, and to the Son,~* and to the Holy Ghost. As it was in the beginning, is now,~* and ever shall be, world without end. Amen. Alleluia.\end{enpars}

\smalltitle{First Antiphon. 6. F.}
\initialscore{an_o_admirabile_commercium_solesmes_1961}

\antiphona{english}{O wondrous interchange! the Creator of mankind, taking upon him a living body, vouchsafed to be born of a pure Virgin: and by his Humanity, which was begotten in no earthly wise, hath made us partakers of his Divinity. \end{enpars}

\rubrique{\begin{enpars}
Sit at the word ``Sede.'' The cantor sings odd verses; all sing even verses, and this for the hymn and Magnificat, when sung entirely in chant, as well.
\end{enpars}}

\psalmus{109}{109_6_alt}{109_6_alt}{109_en}

\rubrique{\begin{enpars}
Clerics in descending order, or a cantor, intone the subsequent antiphons.\end{enpars}}

 \gscore[2. Ant.]{3. a2}{an_quando_natus_es_solesmes_1961}
 \antiphona{english}{When thou wast born all ineffably of a Virgin, then were the Scriptures fulfilled; thou camest down like the dew upon the fleece of wool, to bring salvation unto all mankind; we praise thee, o our God}
\psalmus{112}{112_3a2}{112_3a2_g}{112_en}

  \gscore[3. Ant.]{4. E}{an_rubum_quem_solesmes_1961}
  \antiphona{english}{In the bush which Moses saw unconsumed, we recognize the preservation of thy glorious virginity: holy Mother of God, intercede for us.}
  \psalmus{121}{121_4E}{121_4E}{121_en}

  \gscore[4. Ant.]{1. f}{an_germinavit_radix_solesmes_1961}
\antiphona{english}{The Root of Jesse hath budded, the Star hath come out of Jacob, the Virgin hath borne the Saviour: we praise thee, o our God}
  \psalmus{126}{126_1f}{126_1}{126_en}

 \gscore[5. Ant.]{2. D}{an_ecce_maria_genuit_solesmes_1961}
\antiphona{english}{Lo, Mary hath brought forth the Saviour of whom when John saw him, he did proclaim: Behold the Lamb of God that taketh away the sins of the world (alleluia.)}
\psalmus{147}{147_2}{147_2}{147_en}

\rubrique{\begin{enpars}
Having sung the last antiphon in full, stand when the celebrant does.
\end{enpars}}

\chapterlateng
\scripture{Malach. 3, 1.}

\texteparacol{\lettrine{E}{c}ce ego mitto Angelum meum, et præparábit viam ante fáciem meam.~† Et statim véniet ad templum sanctum suum Dominátor, quem vos quǽritis,~* et Angelus Testaménti, quem vos vultis.

\rr Deo grátias.}{english}{\noindent Behold I send my angel, and he shall prepare the way before my face. And presently the Lord, whom you seek, and the angel of the testament, whom you desire, shall come to his temple. Thanks be to God.}

% Only the usual tone of the hymn is given.

\hymnlateng

\rubrique{The first verse is sung kneeling.}

\gscore[]{1.}{hy_ave_maris_stella_mode_1_LA1949}

\begin{multicols}{2}
\raggedcolumns
\begin{enpars}
Hail, star of the sea,\\
loving Mother of God,\\
and also always a virgin,\\
Happy gate of heaven.

Receiving that Ave\\
from Gabriel's mouth\\
confirm us in peace,\\
Reversing Eva's name.

Break the chains of sinners,\\
Bring light to the blind,\\
Drive away our evils,\\
Ask for all good.

Show thyself to be a mother,\\
May he accept prayers through thee,\\
he who, born for us,\\
Chose to be thine.

O unique virgin,\\
Meek above all,\\
Make us, absolved from sin,\\
Gentle and chaste.

Keep life pure,\\
Make the journey safe,\\
So that, seeing Jesus,\\
We may always rejoice together.

Let there be praise to God the Father,\\
Glory to Christ in the highest,\\
To the Holy Spirit,\\
One honor to all three. Amen
\end{enpars}
\end{multicols}

\rubrique{\begin{enpars}
The cantor sings the verse to which all reply.
\end{enpars}}

\smallscore{versiculus_2_feb}
\begin{enpars} \noindent \vv It was revealed unto Simeon by the Holy Ghost.

\noindent \rr That he should not see death before he had seen the Lord's Christ.
 \end{enpars}

\smalltitle{Antiphon at the Magnificat.}

\rubrique{Sit when the celebrant does, after he sings the first words. Rise at the end.}

\gscore[]{1. D}{an_senex_puerum_portabat_solesmes_1961.gabc}
 
\smallscore{Magnificat_1D_solemnis}

   \smallskip%
    \noindent%
 \begin{enumerate}[label=\arabic*.,%
			leftmargin=0pt, %
			itemindent=0mm, %
			labelsep=3pt, %
			labelwidth=0pt, %
			rightmargin=0pt, %
			parsep=0pt, %
			topsep=0pt, %
			itemsep=0pt,%
			start=3]
    \input{psaumes/Magnificat_1_Solemnis}
    \end{enumerate}

\myrule

\smalltitle{Canticle of the Blessed Virgin Mary.}

\scripture{Luc. 1, 46–55.}

%\begin{paracol}{2}
\begin{multicols}{2}
%\lettrine{M}agníficat~* ánima mea Dóminum.\\ Et exsultávit spíritus meus~* in Deo salutári meo.
%
% Quia respéxit humilitátem ancíllæ suæ:~* ecce enim ex hoc beátam me dicent o\-mnes generatiónes.
%
% Quia fecit mihi magna qui potens est:~* et san\-ctum nomen ejus.
%
% Et misericórdia ejus a progénie in progénies~* timéntibus eum.
%
% Fecit poténtiam in bráchio suo:~* dispérsit supérbos mente cordis sui.
%
% Depósuit poténtes de sede,~* et exaltávit húmiles.
%
% Esuriéntes implévit bonis:~* et dívites dimísit inánes.
%
% Suscépit Israel púerum suum,~* recordátus misericórdiæ suæ.
%
% Sicut locútus est ad patres no\-stros,~* Abraham, et sémini ejus in sǽcula.
%
% Glória Patri, et Fílio,~* et Spirítui San\-cto.
%
% Sicut erat in princípio, et nunc, et semper,~* et in sǽcula sæculórum. Amen.
% 
% \switchcolumn
 \begin{enpars}
My soul \cc~* doth magnify the Lord.

And my spirit hath rejoiced~* in God my Saviour.

Because he hath regarded the humility of his handmaid;~* for behold from henceforth all generations shall call me blessed.

Because he that is mighty, hath done great things to me;~* and holy is his name.

And his mercy is from generation unto generations,~* to them that fear him.

He hath shewed might in his arm:~* he hath scattered the proud in the conceit of their heart.

He hath put down the mighty from their seat,~* and hath exalted the humble.

He hath filled the hungry with good things;~* and the rich he hath sent empty away.

He hath received Israel his servant,~* being mindful of his mercy:

As he spoke to our fathers,~* to Abraham and to his seed for ever.

Glory be to the Father, and to the Son,~* and to the Holy Ghost.

As it was in the beginning, is now,~* and ever shall be, world without end. Amen.
\end{enpars}
% \end{paracol}
\end{multicols}

\rubrique{\begin{enpars}
Sit as the antiphon is repeated. Then the usual dialogue follows. All stand from now until the end of the office, unless one sits for the antiphon of a commemorated Sunday , according to custom.\end{enpars}}

\textebilingue{dv}{dv_eng}

\texteparacol{Orémus.
   
\lettrine{O}{m}nípotens sempitérne Deus, majestátem tuam súpplices exorámus:~† ut sicut unigénitus Fílius tuus hodiérna die cum nostræ carnis substántia in templo est præsentátus;~* ita nos fácias purificátis tibi méntibus præsentári.
Per eúmdem Dóminum.
   
  \rr Amen.}{english}{\noindent Let us pray.
Almighty and everliving God, we humbly beseech thy majesty, that as thy Only-begotten Son was this day presented in the temple in substance of our flesh, so we may be presented unto thee with pure and clean hearts. Through the same Jesus Christ, thy Son, our Lord, who liveth and reigneth with thee in the unity of the Holy Ghost, God, world without end. Amen.}

\rubrique{Commemoration of the preceding feast of Saint Ignatius, bishop and martyr; a commemoration of the Sunday is also made in the appropriate place, whether Sunday be the next day or the preceding day.}

\gscore[Ant.]{1.}{an_cs_qui_vult_venire_solesmes_1961}
%
\antiphona{english}{If any man will come after me, let him deny himself and take up his cross daily and follow me}
%%translation
%
\texteparacol{\vv Justus ut palma florébit.

\rr Sicut cedrus Líbani multiplicábitur.}{english}{
 \noindent  \vv The righteous shall flourish like the palm-tree.

\noindent \rr He shall grow like a cedar in Lebanon.}
%
\texteparacol{Orémus.

\lettrine{I}{n}firmitátem nostram réspice, omnípotens Deus:~† et quia pondus própriæ actiónis gravat,~* beáti lgnátii Mártyris tui atque Pontíficis intercéssio gloriósa nos prótegat. Per Dóminum.}{english}{\noindent Let us pray. Mercifully consider our weakness, o almighty God, and whereas by the burden of our sins we are sore, let Ignatius thy martyr and bishop be mercifully pleased to deliver us from all things which may hurt our bodies, and from all evil thoughts which may defile our souls. Through Jesus Christ, thy Son our Lord, who liveth and reigneth with thee, in the unity of the Holy Ghost, God, world without end. Amen.}

\textebilingue{dv}{dv_eng}

\rubrique{\begin{enpars}
The cantor sings the verse to which all reply. One of the following tones is chosen, at the choirmaster's discretion.
\end{enpars}}

\gscore[]{2.}{BD_Solemnis_I_Vesp}

\noindent \vv Let us bless the Lord.

\noindent \rr Thanks be to God.

\newpage

\rubrique{Or the following tone may be used:}

\gscore[]{1.}{BD_BVM}

\textebilingue{fidelium_animae}{fidelium_animae_en}

\end{document}