% !TEX TS-program = lualatexmk
% !TEX parameter =  --shell-escape

\documentclass[12pt]{book}

\input{commonheaders_vespers_people} %%% will need to fix spacing problem in psalms at a later date.

\begin{document}
 
 \twosided[pc] %% to allow for the left-right switching of columns needed for bilingual liturgical typesetting (like in missals); it's unclear why this goes BEFORE everything else and not when activating the environment. %% It's OK for paper booklets to not switch, but it's probably best practice to switch.
 
  \thispagestyle{empty}
 
 {\centering{\huge\capspace{{PURIFICATION OF THE B.V.M.}}}\par}\header{purification of the b.v.m.}
 
\vspace{0.5\baselineskip}
 
 {\centering{\huge{At II Vespers.}}\par}
 
\smalltitle{Before the Office.}

\rubrique{These prayers are said kneeling. End when the priest stands.}

\textebilingue{orationes_ante_do}{orationes_ante_do_eng}

\smalltitle{\prayers}

\rubrique{The celebrant intones the following verse. All sing the response and doxology.}

\begin{enpars}
\rubrique{Make the sign of the cross at the beginning and at the Magnificat, and bow the head when the Trinity is mentioned, when mention is made of the ``Holy Name'' in some way, and at ``Orémus.''}\end{enpars}

\gscore[]{℣.}{Deus_In_Adjutorium_Festivus}

\noindent
\begin{enpars}
O God, \cc{} come to my assistance. \rr O Lord, make haste to help me. Glory be to the Father, and to the Son,~* and to the Holy Ghost. As it was in the beginning, is now,~* and ever shall be, world without end. Amen. Alleluia.\end{enpars}

\smalltitle{First Antiphon. 3. b.}

\initialscore{an_simeon_justus_solesmes_1961}

\antiphona{english}{Simeon was just and devout, waiting for the consolation of Israel, and the Holy Ghost was upon him}

\rubrique{\begin{enpars}
Sit at the word ``Sede.'' The cantor sings odd verses; all sing even verses, and this for the hymn and Magnificat, when sung entirely in chant, as well.
\end{enpars}}

\psalmus{109}{109_3b}{109_3a_b}{109_en}

\rubrique{\begin{enpars}
Clerics in descending order, or a cantor, intone the subsequent antiphons.\end{enpars}}

\gscore[2. Ant.]{7 .a}{an_responsum_accepit__solesmes_1961} 
 
 \antiphona{english}{It was revealed unto Simeon by the Holy Ghost that he should not see death before he had seen the Lord.
\end{enpars}

\psalmus{112}{112_7a}{112_7}{112_en}

\gscore[3. Ant.]{3. b}{an_accipiens_simeon_solesmes_1961}
  
\antiphona{english}{Simeon took the Child up in his arms, and gave thanks, and blessed the Lord. \end{enpars}

 \psalmus{121}{121_3b}{121_3a_b}{121_en}

\gscore[4. Ant.]{8. c}{an_lumen_ad_revelationem__solesmes_1961}

\antiphona{english}{A light to lighten the Gentiles, and the glory of thy people Israel}

\psalmus{126}{126_8c}{126_8}{126_en}

\gscore[5. Ant.]{8. G}{an_obtulerunt_pro_eo_solesmes_1961}
 
\antiphona{english}{They offered for him unto the Lord, a pair of turtle doves, or two young pigeons}

\psalmus{147}{147_8G}{147_8}{147_en}

\rubrique{\begin{enpars}
Having sung the last antiphon in full, stand when the celebrant does.
\end{enpars}}

\chapterlateng
\scripture{Malach. 3, 1.}

\texteparacol{\lettrine{E}{c}ce ego mitto Angelum meum, et præparábit viam ante fáciem meam.~† Et statim véniet ad templum sanctum suum Dominátor, quem vos quǽritis,~* et Angelus Testaménti, quem vos vultis.

\rr Deo grátias.}{\noindent Behold I send my angel, and he shall prepare the way before my face. And presently the Lord, whom you seek, and the angel of the testament, whom you desire, shall come to his temple. Thanks be to God.}

%\textebilingue{cap_15_aug_old}{cap_15_aug_old_en}

%% Only the usual tone of the hymn is given.

\hymnlateng

\rubrique{The first verse is sung kneeling.}

\grechangenextscorelinedim{3}{spacebeneathtext}{0.1cm}{scalable}
\gscore[]{1.}{hy_ave_maris_stella_mode_1_LA1949}

\textes{ams_en}

\textes{cantor_rubric}

\smallscore{versiculus_2_feb}
\begin{enpars} \noindent \vv It was revealed unto Simeon by the Holy Ghost.

\noindent \rr That he should not see death before he had seen the Lord's Christ.
 \end{enpars}

\smalltitle{Antiphon at the Magnificat.}

\rubrique{Sit when the celebrant does, after he sings the first words. Rise at the end.}

\gscore[]{8. G*}{an_hodie_beata_virgo_solesmes_1961}
 
\smallscore{Magnificat_8_Gstar_Solemnis}

   \smallskip%
    \noindent%
 \begin{enumerate}[label=\arabic*.,%
			leftmargin=0pt, %
			itemindent=0mm, %
			labelsep=3pt, %
			labelwidth=0pt, %
			rightmargin=0pt, %
			parsep=0pt, %
			topsep=0pt, %
			itemsep=0pt,%
			start=3]
    \input{psaumes/Magnificat_8_Solemnis}
    \end{enumerate}

\myrule

\smalltitle{Canticle of the Blessed Virgin Mary.}

\scripture{Luc. 1, 46–55.}

%\begin{paracol}{2}
\begin{multicols}{2}
%\lettrine{M}agníficat~* ánima mea Dóminum.\\ Et exsultávit spíritus meus~* in Deo salutári meo.
%
% Quia respéxit humilitátem ancíllæ suæ:~* ecce enim ex hoc beátam me dicent o\-mnes generatiónes.
%
% Quia fecit mihi magna qui potens est:~* et san\-ctum nomen ejus.
%
% Et misericórdia ejus a progénie in progénies~* timéntibus eum.
%
% Fecit poténtiam in bráchio suo:~* dispérsit supérbos mente cordis sui.
%
% Depósuit poténtes de sede,~* et exaltávit húmiles.
%
% Esuriéntes implévit bonis:~* et dívites dimísit inánes.
%
% Suscépit Israel púerum suum,~* recordátus misericórdiæ suæ.
%
% Sicut locútus est ad patres no\-stros,~* Abraham, et sémini ejus in sǽcula.
%
% Glória Patri, et Fílio,~* et Spirítui San\-cto.
%
% Sicut erat in princípio, et nunc, et semper,~* et in sǽcula sæculórum. Amen.
% 
% \switchcolumn
 \begin{enpars}
My soul \cc~* doth magnify the Lord.

And my spirit hath rejoiced~* in God my Saviour.

Because he hath regarded the humility of his handmaid;~* for behold from henceforth all generations shall call me blessed.

Because he that is mighty, hath done great things to me;~* and holy is his name.

And his mercy is from generation unto generations,~* to them that fear him.

He hath shewed might in his arm:~* he hath scattered the proud in the conceit of their heart.

He hath put down the mighty from their seat,~* and hath exalted the humble.

He hath filled the hungry with good things;~* and the rich he hath sent empty away.

He hath received Israel his servant,~* being mindful of his mercy:

As he spoke to our fathers,~* to Abraham and to his seed for ever.

Glory be to the Father, and to the Son,~* and to the Holy Ghost.

As it was in the beginning, is now,~* and ever shall be, world without end. Amen.
\end{enpars}
% \end{paracol}
\end{multicols}

\rubrique{\begin{enpars}
Sit as the antiphon is repeated. Then the usual dialogue follows. All stand from now until the end of the office, unless one sits for the antiphon of a commemorated Sunday or Lenten feria, according to custom.\end{enpars}}

\textebilingue{dv}{dv_eng}

\texteparacol{Orémus.
   
\lettrine{O}{m}nípotens sempitérne Deus, majestátem tuam súpplices exorámus:~† ut sicut unigénitus Fílius tuus hodiérna die cum nostræ carnis substántia in templo est præsentátus;~* ita nos fácias purificátis tibi méntibus præsentári.
Per eúmdem Dóminum.
   
   ℟. Amen.}
{\noindent Let us pray.
Almighty and everliving God, we humbly beseech thy majesty, that as thy Only-begotten Son was this day presented in the temple in substance of our flesh, so we may be presented unto thee with pure and clean hearts. Through the same Jesus Christ, thy Son, our Lord, who liveth and reigneth with thee in the unity of the Holy Ghost, God, world without end. Amen.}

\rubrique{No commemorations follow, except if there is one to be made of the Sunday.}

\textebilingue{dv}{dv_eng}

\rubrique{\begin{enpars}
The cantor sings the verse to which all reply. One of the following tones is chosen at the choirmaster's discretion.
\end{enpars}}

%%not needed, but kept here since a copy and paste job from Circumcision is needed for I Vespers.
%%%\smalltitle{At I Vespers.}
%%%
%%%\gscore[]{2.}{BD_Solemnis_I_Vesp}


%%\smalltitle{At II Vespers.}

\gscore[]{6.}{BD_Solemnis_mode_6}

\noindent \vv Let us bless the Lord.

\noindent \rr Thanks be to God.

\rubrique{Or the following tone may be used:}

\gscore[]{5.}{BD_Solemnis_mode_5}

\rubrique{Or:}

\gscore[]{1.}{BD_BVM}

\textebilingue{fidelium_animae}{fidelium_animae_en}

\end{document}