% !TEX TS-program = lualatexmk
% !TEX parameter =  --shell-escape

\documentclass[12pt]{book}

\input{commonheaders_vespers_people}

\begin{document}
 
 \twosided[pc] %% to allow for the left-right switching of columns needed for bilingual liturgical typesetting (like in missals); it's unclear why this goes BEFORE everything else and not when activating the environment. %% It's OK for paper booklets to not switch, but it's probably best practice to switch.
 
 \thispagestyle{empty}
 
 {\centering{\huge\capspace{{IV SUNDAY OF ADVENT}}}\par}\header{vespers of the iv sunday of advent.}
 \vspace{\baselineskip}
{\centering{\huge\capspace{{AT VESPERS.}}}\par}
 
\smalltitle{Before the Office.}

\rubrique{These prayers are said kneeling. End when the priest stands.}

\textebilingue{orationes_ante_do}{orationes_ante_do_eng}

\smalltitle{\prayers}

\rubrique{The celebrant intones the following verse. All sing the response and doxology.}

\begin{enpars}
\rubrique{Make the sign of the cross at the beginning and at the Magnificat, and bow the head when the Trinity is mentioned, when mention is made of the ``Holy Name'' in some way, and at ``Orémus.''}\end{enpars}

\gscore[]{℣.}{Deus_In_Adjutorium_Festivus}

\noindent
\begin{enpars}
O God, \cc{} come to my assistance. \rr O Lord, make haste to help me. Glory be to the Father, and to the Son,~* and to the Holy Ghost. As it was in the beginning, is now,~* and ever shall be, world without end. Amen. Alleluia.\end{enpars}

\smalltitle{First Antiphon. 1. g.}

\gscore[]{}{canite_tuba_incipit}

\rubrique{\begin{enpars}
Sit at the word ``Sede.'' The cantor sings odd verses; all sing even verses, and this for the hymn and Magnificat as well.
\end{enpars}}

\psalmus{109}{109_1g}{109_1}{109_en}

\smallscore{an_canite_tuba_solesmes_1961}

\antiphona{english}{Blow ye the trumpet in Zion, for the day of the Lord is nigh at hand: behold, He cometh to save us! Alleluia, Alleluia}

\rubrique{\begin{enpars}
Clerics in descending order, or a cantor, intone the subsequent antiphons. The cantor then continues the psalm.
\end{enpars}}

\gscore[2. Ant.]{1. f}{ecce_veniet_desideratus_incipit}

\psalmus{110}{110_1f}{110_1}{110_en}

\smallscore{an_ecce_veniet_desideratus_solesmes_1961}

\antiphona{english}{Behold, the desire of all nations shall come; and the house of the Lord shall be filled with glory. Alleluia}

\gscore[3. Ant.]{1. g}{erunt_prava_incipit}

\psalmus{111}{111_1g}{111_1}{111_en}

\smallscore{an_erunt_prava_solesmes_1961}

\antiphona{english}{The crooked shall be made straight, and the rough places plain; O Lord, come, and make no tarrying. Alleluia}

\gscore[4. Ant.]{1. f}{dominus_veniet_incipit}

\psalmus{112}{112_1f}{112_1}{112_en}

\smallscore{an_dominus_veniet_solesmes_1961}

\antiphona{english}{The Lord cometh! Go ye out to meet him, and say: How great is his dominion, and of his kingdom there shall be no end! He is the mighty God, the Ruler, the Prince of Peace. Alleluia, Alleluia}

\gscore[5. Ant.]{2. D}{omnipotens_sermo_incipit}

\psalmus{113}{113_2}{113_2}{113_en}

\smallscore{an_omnipotens_sermo_solesmes_1961}

\antiphona{english}{Thine almighty Word, o Lord, shall leap down out of thy royal throne. Alleluia}

\rubrique{\begin{enpars}
Having sung the last antiphon in full, stand when the celebrant does.
\end{enpars}}

\rubrique{\begin{enpars}
The chapter is proper to each Sunday of Advent, while the hymn and verse are always the same.\end{enpars}}

\chapterlateng
\scripture{1 Cor. 4, 1 – 2.}

\texteparacol{\lettrine{F}{r}atres: Sic nos exístimet homo ut minístros Christi,~† et dispensatóres mysteriórum Dei.~* Hic jam quǽritur inter dispensatóres, ut fidélis quis inveniátur}{english}{\noindent Brethren: Let a man so account of us as of the ministers of Christ, and the dispensers of the mysteries of God. Here now it is required among the dispensers, that a man be found faithful.}

\hymnlateng

\gscore[]{4.}{hy_creator_alme_siderum_solesmes_1961}

\begin{multicols}{2}
\begin{enpars}

Bright builder of the heavenly poles,\\
Eternal light of faithful souls,\\
Jesus, Redeemer of mankind,\\
Our humble prayers vouchsafe to mind.

Who, lest the fraud of hell’s black king\\
Should all men to destruction bring,\\
Didst, by an act of generous love,\\
The fainting world’s physician prove.

Who, that thou mightst our ransom pay\\
And wash the stains of sin away,\\
Wouldst from a Virgin’s womb proceed\\
And on the cross a victim bleed.

Whose glorious power, whose saving name\\
No sooner any voice can frame,\\
But heaven and earth and hell agree\\
To honour them with trembling knee.

thee, Christ, who at the latter day\\
Shalt be our Judge, we humbly pray\\
Such arms of heavenly grace to send\\
As may thy Church from foes defend.

Be glory given and honour done\\
To God the Father and the Son\\
And to the Holy Ghost on high,\\
From age to age eternally. Amen.
\end{enpars}
\end{multicols}

\textes{cantor_rubric}

\smallscore{versiculus_adventus}
\begin{enpars} \noindent \vv Drop down dew, ye heavens, from above, and let the clouds rain the Just.

\noindent \rr Let the earth be opened, and bud forth the Saviour.
 \end{enpars}

\rubrique{\begin{enpars}
As this Sunday falls after December 17, the \normaltext{O} antiphon corresponding to the day is sung; there is no antiphon for the Sunday itself.

On December 20 and 21, the Sunday is instead commemorated in the office of Saint Thomas, Apostle.\end{enpars}}

 \rubrique{\begin{enpars}
The antiphon \normaltext{O} is sung in full before and after, or ``doubled''. All remain standing both before and after the canticle, from now until the end of the office.\end{enpars}}

 \rubrique{\begin{enpars}
The celebrant intones the \normaltext{Magnificat} antiphon, and having sung it in full, the cantor intones the \normaltext{Ma\-gnificat.}\end{enpars}}

\myrule \smalltitle{December 18.}

\gscore[]{2. D}{an_o_adonai_solesmes_1961}
\antiphona{english}{O Adonai, and Ruler of the house of Israel, who didst appear unto Moses in the burning bush, and gavest him the law in Sinai; come, to redeem us with an outstretched arm}

\myrule \smalltitle{December 19.}

\gscore[]{2. D}{an_o_radix_jesse_solesmes_1961}
\antiphona{english}{O Root of Jesse, which standest for an ensign of the people, at whom the kings shall shut their mouths, to whom the Gentiles shall seek; come to deliver us, make no tarrying}

\myrule

\smalltitle{December 22.}

\gscore[]{2. D}{an_o_rex_gentium_solesmes_1961}
\antiphona{english}{O King of the Gentiles, yea, and Desire thereof, o Cornerstone that makest of twain one; come to save man, whom thou hast made of the dust of the earth}

\myrule

\smalltitle{December 23.}

\gscore[]{2. D}{an_o_emmanuel_solesmes_1961}
\antiphona{english}{O Emmanuel, our King and our Lawgiver, Longing of the Gentiles, yea, and Salvation thereof; come to save us, o Lord our God}

\myrule

 \rubrique{\begin{enpars}
The celebrant intones the \normaltext{Magnificat} antiphon. The cantor continues the \normaltext{Ma\-gnificat.}\end{enpars}}

\smalltitle{Canticle of the Blessed Virgin Mary.}

\scripture{Luc. 1, 46–55.}

\smallscore{Magnificat_2_Solemnis} %%Magnificat file 

%\gabcsnippet{}%%

\begin{enumerate}[label=\arabic*.,%
			leftmargin=0pt, %
			itemindent=0mm, %
			labelsep=3pt, %
			labelwidth=0pt, %
			rightmargin=0pt, %
			parsep=0pt, %
			topsep=0pt, %
			itemsep=0pt,%
			start=3]
			\input{psaumes/Magnificat_2_Solemnis}
%%Magnificat
%% each line requires \item
\end{enumerate}

\begin{multicols}{2}
%\begin{paracol}{2}
%\lettrine{M}agníficat~* ánima mea Dóminum.\\ Et exsultávit spíritus meus~* in Deo salutári meo.
%
% Quia respéxit humilitátem ancíllæ suæ:~* ecce enim ex hoc beátam me dicent o\-mnes generatiónes.
%
% Quia fecit mihi magna qui potens est:~* et san\-ctum nomen ejus.
%
% Et misericórdia ejus a progénie in progénies~* timéntibus eum.
%
% Fecit poténtiam in bráchio suo:~* dispérsit supérbos mente cordis sui.
%
% Depósuit poténtes de sede,~* et exaltávit húmiles.
%
% Esuriéntes implévit bonis:~* et dívites dimísit inánes.
%
% Suscépit Israel púerum suum,~* recordátus misericórdiæ suæ.
%
% Sicut locútus est ad patres no\-stros,~* Abraham, et sémini ejus in sǽcula.
%
% Glória Patri, et Fílio,~* et Spirítui San\-cto.
%
% Sicut erat in princípio, et nunc, et semper,~* et in sǽcula sæculórum. Amen.
% 
% \switchcolumn
 \begin{enpars}
My soul \cc~* doth magnify the Lord.

And my spirit hath rejoiced~* in God my Saviour.

Because he hath regarded the humility of his handmaid;~* for behold from henceforth all generations shall call me blessed.

Because he that is mighty, hath done great things to me;~* and holy is his name.

And his mercy is from generation unto generations,~* to them that fear him.

He hath shewed might in his arm:~* he hath scattered the proud in the conceit of their heart.

He hath put down the mighty from their seat,~* and hath exalted the humble.

He hath filled the hungry with good things;~* and the rich he hath sent empty away.

He hath received Israel his servant,~* being mindful of his mercy:

As he spoke to our fathers,~* to Abraham and to his seed for ever.

Glory be to the Father, and to the Son,~* and to the Holy Ghost.

As it was in the beginning, is now,~* and ever shall be, world without end. Amen.
\end{enpars}
% \end{paracol}
\end{multicols}

 \rubrique{\begin{enpars}
Then the usual dialogue follows.\end{enpars}}

 \rubrique{\begin{enpars}
The celebrant sings the collect of the day. The Suffrage of All Saints is omitted for all of Advent}}

\texteparacol{\vv Dóminus vobíscum.

\rr Et cum spíritu tuo.

\lettrine{E}{x}cita, quǽsumus Dómine, poténtiam tuam, et veni: et magna nobis virtúte succúrre:~† ut per auxílium grátiæ tuæ, quod nostra peccáta præpédiunt,~* indulgéntia tuæ propitiatiónis accéleret. Qui vivis et regnas.}{english}{\noindent \vv The Lord be with you.

\noindent \rr  And with thy spirit.

\noindent Stir up, we beseech thee, o Lord, thy power, and come; make haste to our aid with thy great might; that, by the help of thy grace, that which is hindered by our sins may be hastened by thy merciful forgiveness. Who livest and reignest with God the Father.}

\texteparacol{\vv Dóminus vobíscum.

\rr Et cum spíritu tuo.}
{english}
{\noindent \vv The Lord be with you.

\noindent \rr  And with thy spirit.}

\textes{cantor_rubric}

\gscore[]{6.}{BD_Dominicis_Adventus_Quadragesimae}

\vv Let us bless the Lord.

\rr Thanks be to God.

\textebilingue{fidelium_animae}{fidelium_animae_en}

%\rubrique{\begin{enpars}
%Then one Our Father is said, unless Compline follows, or even some pious exercises, a sermon, or Benediction of the Blessed Sacrament, in parish and other smaller churches}}

%\rubrique{\begin{enpars}
%Then the office is concluded with the following verses, the Marian antiphon, etc.
%\end{enpars}}


\end{document}