% !TEX TS-program = lualatexmk
% !TEX parameter =  --shell-escape

\documentclass[12pt]{book}

\input{commonheaders_vespers_people} %%% will need to fix spacing problem in psalms at a later date.

\begin{document}
 
 \twosided[pc] %% to allow for the left-right switching of columns needed for bilingual liturgical typesetting (like in missals); it's unclear why this goes BEFORE everything else and not when activating the environment. %% It's OK for paper booklets to not switch, but it's probably best practice to switch.
 
  \thispagestyle{empty}
 
 {\centering{\huge\capspace{{IMMACULATE CONCEPTION\\[0.5\baselineskip] OF THE B.V.M.}}}\par}\header{immaculate conception of the b.v.m.}
 
\smalltitle{Before the Office.}

\rubrique{These prayers are said kneeling. End when the priest stands.}

\textebilingue{orationes_ante_do}{orationes_ante_do_eng}

\smalltitle{\prayers}

\rubrique{The celebrant intones the following verse. All sing the response and doxology.}

\begin{enpars}
\rubrique{Make the sign of the cross at the beginning and at the Magnificat, and bow the head when the Trinity is mentioned, when mention is made of the ``Holy Name'' in some way, and at ``Orémus.''}\end{enpars}

\gscore[]{℣.}{Deus_in_adjutorium_tonus_solemnior}

\noindent
\begin{enpars}
O God, \cc{} come to my assistance. \rr O Lord, make haste to help me. Glory be to the Father, and to the Son,~* and to the Holy Ghost. As it was in the beginning, is now,~* and ever shall be, world without end. Amen. Alleluia}

\rubrique{On the octave day (double major):}

\gscore[]{℣.}{Deus_In_Adjutorium_Festivus}

\smalltitle{First Antiphon. 1. g2.}

\initialscore{an_tota_pulchra_es_maria__solesmes_1961}

\antiphona{english}{Thou art all fair, O Mary, there is no spot of original sin in thee}

\rubrique{\begin{enpars}
Sit at the word ``Sede.'' The cantor sings odd verses; all sing even verses, and this for the hymn and Magnificat, when sung entirely in chant, as well.
\end{enpars}}

\psalmus{109}{109_1g2}{109_1}{109_en}

\rubrique{\begin{enpars}
Clerics in descending order, or a cantor, intone the subsequent antiphons}}

\gscore[2. Ant.]{8. G}{an_vestimentum_tuum_solesmes_1961}
\antiphona{english}{Thy raiment is white as snow, and thy countenance as the sun}
\psalmus{112}{112_8G}{112_8}{112_en}

\gscore[3. Ant.]{8. c}{an_tu_gloria_jerusalem_solesmes_1961}
\antiphona{english}{Thou art the exaltation of Jerusalem, thou art the great glory of Israel, thou art the great rejoicing of our nation. \end{enpars}
 \psalmus{121}{121_8c}{121_8}{121_en}

\gscore[4. Ant.]{7. a}{an_benedicta_es_tu_virgo_solesmes}
\antiphona{english}{Virgin Mary, blessed art thou of the Most High God, above all the women upon the earth}
\psalmus{126}{126_7a}{126_7}{126_en}

\gscore[5. Ant.]{3. a2}{an_trahe_nos_virgo_solesmes_1961}
\antiphona{english}{Draw us, Maiden undefiled, we will run after thee in the odour of thy perfumes}
\psalmus{147}{147_3a2}{147_3a2_g}{147_en}

\rubrique{\begin{enpars}
Having sung the last antiphon in full, stand when the celebrant does.
\end{enpars}}

\chapterlateng
\scripture{Prov. 8, 22 – 24.}

\texteparacol{\lettrine{D}{o}minus possédit me in inítio viárum suárum, ántequam quidquam fáceret a princípio.~† Ab ætérno ordináta sum, et ex antíquis ántequam terra fíeret:~* nondum erant abýssi, et ego jam concépta eram.

\rr Deo grátias.}{\noindent The Lord possessed me in the beginning of his ways, before he made any thing from the beginning. I was set up from eternity, and of old before the earth was made. The depths were not as yet, and I was already conceived. Thanks be to God.}

%\textebilingue{cap_15_aug_old}{cap_15_aug_old_en}

%% Only the usual tone of the hymn is given.

\hymnlateng

\rubrique{The first verse is sung kneeling.}

\gscore[]{1.}{hy_ave_maris_stella_mode_1_LA1949}

\begin{multicols}{2}
\raggedcolumns
\begin{enpars}
Hail, star of the sea,\\
loving Mother of God,\\
and also always a virgin,\\
Happy gate of heaven.

Receiving that Ave\\
from Gabriel's mouth\\
confirm us in peace,\\
Reversing Eva's name.

Break the chains of sinners,\\
Bring light to the blind,\\
Drive away our evils,\\
Ask for all good.

Show thyself to be a mother,\\
May he accept prayers through thee,\\
he who, born for us,\\
Chose to be thine.

O unique virgin,\\
Meek above all,\\
Make us, absolved from sin,\\
Gentle and chaste.

Keep life pure,\\
Make the journey safe,\\
So that, seeing Jesus,\\
We may always rejoice together.

Let there be praise to God the Father,\\
Glory to Christ in the highest,\\
To the Holy Spirit,\\
One honor to all three. Amen
\end{enpars}
\end{multicols}

\rubrique{\begin{enpars}
The cantor sings the verse to which all reply.
\end{enpars}}

\smallscore{versiculus_8_dec}
\begin{enpars}\noindent \vv This day is the Holy Virgin Mary conceived without sin.

\noindent \rr The Virgin's foot hath bruised the serpent's head.
 \end{enpars}

\newpage
\smalltitle{Antiphon at the Magnificat.}

\rubrique{Sit when the celebrant does, after he sings the first words. Rise at the end.}

 \smalltitle{At I Vespers.}

\gscore[]{8. G}{an_beatam_me_dicent_quia_fecit_solesmes_1961}
 
\antiphona{english}{All generations shall call me blessed, because he that is mighty, hath done great things to me, alleluia}
 
 \smallscore{Magnificat_8_G_Solemnis}

   \smallskip%
    \noindent%
 \begin{enumerate}[label=\arabic*.,%
			leftmargin=0pt, %
			itemindent=0mm, %
			labelsep=3pt, %
			labelwidth=0pt, %
			rightmargin=0pt, %
			parsep=0pt, %
			topsep=0pt, %
			itemsep=0pt,%
			start=3]
    \input{psaumes/Magnificat_8_Solemnis}
    \end{enumerate}

\myrule

 \smalltitle{At II Vespers.}
\gscore[]{1. f}{an_hodie_egressa_est_solesmes_1961}
 
\antiphona{english}{This day is a rod sprung from the root of Jesse: this day is Mary conceived without any stain of sin: this day hath she bruised the head of the old serpent, alleluia}
 
 \smallscore{Magnificat_1_f_Solemnis}

    \smallskip%
    \noindent%
 \begin{enumerate}[label=\arabic*.,%
			leftmargin=0pt, %
			itemindent=0mm, %
			labelsep=3pt, %
			labelwidth=0pt, %
			rightmargin=0pt, %
			parsep=0pt, %
			topsep=0pt, %
			itemsep=0pt,%
			start=3]
\item Quia respéxit humilitátem \textit{ancíllæ} \textbf{su}æ:~* ecce enim ex hoc beátam me dicent omnes gene\textit{rati}\textbf{ó}nes.
\item Quia fecit mihi \textit{magna} \textit{qui} \textbf{pot}\textbf{ens} est:~* et sanctum \textit{nomen} \textbf{e}jus.
\item Et misericórdia ejus a progéni\textit{e} \textit{in} \textit{pro}\textbf{gé}\textbf{ni}es~* timén\textit{tibus} \textbf{e}um.
\item Fecit poténtiam in \textit{bráchio} \textbf{su}o:~* dispérsit supérbos mente \textit{cordis} \textbf{su}i.
\item Depósuit po\textit{téntes} \textit{de} \textbf{se}de,~* et exal\textit{távit} \textbf{hú}miles.
\item Esuriéntes \textit{implévit} \textbf{bo}nis:~* et dívites dimí\hspace{0.04em}\textit{sit} \textit{in}\textbf{á}nes.
\item Suscépit Israel \textit{púerum} \textbf{su}um,~* recordátus misericór\textit{diæ} \textbf{su}æ.
\item Sicut locútus est \textit{ad} \textit{pa}\textit{tres} \textbf{no}stros,~* Abraham et sémini e\hspace{0.02em}\textit{jus} \textit{in} \textbf{sǽ}cula.
\item Glória \textit{Pa}\textit{tri,} \textit{et} \textbf{Fí}\textbf{li}o,~* et Spirí\tinyhspace\textit{tui} \textbf{San}cto.
\item Sicut erat in princípio, \textit{et} \textit{nunc,} \textit{et} \textbf{sem}per,~* et in sǽcula sæcu\tinyhspace\textit{lórum.} \textbf{A}men.
    \end{enumerate}

\myrule

\smalltitle{Canticle of the Blessed Virgin Mary.}

\scripture{Luc. 1, 46–55.}

%\begin{paracol}{2}
\begin{multicols}{2}
%\lettrine{M}agníficat~* ánima mea Dóminum.\\ Et exsultávit spíritus meus~* in Deo salutári meo.
%
% Quia respéxit humilitátem ancíllæ suæ:~* ecce enim ex hoc beátam me dicent o\-mnes generatiónes.
%
% Quia fecit mihi magna qui potens est:~* et san\-ctum nomen ejus.
%
% Et misericórdia ejus a progénie in progénies~* timéntibus eum.
%
% Fecit poténtiam in bráchio suo:~* dispérsit supérbos mente cordis sui.
%
% Depósuit poténtes de sede,~* et exaltávit húmiles.
%
% Esuriéntes implévit bonis:~* et dívites dimísit inánes.
%
% Suscépit Israel púerum suum,~* recordátus misericórdiæ suæ.
%
% Sicut locútus est ad patres no\-stros,~* Abraham, et sémini ejus in sǽcula.
%
% Glória Patri, et Fílio,~* et Spirítui San\-cto.
%
% Sicut erat in princípio, et nunc, et semper,~* et in sǽcula sæculórum. Amen.
% 
% \switchcolumn
 \begin{enpars}
My soul \cc~* doth magnify the Lord.

And my spirit hath rejoiced~* in God my Saviour.

Because he hath regarded the humility of his handmaid;~* for behold from henceforth all generations shall call me blessed.

Because he that is mighty, hath done great things to me;~* and holy is his name.

And his mercy is from generation unto generations,~* to them that fear him.

He hath shewed might in his arm:~* he hath scattered the proud in the conceit of their heart.

He hath put down the mighty from their seat,~* and hath exalted the humble.

He hath filled the hungry with good things;~* and the rich he hath sent empty away.

He hath received Israel his servant,~* being mindful of his mercy:

As he spoke to our fathers,~* to Abraham and to his seed for ever.

Glory be to the Father, and to the Son,~* and to the Holy Ghost.

As it was in the beginning, is now,~* and ever shall be, world without end. Amen.
\end{enpars}
% \end{paracol}
\end{multicols}

\rubrique{\begin{enpars}
Sit as the antiphon is repeated. Then the usual dialogue follows. All stand from now until the end of the office, unless one sits for the antiphon of a commemorated Sunday or feria, according to custom}}

\textebilingue{dv}{dv_eng}

   \begin{paracol}{2}
   Orémus.
   
\lettrine{D}{e}us, qui per immaculátam Vírginis Conceptiónem dignum Fílio tuo habitáculum præparásti:~† quǽsumus; ut qui ex morte ejúsdem Fílii tui prævísa, eam ab omni labe præservásti,~* nos quoque mundos ejus intercessióne ad te perveníre concédas. Per eúmdem Dóminum.
   
\rr Amen.
   
\switchcolumn
\begin{enpars}
\noindent Let us pray.
O God, by the Immaculate Conception of the Virgin, thou prepared a worthy habitation for thy Son; we beseech thee, that, as by the foreseen death of thy same Son thou preserved her from all stain of sin, so thou would grant us also, through her intercession, to come to thee with pure hearts. Through the same Jesus Christ, thy Son, our Lord, who liveth and reigneth with thee in the unity of the Holy Ghost, God, world without end. Amen.
\end{enpars}
    \end{paracol}

\textebilingue{dv}{dv_eng}

\myrule

\rubrique{A commemoration is made of the Advent feria or of the Sunday. The antiphon is always proper to the day; the verse is as on Sundays of the season at Vespers, and the collect is taken from the I Sunday of Advent, if the feast falls in the first week of Advent, or from the II Sunday, when it falls on Saturday, on that Sunday, or in the second week.}

\texteparacol{

\vv Roráte cæli désuper, et nubes pluant justum.

\rr Aperiátur terra, et gérminet Salvatórem.

\lettrine{E}{x}cita, quǽsumus Dómine, poténtiam tuam, et veni:~† ut ab imminéntibus peccatórum nostrórum perículis, te mereámur protegénte éripi,~* te liberánte salvári. (Qui vivis et regnas cum Deo Patre.)}{english}{
\noindent \vv Drop down dew, ye heavens, from above, and let the clouds rain the Just.

\noindent \rr Let the earth be opened, and bud forth the Saviour.

\noindent Let us pray. Stir up, O Lord, we pray thee, thy strength, and come among us, that whereas through our sins and wickedness we do justly apprehend thy wrathful judgments hanging over us, thy bountiful grace and mercy may speedily help and deliver us; Who livest and reignest with God the Father.}

\rubrique{Or:}

\texteparacol{
\lettrine{E}{x}cita Dómine corda nostra ad præparándas Unigéniti tui vias:~† ut per ejus advéntum,~* purificátis tibi méntibus servíre mereámur.
(Qui tecum vivit et regnat.)}{english}{\noindent Let us pray. Stir up our hearts, O Lord, to make ready the ways of thine Only-begotten Son, that by his coming our minds being purified, we may the more worthily give up ourselves to thy service: Who with thee liveth and reigneth.}


\rubrique{\begin{enpars}
The cantor sings the verse to which all reply. One of the following tones is chosen according to the Vespers sung, at the choirmaster's discretion.
\end{enpars}}

\smalltitle{At I Vespers.}

\gscore[]{2.}{BD_Solemnis_I_Vesp}

\noindent \vv Let us bless the Lord.

\noindent \rr Thanks be to God.

\smalltitle{At II Vespers.}

\gscore[]{6.}{BD_Solemnis_mode_6}

\rubrique{Or the following tone may be used:}

\gscore[]{5.}{BD_Solemnis_mode_5}

\smalltitle{At either Vespers.}

\gscore[]{1.}{BD_BVM}

\textebilingue{fidelium_animae}{fidelium_animae_en}

\end{document}