% !TEX TS-program = lualatexmk
% !TEX parameter =  --shell-escape
\documentclass[11pt]{book}

\input{commonheaders_vespers_people}

\begin{document}

\twosided[pc] %%% why this goes at beginning, IDK, unclear what p does exactly, but it's necessary for flipping columns correctly. ; it's unclear why this goes BEFORE everything else and not when activating the environment. %% It's OK for paper booklets to not switch, but it's probably best practice to switch.

\thispagestyle{empty}

\bigtitle{Marian Antiphons\\Sung after Mass or a Divine Office.}

\smalltitle{Alma Redemptoris Mater.}

\rubrique{¶ From Vespers of the Saturday before the I Sunday of Advent to Vespers of February 2 inclusive.}

\gscore[]{5.}{an--alma_redemptoris--solesmes_1961}
%\vspace{1ex}

\smalltitle{Simple Tone.}
\gscore[]{5.}{an_alma_redemptoris_simple_solesmes}

\newpage

\begin{sloppypar}
 \begin{paracol}{2}
  \setlength{\columnsep}{1.5em} %EB Garamond, Octavio Pardo; should define this so that there are no magic numbers
\textes{Alma_Redemptoris_en}

%\columnbreak
\switchcolumn
\textes{Alma_Redemptoris_la}
\end{paracol}
\end{sloppypar}
%\end{multicols}
%\smallskip

\newpage

\smalltitle{Ave Regina Cælorum.}

\rubrique{¶ After the Purification, that is, from Compline of February 2 even if the feast is transferred, until Compline of Wednesday in Holy Week.}

\gscore[]{6.}{an--ave_regina_caelorum--solesmes_1961}
%\vspace{1ex}

\smalltitle{Simple Tone.}
\gscore[]{6.}{an_ave_regina_caelorum_simple_tone_solesmes}

\begin{sloppypar}
 \begin{paracol}{2}
  \setlength{\columnsep}{1.5em} %EB Garamond, Octavio Pardo; should define this so that there are no magic numbers
\textes{Ave_Regina_la}

%\columnbreak
\switchcolumn
\textes{Ave_Regina_en}
\end{paracol}
\end{sloppypar}
%\end{multicols}
%\smallskip

\newpage
\smalltitle{Regina Cæli.}

\rubrique{¶ From Compline of Holy Saturday until None of Saturday after Pentecost inclusive.}

\gscore[]{6.}{an--regina_caeli--solesmes}
%\vspace{1ex}

\smalltitle{Simple Tone.}
\gscore[]{6.}{an_regina_caeli_simple_tone_solesmes}

\begin{sloppypar}
 \begin{paracol}{2}
  \setlength{\columnsep}{1.5em} %EB Garamond, Octavio Pardo; should define this so that there are no magic numbers
\textes{Regina_Caeli_la}

%\columnbreak
\switchcolumn
\textes{Regina_Caeli_en}
\end{paracol}
\end{sloppypar}
%\end{multicols}
%\smallskip

\newpage

\smalltitle{Salve Regina.}

\rubrique{¶ From Vespers of the Holy Trinity until None of the Saturday before Advent inclusive.}

\gscore[]{1.}{an--salve_regina--solesmes}
%\vspace{1ex}

\smalltitle{Simple Tone.}
\gscore[]{5.}{an_salve_regina_simple_tone_solesmes}

\begin{sloppypar}
 \begin{paracol}{2}
  \setlength{\columnsep}{1.5em} %EB Garamond, Octavio Pardo; should define this so that there are no magic numbers
\textes{Salve_Regina_la}

%\columnbreak
\switchcolumn
\textes{Salve_Regina_en}
\end{paracol}
\end{sloppypar}
%\end{multicols}

\enddocument