%% !TEX TS-program = lualatexmk
%% !TEX parameter =  --shell-escape

\documentclass[11pt]{book}
\input{commonheaders_vespers_people}

\begin{document}

\thispagestyle{empty}

\bigtitle{Benediction on Sundays of Advent and Lent.}

\rubrique{The cantors intone until the asterisk or the quarter bar.}

\rubrique{\normaltext{O Salutáris VI,} from \normaltext{Cantus Selecti,} 1957; melody of \normaltext{Cónditor alme sidérum} from the office of the Most Holy Redeemer.}

%\grechangestyle{initial}{\fontsize{28}{28}\selectfont}

\gscore[]{4.}{hy_o_salutaris_vi_cantus_selecti_solesmes}

\textes{O_Salutaris_en.tex}

%\begin{multicols}{2}
%\raggedcolumns
%\begin{enpars}
%O saving victim, opening wide\\
%The gate of heaven to man below,\\
%Our foes press on from every side,\\
%Thine aid supply, thy strength bestow.x
%
%
%To thy great name be endless praise,\\
%Immortal Godhead, One in Three;\\
%O grant us endless length of days\\
%In our true native land, with thee.
%Amen.
%\end{enpars}
%\end{multicols}

\begin{enpars}

\textes{Cardinal_Wiseman_Prayer}

\rubrique{On the Second Sunday of each month, instead of the prayer \normaltext{O Blessed Virgin Mary,} the following prayer is said, all kneeling.}

\textes{Second_Sunday_Prayer}

\rubrique{Then follows Benediction of the Most Blessed Sacrament of the Eucharist, in the usual way.}

\rubrique{\normaltext{Tantum ergo VI,} from \normaltext{Cantus Selecti,} 1957; melody of the hymn \normaltext{O quot undis lacrimárum} from the office of the Seven Sorrows of the Blessed Virgin Mary.}

\end{enpars}

\gscore[]{2.}{hy_tantum_ergo_vi_solesmes}

\textes{Tantum_Ergo_en}

\texteparacol{
%\sloppy
\noindent \vv Panem de cælo præstitisti eis.
%\tpalleluia{}\footnote{Atque per octavam Corporis Christi. \textit{As well as throughout the octave of Corpus Christi.}}

\noindent \rr Omne dele\-ctaméntum in se habéntem.
%\tpalleluia{}
%\newpage

  Orémus.
  
  \lettrine{D}{e}us, qui nobis sub Sacraménto mirábili passiónis tuæ memóriam reliquísti:~† tríbue, quǽsumus, ita nos córporis, et sánguinis tui sacra mystéria venerári;~* ut redemptiónis tuæ fru\-ctum in nobis júgiter sentiámus. Qui vivis et regnas in sǽcula sæculórum.   \rr Amen.}{english}
%  \switchcolumn
%\begin{enpars}
%\begin{enpars}
{\noindent \vv Thou hast given them bread from heaven.
%(\textit{P.T.} alleluia.)

\noindent \rr Having all sweetness within it.
%(\textit{P.T.} alleluia.)
%\newpage

\noindent Let us pray. O God, who under a wonderful Sacrament hast left us a memorial of Thy Passion; grant us, we beseech Thee, so to reverence the sacred mysteries of Thy Body and Blood, that we may ever feel within ourselves the fruit of Thy Redemption. Who livest and reignest forever and ever. Amen.}

\textes{laudes_divinae_en}

\rubrique{These are sung alternating between a cantor and the whole congregation, which also sings \normaltext{Fiat, Fiat} at the end.}

\gscore[]{}{laudes_divinae_no_breaks}

\rubrique{The cantors intone until the quarter bar, then the people are welcome to sing the rest of the antiphon as well the second half of each verse and the repeated antiphon.}

\rubrique{From the \normaltext{Westminster Hymnal} edited by Sir R.R. Terry, 1912.}

\gscore[]{6.}{Adoremus_westminster_mode_6}

\textes{Adoremus_in_aeternum_en}

\enddocument