 %%!TEX TS-program = lualatexmk
%% !TEX parameter =  --shell-escape

\documentclass[11pt]{article}
\input{commonheaders_vespers_people}

\begin{document}

\thispagestyle{empty}

\bigtitle{Adoration and Benediction on Major Feasts.}

\rubrique{The cantors intone until the first barline.}

\rubrique{The melodies are those of Corpus Christi in the Roman usage.}

\gscore[]{8.}{o_salutaris_mode_8}

\textes{O_Salutaris_en.tex}

\begin{enpars}

\rubrique{Devotions may follow according to the month or the occasion; on Sundays, the usual prayers may be said. Otherwise, Benediction follows immediately.}

\textes{Cardinal_Wiseman_Prayer}

\rubrique{On the Second Sunday of each month, instead of the prayer \normaltext{O Blessed Virgin Mary,} the following prayer is said, all kneeling.}

\textes{Second_Sunday_Prayer}

\rubrique{Then follows Benediction of the Most Blessed Sacrament of the Eucharist, in the usual way.}
\end{enpars}

\grechangestyle{initial}{\fontsize{28}{28}\selectfont}
\gscore[]{3.}{hy_tantum_ergo_i_solesmes_1961}

\textes{Tantum_Ergo_en}

\texteparacol{
\noindent \vv Panem de cælo præstitisti eis. \tpalleluia{}\footnote{Atque per octavam Corporis Christi. \textit{As well as throughout the octave of Corpus Christi.}}

\noindent \rr Omne dele\-ctaméntum in se habéntem. \tpalleluia{}

  Orémus.
  
\lettrine{D}{e}us, qui nobis sub Sacraménto mirábili passiónis tuæ memóriam reliquísti:~† tríbue, quǽsumus, ita nos córporis, et sánguinis tui sacra mystéria venerári;~* ut redemptiónis tuæ fructum in nobis júgiter sentiámus. Qui vivis et regnas in sǽcula sæculórum.

\rr Amen.}
{english}
{\noindent \vv Thou hast given them bread from heaven. (\textit{P.T.} alleluia.)
\newpage
\noindent \rr Having all sweetness within it. (\textit{P.T.} alleluia.)

\noindent Let us pray. O God, who under a wonderful Sacrament hast left us a memorial of Thy Passion; grant us, we beseech Thee, so to reverence the sacred mysteries of Thy Body and Blood, that we may ever feel within ourselves the fruit of Thy Redemption. Who livest and reignest forever and ever. Amen.}

\textes{laudes_divinae_en}

\rubrique{These are sung alternating between a cantor and the whole congregation, which also sings \normaltext{Fiat, Fiat} at the end.}

\gscore[]{}{laudes_divinae_no_breaks}

\rubrique{The cantors intone until the asterisk; for greater solemnity, the organ may supply the even verses.}

\rubrique{This tone is adapted from chant used by the French royal chapel and at Notre-Dame de Paris.}

\gscore[\textit{Ton roy.}]{6.}{ps_laudate_dominum_ton_royal}

\begin{otherlanguage}{english}\noindent Ps. 116: Praise the Lord, all ye nations: praise him, all ye people. For his mercy is confirmed upon us: and the truth of the Lord remaineth forever. Glory be to the Father, and to the Son, and to the Holy Ghost. As it was in the beginning, is now, and ever shall be, world without end. Amen.\end{otherlanguage}

\end{document}