% !TEX TS-program = lualatexmk
% !TEX parameter =  --shell-escape

\documentclass[11pt]{book}
\input{commonheaders_vespers_people}

\begin{document}

\thispagestyle{empty}

\bigtitle{Benediction on Sundays of June.}

\rubrique{The cantors intone until the asterisk or the quarter bar.}

\rubrique{Alternate medieval melody restored in the Vatican Edition.}

\gscore[]{7.}{hy_o_salutaris_mode_7}

\textes{O_Salutaris_en.tex}

\rubrique{The Litany of the Sacred Heart is said.}

\begin{enpars}

\textes{Cardinal_Wiseman_Prayer}

\rubrique{On the Second Sunday of each month, instead of the prayer \normaltext{O Blessed Virgin Mary,} the following prayer is said, all kneeling.}

\textes{Second_Sunday_Prayer}

\rubrique{Then follows Benediction of the Most Blessed Sacrament of the Eucharist, in the usual way.}

\rubrique{Modern chant, based on the familiar tune attributed to John Francis Wade, ca. 1751.}

\end{enpars}

\gscore[]{5.}{hy_tantum_ergo_iii_solesmes_1961.gabc}

\textes{Tantum_Ergo_en}

\texteparacol{
%\sloppy
\noindent \vv Panem de cælo præstitisti eis.
\tpalleluia{}\footnote{Atque per octavam Corporis Christi. \textit{As well as throughout the octave of Corpus Christi.}}

\noindent \rr Omne dele\-ctaméntum in se habéntem.
\tpalleluia{}
%\newpage

  Orémus.
  
  \lettrine{D}{e}us, qui nobis sub Sacraménto mirábili passiónis tuæ memóriam reliquísti:~† tríbue, quǽsumus, ita nos córporis, et sánguinis tui sacra mystéria venerári;~* ut redemptiónis tuæ fru\-ctum in nobis júgiter sentiámus. Qui vivis et regnas in sǽcula sæculórum.
  
    \rr Amen.}{english}
%  \switchcolumn
%\begin{otherlanguage}{english}
%\begin{enpars}
{\noindent \vv Thou hast given them bread from heaven.
(\textit{P.T.} alleluia.)

\noindent \rr Having all sweetness within it.
(\textit{P.T.} alleluia.)
%\newpage

\noindent Let us pray. O God, who under a wonderful Sacrament hast left us a memorial of Thy Passion; grant us, we beseech Thee, so to reverence the sacred mysteries of Thy Body and Blood, that we may ever feel within ourselves the fruit of Thy Redemption. Who livest and reignest forever and ever. Amen.}
%\end{enpars}
%\end{paracol}

\textes{laudes_divinae_en}

\rubrique{These are sung alternating between a cantor and the whole congregation, which also sings \normaltext{Fiat, Fiat} at the end.}

\gscore[]{}{laudes_divinae_no_breaks}

\begin{otherlanguage}{english}\rubrique{The cantors intone until the half bar, then the people are welcome to sing the rest of the antiphon and the repetitions.}\end{otherlanguage}

\rubrique{Melody adapted, probably by Dom Joseph Pothier, o.s.b.}

\gscore[]{1.}{va--cor_jesu--solesmes}
\begin{otherlanguage}{english}Most Sacred Heart of Jesus, have mercy on us.
\end{otherlanguage}

\enddocument