% !TEX TS-program = lualatexmk
% !TEX parameter =  --shell-escape

\documentclass[11pt]{book}
\input{commonheaders_vespers_people}

\begin{document}

\thispagestyle{empty}

\bigtitle{Benediction on Sundays \textit{per annum} and until Ash Wednesday.}

\rubrique{The cantors intone until the asterisk or the quarter bar.}

\rubrique{Alternate medieval melody restored in the Vatican Edition.}

\gscore[]{7.}{hy_o_salutaris_mode_7}

\textes{O_Salutaris_en.tex}

\begin{enpars}

\textes{Cardinal_Wiseman_Prayer}

\rubrique{On the Second Sunday of each month, instead of the prayer \normaltext{O Blessed Virgin Mary,} the following prayer is said, all kneeling.}

\textes{Second_Sunday_Prayer}

\rubrique{Then follows Benediction of the Most Blessed Sacrament of the Eucharist, in the usual way.}

\rubrique{Modern chant, based on the familiar tune attributed to John Francis Wade, ca. 1751.}

\end{enpars}

\gscore[]{5.}{hy_tantum_ergo_iii_solesmes_1961.gabc}

\textes{Tantum_Ergo_en}
%\newpage
%\begin{multicols}{2}
%\raggedcolumns
%%\begin{otherlanguage}{english}
%\begin{enpars}
%Down in adoration falling,\\
%Lo! the sacred Host we hail;\\
%Lo! o'er ancient forms departing,\\
%Newer rites of grace prevail;\\
%Faith for all defects supplying,\\
%Where the feeble senses fail.
%
%To the everlasting Father,\\
%And the Son who reigns on high,\\
%With the Holy Ghost proceeding\\
%Forth from each eternally,\\
%Be salvation, honour, blessing,\\
%Might, and endless majesty.
%Amen.
%\end{enpars}
%%\end{otherlanguage}
%\end{multicols}

\texteparacol{
%\sloppy
\noindent \vv Panem de cælo præstitisti eis.
%\tpalleluia{}\footnote{Atque per octavam Corporis Christi. \textit{As well as throughout the octave of Corpus Christi.}}

\noindent \rr Omne dele\-ctaméntum in se habéntem.
%\tpalleluia{}
\newpage

  Orémus.
  
  \lettrine{D}{e}us, qui nobis sub Sacraménto mirábili passiónis tuæ memóriam reliquísti:~† tríbue, quǽsumus, ita nos córporis, et sánguinis tui sacra mystéria venerári;~* ut redemptiónis tuæ fru\-ctum in nobis júgiter sentiámus. Qui vivis et regnas in sǽcula sæculórum.
  
  \rr Amen.}{english}
%  \switchcolumn
%\begin{otherlanguage}{english}
%\begin{enpars}
{\noindent \vv Thou hast given them bread from heaven.
%(\textit{P.T.} alleluia.)

\noindent \rr Having all sweetness within it.
%(\textit{P.T.} alleluia.)
\newpage

\noindent Let us pray. O God, who under a wonderful Sacrament hast left us a memorial of Thy Passion; grant us, we beseech Thee, so to reverence the sacred mysteries of Thy Body and Blood, that we may ever feel within ourselves the fruit of Thy Redemption. Who livest and reignest forever and ever. Amen.}
%\end{enpars}
%\end{paracol}

\textes{laudes_divinae_en}

%\begin{multicols}{2}
%\raggedcolumns
%\begin{otherlanguage}{english}
%
%Blessed be God. 
%
%Blessed be His Holy Name. 
%
%Blessed be Jesus Christ, true God and true Man.
% 
%Blessed be the Name of Jesus.
%
%Blessed be His Most Sacred Heart.
%
%Blessed be His Most Precious Blood.
%
%Blessed be Jesus in the Most Holy Sacrament of the Altar.
%
%Blessed be the Holy Spirit, the Paraclete.
%
%Blessed be the great Mother of God, Mary most Holy.
%
%Blessed be her Holy and Immaculate Conception.
%
%\rubrique{The following invocation is said thrice.}
%
%Blessed be her Glorious Assumption.
%
%Blessed be the name of Mary, Virgin and Mother.
%
%Blessed be Saint Joseph, her most chaste spouse.
%
%Blessed be God in His Angels and in His Saints.
%
%\end{otherlanguage}
%\end{multicols}

\rubrique{These are sung alternating between a cantor and the whole congregation, which also sings \normaltext{Fiat, Fiat} at the end.}

\gscore[]{}{laudes_divinae_no_breaks}

\rubrique{The cantors intone until the quarter bar, then the people are welcome to sing the rest of the antiphon as well the second half of each verse and the repeated antiphon.}

\rubrique{Melody adapted ca. 1895, probably by Dom Joseph Pothier, o.s.b.}

\gscore[]{5.}{va_adoremus_in_aeternum_solesmes_1961}
\textes{Adoremus_in_aeternum_en}

\enddocument