% !TEX TS-program = lualatexmk
% !TEX parameter =  --shell-escape

\documentclass[11pt]{article}
\input{commonheaders_vespers_people}

\begin{document}

\thispagestyle{empty}

\bigtitle{Exposition and Benediction on Weekdays of Paschal Time.}

\rubrique{The cantors intone until the first barline.}

\rubrique{\normaltext{O Salutáris III,} in the \normaltext{Liber antiphonarius,} 1949. Melody of the hymn at Lauds of Pentecost.}

\gscore[]{1.}{hy_o_salutaris_hostia_liber_antiphonarius_3_solesmes}

\textes{O_Salutaris_en.tex}

%\begin{otherlanguage}{english}
\begin{enpars}

\rubrique{Devotions are said according to the month.}

\rubrique{On First Friday, the Sacred Heart devotions are said instead.}

{\centering{Prayer Before Benediction on Thursdays.}\par}

\textes{Prayer_Before_Benediction}
\end{enpars}
%\end{otherlanguage}

\rubrique{\normaltext{Tantum ergo IX} from \normaltext{Cantus Selecti,} 1957; melody of \normaltext{Urbs beáta Jerúsalem.}}

\gscore[]{4.}{hy_tantum_ergo_ix_solesmes_1957}

\textes{Tantum_Ergo_en}

   \texteparacol{\vv Panem de cælo præstitísti eis, allelúia.

\noindent \rr Omne dele\-ctaméntum in se habéntem, allelúia.
  
  Orémus.
  
  \lettrine{D}{e}us, qui nobis sub Sacraménto mirábili passiónis tuæ memóriam reliquísti:~† tríbue, quǽsumus, ita nos córporis, et sánguinis tui sacra mystéria venerári;~* ut redem\-ptiónis tuæ fru\-ctum in nobis júgiter sentiámus. Qui vivis et regnas in sǽcula sæculórum. 
  
  \rr Amen.}{english}{
\noindent \vv Thou hast given them bread from heaven, alleluia.

\noindent \rr Having all sweetness within it, alleluia.

\noindent Let us pray. O God, who under a wonderful Sacrament hast left us a memorial of thy Passion; grant us, we beseech thee, so to reverence the sacred mysteries of thy Body and Blood, that we may ever feel within ourselves the fruit of thy Redemption. Who livest and reignest forever and ever. Amen.}

\textes{laudes_divinae_en}

\rubrique{These are sung alternating between a cantor and the whole congregation, which also sings \normaltext{Fiat, Fiat} at the end.}

\gscore[]{}{laudes_divinae_no_breaks}

\rubrique{The cantors sing the odd verses, all others even.}

\rubrique{Psalm tone for psalms without an antiphon in Paschal Time from the \normaltext{Antiphonale Romanum.}}

\gscore[]{}{ps_116_tp}

\begin{otherlanguage}{english}\noindent Ps. 116: Praise the Lord, all ye nations: praise him, all ye people. For his mercy is confirmed upon us: and the truth of the Lord remaineth forever. Glory be to the Father, and to the Son, and to the Holy Ghost. As it was in the beginning, is now, and ever shall be, world without end. Amen.\end{otherlanguage}

\rubrique{On First Friday, after Communion:}

\rubrique{Melody adapted, probably by Dom Joseph Pothier, o.s.b.}

\begin{otherlanguage}{english}\rubrique{The cantors intone until the half bar, then the people are welcome to sing the rest of the antiphon and the repetitions.}\end{otherlanguage}

\gscore[]{1.}{va--cor_jesu--solesmes}

%\noindent
\begin{otherlanguage}{english}Most Sacred Heart of Jesus, have mercy on us.
\end{otherlanguage}

\end{document}