% !TEX TS-program = lualatexmk
% !TEX parameter =  --shell-escape
\documentclass[benediction_assumption_la_en.tex]{subfiles}

\ifcsname preamble@file\endcsname
  \setcounter{page}{\getpagerefnumber{M-benediction_assumption_la_en_marian_antiphons}}
\fi

\begin{document}
 \twosided[pc] 

\chapter*{MARIAN ANTIPHONS.}
\addcontentsline{toc}{chapter}{Marian Antiphons}

\smalltitle{Alma Redemptoris Mater.}

\rubrique{¶ From Vespers of the Saturday before the I Sunday of Advent to Vespers of February 2 inclusive.}

\gscore[]{5.}{an_alma_redemptoris_solesmes_1961}
%%\vspace{1ex}

\smalltitle{Simple Tone.}
\gscore[]{5.}{an_alma_redemptoris_simple_solesmes}

\textes{alma_redemptoris_la_en}

\smalltitle{Ave Regina Cælorum.}

\rubrique{¶ After the Purification, that is, from Compline of February 2 even if the feast is transferred, until Compline of Wednesday in Holy Week.}

\gscore[]{6.}{an_ave_regina_caelorum_solesmes_1961}
%\vspace{1ex}

\smalltitle{Simple Tone.}

\gscore[]{6.}{an_ave_regina_caelorum_simple_tone_solesmes}

\textes{ave_regina_la_en}

\smalltitle{Regina Cæli.}

\rubrique{¶ From Compline of Holy Saturday until None of Saturday after Pentecost inclusive.}

\gscore[]{6.}{an_regina_caeli_solesmes}
%\vspace{1ex}

\smalltitle{Simple Tone.}

\gscore[]{6.}{an_regina_caeli_simple_tone_solesmes}

\textes{regina_caeli_la_en}

\smalltitle{Salve Regina.}

\rubrique{¶ From Vespers of the Holy Trinity until None of the Saturday before Advent inclusive.}

\gscore[]{1.}{an_salve_regina_solesmes}
%\vspace{1ex}

\smalltitle{Simple Tone.}
\gscore[]{5.}{an_salve_regina_simple_tone_solesmes}

\textes{salve_regina_la_en}

%\rubrique{The versicle is sung on a lower note.}
%\label{conclusion} \phantomsection

%\texteparacol{\vv Divínum auxílium máneat semper nobíscum.
%
%\noindent \rr Amen.}
%{english}
%{\noindent \vv May the divine assistance remain always with us.
%
%\noindent \rr Amen.}

\end{document}