% !TEX TS-program = LuaLaTeX+se

\documentclass[12pt]{book}

\input{commonheaders_vespers_people} %%% will need to fix spacing problem in psalms at a later date.

\begin{document}
 
 \twosided[pc] %% to allow for the left-right switching of columns needed for bilingual liturgical typesetting (like in missals); it's unclear why this goes BEFORE everything else and not when activating the environment. %% It's OK for paper booklets to not switch, but it's probably best practice to switch.
 
  \thispagestyle{empty}
 
 {\centering{\huge{CHRIST THE KING.}}\par}\header{christ the king.}
 
\smalltitle{Before the Office.}

\rubrique{These prayers are said kneeling. End when the priest stands.}

\textebilingue{orationes_ante_do}{orationes_ante_do_eng}

\smalltitle{\prayers}

\rubrique{The celebrant intones the following verse. All sing the response and doxology.}

\begin{otherlanguage}{english}
\rubrique{Make the sign of the cross at the beginning and at the Magnificat, and bow the head when the Trinity is mentioned, when mention is made of the ``Holy Name'' in some way, and at ``Orémus.''}\end{otherlanguage}

\gscore[]{℣.}{Deus_in_adjutorium_tonus_solemnior}

\noindent
\begin{otherlanguage}{english}
O God, \cc{} come to my assistance. \rr O Lord, make haste to help me. Glory be to the Father, and to the Son,~* and to the Holy Ghost. As it was in the beginning, is now,~* and ever shall be, world without end. Amen. Alleluia.\end{otherlanguage}

\smalltitle{First Antiphon. 8. G.}
 \initialscore{an_pacificus_vocabitur_solesmes_1961} 
 \begin{otherlanguage}{english} \noindent Ant. He shall be called Peaceable, and his throne shall be most firm for ever. \end{otherlanguage}

\rubrique{\begin{otherlanguage}{english}
Sit at the word ``Sede.'' The cantor sings odd verses; all sing even verses, and this for the hymn and Magnificat, when sung entirely in chant, as well.
\end{otherlanguage}}

\psalmus{109}{109_8G}{109_8G}{109_en}
%

\rubrique{\begin{otherlanguage}{english}
Clerics in descending order, or a cantor, intone the subsequent antiphons.\end{otherlanguage}}
%
 \gscore[2. Ant.]{8. c}{an_regnum_ejus_solesmes_1961}
% 
 \begin{otherlanguage}{english} \noindent Ant. His kingdom is an everlasting kingdom, and all kings shall serve him, and shall obey him. \end{otherlanguage}
%
 \psalmus{110}{110_8c}{110_8c}{110_en}
%
\gscore[3. Ant.]{7. a}{an_ecce_vir_oriens_solesmes_1961}

  \begin{otherlanguage}{english} \noindent Ant. Behold a man, the Orient is his name: he shall sit and rule, and shall speak peace to the Gentiles. \end{otherlanguage}
%
 \psalmus{111}{111_7a}{111_7a}{111_en}
%
 \gscore[4. Ant.]{3. a}{an_dominus_judex_solesmes_1961}
%
\begin{otherlanguage}{english} \noindent Ant. The Lord is our judge, the Lord is our lawgiver: the Lord is our king, he will save us. \end{otherlanguage}
%
  \psalmus{112}{112_3a}{112_3a}{112_en}
%
  \gscore[5. Ant.]{8. G}{an_ecce_dedi_te_solesmes_1961}
% 
\begin{otherlanguage}{english} \noindent Ant. Behold, I have given thee to be the light of the Gentiles, that thou mayst be my salvation even to the farthest part of the earth. \end{otherlanguage}
%
  \psalmus{116}{116_8G}{116_8G}{116_en}

\rubrique{\begin{otherlanguage}{english}
Having sung the last antiphon in full, stand when the celebrant does.
\end{otherlanguage}}

\smalltitle{Chapter.}
\scripture{Col. 1, 12– 13.}

\textebilingue{cap_christi_regis}{cap_christi_regis_en}

\smalltitle{Hymn.}

\gscore[]{1.}{hy_te_saeculorum_principem_solesmes_1961}

\begin{multicols}{2}
\begin{otherlanguage}{english}
To thee, O Prince of all that be,\\
Thou Christ, O King eternally;\\
O Framer of the mind and heart,\\
Our one true Judge we say thou art.

The wicked protest, wail and cry,\\
Christ Jesus’ reign they would deny;\\
Rejoice we at thy glorious name,\\
Thou Highest King we do proclaim.

O Christ! The Source of all our peace,\\
Make all our sinful thoughts to cease;\\
And still in us our loves misplaced,\\
As Thy one sheepfold be we embraced.

For this, hanging on cruel tree,\\
With arms outstretched, for all to see;\\
His heart is pierced by soldier’s spear,\\
Revealing burning love most dear.

From this the altar of the tree\\
Thy blood flows forth from Calvary;\\
As wine to us it doth appear,\\
To thine own heart it draws us near.

Thou Governor of all that be,\\
May all thy creatures honour thee;\\
All those who rule, O Lord renew!\\
Source of all precepts just and true.

To regal glory, all submit,\\
All crowns and honours we do remit;\\
To thy scepter—so sweet and mild!\\
Submit we as a little child.

All glory be, Jesu, to thee,\\
Thy scepter over all that be;\\
All glory, as is ever meet,\\
To Father and to Paraclete. 

Amen.
\end{otherlanguage}
\end{multicols}

\rubrique{\begin{otherlanguage}{english}
The cantor sings the verse to which all reply.
\end{otherlanguage}}

\myrule
 \smalltitle{At I Vespers.}
 
\smallscore{versiculus_1_vesperis_christi_regis}
\begin{otherlanguage}{english} \noindent \vv Given to me is all power.

\noindent \rr In heaven and in earth.
 \end{otherlanguage}
 
 \myrule

 \smalltitle{At II Vespers.}
\smallscore{versiculus_2_vesperis_christi_regis}
\begin{otherlanguage}{english} \noindent \vv His empire shall be multiplied.

\noindent \rr And there shall be no end of peace.
 \end{otherlanguage}
 
\smalltitle{Antiphon at the Magnificat.}

\rubrique{Sit when the celebrant does, after he sings the first words. Rise at the end.}

\myrule

 \smalltitle{At I Vespers.}
 \gscore[]{1. f}{an_dabit_illi_dominus_solesmes_1961}
 
 \begin{otherlanguage}{english} \noindent Ant. The Lord God shall give unto him the throne of David, his father: and he shall reign in the house of Jacob for ever, and of his kingdom there shall be no end, alleluia. \end{otherlanguage}
 
 \myrule

 \smalltitle{At II Vespers.}
 \gscore[]{7. a}{an_habet_in_vestimento_solesmes_1961}
 
 \begin{otherlanguage}{english} \noindent Ant. And he hath on his garment, and on his thigh written: King of kings, and Lord of lords. To him be glory and empire for ever and ever. \end{otherlanguage}

 \rubrique{\begin{otherlanguage}{english}
The Magnificat may be sung entirely in chant or sung alternating between a cantor singing chant and the choir singing a polyphonic or harmonized setting. Please see the final pages for the chant.\end{otherlanguage}}

\smalltitle{Canticle of the Blessed Virgin Mary.}

\scripture{Luc. 1, 46–55.}

\begin{paracol}{2}

\lettrine{M}agníficat~* ánima mea Dóminum.\\ Et exsultávit spíritus meus~* in Deo salutári meo.

 Quia respéxit humilitátem ancíllæ suæ:~* ecce enim ex hoc beátam me dicent o\-mnes generatiónes.

 Quia fecit mihi magna qui potens est:~* et san\-ctum nomen ejus.

 Et misericórdia ejus a progénie in progénies~* timéntibus eum.

 Fecit poténtiam in bráchio suo:~* dispérsit supérbos mente cordis sui.

 Depósuit poténtes de sede,~* et exaltávit húmiles.

 Esuriéntes implévit bonis:~* et dívites dimísit inánes.

 Suscépit Israel púerum suum,~* recordátus misericórdiæ suæ.

 Sicut locútus est ad patres no\-stros,~* Abraham, et sémini ejus in sǽcula.

 Glória Patri, et Fílio,~* et Spirítui San\-cto.

 Sicut erat in princípio, et nunc, et semper,~* et in sǽcula sæculórum. Amen.
 
 \switchcolumn
 \begin{otherlanguage}{english}
My soul \cc~* doth magnify the Lord.

And my spirit hath rejoiced~* in God my Saviour.

Because he hath regarded the humility of his handmaid;~* for behold from henceforth all generations shall call me blessed.

Because he that is mighty, hath done great things to me;~* and holy is his name.

And his mercy is from generation unto generations,~* to them that fear him.

He hath shewed might in his arm:~* he hath scattered the proud in the conceit of their heart.

He hath put down the mighty from their seat,~* and hath exalted the humble.

He hath filled the hungry with good things;~* and the rich he hath sent empty away.

He hath received Israel his servant,~* being mindful of his mercy:

As he spoke to our fathers,~* to Abraham and to his seed for ever.

Glory be to the Father, and to the Son,~* and to the Holy Ghost.

As it was in the beginning, is now,~* and ever shall be, world without end. Amen.
\end{otherlanguage}
 \end{paracol}

\rubrique{\begin{otherlanguage}{english}
Sit as the antiphon is repeated. Then the usual dialogue follows. All stand from now until the end of the office, unless one sits for the antiphon sung for the commemoration of the Sunday or of the saints, according to custom.\end{otherlanguage}}

\textebilingue{dv}{dv_eng}

\textebilingue{or_christi_regis}{or_christi_regis_en}

\textebilingue{dv}{dv_eng}

\rubrique{\begin{otherlanguage}{english}
The cantor sings the verse to which all reply. The tone is chosen at the choirmaster's discretion.
\end{otherlanguage}}

\myrule

\smalltitle{At I Vespers.}

\gscore[]{2.}{BD_Solemnis_I_Vesp}

\vv Let us bless the Lord.

\rr Thanks be to God.

\myrule

\smalltitle{At II Vespers.}

\gscore[]{6.}{BD_Solemnis_mode_6}

\rubrique{Or the following tone may be used:}

\gscore[]{5.}{BD_Solemnis_mode_5}

\textebilingue{fidelium_animae}{fidelium_animae_en}

\smalltitle{Chant for the Magnificat.}

\myrule
 \smalltitle{At I Vespers.}

\smallscore{Magnificat_1_f_Solemnis}

   \smallskip%
    \noindent%
    \begin{itemize}[%
    		label=\null, %
			leftmargin=0pt, %
			itemindent=0mm, %
			labelsep=0pt, %
			labelwidth=0pt, %
			rightmargin=0pt, %
			parsep=0pt, %
			topsep=0pt, %
			itemsep=0pt]%
    \input{psaumes/Magnificat_1_Solemnis}
    \end{itemize}

\myrule

 \smalltitle{At II Vespers.}
\smallscore{Magnificat_7a_Solemnis}

   \smallskip%
    \noindent%
    \begin{itemize}[%
    		label=\null, %
			leftmargin=0pt, %
			itemindent=0mm, %
			labelsep=0pt, %
			labelwidth=0pt, %
			rightmargin=0pt, %
			parsep=0pt, %
			topsep=0pt, %
			itemsep=0pt]%
    \input{psaumes/Magnificat_7}
    \end{itemize}

%\rubrique{\begin{otherlanguage}{english}
%Then one Our Father is said, unless Compline follows, or even some pious exercises, a sermon, or Benediction of the Blessed Sacrament, in parish and other smaller churches.\end{otherlanguage}}

%\rubrique{\begin{otherlanguage}{english}
%Then the office is concluded with the following verses, the Marian antiphon, etc.
%\end{otherlanguage}}

\end{document}