% !TEX TS-program = lualatexmk
% !TEX parameter =  --shell-escape
\documentclass[12pt]{book}

\input{commonheaders_vespers_people} %%% will need to fix spacing problem in psalms at a later date.

\begin{document}
 
 \twosided[pc] %% to allow for the left-right switching of columns needed for bilingual liturgical typesetting (like in missals); it's unclear why this goes BEFORE everything else and not when activating the environment. %% It's OK for paper booklets to not switch, but it's probably best practice to switch.
 
  \thispagestyle{empty}
 
 {\centering{\capspace{\huge{VESPERS OF AN APOSTLE OR \\[0.25\baselineskip]EVANGELIST.}}}\par}\header{vespers of an apostle or evangelist.}
 
%{\centering{\huge\capspace{{ OR EVANGELIST.}}}\par}
 
 \vspace{0.5\baselineskip}
 
{\centering{\huge{At I Vespers.}}\par}
 
\smalltitle{Before the Office.}

\rubrique{These prayers are said kneeling. End when the priest stands.}

\textebilingue{orationes_ante_do}{orationes_ante_do_eng}

\smalltitle{\prayers}

\rubrique{The celebrant intones the following verse. All sing the response and doxology.}

\begin{otherlanguage}{english}
\rubrique{Make the sign of the cross at the beginning and at the Magnificat, and bow the head when the Trinity is mentioned, when mention is made of the ``Holy Name'' in some way, and at ``Orémus.''}\end{otherlanguage}

\gscore[]{℣.}{Deus_In_Adjutorium_Festivus}

\noindent
\begin{otherlanguage}{english}
O God, \cc{} come to my assistance. \rr O Lord, make haste to help me. Glory be to the Father, and to the Son, and to the Holy Ghost. As it was in the beginning, is now, and ever shall be, world without end. Amen. Alleluia.\end{otherlanguage}

\rubrique{After Septuagesima:}

\smallscore{Laus_tibi_Domine}
\noindent Praise be unto Thee, O Lord, King of Eternal Glory.

\smalltitle{First Antiphon. 8. c.}
\initialscore{an_hoc_est_praeceptum_meum_solesmes_1961}

\begin{otherlanguage}{english} \noindent Ant. This is my commandment that you love one another, as I have loved you.\end{otherlanguage}

\rubrique{\begin{otherlanguage}{english}
Sit at the word ``Sede.'' The cantor sings odd verses; all sing even verses, and this for the hymn and Magnificat, when sung entirely in chant, as well.
\end{otherlanguage}}

\psalmus{109}{109_8c}{109_8}{109_en}

\rubrique{\begin{otherlanguage}{english}
Clerics in descending order, or a cantor, intone the subsequent antiphons.\end{otherlanguage}}

 \gscore[2. Ant.]{1. g}{an_majorem_caritatem_solesmes_1961}
% 
 \begin{otherlanguage}{english} \noindent Ant. Greater love than this no man hath, that a man lay down his life for his friends. \end{otherlanguage}
%
\psalmus{110}{110_1g}{110_1g}{110_en}
%
  \gscore[3. Ant.]{1. a3}{an_vos_amici_mei_estis_solesmes_1961}
%  
  \begin{otherlanguage}{english} \noindent Ant. You are my friends if you do the things that I command you, said the Lord. \end{otherlanguage}
%
 \psalmus{111}{111_1a3}{111_1a3}{111_en}
%
\gscore[4. Ant.]{1. f}{an_beati_pacifici_solesmes_1961}
%
\begin{otherlanguage}{english} \noindent Ant. You are my friends if you do the things that I command you, said the Lord.\end{otherlanguage}
%
\psalmus{112}{112_1f}{112_1f}{112_en}

 \gscore[5. Ant.]{1. g}{an_in_patientia_vestra_solesmes_1961}
 
\begin{otherlanguage}{english} \noindent Ant. In your patience you shall possess your souls. \end{otherlanguage}

\psalmus{116}{116_1g}{116_1g}{116_en}

\rubrique{\begin{otherlanguage}{english}
Having sung the last antiphon in full, stand when the celebrant does.
\end{otherlanguage}}

\smalltitle{Chapter.}
\scripture{Ephes. 2, 19 – 20.}

\texteparacol{\lettrine{F}{r}atres: Jam non estis hóspites et ádvenæ:~† sed estis cives sanctórum et doméstici Dei: superædificáti super fundaméntum Apostolórum et Prophetárum,~* ipso summo angulári lápide Christo Jesu.\par
\rr Deo grátias.}{\noindent Brothers: Now therefore you are no more strangers and foreigners; but you are fellow citizens with the saints, and the domestics of God, built upon the foundation of the apostles and prophets, Jesus Christ himself being the chief corner stone. Thanks be to God.}

\smalltitle{Hymn.}

\gscore[]{4.}{hy_exsultet_orbis_gaudiis_solesmes_1961}

\begin{multicols}{2}
\begin{otherlanguage}{english}
Now let the earth with joy resound,\\
And heaven the chant re-echo round;\\
Nor heaven nor earth too high can raise\\
The great Apostles' glorious praise.

O ye who, throned in glory dread,\\
Shall judge the living and the dead,\\
Lights of the world forevermore!\\
To you the suppliant prayer we pour.

Ye close the sacred gates on high;\\
At your command apart they fly:\\
Oh! loose for us the guilty chain\\
We strive to break, and strive in vain.

Sickness and health your voice obey;\\
At your command they go or stay:\\
From sin's disease our souls restore;\\
In good confirm us more and more.

So when the world is at its end,\\
And Christ to judgement shall descend,\\
May we be called those joys to see\\
Prepared from all eternity.

Praise to the Father, with the Son,\\
And Holy Spirit, Three in One;\\
As ever was in ages past,\\
And so shall be while ages last. Amen.

\end{otherlanguage}
\end{multicols}

\rubrique{\begin{otherlanguage}{english}
The cantor sings the verse to which all reply.
\end{otherlanguage}}

\gresetinitiallines{0}
\gabcsnippet{(c3) ℣. In(h) om(h)nem(h) ter(h)ram(h) ex(h)í(h)vit(h) so(h)nus(h) e(h)ó(h)rum.(g_hvGFEfgf.) (::) (Z-)
℟. Et(h) in(h) fi(h)nes(h) or(h)bis(h) ter(h)ræ(h) ver(h)ba(h) e(h)ó(h)rum.(g_hvGFEfgf.) (::)}
\gresetinitiallines{1}

%\smallscore{versiculus_i_vesperis_nativitatis_domini}

\begin{otherlanguage}{english} \noindent \vv Their sound hath gone forth into all the earth.

\noindent \rr And their words unto the ends of the world.
 \end{otherlanguage}
 
\smalltitle{Antiphon at the Magnificat.}

\rubrique{Sit when the celebrant does, after he sings the first words. Rise at the end.}

\gscore[]{1. f}{an_tradent_enim_vos_solesmes_1961}
 
 \begin{otherlanguage}{english} \noindent Ant. For they shall deliver you up to councils: and in the synagogues you shall be beaten: and you shall stand before governors and kings for my sake, for a testimony unto them. \end{otherlanguage}

\gresetinitiallines{0}
\gabcsnippet{(c4) 1. Ma(f)gní(gh)fi(h)cat(h) * (:) á(h)ni(h)ma(h) <i>me</i>(g)a(f) <b>Dó</b>(gh)mi(gr)num.(gf..) (::)
 (Z-)

2. Et(f) ex(gh)sul(h)tá(h)vit(h) <i>spí</i>(hg)<i>ri</i>(ixgi)<i>tus</i>(h hr[ocb:1{]) <b>me</b>(hg[ocb:0}])us(gh..) *(:) in(h) De(h)o(h) sa(h)lu(h)<i>tá</i>(g)<i>ri</i>(f) <b>me</b>(gh//gr)o.(gf..) (::)}
\gresetinitiallines{1}
   \smallskip%
    \noindent%
    \begin{itemize}[%
    		label=\null, %
			leftmargin=0pt, %
			itemindent=0mm, %
			labelsep=0pt, %
			labelwidth=0pt, %
			rightmargin=0pt, %
			parsep=0pt, %
			topsep=0pt, %
			itemsep=0pt]%
\item 3. Quia respéxit humilitátem \textit{ancíllæ} \textbf{su}æ:~* ecce enim ex hoc beátam me dicent omnes gene\textit{rati}\textbf{ó}nes.

\item 4. Quia fecit mihi \textit{magna} \textit{qui} \textbf{pot}\textbf{ens} est:~* et sanctum \textit{nomen} \textbf{e}jus.

\item 5. Et misericórdia ejus a progéni\textit{e} \textit{in} \textit{pro}\textbf{gé}\textbf{ni}es~* timén\textit{tibus} \textbf{e}um.

\item 6. Fecit poténtiam in \textit{bráchio} \textbf{su}o:~* dispérsit supérbos mente \textit{cordis} \textbf{su}i.

\item 7. Depósuit po\textit{téntes} \textit{de} \textbf{se}de,~* et exal\textit{távit} \textbf{hú}miles.

\item 8. Esuriéntes \textit{implévit} \textbf{bo}nis:~* et dívites dimí\kern 0.04em\textit{sit} \textit{in}\textbf{á}nes.

\item 9. Suscépit Israel \textit{púerum} \textbf{su}um,~* recordátus misericór\textit{diæ} \textbf{su}æ.

\item 10. Sicut locútus est \textit{ad} \textit{pa}\textit{tres} \textbf{no}stros,~* Abraham et sémini e\kern 0.02em\textit{jus} \textit{in} \textbf{sǽ}cula.

\item 11. Glória \textit{Pa}\textit{tri,} \textit{et} \textbf{Fí}\textbf{li}o,~* et Spirí\textit{tui} \textbf{San}cto.

\item 12. Sicut erat in princípio, \textit{et} \textit{nunc,} \textit{et} \textbf{sem}per,~* et in sǽcula sæcu\hspace{0.03em}\textit{lórum.} \textbf{A}men.

    \end{itemize}

\smalltitle{Canticle of the Blessed Virgin Mary.}

\scripture{Luc. 1, 46–55.}
\begin{multicols}{2}
%\begin{paracol}{2}
%
%\lettrine{M}agníficat~* ánima mea Dóminum.\\ Et exsultávit spíritus meus~* in Deo salutári meo.
%
% Quia respéxit humilitátem ancíllæ suæ:~* ecce enim ex hoc beátam me dicent o\-mnes generatiónes.
%
% Quia fecit mihi magna qui potens est:~* et san\-ctum nomen ejus.
%
% Et misericórdia ejus a progénie in progénies~* timéntibus eum.
%
% Fecit poténtiam in bráchio suo:~* dispérsit supérbos mente cordis sui.
%
% Depósuit poténtes de sede,~* et exaltávit húmiles.
%
% Esuriéntes implévit bonis:~* et dívites dimísit inánes.
%
% Suscépit Israel púerum suum,~* recordátus misericórdiæ suæ.
%
% Sicut locútus est ad patres no\-stros,~* Abraham, et sémini ejus in sǽcula.
%
% Glória Patri, et Fílio,~* et Spirítui San\-cto.
%
% Sicut erat in princípio, et nunc, et semper,~* et in sǽcula sæculórum. Amen.
 
% \switchcolumn
 \begin{otherlanguage}{english}
My soul \cc~* doth magnify the Lord.

And my spirit hath rejoiced~* in God my Saviour.

Because he hath regarded the humility of his handmaid;~* for behold from henceforth all generations shall call me blessed.

Because he that is mighty, hath done great things to me;~* and holy is his name.

And his mercy is from generation unto generations,~* to them that fear him.

He hath shewed might in his arm:~* he hath scattered the proud in the conceit of their heart.

He hath put down the mighty from their seat,~* and hath exalted the humble.

He hath filled the hungry with good things;~* and the rich he hath sent empty away.

He hath received Israel his servant,~* being mindful of his mercy:

As he spoke to our fathers,~* to Abraham and to his seed for ever.

Glory be to the Father, and to the Son,~* and to the Holy Ghost.

As it was in the beginning, is now,~* and ever shall be, world without end. Amen.
\end{otherlanguage}
% \end{paracol}
\end{multicols}

\rubrique{\begin{otherlanguage}{english}
Sit as the antiphon is repeated. Then the usual dialogue follows. All stand from now until the end of the office.\end{otherlanguage}}

\textebilingue{dv}{dv_eng}

%\textebilingue{oratio_15_aug_old}{oratio_15_aug_old_en}

\rubrique{The collect is proper to the feast.}

\textebilingue{dv}{dv_eng}

\rubrique{\begin{otherlanguage}{english}
The cantor sings the verse to which all reply.
\end{otherlanguage}}

\gscore[]{2.}{BD_Solemnis_I_Vesp}

\noindent \vv Let us bless the Lord.

\noindent \rr Thanks be to God.

\textebilingue{fidelium_animae}{fidelium_animae_en}
    
%\rubrique{\begin{otherlanguage}{english}
%Then one Our Father is said, unless Compline follows, or even some pious exercises, a sermon, or Benediction of the Blessed Sacrament, in parish and other smaller churches.\end{otherlanguage}}

%\rubrique{\begin{otherlanguage}{english}
%Then the office is concluded with the following verses, the Marian antiphon, etc.
%\end{otherlanguage}}

\end{document}