% !TEX TS-program = LuaLaTeX+se
\documentclass[11pt]{book}

\input{commonheaders_vespers_people}

\begin{document}

\twosided[pc] %%% why this goes at beginning, IDK, unclear what p does exactly, but it's necessary for flipping columns correctly. ; it's unclear why this goes BEFORE everything else and not when activating the environment. %% It's OK for paper booklets to not switch, but it's probably best practice to switch.

\thispagestyle{empty}

\bigtitle{At the End of Lauds or Vespers Celebrated in the Absence of a Deacon or Priest.}

\medskip

\rubrique{Before the collect and the verse \normaltext{Benedicámus Dómino:}}

\gresetinitiallines{0}
\gabcsnippet{(c4) ℣. Dó(g)mi(h)ne(h) ex(h)aú(h)di(g) o(g)ra(g)ti(g)ó(g)nem(h) me(gh)am.(g.) (::) <sp>R/</sp>. Et(g) cla(h)mor(h) me(h)us(g) ad(g) te(g) vé(h)ni(h)at.(h.) (::)}
\gresetinitiallines{1}

\begin{enpars}
\noindent \vv O Lord hear my prayer.

\noindent \rr And let my cry come unto Thee.
\end{enpars}

\rubrique{After the verse \normaltext{Fidélium ánimæ} is sung, the Lord's Prayer is said silently. Then the following is sung, also on a lower note.}

\texteparacol{
\noindent \vv Dóminus det nobis suam pacem.

\noindent \rr Et vitam ætérnam. Amen.}
{english}
\noindent \vv May the Lord grant us his peace.

\noindent \rr And life everlasting. Amen.}

\rubrique{\begin{enpars}On Sundays and in Paschal Time, all remain standing, but on weekdays, the choir and congregation kneel, then the officiant kneels after intoning the antiphon. He rises for the collect on feasts and on any weekday which is not a penitential feria.\end{enpars}}

\rubrique{\begin{enpars} The officiant intones the Marian antiphon according to the season, followed by the versicle and collect in the usual way. The versicle \normaltext{Divínum auxílium} follows, \normaltext{\pageref{conclusion}.}\end{enpars}}

\smalltitle{Alma Redemptoris Mater.}

\rubrique{¶ From Vespers of the Saturday before the I Sunday of Advent to Vespers of February 2 inclusive.}

\gscore[]{5.}{an--alma_redemptoris--solesmes_1961}
%%\vspace{1ex}

\smalltitle{Simple Tone.}
\gscore[]{5.}{an--alma_redemptoris_simple_solesmes}

\textes{alma_redemptoris_la_en}

\smalltitle{Ave Regina Cælorum.}

\rubrique{¶ After the Purification, that is, from Compline of February 2 even if the feast is transferred, until Compline of Wednesday in Holy Week.}

\gscore[]{6.}{an--ave_regina_caelorum--solesmes_1961}
%\vspace{1ex}

\smalltitle{Simple Tone.}

\gscore[]{6.}{an--ave_regina_caelorum_simple_tone--solesmes}

\textes{ave_regina_la_en}

\smalltitle{Regina Cæli.}

\rubrique{¶ From Compline of Holy Saturday until None of Saturday after Pentecost inclusive.}

\gscore[]{6.}{an--regina_caeli--solesmes}
%\vspace{1ex}

\smalltitle{Simple Tone.}

\gscore[]{6.}{an--regina_caeli_simple_tone--solesmes}

\textes{regina_caeli_la_en}

\smalltitle{Salve Regina.}

\rubrique{¶ From Vespers of the Holy Trinity until None of the Saturday before Advent inclusive.}

\gscore[]{1.}{an--salve_regina--solesmes}
%\vspace{1ex}

\smalltitle{Simple Tone.}
\gscore[]{5.}{an--salve_regina_simple_tone_solesmes}

\textes{salve_regina_la_en}

\rubrique{The versicle is sung on a lower note.}
\label{conclusion} \phantomsection

\texteparacol{\noindent \vv Divínum auxílium máneat semper nobíscum.

\noindent \rr Amen.}
{english}
{\noindent \vv May the divine assistance remain always with us.

\noindent \rr Amen.}

\end{document}