% !TEX TS-program = lualatexmk
% !TEX parameter =  --shell-escape
\documentclass[12pt]{book}

\input{commonheaders_vespers_people} %%% will need to fix spacing problem in psalms at a later date.

\begin{document}
 
 \twosided[pc] %% to allow for the left-right switching of columns needed for bilingual liturgical typesetting (like in missals); it's unclear why this goes BEFORE everything else and not when activating the environment. %% It's OK for paper booklets to not switch, but it's probably best practice to switch.
 
  \thispagestyle{empty}
 
 {\centering{\capspace{\huge{VESPERS OF SAINT THOMAS, \\[0.25\baselineskip]APOSTLE.}}}\par}\header{vespers of saint thomas, apostle.}
 
%{\centering{\huge\capspace{{ OR EVANGELIST.}}}\par}
 
 \vspace{0.5\baselineskip}
 
{\centering{\huge{At I Vespers.}}\par}
 
\smalltitle{Before the Office.}

\rubrique{These prayers are said kneeling. End when the priest stands.}

\textes{orationes_ante_do_la_en}

\smalltitle{\prayers}

\rubrique{The celebrant intones the following verse. All sing the response and doxology.}

\begin{enpars}
\rubrique{Make the sign of the cross at the beginning and at the Magnificat, and bow the head when the Trinity is mentioned, when mention is made of the ``Holy Name'' in some way, and at ``Orémus.''}\end{enpars}

\gscore[]{℣.}{Deus_In_Adjutorium_Festivus}

\noindent
\begin{enpars}
O God, \cc{} come to my assistance. \rr O Lord, make haste to help me. Glory be to the Father, and to the Son, and to the Holy Ghost. As it was in the beginning, is now, and ever shall be, world without end. Amen. Alleluia.\end{enpars}
\newpage

\firstantiphon{8. c}
\initialscore{an_hoc_est_praeceptum_meum_solesmes_1961}

\antiphona{english}{This is my commandment that you love one another, as I have loved you}

\rubrique{\begin{enpars}
Sit at the word ``Sede.'' The cantor sings odd verses; all sing even verses, and this for the hymn and Magnificat, when sung entirely in chant, as well.
\end{enpars}}

\psalmus{109}{109_8c}{109_8}{109_en}

\rubrique{\begin{enpars}
Clerics in descending order, or a cantor, intone the subsequent antiphons.\end{enpars}}

 \gscore[2. Ant.]{1. g}{an_majorem_caritatem_solesmes_1961}
 \antiphona{english}{Greater love than this no man hath, that a man lay down his life for his friends}
\psalmus{110}{110_1g}{110_1}{110_en}

  \gscore[3. Ant.]{1. a3}{an_vos_amici_mei_estis_solesmes_1961}
  \antiphona{english}{You are my friends if you do the things that I command you, said the Lord}
 \psalmus{111}{111_1a3}{111_1}{111_en}

\gscore[4. Ant.]{1. f}{an_beati_pacifici_solesmes_1961}
\antiphona{english}{Blessed are the peacemakers the clean of heart, they shall see God}
\psalmus{112}{112_1f}{112_1}{112_en}

 \gscore[5. Ant.]{1. g}{an_in_patientia_vestra_solesmes_1961}
\antiphona{english}{In your patience you shall possess your souls}
\psalmus{116}{116_1g}{116_1}{116_en}

\rubrique{\begin{enpars}
Having sung the last antiphon in full, stand when the celebrant does.
\end{enpars}}

\chapterlateng
\scripture{Ephes. 2, 19 – 20.}

\texteparacol{\lettrine{F}{r}atres: Jam non estis hóspites et ádvenæ:~† sed estis cives san\-ctórum et doméstici Dei: superædificáti super fundaméntum Apostolórum et Prophetárum,~* ipso summo angulári lápide Christo Jesu.}{english}{\noindent Brethren: Now therefore you are no more strangers and foreigners; but you are fellow citizens with the saints, and the domestics of God, built upon the foundation of the apostles and prophets, Jesus Christ himself being the chief cornerstone.}

\hymnlateng
%\grechangenextscorelinedim{2}{spacebeneathtext}{0.1cm plus 0cm minus 0cm}{scalable}
%\grechangenextscorelinedim{4}{spacebeneathtext}{0.1cm plus 0cm minus 0cm}{scalable}
%\grechangenextscorelinedim{6}{spacebeneathtext}{0.1cm plus 0cm minus 0cm}{scalable}
%\grechangenextscorelinedim{7}{spacebeneathtext}{0.1cm plus 0cm minus 0cm}{scalable}
\grechangedim{spacebeneathtext}{0.2cm}{scalable}
\gscore[]{4.}{hy_exsultet_caelum_laudibus_textus_antiquus}
\grechangedim{spacebeneathtext}{0cm}{scalable}

\textes{exsultet_orbis_en}

\rubrique{\begin{enpars}
The cantor sings the verse to which all reply.
\end{enpars}}

\smallscore{versiculus_1_vesperis_apostolorum_extra_tp}

\begin{enpars} \noindent \vv Their sound hath gone forth into all the earth.

\noindent \rr And their words unto the ends of the world.
 \end{enpars}
 
\smalltitle{Antiphon at the Magnificat.}

\rubrique{Sit when the celebrant does, after he sings the first words. Rise at the end.}

\gscore[]{8. G}{an_quia_vidisti_me_solesmes_1961}
 
 \antiphona{english}{Because thou hast seen me, Thomas, thou hast believed: blessed are they that have not seen and have believed, alleluia}

\smallscore{Magnificat_8_G_Solemnis}

\magnificattexte{Magnificat_8_Solemnis}

\textes{magnificat_en_multicols}

\rubrique{\begin{enpars}
Sit as the antiphon is repeated. Then the usual dialogue follows. All stand from now until the end of the office.\end{enpars}}

\texteparacol{\vv Dóminus vobíscum.

\rr Et cum spíritu tuo.

\lettrine{D}{a} nobis, quǽsumus Dómine, beáti Apóstoli tui Thomæ solemnitátibus gloriári:~† ut ejus semper et patrocíniis sublevémur;~* et fidem cóngrua devotióne sectémur. Per Dóminum.}
{english}{\noindent \vv The Lord be with you.

\noindent \rr  And with thy spirit.

\noindent Grant us, Lord, we beseech thee, to glory in the solemnity of thy blessed Apostle Thomas, that we may ever be aided by his patronage, and follow his faith with true devotion. Through our Lord.}

\rubrique{Then follows the commemoration of the major feria of Advent, December 20.}

\gscore[Ant.]{2.}{an_o_clavis_david_solesmes_1961}
\antiphona{english}{O Key of David, and Sceptre of the house of Israel that openest, and no man shutteth; and shuttest and no man openeth; come to bring out the prisoners from the prison, and them that sit in darkness, and in the shadow of death}

\texteparacol{
\vv Roráte cæli désuper, et nubes pluant justum.

\rr Aperiátur terra, et gérminet Salvatórem.

\lettrine{E}{x}cita, quǽsumus Dómine, poténtiam tuam, et veni: et magna nobis virtúte succúrre:~† ut per auxílium grátiæ tuæ, quod nostra peccáta præpédiunt,~* indulgéntia tuæ propitiatiónis accéleret. Qui vivis et regnas cum Deo Patre.}{english}{
\noindent \vv Drop down dew, ye heavens, from above, and let the clouds rain the Just.

\noindent \rr Let the earth be opened, and bud forth the Saviour.

\noindent Stir up, we beseech thee, O Lord, thy power, and come; make haste to our aid with thy great might; that, by the help of thy grace, that which is hindered by our sins may be hastened by thy merciful forgiveness. Who livest and reignest with God the Father.}

\myrule
\rubrique{But when the feast falls on a Sunday, such that I Vespers are celebrated instead on that day, then the antiphon is as follows, with the verse and collect as above.}

\gscore[Ant.]{2.}{an_o_oriens_solesmes_1961}
\antiphona{english}{O Day-Spring, brightness of the everlasting Light, Sun of Righteousness; come, to give light to them that sit in darkness, and in the shadow of death}

\myrule

\texteparacol{\vv Dóminus vobíscum.

\rr Et cum spíritu tuo.}
{english}{\noindent \vv The Lord be with you.

\noindent \rr  And with thy spirit.}

\rubrique{\begin{enpars}
The cantor sings the verse to which all reply.
\end{enpars}}

\gscore[]{2.}{BD_Solemnis_I_Vesp}

\noindent \vv Let us bless the Lord.

\noindent \rr Thanks be to God.

\textebilingue{fidelium_animae}{fidelium_animae_en}
    
%\rubrique{\begin{enpars}
%Then one Our Father is said, unless Compline follows, or even some pious exercises, a sermon, or Benediction of the Blessed Sacrament, in parish and other smaller churches.\end{enpars}}

%\rubrique{\begin{enpars}
%Then the office is concluded with the following verses, the Marian antiphon, etc.
%\end{enpars}}

\end{document}