% !TEX TS-program = LuaLaTeX+se

\documentclass[12pt]{book}

\input{commonheaders_vespers_people} %%% will need to fix spacing problem in psalms at a later date.

\begin{document}
 
 \twosided[pc] %% to allow for the left-right switching of columns needed for bilingual liturgical typesetting (like in missals); it's unclear why this goes BEFORE everything else and not when activating the environment. %% It's OK for paper booklets to not switch, but it's probably best practice to switch.
 
  \thispagestyle{empty}
 
 {\centering{\huge{SUNDAY OF THE RESURRECTION.}}\par}\header{sunday of the resurrection.}
 \vspace{0.5\baselineskip}
 
%{\centering{\huge{And at Vespers of the Paschal Octave.}}\par}
 
\smalltitle{Before the Office.}

\rubrique{These prayers are said kneeling. End when the priest stands.}

\textebilingue{orationes_ante_do}{orationes_ante_do_eng}

\smalltitle{\prayers}

\rubrique{The celebrant intones the following verse. All sing the response and doxology.}

\begin{enpars}
\rubrique{Make the sign of the cross at the beginning and at the Magnificat, and bow the head when the Trinity is mentioned, when mention is made of the ``Holy Name'' in some way, and at ``Orémus.''}\end{enpars}

\gscore[]{℣.}{Deus_in_adjutorium_tonus_solemnior}

\noindent
\begin{enpars}
O God, \cc{} come to my assistance. \rr O Lord, make haste to help me. Glory be to the Father, and to the Son,~* and to the Holy Ghost. As it was in the beginning, is now,~* and ever shall be, world without end. Amen. Alleluia.\end{enpars}

\smalltitle{First Antiphon. 8. G.}
 \initialscore{an_angelus_autem_domini_solesmes_1961}

\begin{enpars} \noindent Ant. For an angel of the Lord descended from heaven and coming rolled back the stone and sat upon it, alleluia, alleluia. \end{enpars}

\rubrique{\begin{enpars}
Sit at the word ``Sede.'' The cantor sings odd verses; all sing even verses, and this also for the Magnificat, when sung entirely in chant, as well.
\end{enpars}}

\psalmus{109}{109_8G}{109_8G}{109_en}

\rubrique{\begin{enpars}
Clerics in descending order, or a cantor, intone the subsequent antiphons.\end{enpars}}

 \gscore[2. Ant.]{7. c}{an_et_ecce_terraemotus_solesmes_1961}
% 
 \begin{enpars} \noindent Ant. There was a great earthquake; For an angel of the Lord descended from heaven, alleluia. \end{enpars}
%
\psalmus{110}{110_7c}{110_7c}{110_en}
%
\gscore[3. Ant.]{8. c}{an_erat_autem_aspectus_solesmes_1961}
%  
 \begin{enpars} \noindent Ant. And his countenance was as lightning and his raiment as snow, alleluia, alleluia. \end{enpars}
%
  \psalmus{111}{111_8c}{111_8G}{111_en}
%
\gscore[4. Ant.]{7. a}{an_prae_timore_solesmes_1961}
%
\begin{enpars} \noindent Ant. And for fear of him the guards were struck with terror and became as dead men, alleluia.  \end{enpars}
%
  \psalmus{112}{112_7a}{112_7a}{112_en}
%
\gscore[5. Ant.]{8. G}{an_respondens_autem_angelus_solesmes_1961}
% 
\begin{enpars} \noindent Ant. And the angel answering said to the women: Fear not you: for I know that you seek Jesus, alleluia.\end{enpars}
%
\psalmus{113}{113_8G}{113_8G}{113_en}

\rubrique{\begin{enpars}
Having sung the last antiphon in full, stand when the celebrant does.
\end{enpars}}

\rubrique{The hymn, chapter, and verse are omitted. The following antiphon is sung.}

\gscore[Ant.]{2.}{an_haec_dies_solesmes_1961}
\begin{enpars} \noindent
Ant. This is the day which the Lord hath made: let us be glad and rejoice therein.\end{enpars}
 
\smalltitle{Antiphon at the Magnificat.}

\rubrique{Sit when the celebrant does, after he sings the first words. Rise at the end.}

\gscore[]{3. a}{an_et_respicientes_solesmes_1961}
 
 \begin{enpars} \noindent Ant. And looking they saw the stone rolled back, for it was very great, alleluia.\end{enpars}
 
  \rubrique{\begin{enpars}
The Magnificat may be sung entirely in chant or sung alternating between a cantor singing chant and the choir singing a polyphonic or harmonized setting.\end{enpars}}

% 
\gresetinitiallines{0}
\gabcsnippet{(c4) 1. Ma(g)gní(hj)fi(j)cat(j) *(:) á(j)ni(j)ma(j) me(j)<i>a</i>(h) <b>Dó</b>(j)mi(jr)num.(ih..) (::) (Z-)
2. Et(g) ex(hj)sul(j)tá(j)vit(j) <b>spí</b>(jk)ri(jr)tus(j) (jr[ocb:1;8.7mm])<b>me</b>(ih)us(hj..) *(:) in(j) De(j)o(j) sa(j)lu(j)tá(j)<i>ri</i>(h) me(j//jr)o.(ih..) (::)}
%
\gresetinitiallines{1}
   \smallskip%
    \noindent%
    \begin{itemize}[%
    		label=\null, %
			leftmargin=0pt, %
			itemindent=0mm, %
			labelsep=0pt, %
			labelwidth=0pt, %
			rightmargin=0pt, %
			parsep=0pt, %
			topsep=0pt, %
			itemsep=0pt]%
\item 3. Quia respéxit humilitátem an\textbf{cíl}læ \textbf{su}æ:~* ecce enim ex hoc beátam me dicent omnes genera\kern 0.01 em \textit{ti}\textbf{ó}nes.

\item 4. Quia fecit mihi \textbf{ma}gna qui \textbf{pot}\textbf{ens} est:~* et sanctum no\textit{men} \textbf{e}jus.

\item 5. Et misericórdia ejus a progénie \textbf{in} pro\textbf{gé}\textbf{ni}es~* timénti\textit{bus} \textbf{e}um.

\item 6. Fecit poténtiam in \textbf{brá}chio \textbf{su}o:~* dispérsit supérbos mente cor\textit{dis} \textbf{su}i.

\item 7. Depósuit po\textbf{tén}tes de \textbf{se}de,~* et exaltá\textit{vit} \textbf{hú}miles.

\item 8. Esuriéntes im\textbf{plé}vit \textbf{bo}nis:~* et dívites dimísit \textit{in}\textbf{á}nes.

\item 9. Suscépit Israel \textbf{pú}erum \textbf{su}um,~* recordátus misericórdi\kern 0.02 em\textit{æ} \textbf{su}æ.

\item 10. Sicut locútus est ad \textbf{pa}tres \textbf{no}stros,~* Abraham et sémini ejus \textit{in} \textbf{sǽ}cula.

\item 11. Glória \textbf{Pa}tri, et \textbf{Fí}\textbf{li}o,~* et Spirítu\textit{i} \textbf{San}cto.

\item 12. Sicut erat in princípio, et \textbf{nunc,} et \textbf{sem}per,~* et in sǽcula sæculó\textit{rum}. \textbf{A}men.
    \end{itemize}

\smalltitle{Canticle of the Blessed Virgin Mary.}

\scripture{Luc. 1, 46–55.}
\begin{multicols}{2}
%\begin{paracol}{2}
%
%\lettrine{M}agníficat~* ánima mea Dóminum.\\ Et exsultávit spíritus meus~* in Deo salutári meo.
%
% Quia respéxit humilitátem ancíllæ suæ:~* ecce enim ex hoc beátam me dicent o\-mnes generatiónes.
%
% Quia fecit mihi magna qui potens est:~* et san\-ctum nomen ejus.
%
% Et misericórdia ejus a progénie in progénies~* timéntibus eum.
%
% Fecit poténtiam in bráchio suo:~* dispérsit supérbos mente cordis sui.
%
% Depósuit poténtes de sede,~* et exaltávit húmiles.
%
% Esuriéntes implévit bonis:~* et dívites dimísit inánes.
%
% Suscépit Israel púerum suum,~* recordátus misericórdiæ suæ.
%
% Sicut locútus est ad patres no\-stros,~* Abraham, et sémini ejus in sǽcula.
%
% Glória Patri, et Fílio,~* et Spirítui San\-cto.
%
% Sicut erat in princípio, et nunc, et semper,~* et in sǽcula sæculórum. Amen.
 
% \switchcolumn
 \begin{enpars}
My soul \cc~* doth magnify the Lord.

And my spirit hath rejoiced~* in God my Saviour.

Because he hath regarded the humility of his handmaid;~* for behold from henceforth all generations shall call me blessed.

Because he that is mighty, hath done great things to me;~* and holy is his name.

And his mercy is from generation unto generations,~* to them that fear him.

He hath shewed might in his arm:~* he hath scattered the proud in the conceit of their heart.

He hath put down the mighty from their seat,~* and hath exalted the humble.

He hath filled the hungry with good things;~* and the rich he hath sent empty away.

He hath received Israel his servant,~* being mindful of his mercy:

As he spoke to our fathers,~* to Abraham and to his seed for ever.

Glory be to the Father, and to the Son,~* and to the Holy Ghost.

As it was in the beginning, is now,~* and ever shall be, world without end. Amen.
\end{enpars}
% \end{paracol}
\end{multicols}

\rubrique{\begin{enpars}
Sit as the antiphon is repeated. Then the usual dialogue follows. All stand from now until the end of the office.\end{enpars}}

\textebilingue{dv}{dv_eng}

\texteparacol{Orémus.

\lettrine{D}{e}us, qui hodiérna die per Unigénitum tuum æternitátis nobis áditum deví\-cta morte reserásti:~† vota nostra, quæ præveniéndo aspíras,~* étiam adjuvándo proséquere.
Per eúmdem Dóminum nostrum.

\rr Amen.}{\noindent Let us pray. O God, who on this day, through thy Only-begotten Son, overcame death, and opened to us the gate of everlasting life: as by thy anticipating grace, thou breathest good desires into our hearts, so also, by thy gracious help, bring them to good effect. Through the same Jesus Christ, thy Son, Our Lord, Who liveth and reigneth with thee in the unity of the Holy Ghost, God, world without end. Amen.}

\textebilingue{dv}{dv_eng}

\textes{cantor_rubric}

\smallscore{Benedicamus_domino_alleluia_solesmes_1961}

\noindent \vv Let us bless the Lord.

\noindent \rr Thanks be to God.

\rubrique{This tone may be used today through Vespers of Easter Tuesday inclusively.}

\smallscore{Benedicamus_Dom_resurrectionis_solesmes_1961}

\textebilingue{fidelium_animae}{fidelium_animae_en}
    
%\rubrique{\begin{enpars}
%Then one Our Father is said, unless Compline follows, or even some pious exercises, a sermon, or Benediction of the Blessed Sacrament, in parish and other smaller churches.\end{enpars}}

%\rubrique{\begin{enpars}
%Then the office is concluded with the following verses, the Marian antiphon, etc.
%\end{enpars}}

\end{document}