% !TEX TS-program = LuaLaTeX+se

\documentclass[12pt]{book}

\input{commonheaders_vespers_people} %%% will need to fix spacing problem in psalms at a later date.

\begin{document}
 
 \twosided[pc] %% to allow for the left-right switching of columns needed for bilingual liturgical typesetting (like in missals); it's unclear why this goes BEFORE everything else and not when activating the environment. %% It's OK for paper booklets to not switch, but it's probably best practice to switch.
 
  \thispagestyle{empty}
 
 {\centering{\huge{HOLY FAMILY}}\par}\header{holy family}
 \vspace{0.5\baselineskip}
 
{\centering{\huge{At I Vespers.}}\par}

{\centering{\textit{Or at II Vespers, when the Epiphany falls on Saturday.}}\par}

%{\centering{{\textit{or when I Vespers are impeded by II Vespers of the Epiphany.}}\par}

\smalltitle{Before the Office.}

\rubrique{These prayers are said kneeling. End when the priest stands.}

\textebilingue{orationes_ante_do}{orationes_ante_do_eng}

\smalltitle{\prayers}

\rubrique{The celebrant intones the following verse. All sing the response and doxology.}

\begin{otherlanguage}{english}
\rubrique{Make the sign of the cross at the beginning and at the Magnificat, and bow the head when the Trinity is mentioned, when mention is made of the ``Holy Name'' in some way, and at ``Orémus.''}\end{otherlanguage}

\gscore[]{℣.}{Deus_In_Adjutorium_Festivus}

\noindent
\begin{otherlanguage}{english}
O God, \cc{} come to my assistance. \rr O Lord, make haste to help me. Glory be to the Father, and to the Son,~* and to the Holy Ghost. As it was in the beginning, is now,~* and ever shall be, world without end. Amen. Alleluia.\end{otherlanguage}

\smalltitle{First Antiphon. 1. g.}
\initialscore{an_jacob_autem_genuit_solesmes_1961}

\begin{otherlanguage}{english} \noindent Ant. And Jacob begot Joseph the husband of Mary, of whom was born Jesus, who is called Christ. \end{otherlanguage}

\rubrique{\begin{otherlanguage}{english}
Sit at the word ``Sede.'' The cantor sings odd verses; all sing even verses, and this for the hymn and Magnificat, when sung entirely in chant, as well.
\end{otherlanguage}}

\psalmus{109}{109_1g}{109_1g}{109_en}

\rubrique{\begin{otherlanguage}{english}
Clerics in descending order, or a cantor, intone the subsequent antiphons.\end{otherlanguage}}

 \gscore[2. Ant.]{7. c}{an_angelus_domini_apparuit_holy_family_solesmes_1961}
 
 \begin{otherlanguage}{english} \noindent Ant. The angel of the Lord appeared to Joseph in his sleep, saying: Joseph, son of David, fear not to take unto thee Mary thy wife, for that which is conceived in her, is of the Holy Ghost. \end{otherlanguage}

\psalmus{112}{112_7c}{112_7c}{110_en}

  \gscore[3. Ant.]{7. d}{an_pastores_venerunt_solesmes_1961}
  
  \begin{otherlanguage}{english} \noindent Ant. The shepherds came with haste; and they found Mary and Joseph, and the infant lying in the manger. \end{otherlanguage}

  \psalmus{121}{121_7d}{121_7d}{121_en}

  \gscore[4. Ant.]{2. D}{an_magi_intrantes_solesmes_1961}

\begin{otherlanguage}{english} \noindent Ant. The Magi, entering into the house, found the child with Mary his mother. \end{otherlanguage}

  \psalmus{126}{126_2}{126_2}{126_en}

 \gscore[5. Ant.]{1. f}{an_erat_pater_ejus_solesmes_1961}
 
\begin{otherlanguage}{english} \noindent Ant. His father and mother were wondering at those things which were spoken concerning him. \end{otherlanguage}

\psalmus{147}{147_1f}{147_1f}{147_en}

\rubrique{\begin{otherlanguage}{english}
Having sung the last antiphon in full, stand when the celebrant does.
\end{otherlanguage}}

\smalltitle{Chapter.}
\scripture{Luc. 2, 51.}

\texteparacol{\lettrine{D}{e}scéndit Jesus cum María et Joseph, et venit Názareth,~* et erat súbditus illis.\par
\rr Deo grátias.}{\noindent Jesus went down with Mary and Joseph, and came to Nazareth, and was subject to them. Thanks be to God.}

\smalltitle{Hymn.}

\gscore[]{2.}{hy_o_lux_beata_caelitum_solesmes_1961}

\begin{multicols}{2}
\begin{otherlanguage}{english}
O Blest Light of Saints above,\\
highest hope of mortals,\\
O Jesu, on whose boyhood\\
Home smiled with kindly love.

And thou whose bosom nursed him,\\
O Mary, highly graced,\\
Whose breast gave milk to Jesus,\\
Whose arms thy God embraced.

And thou of all men chosen\\
To guard the Virgin’s fame,\\
To whom God’s Son refused not\\
A Father’s gracious name.

Born for the nation’s healing,\\
Of Jesse’s lineage high,\\
Behold the suppliants kneeling,\\
O hear the sinners’ cry!

The sun returned to evening,\\
Dusks all the twilight air:\\
We, lingering here before you,\\
Pour out our heartfelt prayer.

Thy home was as a garden\\
Made glad with fairest flowers;\\
May life thus blossom sweetly\\
In every home of ours.

Jesus, obedient unto those\\
who were made Thy parents\\
With the Father on high and the Spirit also\\
Always be the glory.
Amen.

\end{otherlanguage}
\end{multicols}

\rubrique{\begin{otherlanguage}{english}
The cantor sings the verse to which all reply.
\end{otherlanguage}}

\smallscore{versiculus_1_vespers_holy_family.gabc}

\begin{otherlanguage}{english} \noindent \vv Blessed are they that dwell in thy house, O Lord.

\noindent \rr They shall praise thee for ever and ever.

 \end{otherlanguage}

\bigrule

\rubrique{In II Vespers.}
\smallscore{versiculus_2_vespers_holy_family.gabc}

\begin{otherlanguage}{english} \noindent \vv All thy children shall be taught of the Lord.

\noindent \rr And great shall be the peace of thy children.
 \end{otherlanguage}
 
 \bigrule
 
%\rubrique{This tone is used on the feast itself, but not during the octave.}
 
\smalltitle{Antiphon at the Magnificat.}

\rubrique{Sit when the celebrant does, after he sings the first words. Rise at the end.}

\gscore[]{8. G}{an_verbum_caro_factum_est_solesmes_1961}
 
 \begin{otherlanguage}{english} \noindent Ant. The Word was made flesh, and dwelt among us, full of grace and truth; and of his fulness we all have received, and grace for grace, alleluia.\end{otherlanguage}
 
 \bigrule

\rubrique{In II Vespers.}
 \gscore[]{8. G}{an_maria_autem_solesmes_1961}
 
 \begin{otherlanguage}{english} \noindent Ant. But Mary kept all these words, pondering them in her heart. \end{otherlanguage}
 
  \bigrule
 
 \smallscore{Magnificat_8G}

   \smallskip%
    \noindent%
    \begin{itemize}[%
    		label=\null, %
			leftmargin=0pt, %
			itemindent=0mm, %
			labelsep=0pt, %
			labelwidth=0pt, %
			rightmargin=0pt, %
			parsep=0pt, %
			topsep=0pt, %
			itemsep=0pt]%
    \input{psaumes/Magnificat_8_Solemnis}
    \end{itemize}

 \rubrique{\begin{otherlanguage}{english}
The Magnificat may be sung entirely in chant or sung alternating between a cantor singing chant and the choir singing a polyphonic or harmonized setting.\end{otherlanguage}}

\smalltitle{Canticle of the Blessed Virgin Mary.}

\scripture{Luc. 1, 46–55.}
\begin{multicols}{2}
%\begin{paracol}{2}
%
%\lettrine{M}agníficat~* ánima mea Dóminum.\\ Et exsultávit spíritus meus~* in Deo salutári meo.
%
% Quia respéxit humilitátem ancíllæ suæ:~* ecce enim ex hoc beátam me dicent o\-mnes generatiónes.
%
% Quia fecit mihi magna qui potens est:~* et san\-ctum nomen ejus.
%
% Et misericórdia ejus a progénie in progénies~* timéntibus eum.
%
% Fecit poténtiam in bráchio suo:~* dispérsit supérbos mente cordis sui.
%
% Depósuit poténtes de sede,~* et exaltávit húmiles.
%
% Esuriéntes implévit bonis:~* et dívites dimísit inánes.
%
% Suscépit Israel púerum suum,~* recordátus misericórdiæ suæ.
%
% Sicut locútus est ad patres no\-stros,~* Abraham, et sémini ejus in sǽcula.
%
% Glória Patri, et Fílio,~* et Spirítui San\-cto.
%
% Sicut erat in princípio, et nunc, et semper,~* et in sǽcula sæculórum. Amen.
 
% \switchcolumn
 \begin{otherlanguage}{english}
My soul \cc~* doth magnify the Lord.

And my spirit hath rejoiced~* in God my Saviour.

Because he hath regarded the humility of his handmaid;~* for behold from henceforth all generations shall call me blessed.

Because he that is mighty, hath done great things to me;~* and holy is his name.

And his mercy is from generation unto generations,~* to them that fear him.

He hath shewed might in his arm:~* he hath scattered the proud in the conceit of their heart.

He hath put down the mighty from their seat,~* and hath exalted the humble.

He hath filled the hungry with good things;~* and the rich he hath sent empty away.

He hath received Israel his servant,~* being mindful of his mercy:

As he spoke to our fathers,~* to Abraham and to his seed for ever.

Glory be to the Father, and to the Son,~* and to the Holy Ghost.

As it was in the beginning, is now,~* and ever shall be, world without end. Amen.
\end{otherlanguage}
% \end{paracol}
\end{multicols}

\rubrique{\begin{otherlanguage}{english}
Sit as the antiphon is repeated. Then the usual dialogue follows. All stand from now until the end of the office.\end{otherlanguage}}

\textebilingue{dv}{dv_eng}

%\textebilingue{oratio_15_aug_old}{oratio_15_aug_old_en}

\texteparacol{Orémus.

\lettrine{D}{o}mine Jesu Christe, qui Maríæ et Joseph súbditus, domésticam vitam ineffabílibus virtútibus consecrásti:~† fac nos, utriúsque auxílio, Famíliæ sanctæ tuæ exémplis ínstrui;~* et consórtium cónsequi sempitérnum. Qui vivis et regnas cum Deo Patre.}{\noindent Let us pray.
Lord Jesus Christ, who subject to Mary and Joseph, didst consecrate family life by thy unspeakable virtues, aid us by their united intercession to profit by the examples of thy Holy Family, and attain to their everlasting companionship. Who livest and reignest with God the Father in the unity of the Holy Spirit, God, world without end.  Amen}

\smalltitle{Commemoration of the Epiphany.}

\rubrique{The octave of the Epiphany is commemorated with an antiphon proper to each day (Jan. 7, etc.)}

\texteparacol{\vv  Reges Tharsis et ínsulæ múnera ófferent.

\rr Reges Arabum et Saba dona addúcent.}{\noindent \vv The kings of Tarshish and of the isles shall bring presents.

\noindent \rr The kings of Arabia and Saba shall offer gifts.}

\texteparacol{Orémus.

\lettrine{D}{e}us, qui hodiérna die Unigénitum tuum géntibus stella duce revelásti:~† concéde propítius; ut qui jam te ex fide cognóvimus,~* usque ad contemplándam spéciem tuæ celsitúdinis perducámur. Per eumdem Dóminum.}{\noindent Let us pray. O God, Who by the leading of a star didst, as on this day, manifest thine Only-begotten Son to the Gentiles, mercifully grant that we, which know thee now by faith, may after this life have the fruition of thy glorious Godhead:}


\smalltitle{Commemoration of the Sunday.}

\gscore[Ant.]{2.}{an_remansit_puer_solesmes_1961}
 \begin{otherlanguage}{english} \noindent Ant. The child Jesus remained in Jerusalem; and his parents knew it not. And thinking that he was in the company, they came a day's journey, and sought him among their kinsfolks and acquaintance. \end{otherlanguage}

 \bigrule

\rubrique{In II Vespers.}

\gscore[Ant.]{8.}{an_fili_quid_fecisti_solesmes_1961}

 \begin{otherlanguage}{english} \noindent Ant. The child Jesus remained in Jerusalem; and his parents knew it not. And thinking that he was in the company, they came a day's journey, and sought him among their kinsfolks and acquaintance. \end{otherlanguage}
 
  \bigrule

\texteparacol{\noindent \vv  Omnes de Saba vénient, allelúia.

\noindent \rr Aurum et thus deferéntes, allelúia.}{\noindent \vv All they from Saba shall come. Alleluia.

\noindent \rr They shall bring gold and incense. Alleluia.}

\texteparacol{Orémus.

\lettrine[findent=0.05em]{V}{o}ta quǽsumus Dómine, supplicántis pópuli cælésti pietáte proséquere:~† ut et quæ agénda sunt, vídeant,~* et ad implénda quæ víderint, convaléscant.
Per Dóminum.}{\noindent Let us pray. Attend, O Lord, we beseech thee, of thy heavenly mercy, to the desires of thy suppliant people; and grant that they may both perceive what they ought to do, and may have strength to fulfill the same. Through Jesus Christ, thy Son our Lord, Who liveth and reigneth with thee, in the unity of the Holy Ghost, God, world without end. Amen.}

\rubrique{\begin{otherlanguage}{english}
The cantor sings the verse to which all reply.
\end{otherlanguage}}

\gscore[]{2.}{BD_Duplex_I_Vesp}

\vv Let us bless the Lord.

\rr Thanks be to God.

 \bigrule
\rubrique{In II Vespers.}
\gscore[]{8.}{BD_Duplex_II_Vesp} 

 \bigrule
 
 \textebilingue{fidelium_animae}{fidelium_animae_en}
%\rubrique{\begin{otherlanguage}{english}
%Then one Our Father is said, unless Compline follows, or even some pious exercises, a sermon, or Benediction of the Blessed Sacrament, in parish and other smaller churches.\end{otherlanguage}}

%\rubrique{\begin{otherlanguage}{english}
%Then the office is concluded with the following verses, the Marian antiphon, etc.
%\end{otherlanguage}}

\end{document}