% !TEX TS-program = lualatexmk
% !TEX parameter =  --shell-escape
\documentclass[12pt]{book}

\input{commonheaders_vespers_people} %%% will need to fix spacing problem in psalms at a later date.

\begin{document}
 
 \twosided[pc] %% to allow for the left-right switching of columns needed for bilingual liturgical typesetting (like in missals); it's unclear why this goes BEFORE everything else and not when activating the environment. %% It's OK for paper booklets to not switch, but it's probably best practice to switch.
 
  \thispagestyle{empty}
  
\feasttitle{MOST HOLY NAME OF JESUS.}

\header{most holy name of jesus}

\doubleclass{II}

\smalltitle{Before the Office.}

\rubrique{These prayers are said kneeling. End when the priest stands.}

\textebilingue{orationes_ante_do}{orationes_ante_do_eng}

\smalltitle{\prayers}

\rubrique{The celebrant intones the following verse. All sing the response and doxology.}

\begin{otherlanguage}{english}
\rubrique{Make the sign of the cross at the beginning and at the Magnificat, and bow the head when the Trinity is mentioned, when mention is made of the ``Holy Name'' in some way, and at ``Orémus.''}\end{otherlanguage}

\gscore[]{℣.}{Deus_In_Adjutorium_Festivus}

\noindent
\begin{otherlanguage}{english}
O God, \cc{} come to my assistance. \rr O Lord, make haste to help me. Glory be to the Father, and to the Son,~* and to the Holy Ghost. As it was in the beginning, is now,~* and ever shall be, world without end. Amen. Alleluia.\end{otherlanguage}

\firstantiphon{8. G}

\initialscore{an_omnis_qui_invocaverit_solesmes_1961}
\antiphona{english}{Whosoever shall call on the name of the Lord shall be saved}
\psalmus{109}{109_8G}{109_8}{109_en}
%\rubrique{Ps. \normaltext{Dixit Dóminus, \pageref{109_8g}.}}

\gscore[2. Ant.]{5. a}{an_sanctum_et_terribile_solesmes_1961}
\antiphona{english}{Holy and terrible is his Name; the fear of the Lord is the beginning of wisdom}
\psalmus{110}{110_5a}{110_5}{110_en}

\gscore[3. Ant.]{3. a2}{an_ego_autem_in_domino_solesmes_1961}
\antiphona{english}{Yet will I rejoice in the Lord I will rejoice in the God of my salvation}

\psalmus{111}{111_3a2}{111_3a2_g}{111_en}

\gscore[4. Ant.]{4. E}{an_a_solis_ortu_solesmes_1961}
\antiphona{english}{From the rising of the sun unto the going down of the same, the Name of the Lord is worthy of praise}
\psalmus{112}{112_4E}{112_4E}{112_en}

\gscore[5. Ant.]{8. c}{an_sacrificabo_hostiam_solesmes_1961}
\antiphona{english}{I will offer the sacrifice of praise, and will call upon the name of the Lord}
\psalmus{115}{115_8c}{115_8}{115_en}

\chapterlateng
\scripture{Phil. 2, 8 – 10.}

%%this text probably needs to be in its own file!

\texteparacol{\lettrine{F}{r}atres: Christus humiliávit semetípsum, factus obédiens usque ad mortem, mortem autem crucis.~† Pro\-pter quod et Deus exaltávit illum, et donávit illi nomen, quod est super omne nomen,~* ut in nómine Jesu omne genu flectátur.}{english}{He humbled himself, becoming obedient unto death, even to the death of the cross. For which cause God also hath exalted him, and hath given him a name which is above all names: That in the name of Jesus every knee should bow, of those that are in heaven, on earth, and under the earth.}

\hymnlateng

%%changed pœnitentibus to pæ-
\gscore[]{1.}{hy_jesu_dulcis_memoria_solesmes_1957}
\textes{jesu_dulcis_memoria_en}

\smallscore{versiculus_holy_name}
%\vv Sit nomen Dómini benedíctum, allelúia.
%
%\rr Ex hoc nunc et usque in sǽculum, allelúia.

   \begin{enpars}\noindent \vv Blessed be the name of the Lord, alleluia.

\noindent \rr  From this time forth, and for evermore, alleluia.\end{enpars}

\myrule

\smalltitle{At I Vespers.}
\gscore[Ad Magnif.]{Ant. 8. G}{an_fecit_mihi_magna_solesmes_1961}
\antiphona{english}{He that is mighty hath done to me great things, and holy is his Name, alleluia}

 \smallscore{Magnificat_8_G_Solemnis}
 
\magnificattexte{Magnificat_8_Solemnis}

\myrule

\gscore[Ad Magnif.]{Ant. 1. g}{an_vocabis_nomen_ejus_solesmes_1961}
\antiphona{english}{Thou shalt call his name Jesus, for he shall save his people from their sins, alleluia}

 \smallscore{Magnificat_1_g_Solemnis}
 
\magnificattexte{Magnificat_1_solemn}

\collectlateng

\texteparacol{\lettrine{D}{e}us, qui unigénitum Fílium tuum constituísti humáni géneris Salvatórem, et Jesum vocári jussísti:~† concéde propítius; ut cujus sanctum nomen venerámur in terris,~* ejus quoque aspéctu perfruámur in cælis. Per eúmdem Dóminum.}{english}{\noindent O God, who hast appointed thine Only-Begotten Son to be the Saviour of mankind, and hast commanded that his name should be called Jesus, mercifully grant that we who here on earth do worship that most holy Name may be made glad in heaven by his presence.
Through the same Jesus Christ.}

\rubrique{At either Vespers, when they occur in the evening of January 4, a commemoration is made of the following vigil of the Epiphany.}

\gscore[Ant]{6.}{an_puer_jesus_solesmes_1961}
\antiphona{english}{The child Jesus advanced in wisdom and age, and grace with God and men}

\texteparacol{\vv Notum fecit Dóminus, allelúia.

\rr Salutáre suum, allelúia.

\lettrine{O}{m}nípotens sempitérne Deus, dírige actus nostros in beneplácito tuo:~† ut in nómine dilécti Fílii tui~* mereámur bonis opéribus abundáre:
Qui tecum vivit et regnat in unitáte.}{english}
{\noindent \vv The Lord hath made known, alleluia.

\noindent \rr His salvation, alleluia.

\noindent Almighty and everlasting God, do Thou order all our actions in conformity with thy good pleasure, that through the name of thy well-beloved Son, we may worthily abound in all good works. Who with thee liveth and reigneth, in the unity.}

\texteparacol{\vv Dóminus vobíscum.

\rr Et cum spíritu tuo.}
{english}
{\noindent \vv The Lord be with you.

\noindent \rr  And with thy spirit.}

\rubrique{\begin{otherlanguage}{english}
The cantor sings the verse to which all reply.
\end{otherlanguage}}

\myrule

\smalltitle{At I Vespers.}

\gscore[]{2.}{BD_Solemnis_I_Vesp}

\myrule

\gscore[]{6.}{BD_Solemnis_mode_6}

\rubrique{Or:}

\gscore[]{5.}{BD_Solemnis_mode_5}

\noindent \vv Let us bless the Lord.

\noindent \rr Thanks be to God.

\textebilingue{fidelium_animae}{fidelium_animae_en}

\end{document}
