% !TEX TS-program = lualatexmk
% !TEX parameter =  --shell-escape
\documentclass[11pt]{book}

\input{commonheaders_vespers_people}

\begin{document}
\thispagestyle{empty}

%%need to fix rule from sticking out

{\centering{\large{II Sunday after the Epiphany.}\par}} %%Sunday

{\centering{January 19, 2025.}\par} %%date Month NN, YYYY

\rubrique{Antiphons, psalms, chapter, and hymn per annum. Collect of the day. Commemorations of Saints Fabian and Sebastian, martyrs.} %% or Suffrage of All Saints

%%changes for Easter time; Advent will probably require a booklet and therefore ONLY commemorations

%\chapterlateng

%%for gesima, Lent, Advent, Easter etc.

\smalltitle{Antiphona ad Magnificat.} %%add mode number and final letter, e.g. 4. A.
\gscore[]{1. f}{an_deficiente_vino_solesmes_1961}

\antiphona{english}{The wine failing, Jesus commanded the water pots to be filled with water, which was changed into wine, alleluia. \end{enpars} %%translation

\gresetinitiallines{0}
\gabcsnippet{(c4)1. Ma(f)gní(gh)fi(h)cat(h.) *(:) á(h)ni(h)ma(h) <i>me</i>(g)<i>a</i>(f) <b>Dó</b>(gh)mi(g)num.(gf..) (::) (Z-)
2. Et(f) ex(gh)sul(h)tá(h)vit(h) <b>spí</b>(ixi)ri(h)tus(h) <b>me</b>(g)us(h.) *(:) in(h) De(h)o(h) sa(h)lu(h)<i>tá</i>(g)<i>ri</i>(f) <b>me</b>(gh)o.(gf..) (::)}
\gresetinitiallines{1}

\begin{enumerate}[label=\arabic*.,%
			leftmargin=0pt, %
			itemindent=0mm, %
			labelsep=3pt, %
			labelwidth=0pt, %
			rightmargin=0pt, %
			parsep=0pt, %
			topsep=0pt, %
			itemsep=0pt,%
			start=3]
%%Magnificat
%% each line requires \item
\item Quia respéxit humilitátem an\textbf{cíl}læ \textbf{su}æ:~* ecce enim ex hoc beátam me dicent omnes gene\textit{rati}\textbf{ó}nes.

\item Quia fecit mihi \textbf{ma}gna qui \textbf{pot}ens est:~* et sanctum \textit{nomen} \textbf{e}jus.

\item Et misericórdia ejus a progénie \textbf{in} pro\textbf{gé}nies~* timén\textit{tibus} \textbf{e}um.

\item Fecit poténtiam in \textbf{brá}chio \textbf{su}o:~* dispérsit supérbos mente \textit{cordis} \textbf{su}i.

\item Depósuit po\textbf{tén}tes de \textbf{se}de,~* et exal\textit{távit} \textbf{hú}miles.

\item Esuriéntes im\textbf{plé}vit \textbf{bo}nis:~* et dívites dimí\kern 0.02em\textit{sit} in\textbf{á}nes.

\item Suscépit Israel \textbf{pú}erum \textbf{su}um,~* recordátus misericór\textit{diæ} \textbf{su}æ.

\item Sicut locútus est ad \textbf{pa}tres \textbf{no}stros,~* Abraham et sémini e\kern 0.02em\textit{jus} \textit{in} \textbf{sǽ}cula.

\item Glória \textbf{Pa}tri, et \textbf{Fí}lio,~* et Spirí\tinyhspace\textit{tui} \textbf{San}cto.

\item Sicut erat in princípio, et \textbf{nunc,} et \textbf{sém}per, ~* et in sǽcula sæcu\tinyhspace\textit{lórum.} \textbf{A}men.
\end{enumerate}

\collectlateng
\smallskip

\texteparacol{Orémus.

  \lettrine{O}{m}nípotens sempitérne Deus, qui cæléstia simul et terréna moderáris:~† supplicatiónes pópuli tui cleménter exáudi;~* et pacem tuam nostris concéde tempóribus.
Per Dóminum.

\rr Amen.}{\noindent Let us pray. \noindent Let us pray.
Almighty and eternal God, as thou govern all things in heaven and on earth; in thy mercy hear the supplication of thy people, and grant thy peace in our times.
Through Jesus Christ, thy Son our Lord, who liveth and reigneth with thee, in the unity of the Holy Ghost, God, world without end.
Amen.}
   
    %%repeat for collects of commemorations
    \gscore[Ant.]{8.}{an_cs_istorum_est_enim_solesmes_1961}
    
    \antiphona{english}{For theirs is the Kingdom of heaven who love not their lives in this world, and have attained unto the reward of the Kingdom, and have washed their robes in the blood of the Lamb}

\texteparacol{
   \vv Lætámini in Dómino et exultáte justi.

\rr Et gloriámini ómnes ré\-cti córde.}{english}{
    \noindent  \vv  Be glad in the Lord, and rejoice, ye just.
         
\noindent  \rr And glory, all ye right of heart.}

\collectlateng
\smallskip

\texteparacol{Orémus.

\lettrine{I}{n}firmitátem nostram réspice, omnípotens Deus:~† et quia pondus própriæ actiónis gravat,~* beatórum Mártyrum tuórum Fabiáni et Sebastiáni intercéssio gloriósa nos prótegat.
Per Dóminum.

\rr Amen.}{\noindent Let us pray.
Be mindful of our weakness, almighty God, and since the burden of our sins weighs heavily upon us, may the glorious intercession of thy holy Martyrs Fabian and Sebastian sustain us.
Through Jesus Christ, thy Son our Lord, who liveth and reigneth with thee, in the unity of the Holy Ghost, God, world without end. Amen.}
%%the last collect ends with the conclusion and Amen.




\end{document}

