% !TEX TS-program = lualatexmk
% !TEX parameter =  --shell-escape

\documentclass[12pt]{book}

\input{commonheaders_vespers_people} %%% will need to fix spacing problem in psalms at a later date.

\begin{document}
 
 \twosided[pc] %% to allow for the left-right switching of columns needed for bilingual liturgical typesetting (like in missals); it's unclear why this goes BEFORE everything else and not when activating the environment. %% It's OK for paper booklets to not switch, but it's probably best practice to switch.

\thispagestyle{empty}

\rubrique{The celebrant intones, and the cantors continue. The people may join the schola on alternating verses, which begin at double bar lines.}

\gscore[]{3.}{hy_te_deum_solemn_solesmes}

\begin{multicols}{2}
\begin{otherlanguage}{english}

We praise thee, O God; we acknowledge thee to be the Lord.\\
All the earth doth worship thee, the Father everlasting.

To thee all angels cry aloud, the heavens and all the powers therein.\\
To thee cherubin and seraphin continually do cry,\\
Holy, Holy, Holy, Lord God of Sabaoth;

Heaven and earth are full of the majesty of thy glory.\\
The glorious company of the apostles praise thee.\\
The goodly fellowship of the prophets praise thee.\\
The noble army of martyrs praise thee.

The holy Church throughout all the world doth acknowledge thee:\\
the Father of an infinite majesty;\\
thine honourable, true and only Son;\\
also the Holy Ghost the Comforter.

Thou art the King of glory, O Christ.\\
Thou art the everlasting Son of the Father.\\
When thou tookest upon thee to deliver man,\\
thou didst not abhor the Virgin’s womb.

When thou hadst overcome the sharpness of death,\\
thou didst open the kingdom of heaven to all believers.

Thou sittest at the right hand of God, in the glory of the Father.\\
We believe that thou shalt come to be our judge.\\
We therefore pray thee, help thy servants,\\
whom thou hast redeemed with thy precious blood.

Make them to be numbered with thy saints in glory everlasting.\\
O Lord, save thy people and bless thine heritage.\\
Govern them and lift them up for ever.

Day by day we magnify thee;\\
and we worship thy name, ever world without end.

Vouchsafe, O Lord, to keep us this day without sin.\\
O Lord, have mercy upon us, have mercy upon us.\\
O Lord, let thy mercy lighten upon us, as our trust is in thee.\\
O Lord, in thee have I trusted; let me never be confounded.

\end{otherlanguage}
\end{multicols}

\texteparacol{\noindent \vv Benedicámus Patrem et Fílium cum San\-cto Spíritu.

\noindent \rr Laudémus et superexaltémus eum in sǽcula.

\noindent \vv Benedí\-ctus es Dómine in firmaménto cæli.

\noindent \rr Et laudábilis, et gloriósus, et superexaltátus in sǽcula.

\noindent \vv Dómine exáudi oratiónem meam.

\noindent \rr Et clamor meus ad te véniat.}{\noindent \vv Let us bless the Father and the Son with the Holy Ghost.

\noindent \rr Let us praise and exalt him above all forever.

\noindent \vv Blessed art thou, o Lord, in the firmament of heaven.

\noindent \rr And worthy of praise, and glorious, and above all forever.

\noindent \vv O Lord hear my prayer.

\noindent \rr And let my cry come unto thee.}

\textebilingue{dv}{dv_eng}

\texteparacol{Orémus. 

\lettrine{D}{e}us, cujus misericórdiæ non est númerus, et bonitátis infinítus est thesáurus:~† piíssimæ majestáti tuæ pro collátis donis grátias ágimus, tuam semper cleméntiam exorántes;~* ut qui peténtibus postuláta concédis, eósdem non déserens, ad prǽmia futúra dispónas. Per Christum Dóminum nostrum. \rr Amen.}{\noindent Let us pray. O God, whose mercies are numberless and whose treasurey
goodness has no end, we give thanks to
Thy most gracious Majesty for the gifts
Thou hast bestowed, beseeching Thy
mercy, that as Thou grantest the petitions
of those who ask, so not forsaking them,
Thou wilt prepare them for rewards to
come. Through Christ our Lord. Amen.}


\end{document}