% !TEX TS-program = lualatexmk
% !TEX parameter =  --shell-escape
\documentclass[12pt]{book}

\input{commonheaders_vespers_people}
\begin{document}
 
 \twosided[pc] %% to allow for the left-right switching of columns needed for bilingual liturgical typesetting (like in missals); it's unclear why this goes BEFORE everything else and not when activating the environment. %% It's OK for paper booklets to not switch, but it's probably best practice to switch.
 
  \thispagestyle{empty}
 
 %%need to put readings of 2nd and 3rd nocturns in paracol

 {\centering{\huge\capspace{{HOLY SATURDAY.}}}\par}\header{tenebræ of holy saturday.}
 
 \halfbskipvspace
 
 \doubleclass{I}
 
\smalltitle{Before the Office.}

\rubrique{These prayers are said kneeling. End when the priest stands.}

\textes{orationes_ante_do_la_en}

\rubrique{These prayers are said standing. End when the cantor intones the first antiphon.}

\textes{pater_ave_credo_la_en}

%%\subsection[Ad Matutinum.]{ad matutinum.}

%%\textes{thursday_rubric}

%% \baselineskipvspace
 
 \matins
 
\nocturn{i}

\firstantiphon{8. G.}

\initialscore{an_in_pace_in_idipsum}
\antiphona{english}{I will both lay me down in peace, and sleep}
\psalmus{4}{4_8G}{4_8G}{4_en_tenebrae}

%%\textes{repeat_rubric}

\gscore[2. Ant.]{4. E}{an_habitabit}
\antiphona{english}{He shall abide in thy tabernacle: he shall dwell in thy holy hill}
\psalmus{14}{14_4E}{14_4E}{14_en_tenebrae}

\gscore[3. Ant.]{7. c}{an_caro_mea_requiescet}
\antiphona{english}{My flesh shall rest in hope}
\psalmus{15}{15_7c}{15_7c}{15_en_tenebrae}

%%\negvariablevspace{0.25}

\smallscore{vr_in_pace_in_idipsum}
\versiculus{english}{I will both lay me down in peace}{And sleep}

%\newpage
%%\textes{absolution_rubric}
\silently

%\baselineskipvspace
%\newpage
\altlarge
\lectionscore[Cap. 3, 22 – 30.]{lect_1_sab_en}
\altitshape

\begin{enpars}
Heth. The mercies of the Lord that we are not consumed: because his commiserations have not failed.

Heth. They are new every morning, great is thy faithfulness.

Heth. The Lord is my portion, said my soul: therefore will I wait for him.

Teth. The Lord is good to them that hope in him, to the soul that seeketh him.

Teth. It is good to wait with silence for the salvation of God.

Teth. It is good for a man, when he hath borne the yoke from his youth.

Jod. He shall sit solitary, and hold his peace: because he hath taken it up upon himself.

Jod. He shall put his mouth in the dust, if so be there may be hope.

Jod. He shall give his cheek to him that striketh him, he shall be filled with reproaches.

Jerusalem, Jerusalem, return to the Lord thy God.
\end{enpars}

\resplateng{I}
\gscore[]{4.}{re_sicut_ovis}
%
\begin{enpars}
He hath been brought as a lamb to the slaughter, and while he was evil entreated he opened not his mouth: he was delivered up to death
* That he might quicken his people.
℣.\@  He hath poured out his soul unto death, and he was numbered with the transgressors.
*  That he.
\end{enpars}

\lectionlateng{II}
\lectionscore[Cap. 4, 1 – 6.]{lect_2_sab}
\begin{enpars}
Aleph. How is the gold become dim, the finest colour is changed, the stones of the sanctuary are scattered in the top of every street?

Beth. The noble sons of Sion, and they that were clothed with the best gold: how are they esteemed as earthen vessels, the work of the potter's hands?

Ghimel. Even the sea monsters have drawn out the breast, they have given suck to their young: the daughter of my people is cruel, like the ostrich in the desert.

Daleth. The tongue of the sucking child hath stuck to the roof of his mouth for thirst: the little ones have asked for bread, and there was none to break it unto them.

He. They that were fed delicately have died in the streets; they that were brought up in scarlet have embraced the dung.

Vau. And the iniquity of the daughter of my people is made greater than the sin of Sodom, which was overthrown in a moment, and hands took nothing in her.

Jerusalem, Jerusalem, return to the Lord thy God.
\end{enpars}

\resplateng{II}
\gscore[]{5.}{re_jerusalem_surge}
\begin{enpars}
Jerusalem, aarise, and put off thy garments of rejoicing: cover thee with sackcloth and ashes
* For in the midst of thee the Saviour of Israel hath been slain.
℣.\@  Let thy tears run down like a river day and night, and let not the apple of thine eye cease.
*  For in the midst of thee.
\end{enpars}

\lectionlateng{III}
\lectionscore[Cap. 5, 1 – 11.]{lect_3_sab}
\begin{enpars}
Remember, O Lord, what is come upon us: consider and behold our reproach. Our inheritance is turned to aliens: our houses to strangers. We are become orphans without a father: our mothers are as widows.
We have drunk our water for money: we have bought our wood.
We were dragged by the necks, we were weary and no rest was given us.
%:
We have given our hand to Egypt, and to the Assyrians, that we might be satisfied with bread.
Our fathers have sinned, and are not: and we have borne their iniquities.
Servants have ruled over us: there was none to redeem us out of their hand.
We fetched our bread at the peril of our lives, because of the sword in the desert.
Our skin was burnt as an oven, by reason of the violence of the famine.
They oppressed the women in Sion, and the virgins in the cities of Juda.
Jerusalem, Jerusalem, return to the Lord thy God.
\end{enpars}

\orrubrique

\lectionscore[]{lect_3_sab_ad_lib}

\resplateng{III}
\gscore[]{5.}{re_plange}

\begin{enpars}
Lament, my people, like a virgin girded with sackcloth for the husband of her youth, howl, ye shepherds, in sackcloth and ashes
* For the day of the Lord is at hand, and it is great and very terrible.
℣.\@  Gird yourselves, ye Priests, and howl, ye ministers of the altar: cast up ashes upon you.
℟.\@  Lament.
\end{enpars}

\nocturn{ii}

\gscore[1. Ant.]{5. a}{an_elevamini}
\antiphona{english}{Be ye lift up, ye everlasting doors, and the King of glory shall come inr}
\psalmus{23}{23_5}{23_5}{23_en_tenebrae}

\gscore[2. Ant.]{4. E}{an_credo_videre_sab_sancto}
\antiphona{english}{I believe that I shall yet see the goodness of the Lord in the land of the living}
\psalmus{26}{26_4E}{26_4E}{26_en_tenebrae}

\gscore[3. Ant.]{8. G}{an_domine_abstraxisti}
\antiphona{english}{O Lord, thou hast brought up my soul from the grave}
\psalmus{29}{29_8G}{29_8G}{29_en_tenebrae}

\smallscore{vr_tu_autem}
\versiculus{english}{ But thou, O Lord, be merciful unto me}{And raise me up; and I will requite them}

\silently

%%not ideal to not rename macro but am pressed for time
\lectionlateng{IV}
\textes{sab_lesson_4_la_en}

\resplateng{IV}
\gscore[]{7.}{re_recessit_pastor}
\begin{enpars}
Our shepherd, even the fountain of living waters, is gone from us; he passed away, and the sun was darkened.
* For now hath our Saviour bound him captive, which bound the first man captive; this day hath He burst the gates and bars of death.
℣.\@ The bands of hell he hath utterly abolished, and hath done away the power of the devil.
* For now hath our Saviour.
\end{enpars}

\lectionlateng{V}
\textes{sab_lesson_5_la_en} 

\resplateng{V}
\gscore[]{8.}{re_o_vos_omnes}
\begin{enpars}
O all ye that pass by, behold and see;
* If there be any sorrow like unto my sorrow.
℣.\@ O all ye nations, behold, and see my sorrow,
* If there be.
\end{enpars}

\lectionlateng{VI}
\textes{sab_lesson_6_la_en}

\resplateng{VI}
\gscore[]{4.}{re_ecce_quomodo_moritur}
\begin{enpars}
Behold how the righteous dieth, and no man taketh it to heart; and the just are taken away, and none considereth. From the midst of sinners is the righteous translated;
* And his memory is in peace.
℣.\@ As a lamb before his shearers is dumb, so he opened not his mouth; he was taken from prison and from judgment.
* And his memory.
℟.\@ Behold how the righteous.
\end{enpars}

\nocturn{iii}

\gscore[1. Ant.]{8. G}{an_deus_adjuvat_me}
\antiphona{english}{God is my helper, and the Lord upholdeth my soul}
\psalmus{53}{53_8G}{53_8G}{53_en_tenebrae} 

\gscore[2. Ant.]{7. a}{an_in_pace_factus_est}
\antiphona{english}{His place is in peace and his abode in Sion}
\psalmus{75}{75_7a}{75_7a}{75_en_tenebrae}

\gscore[3. Ant.]{4. d}{an_factus_sum}
\antiphona{english}{I am as a man that hath no strength, lying nerveless among the dead}
\psalmus{87}{87_4_d}{87_4_d}{87_en_tenebrae}

\smallscore{vr_in_pace_factus_est}
\versiculus{english}{His place is in peace}{And his abode in Sion}

\silently

\lectionlateng{VII}
\textes{sab_lesson_7_la_en}

\resplateng{VII}
\gscore[]{7.}{re_astiterunt_reges}
\begin{enpars}
The kings of the earth set themselves, and the rulers take counsel together
* Against the Lord, and against his Anointed.
℣.\@ Why do the heathen rage? and the people imagine a vain thing?
* Against the Lord.
\end{enpars}

\lectionlateng{VIII}
\textes{sab_lesson_8_la_en}

\resplateng{VIII}
\gscore[]{4.}{re_aestimatus_sum}
\begin{enpars}
I am counted with them that go down into the pit.
* I am as a man that hath no strength, lying nerveless among the dead.
℣.\@ They have laid me in the lowest pit, in darkness, and in the shadow of death.
* I am as a man.
\end{enpars}

\lectionlateng{IX}
\textes{sab_lesson_9_la_en}
%\nopagebreak

\resplateng{IX}
\gscore[]{2.}{re_sepulto_domino}
\begin{enpars}
After that the Lord was buried, they sealed the sepulchre, rolling a stone to the door of the sepulchre
* Setting a watch to keep him.
℣.\@ The chief priests came together unto Pilate, and made that request unto him.
* Setting a watch.
℟.\@ After that the Lord.
\end{enpars}

\baselineskipvspace
 \lauds
 
\gscore[1. Ant.]{4. c}{an_o_mors}
 \antiphona{english}{O death, I will be thy death; O grave, I will be thy destruction}
\psalmus{50}{50_4c}{50_4c}{50_en_tenebrae}

\gscore[2. Ant.]{4. A*}{an_plangent_eum}
\antiphona{english}{They shall mourn for him as one mourneth for his only son, for the innocent Lord hath been put to death}
\psalmus{91}{91_4_Astar}{91_4_Astar}{91_en_tenebrae}

\gscore[3. Ant.]{7. b}{an_attendite_universi}
\antiphona{english}{O all ye nations, behold and see my sorrow}
\psalmus{63}{63_7b}{63_7b}{63_en_tenebrae}

\gscore[4. Ant.]{2. D}{an_a_porta_inferi}
\antiphona{english}{O Lord, deliver my soul from the gates of the grave}
\canticum{Ezechiæ}{Ez. 38, 10 – 20.}{Canticum_Ezechiae_2}{Canticum_Ezechiae_2}{canticum_ezechiae_en_tenebrae}

\gscore[5. Ant.]{8. c}{an_o_vos_omnes}
\antiphona{english}{O all ye that pass by behold, and see if there be any sorrow like unto my sorrow}
\psalmus{150}{150_8c}{150_8c}{150_en_tenebrae}

%\textes{triduum_cap_rubric}

\smallscore{vr_caro_mea}
\versiculus{english}{My flesh shall rest in hope}{Neither wilt thou suffer thine Holy One to see corruption}

\smalltitle{At the Benedictus, Antiphon.}

\gscore[]{1. g}{an_mulieres_sedentes}
\antiphona{english}{There were women sitting over against the sepulchre weeping, and making lamentation for the Lord}

\canticum{Zachariæ}{Luc. 1, 68 – 79.}{Benedictus_1g_solemn}{Benedictus_1g_solemn}{benedictus_tenebrae_en}

%%\phantomsection\label{Christus factus est}
\gscore{5.}{an_christus_factus_est}
%%\textes{rubrique_secunda_nocte}

\rubrique{On the second night the following is added:}
\smallscore{an_mortem_autem}
%%\textes{rubrique_tertia_nocte}

\rubrique{On the third night the following is added:}

\smallscore{an_propter_quod}

%\textes{miserere_rubric}

\silently

\rubrique{Then Ps.\@ \normaltext{Miserére,} as on the previous nights.}

\specialpstexte{50_recto_tono}{50_en_tenebrae}

\textes{respice_triduum_la_en}

%\rubrique{All of the Hours of the Triduum are concluded thusly, until None of Holy Saturday inclusive.}

\rubrique{The office concludes in the same manner as on the previous nights.}

%%\rubrique{After the oration is finished, there is a little din and noise: soon the lit candle is brought out from under the altar, and all rise and depart in silence.}
\end{document}