% !TEX TS-program = lualatexmk
% !TEX parameter =  --shell-escape
\documentclass[12pt]{book}

\input{commonheaders_vespers_people}
\begin{document}
 
 \twosided[pc] %% to allow for the left-right switching of columns needed for bilingual liturgical typesetting (like in missals); it's unclear why this goes BEFORE everything else and not when activating the environment. %% It's OK for paper booklets to not switch, but it's probably best practice to switch.
 
  \thispagestyle{empty}
 
 %%need to put readings of 2nd and 3rd nocturns in paracol

 {\centering{\huge\capspace{{FRIDAY \\[0.25\baselineskip]OF THE PREPARATION.}}}\par}\header{tenebræ of friday of the preparation.}
 
 \halfbskipvspace
 
 \doubleclass{I}
 
\smalltitle{Before the Office.}

\rubrique{These prayers are said kneeling. End when the priest stands.}

\textes{orationes_ante_do_la_en}

\rubrique{These prayers are said standing. End when the cantor intones the first antiphon.}

\textes{pater_ave_credo_la_en}

%\textes{thursday_rubric}

 \baselineskipvspace
 
 \matins
 
\nocturn{i}

\firstantiphon{8. G.}

%\begin{enpars}
%\end{enpars}
\initialscore{an_astiterunt_reges}
\antiphona{english}{The kings of the earth set themselves, and the rulers take counsel together, against the Lord, and against his Anointed}
\psalmus{2}{2_8G}{2_8G}{2_en_tenebrae}

%
\gscore[2. Ant.]{8. G}{an_diviserunt_sibi}
\antiphona{english}{They part my garments among them, and cast lots upon my vesture}
\psalmus{21}{21_8G}{21_8G}{21_en_tenebrae}

\gscore[3. Ant.]{8. G}{an_insurrexerunt_in_me}
\antiphona{english}{False witnesses are risen up against me, and iniquity hath belied itself}
\psalmus{26}{26_8G}{26_8G}{26_en_tenebrae}

%%%\negvariablevspace{0.25}

\smallscore{vr_diviserunt}
\versiculus{english}{They part my garments among them}{And cast lots upon my vesture}

%\newpage
%%\textes{absolution_rubric}
\silently

%\baselineskipvspace
\newpage
\altlarge
\lectionscore[Cap. 2, 8–15.]{lect_1_fer_vi_en}
\altitshape

\begin{enpars}
Heth. The Lord hath purposed to destroy the wall of the daughter of Sion: he hath stretched out his line, and hath not withdrawn his hand from destroying: and the bulwark hath mourned, and the wall hath been destroyed together.

Teth. Her gates are sunk into the ground: he hath destroyed, and broken her bars: her king and her princes are among the Gentiles: the law is no more, and her prophets have found no vision from the Lord.

Jod. The ancients of the daughter of Sion sit upon the ground, they have held their peace: they have sprinkled their heads with dust, they are girded with haircloth, the virgins of Jerusalem hang down their heads to the ground.

Caph. My eyes have failed with weeping, my bowels are troubled: my liver is poured out upon the earth, for the destruction of the daughter of my people, when the children, and the sucklings, fainted away in the streets of the city.

Jerusalem, Jerusalem, return to the Lord thy God.
\end{enpars}

\resplateng{I}
\gscore[]{3.}{re_omnes_amici_mei}

\begin{enpars}
All my friends have forsaken me, and mine enemies have prevailed against me; he whom I loved hath betrayed me.
* Mine enemy sharpeneth his eyes upon me; he breaketh me with breach upon breach: and in my thirst they gave me vinegar to drink.
℣.\@  I am numbered with the transgressors; and my life is not spared.
* Mine enemy.
\end{enpars}

\lectionlateng{II}
\lectionscore[]{lect_2_fer_vi}
\begin{enpars}
Lamed. They said to their mothers: Where is corn and wine? when they fainted away as the wounded in the streets of the city: when they breathed out their souls in the bosoms of their mothers.

Mem. To what shall I compare thee? or to what shall I liken thee, O daughter of Jerusalem? to what shall I equal thee, that I may comfort thee, O virgin daughter of Sion? for great as the sea is thy destruction: who shall heal thee?

Nun. thy prophets have seen false and foolish things for thee: and they have not laid open thy iniquity, to excite thee to penance: but they have seen for thee false revelations and banishments.

Samech. All they that passed by the way have clapped their hands at thee: they have hissed, and wagged their heads at the daughter of Jerusalem, saying: Is this the city of perfect beauty, the joy of all the earth?

Jerusalem, Jerusalem, return to the Lord thy God.
\end{enpars}

\resplateng{II}
\gscore[]{2.}{re_velum_templi}
\begin{enpars}
The veil of the Temple was rent in twain, from the top to the bottom,
* And all the earth did quake: the thief on the cross cried, saying: Lord, remember me when thou comest into thy kingdom!
℣.\@  The rocks rent, and the graves were opened, and many bodies of the saints, which slept, arose.
℟.\@  And all the earth.
\end{enpars}

\lectionlateng{III}
\lectionscore[Cap. 3, 1 – 9.]{lect_3_fer_vi}
\begin{enpars}
Aleph. I am the man that see my poverty by the rod of his indignation.

Aleph. He hath led me, and brought me into darkness, and not into light.

Aleph. Only against me he hath turned, and turned again his hand all the day.

Beth. My skin and my flesh he hath made old, he hath broken my bones.

Beth. He hath built round about me, and he hath compassed me with gall and labour.

Beth. He hath set me in dark places as those that are dead for ever.

Ghimel. He hath built against me round about, that I may not get out: he hath made my fetters heavy.

Ghimel. Yea, and when I cry, and entreat, he hath shut out my prayer.

Ghimel. He hath shut up my ways with square stones, he hath turned my paths upside down.

Jerusalem, Jerusalem, return to the Lord thy God.
\end{enpars}

\resplateng{III}
\gscore[]{8.}{re_vinea_mea}
%% p and Q touch notes below

\begin{enpars}
I had planted thee a noble vineyard;
* How then art thou turned into a degenerate plant, which willest that Barabbas should be released unto thee, and that I should be crucified.
℣.\@  I fenced thee, and gathered out the stones from thee, and built a tower in the midst of the land.
* How then.
℟.\@  I had planted.
\end{enpars}

\nocturn{ii}

\gscore[1. Ant.]{8. G}{an_vim_faciebant}
\antiphona{english}{They that sought after my life have used violence again me}
\psalmus{37}{37_8G}{37_8G}{37_en_tenebrae}

\gscore[2. Ant.]{4. A*}{an_confundantur_et_revereantur}
\antiphona{english}{Let them be ashamed and confounded together that seek after my soul, to destroy it}
\psalmus{39}{39_4_Astar}{39_4_Astar}{39_en_tenebrae}

\gscore[3. Ant.]{4. A*}{an_alieni_insurrexerunt}
\antiphona{english}{Strangers are risen up against me, and oppressors seek after my soul}
\psalmus{53}{53_4_Astar}{53_4_Astar}{53_en_tenebrae}

\smallscore{vr_insurrexerunt}
\versiculus{english}{False witnesses are risen up against me}{And iniquity hath belied itself}

\silently

%%not ideal to not rename macro but am pressed for time
\lectionlateng{IV}
\textes{fer_vi_para_lesson_4_la_en}

\resplateng{IV}
\gscore[]{8.}{re_tamquam_ad_latronem}
\begin{enpars}
Are ye come out, as against a thief, with swords and staves, for to take me?
* I sat daily with you, teaching in the Temple, and ye laid no hold on me; and, now when ye have scourged me, ye lead me away to crucify me!
℣.\@  And when they had laid hands on Jesus, and taken him, he said unto them:
* I sat daily.
\end{enpars}

\lectionlateng{V}
\textes{fer_vi_para_lesson_5_la_en} 

\resplateng{V}
\gscore[]{7.}{re_tenebrae}
\begin{enpars}
The Jews crucified Jesus: and there was darkness (over all the land, unto the ninth hour): and about the ninth hour Jesus cried with a loud voice, (saying): My God, why hast thou forsaken me?
* And he bowed his head, and gave up the ghost.
℣.\@  When Jesus had cried with a loud voice, He said: Father, into thy hands I commend my spirit.
* And he bowed.
\end{enpars}

\lectionlateng{VI}
\textes{fer_vi_para_lesson_6_la_en}

\resplateng{VI}
\gscore[]{8.}{re_animam_meam}
\begin{enpars}
I have given the dearly-beloved of my soul into the hand of her enemies and mine heritage is become unto me as a lion in the forest; the enemy crieth out against me, saying: Assemble yourselves together, hasten to devour him: they have made my portion a desolate wilderness, and the whole land mourneth unto me:
* Because there is none found that will know me, nor do well.
℣.\@  There be risen up against me such as breathe out cruelty, and they have not spared my soul.
* Because.
℟.\@  I have given.
\end{enpars}

\nocturn{iii}

\gscore[1. Ant.]{1. f}{an_ab_insurgentibus}
\antiphona{english}{O Lord, defend me from them that rise up against me, for they lie in wait for my life}
\psalmus{58}{58_1f}{58_1f}{58_en_tenebrae}

\gscore[2. Ant.]{8. G}{an_longe_fecisti}
\antiphona{english}{Thou hast put away mine acquaintance far from me; I am shut up, and cannot come forth}
\psalmus{87}{87_8G}{87_8G}{87_en_tenebrae}

\gscore[3. Ant.]{8. G}{an_captabunt}
\antiphona{english}{They gather themselves together against the soul of the righteous, and condemn the innocent blood}
\psalmus{93}{93_8G}{93_8G}{93_en_tenebrae}

\smallscore{vr_locuti_sunt}
\versiculus{english}{They have spoken against me with a lying tongue}{They compassed me about also with words of hatred, and fought against me without a cause}

\silently

\lectionlateng{VII}
\textes{fer_vi_para_lesson_7_la_en}

\resplateng{VII}
\gscore[]{6.}{re_tradiderunt}
\begin{enpars}
They have turned me over into the hands of the wicked: they also have numbered me with the transgressors, neither have they spared my life: the mighty are gathered together against me;
* And stand up against me like giants.
℣.\@  Strangers are risen up against me, and oppressors seek after my soul.
* And stand up.
\end{enpars}

\lectionlateng{VIII}
\textes{fer_vi_para_lesson_8_la_en}

\resplateng{VIII}
\gscore[]{7.}{re_jesum_tradidit}
\begin{enpars}
That wicked one betrayed Jesus to the chief-priests and elders of the people
* But Peter followed him afar off, to see the end.
℣.\@  And they led him away to Caiaphas the high priest, where the scribes and Pharisees were assembled.
* But Peter.
\end{enpars}

\lectionlateng{IX}
\textes{fer_vi_para_lesson_9_la_en}
%\nopagebreak

\resplateng{IX}
\gscore[]{5.}{re_caligaverunt}
\begin{enpars}
Mine eyes do fail with tears, because the Comforter that should relieve me is far from me. Behold, O all ye nations
* If there be any sorrow like unto my sorrow.
℣.\@  O all ye that pass by, behold, and see
* If there be.
℟.\@ Mine eyes.
\end{enpars}

% \baselineskipvspace
\newpage

 \lauds
 
\gscore[1. Ant.]{7. c}{an_proprio_filio_suo}
 \antiphona{english}{God spared not his own Son but delivered Him up for us all}
\psalmus{50}{50_7c}{50_7c}{50_en_tenebrae}

\gscore[2. Ant.]{4. E}{an_anxiatus_est}
\antiphona{english}{My spirit is overwhelmed within me; my heart within me is troubled}
\psalmus{142}{142_4E}{142_4E}{142_en_tenebrae}

\gscore[3. Ant.]{1. f}{an_ait_latro}
\antiphona{english}{One thief said unto the other: We indeed receive the due reward of our deeds, but what hath this man done? Lord, remember me, when thou comest into thy kingdom}
\psalmus{84}{84_1f}{84_1f}{84_en_tenebrae}

\gscore[4. Ant.]{1. f}{an_dum_conturbata}
\antiphona{english}{Lord, when my soul is troubled, thou wilt remember mercy}
\canticum{Habacuc}{Cap. 3, 2–19.}{Canticum_Habacuc_1f}{Canticum_Habacuc_1f}
{canticum_habucuc_tenebrae}

\gscore[5. Ant.]{8. G}{an_memento_mei}
\antiphona{english}{Lord, remember me when thou comest into thy kingdom}
\psalmus{147}{147_8G}{147_8_tenebrae}{147_en_tenebrae}

%%\textes{triduum_cap_rubric}

\smallscore{vr_collocavit}
\versiculus{english}{He hath set me in dark places}{As they that be dead of old}

\smalltitle{At the Benedictus, Antiphon.}

\gscore[]{1. g}{an_posuerunt_super_caput}
\antiphona{english}{They set up over his head his accusation written: Jesus of Nazareth, King of the Jews}

\canticum{Zachariæ}{Luc. 1, 68–79.}{Benedictus_1g_solemn}{Benedictus_1g_solemn}{benedictus_tenebrae_en}

%%\phantomsection\label{Christus factus est}
\gscore{5.}{an_christus_factus_est}
\begin{enpars}
Christ became obedient for us unto death.
\end{enpars}
%%\textes{rubrique_secunda_nocte}

\rubrique{On the second night the following is added:}

\smallscore{an_mortem_autem}
\begin{enpars}
Even to the death of the cross.
\end{enpars}
%%\textes{rubrique_tertia_nocte}
%%\smallscore{an_propter_quod}

%\textes{miserere_rubric}

\silently

\specialpstexte{50_recto_tono}{50_en_tenebrae}

\textes{respice_triduum_la_en}

%\rubrique{All of the Hours of the Triduum are concluded thusly, until None of Holy Saturday inclusive.}

\rubrique{The office concludes in the same manner as on the previous night.}

%\rubrique{After the oration is finished, there is a little din and noise: soon the lit candle is brought out from under the altar, and all rise and depart in silence.}
\end{document}