% !TEX TS-program = lualatexmk
% !TEX parameter =  --shell-escape
\documentclass[12pt]{book}

\input{commonheaders_vespers_people}
\begin{document}
 
 \twosided[pc] %% to allow for the left-right switching of columns needed for bilingual liturgical typesetting (like in missals); it's unclear why this goes BEFORE everything else and not when activating the environment. %% It's OK for paper booklets to not switch, but it's probably best practice to switch.
 
  \thispagestyle{empty}
 
 %%need to put readings of 2nd and 3rd nocturns in paracol

 {\centering{\huge\capspace{{THURSDAY \\[0.25\baselineskip]OF THE LORD'S SUPPER.}}}\par}\header{tenebræ of thursday of the lord's supper.}
 
 \halfbskipvspace
 
 \doubleclass{I}
 
\smalltitle{Before the Office.}

\rubrique{These prayers are said kneeling. End when the priest stands.}

\textes{orationes_ante_do_la_en}

\rubrique{These prayers are said standing. End when the cantor intones the first antiphon.}

\textes{pater_ave_credo_la_en}

%\subsection[Ad Matutinum.]{ad matutinum.}

\textes{thursday_rubric}

 \baselineskipvspace
 
 \matins
 
\nocturn{i}

\firstantiphon{8. c.}

\initialscore{an_zelus_domus_tuae}
\antiphona{english}{The zeal of thine house hath eaten me up, and the reproaches of them that reproached thee are fallen upon me}
\psalmus{68}{68_8c}{68_8c}{68_en_tenebrae}

\textes{repeat_rubric}

\gscore[2. Ant.]{8. c}{an_avertantur_retrorsum}
\antiphona{english}{Let them be turned backward and put to confusion that desire my hurt}
\psalmus{69}{69_8c}{69_8c}{69_en_tenebrae}


\gscore[3. Ant.]{8. c}{an_deus_meus_eripe_me}
\antiphona{english}{Deliver me, my God, out of the hand of the wicked}
\psalmus{70}{70_8c}{70_8c}{70_en_tenebrae}

\negvariablevspace{0.25}

\smallscore{vr_avertantur_retrorsum}
\versiculus{english}{ Let them be turned backward and put to confusion}{That desire my hurt}

%\newpage
\textes{absolution_rubric}

%\baselineskipvspace
\newpage
\altlarge
\lectionscore[Cap. 1, 1–14.]{lect_1_fer_v_en}
\altitshape

\begin{enpars}
Aleph. How doth the city sit solitary that was full of people! how is the mistress of the Gentiles become as a widow: the princes of provinces made tributary!

Beth. Weeping she hath wept in the night, and her tears are on her cheeks: there is none to comfort her among all them that were dear to her: all her friends have despised her, and are become her enemies.

Ghimel. Juda hath removed her dwelling place because of her affliction, and the greatness of her bondage: she hath dwelt among the nations, and she hath found no rest: all her persecutors have taken her in the midst of straits.

Daleth. The ways of Sion mourn, because there are none that come to the solemn feast: all her gates are broken down: her priests sigh: her virgins are in affliction, and she is oppressed with bitterness.

He. Her adversaries are become her lords, her enemies are enriched: because the Lord hath spoken against her for the multitude of her iniquities: her children are led into captivity: before the face of the oppressor.

Jerusalem, Jerusalem, return to the Lord thy God.
\end{enpars}

\resplateng{I}
\gscore[]{8.}{re_in_monte_oliveti}

\begin{enpars}
At the Mount of Olives, he prayed unto the Father: o my Father, if it be possible, let this cup pass from me. * The spirit indeed is willing, but the flesh is weak.
℣.\@ Watch and pray, that ye enter not into temptation.
* The spirit indeed.
\end{enpars}

\lectionlateng{II}
\lectionscore[]{lect_2_fer_v}
\begin{enpars}
Vau. And from the daughter of Sion all her beauty is departed: her princes are become like rams that find no pastures: and they are gone away without strength before the face of the pursuer.

Zain. Jerusalem hath remembered the days of her affliction, and prevarication of all her desirable things which she had from the days of old, when her people fell in the enemy's hand, and there was no helper: the enemies have seen her, and have mocked at her sabbaths.

Heth. Jerusalem hath grievously sinned, therefore is she become unstable: all that honoured her have despised her, because they have seen her shame: but she sighed and turned backward.

Teth. Her filthiness is on her feet, and she hath not remembered her end: she is wonderfully cast down, not having a comforter: behold, O Lord, my affliction, because the enemy is lifted up.

Jerusalem, Jerusalem, return to the Lord thy God.
\end{enpars}

\resplateng{II}
\gscore[]{8.}{re_tristis_est}
\begin{enpars}
my soul is exceeding sorrowful, even unto death: tarry ye here and watch with me yet a little while, and ye shall see the multitude close me in. * Ye shall flee; and I will go to be offered a sacrifice for you.
℣.\@ Behold, the hour is at hand, and the Son of man is betrayed into the hands of sinners.
℟.\@ Ye shall flee.
\end{enpars}

\lectionlateng{III}
\lectionscore[]{lect_3_fer_v}
\begin{enpars}
Jod. The enemy hath put out his hand to all her desirable things: for she hath seen the Gentiles enter into her sanctuary, of whom thou gavest commandment that they should not enter into thy church.

Caph. All her people sigh, they seek bread: they have given all their precious things for food to relieve the soul: see, O Lord, and consider, for I am become vile.

Lamed. O all ye that pass by the way, attend, and see if there be any sorrow like to my sorrow: for he hath made a vintage of me, as the Lord spoke in the day of his fierce anger.

mem. From above he hath sent fire into my bones, and hath chastised me: he hath spread a net for my feet, he hath turned me back: he hath made me desolate, wasted with sorrow all the day long.

Nun. The yoke of my iniquities hath watched: they are folded together in his hand, and put upon my neck: my strength is weakened: the Lord hath delivered me into a hand out of which I am not able to rise.

Jerusalem, Jerusalem, return to the Lord thy God.
\end{enpars}

\resplateng{III}
\gscore[]{5.}{re_ecce_vidimus}
\begin{enpars}
Behold, when we shall see him, he hath no form nor comeliness: there is no beauty in him: this is he which hath borne our griefs and carried our sorrows; but he was wounded for our transgressions.
* And with his stripes we are healed.
℣.\@ Surely he hath borne our griefs and carried our sorrows.
* And with his stripes. ℟.\@ Behold.
\end{enpars}

\nocturn{ii}

\gscore[1. Ant.]{7. c}{an_liberavit_dominus}
\antiphona{english}{The Lord shall deliver the needy from the strong: the poor also, that hath no helper}
\psalmus{71}{71_7c}{71_7c}{71_en_tenebrae}

\gscore[2. Ant.]{8. c}{an_cogitaverunt_impii}
\antiphona{english}{The ungodly think and speak wickedness: they speak loftily concerning oppression}
\psalmus{72}{72_8c}{72_8c}{72_en_tenebrae}

\gscore[3. Ant.]{1. g}{an_exsurge_domine}
\antiphona{english}{Arise, o Lord, and judge my cause}
\psalmus{73}{73_1g}{73_1g}{73_en_tenebrae}

\smallscore{vr_deus_meus}
\versiculus{english}{Deliver me, o my God, out of the hand of the wicked}{Out of the hand of the unrighteous and cruel man}

\silently

%%not ideal to not rename macro but am pressed for time
\lectionlateng{IV}
\textes{fer_v_cd_lesson_4_la_en}

\rubrique{The responsories of the second and third nocturn are settings by the great Spanish priest and composer, \normaltext{Tomás Luis de Victoria, c. 1548 – 1611.}}

\resplateng{IV}
%\gscore[]{8.}{re_amicus_meus}
\texteparacol{\lettrine{A}{m}micus meus ósculi me trádidit signo: Quem osculátus fúero, i\-pse est, tenéte eum: hoc malum fecit signum, qui per ósculum adimplévit homicídium.
* Infélix prætermísit prétium sánguinis, et in fine láqueo se suspéndit.
℣.\@ Bonum erat ei, si natus non fuísset homo ille.
* Infélix.}{english}{\noindent
Mine own friend hath betrayed me by the sign of a kiss: Whomsoever I shall kiss, That same is he; hold him fast. This was the traitorous sign which he gave, even he who murdered with a kiss.
* Woe unto that man! He cast down the price of blood, and went, and hanged himself.
℣.\@ It had been good for that man if he had not been born.
* Woe.}


\lectionlateng{V}
\textes{fer_v_cd_lesson_5_la_en} 
%% watch for page division here

\resplateng{V}
%\gscore[]{2.}{re_judas_mercator}
\texteparacol{\lettrine{J}{u}das mercátor péssimus ósculo pétiit Dóminum: ille ut agnus ínnocens non negávit Judæ ósculum:
* Denariórum número Christum Judǽis trádidit.
℣.\@ Mélius illi erat, si natus non fuísset.
* Denariórum.}{english}
{\noindent The vile trader Judas came to the Lord to kiss him, and he, as a guileless lamb, refused not a kiss to Judas,
* Who, for a certain number of pence, betrayed Christ to the Jews.
℣.\@ It had been good for that man if he had not been born.
* Who.}

\lectionlateng{VI}
\textes{fer_v_cd_lesson_6_la_en}

\resplateng{VI}
%\gscore[]{5.}{re_unus_ex_discipulis}
\texteparacol{\lettrine{U}{n}us ex discípulis meis tradet me hódie: Væ illi per quem tradar ego:
* Mélius illi erat, si natus non fuísset.
℣.\@ Qui intíngit mecum manum in parópside, hic me traditúrus est in manus peccatórum.
* Mélius.
℟.\@ Unus ex.}{english}
{\noindent One of my disciples shall betray me this night. Woe unto that man by whom I am betrayed!
* It had been good for that man if he had not been born.
℣.\@ He that dippeth his hand with me in the dish, the same shall betray me into the hands of sinners.
*  It had been good. ℟.\@  One.}

\nocturn{iii}

\gscore[1. Ant.]{7. c}{an_dixi_iniquis}
\antiphona{english}{I said unto the wicked: Speak not wickedness against God}
\psalmus{74}{74_7c}{74_7c}{74_en_tenebrae}

\gscore[2. Ant.]{8. c}{an_terra_tremuit}
\antiphona{english}{The earth trembled and was still, when God arose to judgment}
\psalmus{75}{75_8c}{75_8c}{75_en_tenebrae}

\gscore[3. Ant.]{7. a}{an_in_die_tribulationis}
\antiphona{english}{In the day of my trouble I sought God with my hands}
\psalmus{76}{76_7a}{76_7a}{76_en_tenebrae}

\smallscore{vr_exsurge_domine}
\versiculus{english}{Arise, O Lord}{And judge thou my cause}

\silently

\lectionlateng{VII}
\textes{fer_v_cd_lesson_7_la_en}

\resplateng{VII}
%\gscore[]{7.}{re_eram_quasi_agnus}
\texteparacol{\lettrine{E}{r}am quasi agnus ínnocens: ductus sum ad immolándum, et nesciébam: consílium fecérunt inimíci mei advérsum me, dicéntes:
* Veníte, mittámus lignum in panem ejus, et eradámus eum de terra vivéntium.
℣.\@ Omnes inimíci mei advérsum me cogitábant mala mihi: verbum iníquum mandavérunt advérsum me, dicéntes.
* Veníte.}{english}{\noindent
I was like a gentle lamb that is brought to the slaughter, and I knew not that mine enemies had devised devices against me, saying:
* Come, let us put (poison of a deadly) tree into his bread, and let us cut him off from the land of the living.
℣.\@ All they that hate me devised my hurt against me: they plotted together to do me evil, saying: * Come.}

\lectionlateng{VIII}
\textes{fer_v_cd_lesson_8_la_en}

\resplateng{VIII}
%\gscore[]{7.}{re_una_hora}
\texteparacol{\lettrine{U}{n}a hora non potuístis vigiláre mecum, qui exhortabámini mori pro me?
* Vel Judam non vidétis, quómodo non dormit, sed festínat trádere me Judǽis.
℣.\@ Quid dormítis? súrgite, et oráte, ne intrétis in tentatiónem.
* Vel.}{english}{\noindent
 Could ye not watch with me one hour, ye that called one on the other to die for me?
* Or see ye not Judas, how that he sleepeth not, but maketh haste to betray me to the Jews?
℣.\@ Why sleep ye? Rise, and pray, lest ye enter into temptation.
* Or see ye not Judas.}

\lectionlateng{IX}
\textes{fer_v_cd_lesson_9_la_en}
%\nopagebreak

\resplateng{IX}
%\gscore[]{1.}{re_seniores_populi}
\texteparacol{\lettrine{S}{e}nióres pópuli consílium fecérunt,
* Ut Jesum dolo tenérent, et occíderent: cum gládiis et fústibus exiérunt tamquam ad latrónem.
℣.\@ Collegérunt pontífices et pharisǽi concílium.
* Ut Jesum.
℟.\@ Senióres}{english}{\noindent
The elders of the people consulted
* That they might take Jesus by subtilty, and kill Him they came out, as against a thief, with swords and staves.
℣.\@ The chief Priests and the Pharisees gathered a council. * That they might take Jesus. ℟.\@ The elders.}

% \baselineskipvspace
 \lauds
 
 \gscore[1. Ant.]{8. G}{an_justificeris_domine}
 \antiphona{english}{O Lord, thou shalt be justified when thou speakest, and be clear when thou art judged}
\psalmus{50}{50_8G}{50_8G}{50_en_tenebrae}

\gscore[2. Ant.]{2. D}{an_dominus_tamquam_ovis}
\antiphona{english}{The Lord was brought as a lamb to the slaughter, and he opened not his mouth}
\psalmus{89}{89_2}{89_2}{89_en_tenebrae}

\gscore[3. Ant.]{8. G}{an_contritum_est_cor_meum}
\antiphona{english}{Mine heart is broken, within me all my bones tremble}
\psalmus{35}{35_8G}{35_8G}{35_en_tenebrae}

\gscore[4. Ant.]{4. A*}{an_exhortatus_es}
\antiphona{english}{O Lord, thou hast spoken unto us in thy strength, and in thy holy banquet}
\canticum{Moysi}{Exodi 15, 1–19.}{Canticum_Moysis_Exod_4_alt_Astar}{Canticum_Moysis_Exod_4_A}{canticle_moses_en_tenebrae}

\gscore[5. Ant.]{2. D}{an_oblatus_est}
\antiphona{english}{He was offered up because he willed it and he bore our sins}
\psalmus{146}{146_2}{146_2}{146_en_tenebrae}

\textes{triduum_cap_rubric}

\smallscore{vr_homo_pacis}
\versiculus{english}{Mine own familiar friend, in whom I trusted}{Which did eat of my bread, hath lifted up his heel against me}

\smalltitle{At the Benedictus, Antiphon.}

\gscore[]{1. g}{an_traditor_autem}
\antiphona{english}{Now he that betrayed him gave them a sign, saying: Whomsoever I shall kiss, that same is he: hold him fast}

\smallscore{Benedictus_1g_solemn}
\canticum{Zachariæ}{Luc. 1, 68–79}{Benedictus_1g_solemn}{Benedictus_1g_solemn}{benedictus_tenebrae_en}

%%\phantomsection\label{Christus factus est}
\gscore{5.}{an_christus_factus_est}
%%\textes{rubrique_secunda_nocte}
%%\smallscore{an_mortem_autem}
%%\textes{rubrique_tertia_nocte}
%%\smallscore{an_propter_quod}

\textes{miserere_rubric}

\specialpstexte{50_recto_tono}{50_en_tenebrae}

\textes{respice_triduum_la_en}

\rubrique{All of the Hours of the Triduum are concluded thusly, until None of Holy Saturday inclusive.}

\rubrique{At Lauds, however, after the oration is finished, there is a little din and noise: soon the lit candle is brought out from under the altar, and all rise and depart in silence.}
\end{document}