% !TEX TS-program = lualatexmk
% !TEX parameter =  --shell-escape

\documentclass[12pt]{book}

\input{commonheaders_vespers_people} %%% will need to fix spacing problem in psalms at a later date.

\begin{document}
 
 \twosided[pc] %% to allow for the left-right switching of columns needed for bilingual liturgical typesetting (like in missals); it's unclear why this goes BEFORE everything else and not when activating the environment. %% It's OK for paper booklets to not switch, but it's probably best practice to switch.

\thispagestyle{empty}

\bigtitle{Invocation of the Holy Spirit on January 1 and on other occasions.\footnote{26. § 1. Plenaria indulgentia conceditur christifideli qui, in ecclesia vel oratorio, devote interfuerit sollemni cantui vel recitationi: 1° 44 hymni Veni, Creator, vel prima anni die ad divinam opem pro totius anni decursu implorandam; vel in sollemnitate Pentecostes.}}

\rubrique{Kneel for the first verse, from \normaltext{Veni Creátor} to \normaltext{pectóra} inclusive.}

\rubrique{The celebrant intones, and the cantors continue. The people may join the schola on alternating verses, which begin at double bar lines.}

\gscore[]{8.}{hy_vcs_dox_votiva}

\begin{multicols}{2}
\raggedcolumns
\begin{enpars}
Come, creator Spirit,\\
visit the souls of thine own,\\
fill with heavenly grace\\
the breasts that thou hast created.

Thou art called Paraclete,\\
gift of the most high God,\\
living water, flame, charity, and spiritual anointing.

Thou, who art sevenfold in thy gifts,\\
finger of God's right hand,\\
thou who wast rightly promised by the Father,\\
enrich our throats with speech.

Inflame the light of our senses,\\
pour love into our hearts;\\
the weakness of our bodies\\
strengthen with lasting power.

Drive the enemy far back,\\
and at once grant us peace:\\
with thou going ahead of us,\\
may we avoid all harm.

Through thee may we know the Father,\\
and recognize the Son;\\
and may we always believe in thee, Spirit of both.

Let there be praise to the Father with the Son,\\ and likewise to the Holy Paraclete;\\
and may the Son send to us\\
the gift of the Holy Spirit. Amen.
\end{enpars}
\end{multicols}

\texteparacol{\vv Emítte Spíritum tuum, et creabúntur.

 \rr Et renovábis fáciem terræ.

 \vv Dóminus vobíscum.

\rr Et cum spíritu tuo.

Orémus. 

\lettrine{D}{e}us, qui corda fidélium Sancti Spíritus illustratióne docuísti:~† da nobis in eódem Spíritu re\-cta sápere,~* et de ejus semper consolatióne gaudére. Per Christum Dóminum nostrum.

\rr Amen.}{english}{\noindent \vv Send forth thy Spirit, and they shall be created.

\noindent \rr And thou shalt renew the face of the earth.

\noindent \rr The Lord be with you.

\noindent \rr And with thy spirit.

\noindent Let us pray. O God, who didst teach the hearts of thy faithful people, by the sending to them the light of thy Holy Spirit, grant us by the same Spirit to have a right judgment in all things, and evermore to rejoice in his holy comfort. Through Christ our Lord. Amen.}


\end{document}
