% !TEX TS-program = lualatexmk
% !TEX parameter =  --shell-escape

\documentclass[12pt]{book}

\input{commonheaders_vespers_people} %%% will need to fix spacing problem in psalms at a later date.

\begin{document}
 
 \twosided[pc] %% to allow for the left-right switching of columns needed for bilingual liturgical typesetting (like in missals); it's unclear why this goes BEFORE everything else and not when activating the environment. %% It's OK for paper booklets to not switch, but it's probably best practice to switch.

\thispagestyle{empty}

\rubrique{Kneel for the first verse, from \normaltext{Veni Creátor} to \normaltext{pectóra} inclusive.}

\rubrique{The celebrant intones, and the cantors continue. The people may join the schola on alternating verses, which begin at double bar lines.}

\rubrique{Vowels in italics are not pronounced; the consonant is pronounced with the following vowel as in v. 4, or the usual melody is sung, which is the case in v. 6.}

\gscore[]{8.}{hy_veni_creator_spiritus_solesmes}

\begin{multicols}{2}
\raggedcolumns
\begin{otherlanguage}{english}
Come, creator Spirit,\\
visit the souls of thine own,\\
fill with heavenly grace\\
the breasts that thou hast created.

Thou art called Paraclete,\\
gift of the most high God,\\
living water, flame, charity, and spiritual anointing.

Thou, who art sevenfold in thy gifts,\\
finger of God's right hand,\\
thou who wast rightly promised by the Father,\\
enrich our throats with speech.

Inflame the light of our senses,\\
pour love into our hearts;\\
the weakness of our bodies\\
strengthen with lasting power.

Drive the enemy far back,\\
and at once grant us peace:\\
with thou going ahead of us,\\
may we avoid all harm.

Through thee may we know the Father,\\
and recognize the Son;\\
and may we always believe in thee, Spirit of both.

To God the Father be glory,\\
and to the Son, who from death rose again,\\
and to the Holy Spirit, forever and ever. Amen.

\end{otherlanguage}
\end{multicols}

\texteparacol{\vv Emítte Spíritum tuum, et creabúntur.

\noindent \rr Et renovábis fáciem terræ.

\noindent \vv Dóminus vobíscum.

\noindent \rr Et cum spíritu tuo.}{\noindent \vv Send forth thy Spirit, and they shall be created.

\noindent \rr And thou shalt renew the face of the earth.

\noindent \rr The Lord be with you.

\noindent \rr And with thy spirit.}

\textebilingue{dv}{dv_eng}

\texteparacol{Orémus. 

\lettrine{D}{e}us, qui corda fidélium Sancti Spíritus illustratióne docuísti:~† da nobis in eódem Spíritu re\-cta sápere,~* et de ejus semper consolatióne gaudére. \rr Amen.}{\noindent Let us pray. O God, who didst teach the hearts of thy faithful people, by the sending to them the light of thy Holy Spirit, grant us by the same Spirit to have a right judgment in all things, and evermore to rejoice in his holy comfort. Through Christ our Lord. Amen.}


\end{document}