% !TEX TS-program = lualatexmk
% !TEX parameter =  --shell-escape

\documentclass[vespers_book_la_eng.tex]{subfiles}

\ifcsname preamble@file\endcsname
  \setcounter{page}{\getpagerefnumber{M-vespers_book_la_eng_sundays_pe_gesima}}
\fi

\begin{document}

\phantomsection
\chapter*{SUNDAYS AFTER THE EPIPHANY UNTIL LENT.}\header{sundays after the epiphany until lent.}
\addcontentsline{toc}{chapter}{Sundays after the Epiphany until Lent.}

\bigtitle{II Sunday after the Epiphany.}
\addcontentsline{toc}{section}{II Sunday after the Epiphany} \phantomsection

\rubrique{All as in the psalter of Sunday, \normaltext{\pageref{M-vespers_book_la_eng_sunday_psalms},} except as given below.}

\gscore[Ad Magnif.]{Ant. 1. f}{an_deficiente_vino_solesmes_1961}
\antiphona{english}{The wine failing, Jesus commanded the water pots to be filled with water, which was changed into wine, alleluia}

\collectlateng

\texteparacol{\lettrine{O}{m}nípotens sempitérne Deus, qui cæléstia simul et terréna moderáris:~† supplicatiónes pópuli tui cleménter exáudi;~* et pacem tuam nostris concéde tempóribus. Per Dóminum.}{english}{\noindent 
Almighty and eternal God, as thou govern all things in heaven and on earth; in thy mercy hear the supplication of thy people, and grant thy peace in our times.
Through our Lord.}
   
\label{3_sunday_pe}
\bigtitle{III Sunday after the Epiphany.}
\addcontentsline{toc}{section}{III Sunday after the Epiphany} \phantomsection

\gscore[Ad Magnif.]{Ant. 1. g2}{an_domine_si_tu_vis_solesmes_1961}
\antiphona{english}{Lord, if Thou wilt, Thou canst make me clean; and Jesus saith to him: I will; be thou clean}

\collectlateng

\texteparacol{\lettrine{O}{m}nípotens sempitérne Deus, infirmitátem nostram propítius réspice:~† atque ad protegéndum nos~* déxteram tuæ majestátis exténde.
Per Dóminum.}{english}{\noindent 
Almighty and everlasting God, mercifully look upon our infirmities, and in all our dangers and necessities stretch forth the right hand of thy majesty to help and defend us.
Through our Lord.}

\label{4_sunday_pe}
\bigtitle{IV Sunday after the Epiphany.}
\addcontentsline{toc}{section}{IV Sunday after the Epiphany} \phantomsection

\gscore[Ad Magnif.]{Ant. 1. g2}{an_domine_salva_nos_solesmes_1961}
\antiphona{english}{ Lord, save us: we perish; give the word, O God, and let there be a great calm}

\collectlateng

\texteparacol{\lettrine{D}{e}us, qui nos in tantis perículis constitútos, pro humána scis fragilitáte non posse subsístere:~† da nobis salútem mentis et córporis;~* ut ea, quæ pro peccátis nostris pátimur, te adjuvánte vincámus. Per Dóminum.}{english}{\noindent  O God, who knowest us to be set in the midst of so many and great dangers, that, by reason of the frailty of our nature, we cannot always stand upright; grant to us such health of mind and body, that by thy strength and protection we may overcome all evils, whereby for our sins we are justly afflicted.
Through our Lord.}

\label{5_sunday_pe}
\bigtitle{V Sunday after the Epiphany.}
\addcontentsline{toc}{section}{V Sunday after the Epiphany} \phantomsection

\gscore[Ad Magnif.]{Ant. 1. g}{an_colligite_primum_solesmes_1961}
\antiphona{english}{Gather ye together first the tares, and bind them in bundles to burn them; but gather the wheat into my barn, saith the Lord}

\collectlateng

\texteparacol{\lettrine{F}{a}míliam tuam, quǽsumus Dómine, contínua pietáte custódi:~† ut quæ in sola spe grátiæ cœléstis innítitur,~* tua semper prote\-ctióne muniátur.
Per Dóminum.}{english}{\noindent O Lord, we beseech thee to keep thy family continually in godliness, that they who do lean only upon the hope of thy heavenly grace, may evermore be defended by thy mighty power. Through our Lord.}

\label{6_sunday_pe}
\bigtitle{VI Sunday after the Epiphany.}
\addcontentsline{toc}{section}{VI Sunday after the Epiphany} \phantomsection

\gscore[Ad Magnif.]{Ant. 1. a3}{an_simile_est_fermento_solesmes_1961}
\antiphona{english}{The kingdom of heaven is like unto leaven, which a woman took and hid in three measures of meal till the whole was leavened}

\collectlateng

\texteparacol{\lettrine{P}{r}æsta, quǽsumus omnípotens Deus:~† ut semper rationabília meditántes,~* quæ tibi sunt plácita, et dictis exsequámur et factis. Per Dóminum.}{english}{\noindent Grant us, we beseech thee, O Almighty God, ever to think such things as be reasonable, and in every word and work of ours, to do always that is well pleasing in thy sight. Through our Lord.}

%\biggerrule

\bigtitle{Septuagesima.}
\addcontentsline{toc}{section}{Septuagesima} \phantomsection

\rubrique{On this and the following two Sundays, all as in the psalter of Sunday, \normaltext{\pageref{M-vespers_book_la_eng_sunday_psalms},} except as given below.}

\chapterlateng

\scripture{1 Cor. 9, 24.}

\texteparacol{\lettrine{F}{r}atres: Nescítis quod ii qui in stádio currunt,~† omnes quidem currunt, sed unus áccipit bravíum? Sic cúrrite ut comprehendátis.}{english}{\noindent Know you not that they that run in the race, all run indeed, but one receiveth the prize? So run that you may obtain.}

\rubrique{
Hymn \normaltext{Lucis creátor óptime, \pageref{M-hy_Lucis_Creator_optime}.} \normaltext{℣. Dirigátur, \pageref{M-dirigatur}.}}

\gscore[Ad Magnif.]{Ant. 7. a}{an_dixit_paterfamilias_solesmes}
\antiphona{english}{The householder saith unto his labourers: Why stand ye here all the day idle? But they answered and said: Because no man hath hired us. Go ye also into my vineyard, and whatsoever is right, that will I give you}

\collectlateng

\texteparacol{\lettrine{P}{r}eces pópuli tui, quǽsumus, Dómine, cleménter exáudi:~† ut, qui juste pro peccátis nostris afflígimur,~* pro tui nóminis glória misericórditer liberémur. Per Dóminum.}{english}
{\noindent O Lord, we beseech thee favourably to hear the prayers of thy people; that we, who are justly punished for our offences, may be mercifully delivered by thy goodness, for the glory of thy Name. Through our Lord.}

\bigtitle{Sexagesima.}
\addcontentsline{toc}{section}{Sexagesima} \phantomsection

\chapterlateng

\scripture{2 Cor. 11, 19 – 20.}

\texteparacol{\lettrine{F}{r}atres: Libénter suffértis insipiéntes, cum sitis ipsi sapiéntes:~† sustinétis enim si quis vos in servitútem rédigit, si quis dévorat, si quis áccipit, si quis extóllitur,~* si quis in fáciem vos cædit.}{english}{\noindent Brethren: for you gladly suffer the foolish; whereas yourselves are wise. For you suffer if a man bring you into bondage, if a man devour you, if a man take from you, if a man be lifted up, if a man strike you on the face.}

\rubrique{
Hymn \normaltext{Lucis creátor óptime, \pageref{M-hy_Lucis_Creator_optime}.} \normaltext{℣. Dirigátur, \pageref{M-dirigatur}.}}

\gscore[Ad Magnif.]{Ant. 6. F}{an_vobis_datum_est_solesmes_1961}
\antiphona{english}{Jesus said to his disciples: Unto you it is given to know the mysteries of the kingdom of God, but to others in parables}

\collectlateng

\texteparacol{\lettrine{D}{e}us, qui cónspicis, quia ex nulla nostra actióne confídimus:~† concéde propítius; ut, contra advérsa ómnia,~* Doctóris géntium prote\-ctióne muniámur. Per Dóminum.}{english}
{\noindent
O God, who seest that we put not our trust in anything that we do, mercifully grant that through the protection of the Teacher of the Gentiles we may be defended against all adversity. Through our Lord.}

\bigtitle{Quinquagesima.}
\addcontentsline{toc}{section}{Quinquagesima} \phantomsection

\chapterlateng

\scripture{1 Cor. 13, 1.}

\texteparacol{\lettrine{F}{r}atres: Si linguis hóminum loquar, et Angelórum,~† caritátem autem non hábeam,~* factus sum velut æs sonans, aut cýmbalum tínniens.}{english}{\noindent Brethren: If I speak with the tongues of men, and of angels, and have not charity, I am become as sounding brass, or a tinkling cymbal.}

\rubrique{
Hymn \normaltext{Lucis creátor óptime, \pageref{M-hy_Lucis_Creator_optime}.} \normaltext{℣. Dirigátur, \pageref{M-dirigatur}.}}

\gscore[Ad Magnif.]{Ant. 1. D}{an_stans_autem_jesus_solesmes_1961}
\antiphona{english}{And Jesus stood, and commanded the blind man to be brought unto him, and He asked him, saying: What wilt thou that I shall do unto thee? Lord, that I may receive my sight. And Jesus said unto him: Receive thy sight, thy faith hath saved thee. And immediately he received his sight, and followed him, glorifying God}

\collectlateng

\texteparacol{\lettrine{P}{r}eces nostras, quǽsumus Dómine, cleménter exáudi:~† atque a peccatórum vínculis absolútos,~* ab omni nos adversitáte custódi. Per Dóminum.}{english}{\noindent O Lord, we beseech thee, mercifully hear our prayers; loose us from the chains of our sins and keep us from all adversity.
Through our Lord.}

\end{document}