% !TEX TS-program = lualatexmk
% !TEX parameter =  --shell-escape

\documentclass[vespers_book_la_eng.tex]{subfiles}

\ifcsname preamble@file\endcsname
  \setcounter{page}{\getpagerefnumber{M-vespers_book_la_eng_christmas_jan_5}}
\fi

\begin{document}
\addcontentsline{toc}{chapter}{Nativity of the Lord through January 5}\phantomsection
\festum{nativity of the lord}
\addcontentsline{toc}{section}{Nativity of the Lord}\phantomsection

\header{christmas through january 5}

\doubleclass{I}[with a Privileged Octave of the III Order]

\atvespers{i}

\firstantiphon{8. G}

\initialscore{an_rex_pacificus_solesmes_1961}
\antiphona{english}{The King of Peace is magnified, whose face the whole earth desireth}
\psalmus{109}{109_8G}{109_8}{109_en}

 \gscore[2. Ant.]{7. a}{an_magnificatus_est_solesmes_1961}
 \antiphona{english}{The Peaceful King exceeded all the kings of the whole earth}
\psalmus{110}{110_7a}{110_7}{110_en}

  \gscore[3. Ant.]{8. G}{an_completi_sunt_solesmes_1961}
  \antiphona{english}{The days of Mary were accomplished that she should bring forth her first-born Son}
  \psalmus{111}{111_8G}{111_8}{111_en}

  \gscore[4. Ant.]{8. G}{an_scitote_quia_prope_solesmes_1961}
\antiphona{english}{Know ye that the kingdom of God is at hand? Amen, I say unto you, it will not tarry}
  \psalmus{112}{112_8G}{112_8}{112_en}

 \gscore[5. Ant.]{1. g}{an_levate_capita_vestra_solesmes_1961}
\antiphona{english}{Lift up your heads; behold, your redemption draweth nigh}
\psalmus{116}{116_1g}{116_1}{116_en}

\chapterlateng
\scripture{Tit. 3, 4 – 5.}

\texteparacol{\lettrine{A}{p}páruit benígnitas et humánitas Salvatóris nostri Dei:~† non ex opéribus justítiæ, quæ fécimus nos,~* sed secúndum suam misericórdiam salvos nos fecit.}{english}{\noindent But when the goodness and kindness of God our Saviour appeared: Not by the works of justice, which we have done, but according to his mercy, he saved us.}

\smalltitle{Hymn. Mode I.} \label{hy_jesu_redemptor_omnium_nativitatis_solesmes_1961}

\rubrique{At the words \normaltext{Jesu Redémptor ómnium,} the celebrant extends and then closes his hands while turning to the cross.}

\initialscore{hy_jesu_redemptor_omnium_nativitatis_solesmes_1961}

\textes{jesu_redemptor_omnium_en}

\smallscore{versiculus_i_vesperis_nativitatis_domini}

\begin{enpars} \noindent \vv Tomorrow the iniquity of the earth shall be blotted out.

\noindent \rr And the Saviour of the world shall reign over us.
\end{enpars} 
 
 \gscore[Ad Magnif.]{Ant. 8. G}{an_cum_ortus_fuerit_solesmes_1961}
  \antiphona{english}{When the sun shall have risen from heaven, ye shall see the King of kings proceeding from the Father, as a bridegroom from his chamber}
  
\rubrique{Collect \normaltext{Concéde, quǽsumus} as in II Vespers, below, \normaltext{\pageref{oratio_25_dec}.}}
  
%  \collectlateng
  
%  \texteparacol{
%\lettrine{C}{o}ncéde, quǽsumus, omnípotens Deus:~* ut nos Unigéniti tui nova per carnem Natívitas líberet; quos sub peccáti jugo vetústa sérvitus tenet.
%Per eúmdem Dóminum.}{english}{\noindent Grant, we beseech thee, almighty God, that the new birth of thy Only-begotten Son as man may set us free, who are held by the old bondage under the yoke of sin. Through the same Lord.}

\atvespers{ii} \label{2_vesperas_dec_25}

\firstantiphon{1. g}

\initialscore{an_tecum_principium_solesmes_1961}
\antiphona{english}{Thine shall be the dominion in the day of thy power, amid the brightness of the saints, from the womb, before the day-star, have I begotten thee}
\psalmus{109}{109_1g}{109_1}{109_en}

\gscore[2. Ant.]{7. a}{an_redemptionem_misit_solesmes_1961} %%should EUOUAE be there if the psalm isn't but it's only 1 psalm referenced elsewhere?
 \antiphona{english}{The Lord sent redemption unto his people, He hath commanded his covenant for ever}
\psalmus{110}{110_7a}{110_7} {110_en}

\gscore[3. Ant.]{7. b}{an_exortum_est_nativitate_solesmes_1961}
  \antiphona{english}{Unto the upright there ariseth light in the darkness; the Lord is gracious, and full of compassion, and righteous}
\psalmus{111}{111_7b}{111_7}{111_en}

\gscore[4. Ant.]{4. A*}{an_apud_dominum_misericordia_solesmes_1961}
\antiphona{english}{With the Lord there is mercy, and with him is plenteous redemption}
\psalmus{129}{129_4A_A_star}{129_4A_A_star}{129_en}

\gscore[5. Ant.]{8. G}{an_de_fructu_ventris_tui_solesmes_1961}
\antiphona{english}{Of the fruit of thy body will I set upon thy throne}
\psalmus{131}{131_8G}{131_8}{131_en}

\chapterlateng

\scripture{Heb. 1, 1 – 2.}

\texteparacol{\lettrine{M}{u}ltifáriam multísque modis olim Deus loquens pátribus in prophétis:~† novíssime diébus istis locútus est nobis in Fílio, quem constítuit hærédem universórum,~* per quem fecit et sǽcula}{english}{\noindent God, who, at sundry times and in diverse manners, spoke in times past to the fathers by the prophets, last of all, in these days hath spoken to us by his Son, whom he hath appointed heir of all things, by whom also he made the world.}

\rubrique{Hymn \normaltext{Jesu Redémptor ómnium, {\pageref{hy_jesu_redemptor_omnium_nativitatis_solesmes_1961}.}}}

\label{versiculus_2_vesperis_nativitatis_domini}
\smallscore{versiculus_ii_vesperis_nativitatis_domini}

\begin{enpars} \noindent \vv The Lord hath made known, alleluia.

\noindent \rr His salvation, alleluia.
\end{enpars}

\label{an_hodie_christus_natus_est_solesmes_1961}

\gscore[Ad Magnif.]{Ant. 1. g2}{an_hodie_christus_natus_est_solesmes_1961}
 \antiphona{english}{Today the Christ is born: Today the Saviour is appeared; Today the Angels sing praise in the earth and the Archangels rejoice; Today the righteous are glad and say: Glory to God in the highest. Alleluia}
 
 \collectlateng \label{oratio_25_dec}
 \texteparacol{
\lettrine{C}{o}ncéde, quǽsumus omnípotens Deus:~* ut nos Unigéniti tui nova per carnem Natívitas líberet; quos sub peccáti jugo vetústa sérvitus tenet.
Per eúmdem Dóminum.}{english}{\noindent Grant, we beseech thee, almighty God, that the new birth of thy Only-begotten Son as man may set us free, who are held by the old bondage under the yoke of sin.
Through the same Lord.}

\rubrique{Then a commemoration of Saint Stephen, Protomartyr is made.} \label{an_stephanus_autem_solesmes_1961}

\gscore[Ant.]{8. C}{an_stephanus_autem_solesmes_1961}
 \antiphona{english}{And Stephen, full of grace and fortitude, did great signs among the people}

%%These verses probably all need to be macros or tex files inserted with \input

\texteparacol{\vv Glória et honóre coronásti eum, Dómine.

\rr Et constituísti eum super ópera mánuum tuárum.}{english}{\noindent \vv Thou hast crowned him with glory and honour, O Lord.

\noindent \rr And madest him to have dominion over the works of thy hands.}

\rubrique{Collect \normaltext{Da nobis, quǽsumus,} as below, \normaltext{ \pageref{oratio_s_stephani_dec_26}.}}

\litdateen{December 26}
\festum{saint stephen, protomartyr}
%%to verify
\addcontentsline{toc}{section}{Saint Stephen, Protomartyr}\phantomsection

\doubleclass{II}[with a Simple Octave]

\rubrique{\begin{enpars}Antiphon \normaltext{Tecum princípium} with the remaining antiphons and psalms of the Nativity, as above, \normaltext{\pageref{2_vesperas_dec_25}.} And it is thus said also on the other feasts within this occurring octave.\end{enpars}}

\rubrique{\begin{enpars}On the three days after Christmas, Vespers of the feast begin from the chapter.\end{enpars}}

\chapterlateng

 \texteparacol{\lettrine{S}{t}éphanus autem plenus grátia et fortitúdine,~* faciébat prodígia et signa magna in pópulo.}{english}
 {\noindent And Stephen, full of grace and fortitude, did great wonders and signs among the people.}

\label{deus_tuorum_militum_nat}

\hymnlateng

\gscore[]{1.}{hy_deus_tuorum_militum_st_stephen_solesmes_1961}
\textes{deus_tuorum_dox_nat_en}

%\vv Stéphanus vidit cælos apértos.
%
%\rr Vidit et introívit: beátus homo, cui cæli patébant.

\smallscore{versiculus_26_dec}
\begin{enpars}
 \noindent \vv
 Stephen saw the heavens opened.

\noindent \rr He saw and entered: blessed man, to whom the heavens were opened.
\end{enpars}

\gscore[Ad Magnif.]{Ant. 8. G}{an_sepelierunt_stephanum_solesmes}
 \antiphona{english}{Devout men buried Stephen and made great mourning over him}

 \collectlateng \label{oratio_s_stephani_dec_26}
 \texteparacol{
\lettrine{D}{a} nobis, quǽsumus Dómine, imitári quod cólimus, ut discámus et inimícos dilígere:~† quia ejus natalítia celebrámus,~* qui novit étiam pro persecutóribus exoráre Dóminum nostrum Jesum Christum Fílium tuum: Qui tecum vivit et regnat in unitáte.}{english}{\noindent Grant us, we beseech thee, O Lord, so to imitate what we revere, that we may learn to love even our enemies; for we celebrate the heavenly birthday of him who knew how to pray for his very persecutors to our Lord Jesus Christ:
Who with thee liveth and reigneth in the unity.}

\rubrique{
\begin{enpars}
 Then a commemoration of Saint John, Apostle and Evangelist, is made.\end{enpars}}
 
\gscore[Ant.]{1.}{an_iste_est_johannes_solesmes_1961}
 \antiphona{english}{This is John who leaned at supper on the bosom of the Lord. Blessed Apostle, to whom heavenly secrets were revealed}

\texteparacol{\vv Valde honorándus est beátus Joánnes.

\rr Qui supra pectus Dómini in cœna recúbuit.}{english}
{\noindent \vv Right worthy of honour is the blessed Apostle John.

\noindent\rr Who leaned on the Lord's bosom at the Last Supper}

\rubrique{Collect \normaltext{Ecclésiam,} as below, \normaltext{ \pageref{oratio_s_joannis_dec_27}.}}

\rubrique{Then a commemoration of the preceding day in the Octava of the Nativity is made. Ant. \normaltext{Hódie, \pageref{an_hodie_christus_natus_est_solesmes_1961}, ℣. Notum fecit, \pageref{versiculus_2_vesperis_nativitatis_domini},} Oratio \normaltext{Concéde, quǽsumus, \pageref{oratio_25_dec}.}}

\litdateen{December 27}
\festum{saint john, apostle and evangelist}
%%to verify
\addcontentsline{toc}{section}{Saint John, Apostle and Evangelist}\phantomsection

\doubleclass{II}[with a Simple Octave]

\chapterlateng

\scripture{Eccli. 15, 1 – 2}

\texteparacol{\lettrine{Q}{u}i timet Deum, fáciet bona:~† et qui cóntinens est justítiæ, apprehéndet illam,~* et obviábit illi quasi mater honorificáta.}{english}{\noindent He that feareth God, will do good: and he that possesseth justice, shall lay hold on her, and she will meet him as an honourable mother.}

\hymnlateng

\gscore[]{1.}{hy_exsultet_orbis_gaudiis_st_john_solesmes_1961}

\textes{exsultet_orbis_nat}

\smallscore{versiculus_27_dec}

\begin{enpars}
 \noindent \vv Right worthy of honour is the blessed Apostle John.

\noindent \rr Who leaned on the Lord's bosom at the Last Supper.
\end{enpars}

\gscore[Ad Magnif.]{Ant. 6. F}{an_exiit_sermo_solesmes_1961}
 \antiphona{english}{A saying went abroad among the brethren, that that disciple should not die; and Jesus did not say that he should not die, but so will I have him remain till I come}

\collectlateng \label{oratio_s_joannis_dec_27}

\texteparacol{\lettrine{E}{c}clésiam tuam Dómine benígnus illústra:~† ut beáti Joánnis Apóstoli tui et Evangelístæ illumináta doctrínis,~* ad dona pervéniat sempitérna.
Per Dóminum.}{english}{\noindent Shine upon thy Church, O Lord, in thy goodness, that, enlightened by the teachings of Blessed John, thy Apostle and Evangelist, she may attain to everlasting gifts. Through our Lord.}

\rubrique{
\begin{enpars}
A commemoration of the Holy Innocents is made.\end{enpars}}

\gscore[Ant.]{1.}{an_hi_sunt_qui_cum_mulieribus_solesmes_1961}

\texteparacol{\vv Heródes irátus occídit multos púeros.

\rr In Béthlehem Judæ civitáte David.}{english}{\noindent \vv  Herod being wroth, slew many children.

\noindent \rr  In Bethlehem of Juda, the city of David.
}

\rubrique{Collect \normaltext{Deus, cujus hodiérna,} as below, \normaltext{ \pageref{oratio_ss_innocentium_dec_28}.}}

\rubrique{Then a commemoration of the preceding day in the Octava of the Nativity is made. Ant. \normaltext{Hódie, \pageref{an_hodie_christus_natus_est_solesmes_1961}, ℣. Notum fecit, \pageref{versiculus_2_vesperis_nativitatis_domini},} Oratio \normaltext{Concéde, quǽsumus, \pageref{oratio_25_dec}.}}

\newpage
\litdateen{December 28}
\festum{feast of the holy innocents}
\doubleclass{II}[with a Simple Octave]
%%to verify
\addcontentsline{toc}{section}{Holy Innocents}\phantomsection

\chapterlateng

\scripture{Apoc. 14, 1}

\texteparacol{\lettrine{V}{i}di supra montem Sion Agnum stántem,~† et cum eo centum quadragínta quátuor míllia,~* habéntes nomen ejus, et nomen Patris ejus scriptum in fróntibus suis.}{english}{\noindent Lo, a lamb stood upon mount Sion, and with him an hundred forty-four thousand, having his name, and the name of his Father, written on their foreheads.}

\hymnlateng

\gscore[]{1.}{hy_salvete_flores_martyrum_solesmes_1961}
\textes{salvete_flores_nat}

\smallscore{versiculus_28_dec}

\begin{enpars}
 \noindent \vv From beneath the throne of God all the saints cry out.

\noindent \rr Avenge thou our blood, O our God.
\end{enpars}

\gscore[Ad Magnif.]{Ant. 2. D}{an_innocentes_pro_christo_solesmes_1961}
 \antiphona{english}{Innocent babies were slain for Christ, sucklings were killed by a wicked king. Now they follow the Lamb without spot and cry without ceasing, Glory be to thee, O Lord}
 
 \collectlateng \label{oratio_ss_innocentium_dec_28}
 
 \texteparacol{\lettrine{D}{e}us, cujus hodiérna die præcónium Innocéntes Mártyres non loquéndo, sed moriéndo conféssi sunt:~† ómnia in nobis vitiórum mala mortífica; ut fidem tuam, quam lingua nostra lóquitur,~* étiam móribus vita fateátur. Per Dóminum.
Per Dóminum.}{english}{\noindent O God, whose praise the martyred Innocents on this day confessed, not by speaking, but by dying, destroy all the evils of sin in us, that our life also may proclaim in deeds, thy faith which our tongues profess. Through our Lord.}

\rubrique{Ant. \normaltext{Iste Sanctus} from the Common.}

%%These verses probably all need to be macros or tex files inserted with \input

\texteparacol{\vv Glória et honóre coronásti eum, Dómine.

\rr Et constituísti eum super ópera mánuum tuárum.}{english}{\noindent \vv Thou hast crowned him with glory and honour, O Lord.

\noindent \rr And madest him to have dominion over the works of thy hands.}

\collectlateng

 \texteparacol{\lettrine{D}{e}us, pro cujus Ecclésia gloriósus Póntifex Thomas gládiis impiórum occúbuit:~† præsta, quǽsumus; ut omnes qui ejus implórant auxílium,~* petitiónis suæ salutárem consequántur effé\-ctum.
(Per Dóminum.)}{english}{\noindent O God, in defence of whose church the glorious Bishop Thomas fell by the swords of wicked men, grant, we beseech thee, that all that ask his help, may obtain wholesome fruit of their petition. (Through our Lord.)}
 
\rubrique{Then a commemoration of the preceding day in the Octave of the Nativity is made. Ant. \normaltext{Hódie, \pageref{an_hodie_christus_natus_est_solesmes_1961}, ℣. Notum fecit, \pageref{versiculus_2_vesperis_nativitatis_domini},} Oratio \normaltext{Concéde, quǽsumus, \pageref{oratio_25_dec}.}}

\bigtitle{Sunday within the octave of the Nativity.}
\addcontentsline{toc}{section}{Sunday within the octave of the Nativity}\phantomsection

\atvespersminor{I}

\chapterlateng \label{cap_dom_oct_nat}

\scripture{Gal. 4, 1 – 2}

\texteparacol{\lettrine{F}{r}atres: Quanto témpore hæres párvulus est, nihil differt a servo, cum sit dóminus ómnium:~† sed sub tutóribus et actóribus est,~* usque ad præfínitum tempus a patre.
}{english}{\noindent But when the goodness and kindness of God our Saviour appeared: Not by the works of justice, which we have done, but according to his mercy, he saved us.}

\rubrique{Hymn \normaltext{Jesu Redémptor ómnium,} as at I Vespers of the Nativity, \normaltext{\pageref{hy_jesu_redemptor_omnium_nativitatis_solesmes_1961}.}}

\label{versiculus_dom_infra_oct_nat}

\smallscore{versiculus_dom_infra_oct_nat}

\begin{enpars} \noindent \vv The Word was made flesh, alleluia.

\noindent \rr And dwelt among us, alleluia.
\end{enpars}

\gscore[Ad Magnif.]{Ant. 8. G}{an_dum_medium_silentium_solesmes_1961}
\antiphona{english}{While all things were in quiet silence and that night was in the midst of her swift course, thine Almighty Word, O Lord, leapt down out of thy Royal Throne}

\collectlateng \label{oratio_dom_infra_oct_nat}

\texteparacol{
\lettrine{O}{m}nípotens sempitérne Deus, dírige actus nostros in beneplácito tuo:~† ut in nómine dilécti Fílii tui~* mereámur bonis opéribus abundáre:
Qui tecum vivit et regnat in unitáte.}{english}
{\noindent Almighty and everlasting God, do Thou order all our actions in conformity with thy good pleasure, that through the name of thy well-beloved Son, we may worthily abound in all good works. Who with thee liveth and reigneth, in the unity.}

\rubrique{Commemoration of the preceding day in the Octave of the Nativity: Ant. \normaltext{Hódie, \pageref{an_hodie_christus_natus_est_solesmes_1961}, ℣. Notum fecit, \pageref{versiculus_2_vesperis_nativitatis_domini},} Oratio \normaltext{Concéde, \pageref{oratio_25_dec}.}}

\atvespersminor{II}

\rubrique{Antiphons and psalms of II Vespers of the Nativity, \normaltext{\pageref{2_vesperas_dec_25}.}}

\rubrique{Chapter \normaltext{Quanto témpore} as at I Vespers, \normaltext{\pageref{cap_dom_oct_nat}.}}

\rubrique{Hymn \normaltext{Jesu Redémptor ómnium, \pageref{hy_jesu_redemptor_omnium_nativitatis_solesmes_1961}.}}

\rubrique{ \normaltext{℣. Verbum,} in the ordinary tone as in I Vespers, \normaltext{\pageref{versiculus_dom_infra_oct_nat}.}}

\gscore[Ad Magnif.]{Ant. 6. F}{an_puer_jesus_solesmes_solesmes_1961}
\antiphona{english}{The child Jesus advanced in wisdom and age, and grace with God and men}

\rubrique{Collect \normaltext{Omnípotens sempitérne Deus,} as above, \normaltext{\pageref{oratio_dom_infra_oct_nat}.}}

\rubrique{Commemoration of the preceding day in the Octave of the Nativity: Ant. \normaltext{Hódie, \pageref{an_hodie_christus_natus_est_solesmes_1961}, ℣. Notum fecit, \pageref{versiculus_2_vesperis_nativitatis_domini},} Oratio \normaltext{Concéde, \pageref{oratio_25_dec}.}}

%\litdateen{December 29}
%
%\bigtitle{Saint Thomas Becket, Bishop and Martyr}
%
%\addcontentsline{toc}{section}{Saint Thomas Becket}\phantomsection


%\litdateen{December 30}
%
%\bigtitle{VI Day in the Octave of the Nativity.}
%
%\addcontentsline{toc}{section}{VI Day in the Octave of the Nativity.}\phantomsection
%
%\semidouble
%
%\rubrique{\begin{enpars}Antiphon \normaltext{Tecum princípium} with the remaining antiphons and psalms of the Nativity, as above, with the antiphons being doubled, \normaltext{\pageref{2_vesperas_dec_25}. From the chapter, Vespers are of the following feast of Saint Silvester, as in the common.} 
\newpage
\festum{circumcision of the lord}
\addcontentsline{toc}{section}{Circumcision of the Lord}\phantomsection

\doubleclass{II}

\atvespers{i} \label{1_vespers_jan_1}

\firstantiphon{6. F}

\initialscore{an_o_admirabile_commercium_solesmes_1961}
\antiphona{english}{O wondrous interchange! the Creator of mankind, taking upon him a living body, vouchsafed to be born of a pure Virgin: and by his Humanity, which was begotten in no earthly wise, hath made us partakers of his Divinity}
\psalmus{109}{109_6_alt}{109_6_alt}{109_en}

 \gscore[2. Ant.]{3. a2}{an_quando_natus_es_solesmes_1961}
 \antiphona{english}{When thou wast born all ineffably of a Virgin, then were the Scriptures fulfilled; thou camest down like the dew upon the fleece of wool, to bring salvation unto all mankind; we praise thee, O our God}
\psalmus{112}{112_3a2}{112_3a2_g}{112_en}

  \gscore[3. Ant.]{4. E}{an_rubum_quem_solesmes_1961}
  \antiphona{english}{In the bush which Moses saw unconsumed, we recognize the preservation of thy glorious virginity: holy Mother of God, intercede for us}
  \psalmus{121}{121_4E}{121_4E}{121_en}

  \gscore[4. Ant.]{1. f}{an_germinavit_radix_solesmes_1961}
\antiphona{english}{The Root of Jesse hath budded, the Star hath come out of Jacob, the Virgin hath borne the Saviour: we praise thee, O our God}
  \psalmus{126}{126_1f}{126_1}{126_en}

 \gscore[5. Ant.]{2. D}{an_ecce_maria_genuit_solesmes_1961}
\antiphona{english}{Lo, Mary hath brought forth the Saviour of whom when John saw him, he did proclaim: Behold the Lamb of God that taketh away the sins of the world, alleluia}
\psalmus{147}{147_2}{147_2}{147_en}

\chapterlateng
\scripture{Tit. 2, 11–12.}

\texteparacol{\lettrine{A}{p}páruit grátia Dei Salvatóris nostri ómnibus homínibus,~† erúdiens nos, ut abnegántes impietátem et sæculária desidéria,~* sóbrie, et juste, et pie vivámus in hoc sǽculo.}{english}{\noindent The grace of God that bringeth salvation hath appeared to all men, teaching us that, denying ungodliness and worldly lusts, we should live soberly, righteously, and godly in this present world.}

\rubrique{Hymn \normaltext{Jesu Redémptor ómnium,} as at I Vespers of the Nativity, \normaltext{\pageref{hy_jesu_redemptor_omnium_nativitatis_solesmes_1961}.}}

\rubrique{\normaltext{℣. Verbum,} as on the Sunday within the octave of the Nativity, \normaltext{\pageref{versiculus_dom_infra_oct_nat}.}}

%\smallscore{versiculus_dom_infra_oct_nat}
%
%\begin{enpars} \noindent \vv The Word was made flesh, alleluia.
%
%\noindent \rr And dwelt among us, alleluia.
%\end{enpars}

\gscore[Ad Magnif.]{Ant. 8. G}{an_propter_nimiam_solesmes_1961}
 \antiphona{english}{God, for his great love wherewith he loved us, sent his own Son in the likeness of sinful flesh. Alleluia}
 
 \collectlateng \label{oratio_1_jan} \phantomsection
 
 \texteparacol{\lettrine{D}{e}us, qui salútis ætérnæ, beátæ Maríæ virginitáte fecúnda, humáno géneri prǽmia præstitísti:~† tríbue, quǽsumus; ut ipsam pro nobis intercédere sentiámus,~* per quam merúimus auctórem vitæ suscípere, Dóminum nostrum Jesum Christum Fílium tuum. Qui tecum vivit et regnat.}{english}
 {\noindent O God, who, by the fruitful virginity of the Blessed Mary, hast given unto mankind the rewards of everlasting life; grant, we beseech thee, that we may continually feel the might of her intercession, through whom we have worthily received the Author of our life, our Lord Jesus Christ, thy Son: Who with thee liveth and reigneth.}

\atvespers{ii}

\allexcept

\smallscore{versiculus_2_vesperis_jan_1}

\begin{enpars} \noindent \vv The Lord hath made known, alleluia.

\noindent \rr His salvation, alleluia.
\end{enpars}

%\label{an_magnum_haereditatis_solesmes_1961}

\gscore[Ad Magnif.]{Ant. 2. A}{an_magnum_haereditatis_solesmes_1961}
 \antiphona{english}{Herein is a great mystery of heirship; the womb of her that knew not a man is become the temple of God; in taking of her flesh, He was not defiled; all nations shall come and say: Glory be to thee, O Lord}

\rubrique{Collect \normaltext{Deus qui salútis,} as above,  \normaltext{\pageref{oratio_1_jan}.}}

\rubrique{And a commemoration of the following day is not made.}

\litdateen{Sunday between the Circumcsion and the Epiphany}

\festum{most holy name of jesus}
\addcontentsline{toc}{section}{Holy Name Of Jesus}\phantomsection

\doubleclass{II}

\rubrique{When this feast is celebrated on January 2, at II Vespers of the preceding day, no commemoration is made of the Most Holy Name.}

\atvespers{i}

\firstantiphon{8. G}

\initialscore{an_omnis_qui_invocaverit_solesmes_1961}
\antiphona{english}{Whosoever shall call on the name of the Lord shall be saved}
\psalmus{109}{109_8G}{109_8}{109_en}
%\rubrique{Ps. \normaltext{Dixit Dóminus, \pageref{109_8g}.}}

\gscore[2. Ant.]{5. a}{an_sanctum_et_terribile_solesmes_1961}
\antiphona{english}{Holy and terrible is his Name; the fear of the Lord is the beginning of wisdom}
\psalmus{110}{110_5a}{110_5}{110_en}

\gscore[3. Ant.]{3. a2}{an_ego_autem_in_domino_solesmes_1961}
\antiphona{english}{Yet will I rejoice in the Lord I will joy in the God of my salvation}
\label{111_3a2}
\psalmus{111}{111_3a2}{111_3a2_g}{111_en}

\gscore[4. Ant.]{4. E}{an_a_solis_ortu_solesmes_1961}
\antiphona{english}{From the rising of the sun unto the going down of the same the Lord's Name is to be praised}
%\rubrique{Ps. \normaltext{Laudáte Púeri, \pageref{M-112_4E}.}}

%\label{112_4E} 
\psalmus{112}{112_4E}{112_4E}{112_en}

\gscore[5. Ant.]{8. c}{an_sacrificabo_hostiam_solesmes_1961}
\antiphona{english}{I will offer the sacrifice of thanksgiving, and will call upon the name of the Lord}
\psalmus{115}{115_8c}{115_8}{115_en}

\chapterlateng
\scripture{Phil. 2, 8 – 10}

%%this text probably needs to be in its own file!

\texteparacol{\lettrine{F}{r}atres: Christus humiliávit semetípsum, factus obédiens usque ad mortem, mortem autem crucis.~† Propter quod et Deus exaltávit illum, et donávit illi nomen, quod est super omne nomen,~* ut in nómine Jesu omne genu flectátur.}{english}{He humbled himself, becoming obedient unto death, even to the death of the cross. For which cause God also hath exalted him, and hath given him a name which is above all names: That in the name of Jesus every knee should bow, of those that are in heaven, on earth, and under the earth.}

\hymnlateng

%%changed pœnitentibus to pæ-
\gscore[]{1.}{hy_jesu_dulcis_memoria_solesmes_1957}
\textes{jesu_dulcis_memoria_en}

\smallscore{versiculus_holy_name}
%\vv Sit nomen Dómini benedíctum, allelúia.
%
%\rr Ex hoc nunc et usque in sǽculum, allelúia.

   \begin{enpars}\noindent \vv Blessed be the name of the Lord, alleluia.

\noindent \rr  From this time forth, and for evermore, alleluia.\end{enpars}

\gscore[Ad Magnif.]{Ant. 8. G}{an_fecit_mihi_magna_solesmes_1961}
\antiphona{english}{He that is mighty hath done to me great things, and holy is his Name, alleluia}

\collectlateng

\texteparacol{\lettrine{D}{e}us, qui unigénitum Fílium tuum constituísti humáni géneris Salvatórem, et Jesum vocári jussísti:~† concéde propítius; ut cujus sanctum nomen venerámur in terris,~* ejus quoque aspéctu perfruámur in cælis. Per eúmdem Dóminum.}{english}{\noindent O God, who hast appointed thine Only-Begotten Son to be the Saviour of mankind, and hast commanded that his name should be called Jesus, mercifully grant that we who here on earth do worship that most holy Name may be made glad in heaven by his presence.
Through the same Jesus Christ}

\atvespers{ii}
\rubrique{When this feast is celebrated on January 5, Vespers of the following feast of the Epiphany are said, with a commemoration of the preceding feast.}

\allexcept

\gscore[Ad Magnif.]{Ant. 1. g}{an_vocabis_nomen_ejus_solesmes_1961}
\antiphona{english}{Thou shalt call his name Jesus, for he shall save his people from their sins, alleluia}

\end{document}