
% !TEX TS-program = lualatexmk
% !TEX parameter =  --shell-escape

\documentclass[vespers_book_la_eng.tex]{subfiles}

\ifcsname preamble@file\endcsname
  \setcounter{page}{\getpagerefnumber{M-vespers_book_la_eng_easter_octave}}
\fi

\begin{document}

% \chapter*{SUNDAY OF THE RESURRECTION.}
\festum{sunday of the resurrection}
\addcontentsline{toc}{chapter}{Sunday of the Resurrection and its Octave}\phantomsection
\header{sunday of the resurrection and its octave.}

\doubleclass{I}[with a Privileged Octave of the I Order]

\firstantiphon{8. G}

 \initialscore{an_angelus_autem_domini_solesmes_1961}
\antiphona{english}{For an angel of the Lord descended from heaven and coming rolled back the stone and sat upon it, alleluia, alleluia}
\psalmus{109}{109_8G}{109_8}{109_en}

 \gscore[2. Ant.]{7. c}{an_et_ecce_terraemotus_solesmes_1961}
 \antiphona{english}{ There was a great earthquake; For an angel of the Lord descended from heaven, alleluia}
\psalmus{110}{110_7c}{110_7}{110_en}

\gscore[3. Ant.]{8. c}{an_erat_autem_aspectus_solesmes_1961}
 \antiphona{english}{ And his countenance was as lightning and his raiment as snow, alleluia, alleluia}
  \psalmus{111}{111_8c}{111_8}{111_en}

\gscore[4. Ant.]{7. a}{an_prae_timore_solesmes_1961}
\antiphona{english}{And for fear of him the guards were struck with terror and became as dead men, alleluia}
  \psalmus{112}{112_7a}{112_7}{112_en}

\gscore[5. Ant.]{8. G}{an_respondens_autem_angelus_solesmes_1961}
\antiphona{english}{ And the angel answering said to the women: Fear not you: for I know that you seek Jesus, alleluia}
\psalmus{113}{113_8G}{113_8}{113_en}

\rubrique{The hymn, chapter, and versicle are omitted. The following antiphon is sung.}

\gscore[Ant.]{2.}{an_haec_dies_solesmes_1961}
\antiphona{english}{This is the day which the Lord hath made: let us be glad and rejoice therein}

\gscore[Ad Magnif.]{Ant. 3. a}{an_et_respicientes_solesmes_1961}
 \antiphona{english}{And looking they saw the stone rolled back, for it was very great, alleluia}

\collectlateng

\texteparacol{\lettrine{D}{e}us, qui hodiérna die per Unigénitum tuum æternitátis nobis áditum devícta morte reserásti:~† vota nostra, quæ præveniéndo aspíras,~* étiam adjuvándo proséquere.
Per eúmdem Dóminum nostrum.}{english}{\noindent O God, who on this day, through thy Only-begotten Son, overcame death, and opened to us the gate of everlasting life: as by thy anticipating grace, thou breathest good desires into our hearts, so also, by thy gracious help, bring them to good effect. Through the same Lord.}

\smallscore{Benedicamus_domino_alleluia_solesmes_1961}

\rubrique{This tone may be used at Vespers today and on the following two days.}

\smallscore{Benedicamus_Dom_resurrectionis_solesmes_1961}
  
    \rubrique{\begin{enpars}Within the Paschal Octave, until Vespers of Saturday exclusive, all is said as on the day, except the Collect and the Antiphon at the Magnificat. On Monday and Tuesday, the office is of double rite; on the other days, semidouble.\end{enpars}}

\bigtitle{Easter Monday.}
\addcontentsline{toc}{section}{Easter Monday} \phantomsection

\gscore[Ad. Magnif.]{Ant. 8. G*}{an_qui_sunt_hi_sermones_solesmes_1961}
\antiphona{english}{What manner of communications are these that ye have one to another, as ye walk, and are sad? Alleluia}

\collectlateng

\texteparacol{\lettrine{D}{e}us, qui solemnitáte pascháli mundo remédia contulísti~† pópulum tuum, quǽsumus, cælésti dono proséquere;~* ut et perféctam libertátem cónsequi mereátur, et ad vitam profíciat sempitérnam. Per Dóminum nostrum.}{english}{\noindent O God, who dost heal the sick world by the solemn gladness of the Passover, continue, we beseech thee, to pour forth thine heavenly gifts upon thy people, till the same shall bring them into perfect liberty, and finally avail them unto life everlasting.
Through our Lord.}

\bigtitle{Easter Tuesday.}
\addcontentsline{toc}{section}{Easter Tuesday} \phantomsection

\gscore[Ad. Magnif.]{Ant. 8. G}{an_videte_manus_meas_solesmes_1961}
\antiphona{english}{Behold My Hands and my Feet, for it is I myself, alleluia, alleluia}

\collectlateng

\texteparacol{\lettrine{D}{e}us, qui Ecclésiam tuam novo semper fœtu multíplicas:~† concéde fámulis tuis, ut sacraméntum vivéndo téneant,~* quod fide percepérunt.
Per Dóminum nostrum.}{english}{\noindent O God, who art ever multiplying the children of thy church, grant unto the same thy servants that they may lead the rest of their lives according to this beginning wherein thou hast given them faith to receive the sacrament of the new birth.
Through our Lord.}

\bigtitle{Easter Wednesday.}
\addcontentsline{toc}{section}{Easter Wednesday} \phantomsection

\gscore[Ad. Magnif.]{Ant. 8. G}{an_dixit_jesus_discipulis_solesmes_1961}
\antiphona{english}{Jesus saith unto his disciples: Bring of the fish which ye have now caught. Simon Peter went up, and drew the net to land, full of great fishes, alleluia}

\collectlateng

\texteparacol{\lettrine{D}{e}us, qui nos resurrectiónis Domínicæ ánnua solemnitáte lætíficas:~† concéde propítius; ut per temporália festa quæ ágimus,~* perveníre ad gaúdia ætérna mereámur.
Per eúmdem Dóminum nostrum.}{english}{\noindent O God, who dost every year fill us with holy gladness for the rising-again of the Lord, mercifully grant that these feast days which we are now keeping here in time, may be to us a mean whereby in the end we may worthily attain unto those pleasures which are at thy right hand for evermore.
Through the same Lord.}

%\rubrique{Deinde dicuntur, commemorationes, si quæ faciendæ sunt, cum reliquis ut supra in die et ut in Ordinarium.}

%%this rubric is LU, I believe.

\bigtitle{Easter Thursday.}
\addcontentsline{toc}{section}{Easter Thursday} \phantomsection

\gscore[Ad. Magnif.]{Ant. 7. b}{an_tulerunt_dominum_solesmes_1961}
\antiphona{english}{They have taken away my Lord, and I know not where they have laid him: if thou have borne him hence, tell me, alleluia, and I will take him away, alleluia}
 
\collectlateng

\texteparacol{\lettrine{D}{e}us, qui diversitátem géntium in confessióne tui nóminis adunásti:~† da, ut renátis fonte baptísmatis,~* una sit fides méntium et píetas actiónum.
Per Dóminum nostrum.}{english}{\noindent O God, who dost make all nations, how diverse soever they be, to become one family in giving of praise to thy Name, grant unto all them that are born again in the fountain of baptism to live ever in oneness of faith, and godliness of works.
Through our Lord.}

\bigtitle{Easter Friday.}
\addcontentsline{toc}{section}{Easter Friday} \phantomsection

\gscore[Ad. Magnif.]{Ant. 8. G}{an_data_est_mihi_solesmes_1961}
\antiphona{english}{All power is given unto Me in heaven and in earth, alleluia}

\collectlateng

\texteparacol{\lettrine{O}{m}nípotens sempitérne Deus, qui paschále sacraméntum in reconciliatiónis humánæ f\ooalign{\hss\kern.2ex\'{}\hss\crcr\oe}dere contulísti:~† da méntibus nostris; ut quod professióne celebrámus, ~* imitémur efféctu. Per Dóminum nostrum.}{english}{\noindent Almighty everlasting God, in the paschal sacrament thou instituted the covenant, whereby thou forgavest mankind; grant to our souls, that what we outwardly profess we may show forth in our deeds.
Through our Lord.}

\end{document}