% !TEX TS-program = lualatexmk
% !TEX parameter =  --shell-escape
\documentclass[12pt]{book}

\input{commonheaders_vespers_people} %%% will need to fix spacing problem in psalms at a later date.

\begin{document}
 
 \twosided[pc] %% to allow for the left-right switching of columns needed for bilingual liturgical typesetting (like in missals); it's unclear why this goes BEFORE everything else and not when activating the environment. %% It's OK for paper booklets to not switch, but it's probably best practice to switch.
 
  \thispagestyle{empty}
 
\feasttitle{PENTECOST.} \header{pentecost.}
 
\smalltitle{Before the Office.}

\rubrique{These prayers are said kneeling. End when the priest stands.}

\textebilingue{orationes_ante_do}{orationes_ante_do_eng}

\smalltitle{\prayers}

\rubrique{The celebrant intones the following verse. All sing the response and doxology.}

\begin{enpars}
\rubrique{Make the sign of the cross at the beginning and at the Magnificat, and bow the head when the Trinity is mentioned, when mention is made of the ``Holy Name'' in some way, and at ``Orémus.''}\end{enpars}

%\rubrique{On the feast itself:}

%%to insert if the octave is included

\gscore[]{℣.}{Deus_in_adjutorium_tonus_solemnior}

\noindent
\begin{enpars}
O God, \cc{} come to my assistance. \rr O Lord, make haste to help me. Glory be to the Father, and to the Son,~* and to the Holy Ghost. As it was in the beginning, is now,~* and ever shall be, world without end. Amen. Alleluia.\end{enpars}



%\myrule %%to insert if the octave is included

%\rubrique{On the other days within the octave:}

%\gscore[]{℣.}{Deus_In_Adjutorium_Festivus}

%\myrule %%to insert if the octave is included


%\rubrique{The office is of double rite through Tuesday inclusive. On all other days, the office is of semidouble rite.}

\smalltitle{First Antiphon. 3. a2.}
\initialscore{an_dum_complerentur_solesmes_1961}

\antiphona{english}{When the day of Pentecost was fully come they were all with one accord in one place, alleluia}

\rubrique{\begin{enpars}
Sit at the word ``Sede.'' The cantor sings odd verses; all sing even verses, and this for the hymn and Magnificat, when sung entirely in chant, as well.
\end{enpars}}

\psalmus{109}{109_3a2}{109_3a2_g}{109_en}

\rubrique{\begin{enpars}
Clerics in descending order, or a cantor, intone the subsequent antiphons}}

 \gscore[2. Ant.]{8. G}{an_spiritus_domini_replevit_solesmes}
 \antiphona{english}{The Spirit of the Lord filleth the world, alleluia}
\psalmus{110}{110_8G}{110_8G}{110_en}

  \gscore[3. Ant.]{8. G}{an_repleti_sunt_omnes_solesmes_1961}
  \antiphona{english}{They were all filled with the Holy Ghost, and began to speak, alleluia, alleluia}
 \psalmus{111}{111_8G}{111_8G}{111_en}

\gscore[4. Ant.]{1. a3}{an_fontes_et_omnia_solesmes_1961}
\antiphona{english}{See ye fountains, and all that move in the waters, ascribe ye praise to God, alleluia}
\psalmus{112}{112_1a3}{112_1a3}{112_en}

 \gscore[5. Ant.]{7. c2}{an_loquebantur_ant_solesmes_1961}
\antiphona{english}{The Apostles spoke in diverse tongues the wonderful works of God, alleluia, alleluia, alleluia}
\psalmus{113}{113_7c2}{113_7c2}{113_en}

\myrule
\smalltitle{At I Vespers.}
\psalmus{116}{116_7c2}{116_7c2}{116_en}
\myrule

\rubrique{\begin{enpars}
Having sung the last antiphon in full, stand when the celebrant does.
\end{enpars}}

\bigtitle{Chapter.}

 \scripture{Act. 2, 1 – 2.}

\texteparacol{\lettrine{C}{u}m compleréntur dies Pentecóstes, erant omnes discípuli páriter in eódem loco:~† et factus est repénte de cælo sonus, tamquam adveniéntis spíritus veheméntis,~* et replévit totam domum, ubi erant sedéntes.}{\noindent When the day of Pentecost was fully come, the disciples were all with one accord in one place and suddenly there came a sound from heaven, as of a rushing mighty wind and it filled all the house where they were sitting.}

\hymnlateng

\rubrique{Kneel for the first verse, from \normaltext{Veni Creátor} to \normaltext{pectóra} inclusive.}

\gscore[]{8.}{hy_veni_creator_spiritus_solesmes}

\begin{multicols}{2}
\begin{enpars}
\raggedcolumns
Come, creator Spirit,\\
visit the souls of thine own,\\
fill with heavenly grace\\
the breasts that thou hast created.

Thou art called Paraclete,\\
gift of the most high God,\\
living water, flame, charity, and spiritual anointing.

Thou, who art sevenfold in thy gifts,\\
finger of God's right hand,\\
thou who wast rightly promised by the Father,\\
enrich our throats with speech.

Inflame the light of our senses,\\
pour love into our hearts;\\
the weakness of our bodies\\
strengthen with lasting power.

Drive the enemy far back,\\
and at once grant us peace:\\
with thou going ahead of us,\\
may we avoid all harm.

Through thee may we know the Father,\\
and recognize the Son;\\
and may we always believe in thee, Spirit of both.

To God the Father be glory,\\
and to the Son, who from death rose again,\\
and to the Holy Spirit, forever and ever. Amen.

\end{enpars}
\end{multicols}

\textes{cantor_rubric}

\myrule

\smalltitle{At I Vespers.}
   \smallscore{versiculus_i_vesperis_pent}
   
   \begin{enpars} \noindent \vv The Apostles spoke in diverse tongues, alleluia.

\noindent \rr  The wonderful works of God, alleluia.\end{enpars}
\myrule

%\rubrique{On the feast, the solemn tone is used, but on the days within the octave, the ordinary tone is used:}
%\label{versiculus}\phantomsection
   \smallscore{versiculus_ii_vesperis_pent}
   
   \begin{enpars} \noindent \vv They were all filled with the Holy Ghost, alleluia.

\noindent \rr And began to speak, alleluia.\end{enpars}
 
\smalltitle{Antiphon at the Magnificat.}

\rubrique{Sit when the celebrant does, after he sings the first words. Rise at the end. Otherwise, remain standing.}

\myrule

\smalltitle{At I Vespers.}

%\label{an_pater_manifestavi_solesmes_1961} \phantomsection
\gscore[]{1. D}{an_non_vos_relinquam_solesmes_1961}
 
 \antiphona{english}{I will not leave you orphans, alleluia; I go away, and come again unto you, alleluia, and your heart shall rejoice, alleluia}
 
 \rubrique{The canticle is as at II Vespers, \normaltext{\pageref{mag_1D}.}}

\myrule
%
% 
 \smalltitle{At II Vespers.}
 \gscore[]{1. D}{an_hodie_completi_sunt_solesmes_1961}
% 
 \antiphona{english}{This day the day of Pentecost is fully come, alleluia. This day the Holy Ghost appeared in fire unto the disciples, and gave unto them gifts of grace. He sent them into all the world, to preach and to testify that he who believeth, and is baptized, shall be saved, alleluia}
% 
\phantomsection
 \label{mag_1D}
% 
 \smallscore{Magnificat_1D_solemnis}

   \smallskip%
    \noindent%
    \begin{enumerate}[%
    		label=\arabic*., %
			leftmargin=0pt, %
			itemindent=0mm, %
			labelsep=3pt, %
			labelwidth=0pt, %
			rightmargin=0pt, %
			parsep=0pt, %
			topsep=0pt, %
			itemsep=0pt,
			start=3]%
    \input{psaumes/Magnificat_1_Solemnis}
    \end{enumerate}

  

\smalltitle{Canticle of the Blessed Virgin Mary.}

\scripture{Luc. 1, 46–55.}

\begin{multicols}{2}
 \begin{enpars}
My soul \cc~* doth magnify the Lord.

And my spirit hath rejoiced~* in God my Saviour.

Because he hath regarded the humility of his handmaid;~* for behold from henceforth all generations shall call me blessed.

Because he that is mighty, hath done great things to me;~* and holy is his name.

And his mercy is from generation unto generations,~* to them that fear him.

He hath shewed might in his arm:~* he hath scattered the proud in the conceit of their heart.

He hath put down the mighty from their seat,~* and hath exalted the humble.

He hath filled the hungry with good things;~* and the rich he hath sent empty away.

He hath received Israel his servant,~* being mindful of his mercy:

As he spoke to our fathers,~* to Abraham and to his seed for ever.

Glory be to the Father, and to the Son,~* and to the Holy Ghost.

As it was in the beginning, is now,~* and ever shall be, world without end. Amen.
\end{enpars}
% \end{paracol}
\end{multicols}

\rubrique{\begin{enpars}
Sit as the antiphon is repeated. Then the usual dialogue follows. All stand from now until the end of the office.\end{enpars}}

\textebilingue{dv}{dv_eng}

%\textebilingue{oratio_15_aug_old}{oratio_15_aug_old_en}

   \begin{paracol}{2}
   Orémus.
   
\label{oratio_dom_pent}\phantomsection

\lettrine{D}{e}us, qui hodiérna die corda fidélium Sancti Spíritus illustratióne docuísti:~† da nobis in eódem Spíritu re\-cta sápere,~* et de ejus semper consolatióne gaudére. Per Dóminum…in unitáte ejúsdem Spíritus Sancti Deus.

   ℟. Amen.
   
\switchcolumn
\begin{enpars}
\noindent Let us pray.
O God, Who on this day didst teach the hearts of thy faithful people, by the sending to them the light of thine Holy Spirit, grant us by the same Spirit to have a right judgment in all things, and evermore to rejoice in his holy comfort.
Through Jesus Christ thy Son our Lord, who liveth and reigneth with thee in the unity of the same Holy Ghost, God, world without end.
 Amen.
\end{enpars}
    \end{paracol}
    
%\rubrique{During the octave, the office is the same except for the Magnificat antiphon and the collect.}

\textebilingue{dv}{dv_eng}

\textes{cantor_rubric}

\myrule

\smalltitle{At I Vespers.}

\gscore[]{2.}{BD_Solemnis_I_Vesp}

\myrule

%\gscore[]{5.}{BD_Solemnis_mode_6}
%
%\rubrique{Or:}

\gscore[]{5.}{BD_Solemnis_mode_5}

\noindent \vv Let us bless the Lord.

\noindent \rr Thanks be to God.

\textebilingue{fidelium_animae}{fidelium_animae_en}

%\rubrique{During the octave, the office is the same except for the Magnificat antiphon and the collect.}

\end{document}