% !TEX TS-program = lualatexmk
% !TEX parameter =  --shell-escape

\documentclass[vespers_book_la_eng.tex]{subfiles}

\ifcsname preamble@file\endcsname
  \setcounter{page}{\getpagerefnumber{M-vespers_book_la_eng_ascension_pentecost}}
\fi

\begin{document}

% \chapter*{FEAST OF THE ASCENSION.}

%\litdateen{Thursday after the V Sunday after Easter}
\festum{feast of the ascension}
\addcontentsline{toc}{chapter}{Feast of the Ascension through the octave of Pentecost}\phantomsection
\addcontentsline{toc}{section}{Feast of the Ascension}\phantomsection
\header{feast of the ascension through the octave of pentecost.}

\doubleclass{I}[with a Privileged Octave of the III Order]
 
\atvespers{i}
\firstantiphon{7. a}
\initialscore{an_viri_galilaei_solesmes_1961}
\antiphona{english}{Ye men of Galilee, why stand ye gazing up into heaven? This same Jesus, who is taken up from you into heaven, shall so come in like manner, alleluia}
\psalmus{109}{109_7a}{109_7}{109_en}

 \gscore[2. Ant.]{8. G*}{an_cumque_intuerentur_solesmes_1961}
 \antiphona{english}{And while they looked steadfastly towards heaven, as He went up, they said, alleluia}
\psalmus{110}{110_8Gstar}{110_8}{110_en}

  \gscore[3. Ant.]{4. A1*}{an_elevatis_manibus_solesmes_1961}
  \antiphona{english}{He lifted up his Hands, and blessed them, and was carried up into heaven, alleluia}
 \psalmus{111}{111_4A_A_star}{111_4A_A_star}{111_en}

\gscore[4. Ant.]{8. G*}{an_exaltate_regem_solesmes_1961}
\antiphona{english}{Extol the King of kings, and ascribe praise to God, alleluia}
\psalmus{112}{112_8Gstar}{112_8}{112_en}

 \gscore[5. Ant.]{8. G}{an_videntibus_illis_solesmes_1961}
\antiphona{english}{While they beheld, He was taken up, and a cloud received him in heaven, alleluia}
\psalmus{116}{116_8G}{116_8}{116_en}

\chapterlateng
 \scripture{Act. 1, 1 – 2.}

\texteparacol{\lettrine{P}{r}imum quidem sermónem feci de ómnibus, o Theóphile,~† quæ cœpit Jesus fácere et docére usque in diem, qua præcípiens Apóstolis per Spíritum San\-ctum, quos elégit,~* assúmptus est}{english}{\noindent The former treatise I made, O Theophilus, of all things which Jesus began to do and to teach, Until the day on which, giving commandments by the Holy Ghost to the apostles whom he had chosen, he was taken up.}

\hymnlateng
\label{hy_salutis_humanae_sator_solesmes_1961}
\gscore[Ad Magnif.]{4.}{hy_salutis_humanae_sator_solesmes_1961}
\textes{salutis_humanae_sator_en.tex}

\textes{octave_versiculus_rubrique}

\label{ascendit}\phantomsection
   \smallscore{versiculus_in_vesperis_ascensionis}
   
\begin{enpars}\noindent \vv God is ascended with jubilee, alleluia.

\noindent \rr  And the Lord with the sound of trumpet, alleluia.
\end{enpars}

\label{an_pater_manifestavi_solesmes_1961} \phantomsection
\gscore[Ad Magnif.]{Ant. 6. F}{an_pater_manifestavi_solesmes_1961}
 
 \antiphona{english}{Father, I have manifested thy Name unto the men which Thou gavest Me but now I pray for them, not for the world, because I come to thee, alleluia}
 
 \label{oratio_ascensionis}\phantomsection
 
 \collectlateng
\texteparacol
{\lettrine{C}{o}ncéde, quǽsumus omnípotens Deus:~† ut qui hodiérna die Unigénitum tuum Redemptórem nostrum ad cælos ascendísse crédimus,~* i\-psi quoque mente in cæléstibus habitémus. Per eúmdem Dóminum.}{english}{
\noindent Let us pray.
Grant, we beseech thee, Almighty God, that just as we do believe thine Only-Begotten Son, our Saviour, to have this day ascended into the heavens, so we may also in heart and mind thither ascend, and with him continually dwell.
Through the same Lord.
}
 
\atvespers{ii}
 
 \allexcept
 
 \label{o_rex_gloriae}\phantomsection
 \gscore[Ad Magnif.]{Ant. 2. D}{an_o_rex_gloriae_solesmes_1961}
 
 \antiphona{english}{O King of glory, Lord of hosts, who hast this day exalted thine own self, with great triumph, above all the heavens, leave us not orphans but send unto us the promise of the Father, even the Spirit of truth, alleluia}

    \bigtitle{Saturday within the octave of the Ascension.}
\addcontentsline{toc}{section}{Saturday within the octave of the Ascension} \phantomsection
    
    \label{cap_dom_ascen}
    \chapterlateng
    \scripture{I Petr. 4, 7 – 8.}

\texteparacol{\lettrine{C}{a}ríssimi: Estóte prudéntes, et vigiláte in oratiónibus.~† Ante ómnia autem mútuam in vobismetípsis caritátem contínuam habéntes:~* quia cáritas óperit multitúdinem peccatórum.}{english}{\noindent Beloved: Be prudent, therefore, and watch in prayers. But before all things have a constant mutual charity among yourselves: for charity covereth a multitude of sins.}

\rubrique{Hymn \normaltext{Salútis humanæ sator, \pageref{hy_salutis_humanae_sator_solesmes_1961}.}}

\label{versiculus_dom_ascenscionis}
 \smallscore{versiculus_dom_ascenscionis}

\begin{enpars}
\noindent \vv The Lord in heaven, alleluia.

\noindent \rr  Hath prepared his throne, alleluia.
 \end{enpars}
 
     \gscore[Ad Magnif.]{Ant. 8. G}{an_cum_venerit_paraclitus_solesmes_1961}
    \antiphona{english}{But when the Comforter is come whom I will send unto you from the Father, He shall testify of Me, alleluia}
   
\label{oratio_dom_ascensionis}\phantomsection

\collectlateng
\texteparacol
{\lettrine{O}{m}nípotens sempitérne Deus:~† fac nos tibi semper et devótam gérere voluntátem,~* et majestáti tuæ sincéro corde servíre. Per Dóminum.}{english}
{\noindent Let us pray.
O Almighty and everlasting God, grant that our will be ever meekly subject unto thy will, and our heart ever honestly ready to serve thy majesty.
Through our Lord.}
 
     \rubrique{On Saturday, then the preceding day is commemorated, ant. \normaltext{O Rex glóriæ, \pageref{o_rex_gloriae}. ℣. Ascendit, \pageref{ascendit},} Oratio \normaltext{Concéde, \pageref{oratio_ascensionis}.}}
     
         \bigtitle{Sunday within the octave of the Ascension.}
\addcontentsline{toc}{section}{Sunday within the octave of the Ascension} \phantomsection

\rubrique{Chapter \normaltext{Estóte prudéntes, as on Saturday, \pageref{cap_dom_ascen}.}}

\rubrique{Hymn \normaltext{Salútis humanæ sator, \pageref{hy_salutis_humanae_sator_solesmes_1961}.}}

\rubrique{℣. \normaltext{Dóminus, \pageref{versiculus_dom_ascenscionis}.}}

     \gscore[Ad Magnif.]{Ant. 8. G}{an_haec_locutus_sum_vobis_solesmes_1961}
        \antiphona{english}{These things have I told you, that, when the time shall come, ye may remember that I told you of them, alleluia}
        
        \rubrique{Collect \normaltext{Omnípotens sempitérne Deus,} as above, \normaltext{\pageref{oratio_dom_ascensionis}.}}

\rubrique{\begin{enpars}On Sunday, then the following day is commemorated, ant. \normaltext{Pater manifestávi, \pageref{an_pater_manifestavi_solesmes_1961}. ℣. Ascendit, \pageref{ascendit},} collect \normaltext{Concéde, \pageref{oratio_ascensionis}.}\end{enpars}}

\rubrique{\begin{enpars}If however a feast of IX lessons is instead celebrated, then the preceding day is commemorated, ant. \normaltext{O Rex glóriæ, \pageref{o_rex_gloriae}. ℣. Ascendit, \pageref{ascendit},} collect \normaltext{Concéde, \pageref{oratio_ascensionis}.}\end{enpars}}

\rubrique{\begin{enpars}Feasts of III lessons are kept as commemorations only from now until Pentecost, and the office is as on the Ascension, unless a feast of IX lessons is celebrated.\end{enpars}}

\rubrique{\begin{enpars} On Wednesday, all is as at I Vespers, and the office is of double major rite. On Thursday, all is as at II Vespers, and no commemorations of the following day are made.\end{enpars}}

\festum{pentecost}
\addcontentsline{toc}{section}{Pentecost}\phantomsection
\doubleclass{I}[with a Privileged Octave of the I Order]

\atvespers{i}

\lastplace{II}{116}

\psalmus{116}{116_7c2}{116_7}{116_en}

\label{cap_pent}

\chapterlateng

 \scripture{Act. 2, 1 – 2.}

\texteparacol{\lettrine{C}{u}m compleréntur dies Pentecóstes, erant omnes discípuli páriter in eódem loco:~† et factus est repénte de cælo sonus, tamquam adveniéntis spíritus veheméntis,~* et replévit totam domum, ubi erant sedéntes.}{english}{\noindent When the day of Pentecost was fully come, the disciples were all with one accord in one place and suddenly there came a sound from heaven, as of a rushing mighty wind and it filled all the house where they were sitting.}

\hymnlateng \label{veni_creator}

\rubrique{Kneel for the first verse, from \normaltext{Veni Creátor} to \normaltext{pectóra} inclusive.}

\gscore[]{8.}{hy_veni_creator_spiritus_solesmes}

\textes{veni_creator_en}

   \smallscore{versiculus_i_vesperis_pent}
   
   \begin{enpars} \noindent \vv The Apostles spoke in diverse tongues, alleluia.

\noindent \rr  The wonderful works of God, alleluia.\end{enpars}

\gscore[Ad Magnif.]{Ant. 1. D}{an_non_vos_relinquam_solesmes_1961}
 
\antiphona{english}{I will not leave you orphans, alleluia; I go away, and come again unto you, alleluia, and your heart shall rejoice, alleluia}

\collectlateng
\label{or_dom_pent}

\texteparacol{\lettrine{D}{e}us, qui hodiérna die corda fidélium Sancti Spíritus illustratióne docuísti:~† da nobis in eódem Spíritu re\-cta sápere,~* et de ejus semper consolatióne gaudére.
Per Dóminum…in unitáte ejúsdem Spíritus Sancti Deus.}{english}
{\noindent O God, who on this day didst teach the hearts of thy faithful people, by the sending to them the light of thine Holy Spirit, grant us by the same Spirit to have a right judgment in all things, and evermore to rejoice in his holy comfort.
Through Jesus Christ…in the unity of the same Holy Ghost.}

\atvespers{ii}

\firstantiphon{3. a2}
\initialscore{an_dum_complerentur_solesmes_1961}

\antiphona{english}{When the day of Pentecost was fully come they were all with one accord in one place, alleluia}

\psalmus{109}{109_3a2}{109_3a2_g}{109_en}

 \gscore[2. Ant.]{8. G}{an_spiritus_domini_replevit_solesmes}
\antiphona{english}{The Spirit of the Lord filleth the world, alleluia}
\psalmus{110}{110_8G}{110_8}{110_en}

  \gscore[3. Ant.]{8. G}{an_repleti_sunt_omnes_solesmes_1961}
 \antiphona{english}{They were all filled with the Holy Ghost, and began to speak, alleluia, alleluia}
 \psalmus{111}{111_8G}{111_8}{111_en}

\gscore[4. Ant.]{1. a3}{an_fontes_et_omnia_solesmes_1961}
\antiphona{english}{See ye fountains, and all that move in the waters, ascribe ye praise to God, alleluia}
\psalmus{112}{112_1a3}{112_1}{112_en}

 \gscore[5. Ant.]{7. c2}{an_loquebantur_ant_solesmes_1961}
\antiphona{english}{The Apostles spoke in diverse tongues the wonderful works of God, alleluia, alleluia, alleluia}
\psalmus{113}{113_7c2}{113_7}{113_en}

%\myrule
%\smalltitle{At I Vespers.}
%\psalmus{116}{116_7c2}{116_7c2}{116_en}
%\myrule

%\allexcept

\caphymnen{I}

   \smallscore{versiculus_ii_vesperis_pent}
   
   \begin{enpars}\noindent \vv They were all filled with the Holy Ghost, alleluia.

\noindent \rr And began to speak, alleluia.\end{enpars}

 \gscore[Ad Magnif.]{1. D}{an_hodie_completi_sunt_solesmes_1961}
\antiphona{english}{This day the day of Pentecost is fully come, alleluia. This day the Holy Ghost appeared in fire unto the disciples, and gave unto them gifts of grace. He sent them into all the world, to preach and to testify that he who believeth, and is baptized, shall be saved, alleluia}

\rubrique{Collect \normaltext{Deus, qui hodiérna, \pageref{or_dom_pent}} as at I Vespers.}

\rubrique{Within the octave of Pentecost, the office is said as on the day, except for the antiphon at the Magnificat and the collect. On Monday and Tuesday, the office is of dobule rite; on the other days, semidouble.}

\bigtitle{Pentecost Monday.}
\addcontentsline{toc}{section}{Pentecost Monday}\phantomsection

\gscore[Ad Magnif.]{Ant. 3. a}{an_si_quis_diligit_me_solesmes_1961}
\antiphona{english}{If a Man love Me he will keep My word and My Father will love him, and We will come unto him, and make Our abode with him. Alleluia}

\collectlateng

\texteparacol{\lettrine{D}{e}us, qui Apóstolis tuis San\-ctum dedísti Spíritum:~† concéde plebi tuæ piæ petitiónis efféctum:~* ut quibus dedísti fidem, largiáris et pacem.
Per Dóminum…in unitáte ejúsdem Spíritus.}{english}{\noindent O God, who didst give thine Holy Spirit unto thine Apostles, grant unto thy people the effectual fruit of their godly supplications, and as thou hast given them faith, give them also peace. Through Jesus Christ…in the unity of the same Holy Ghost.}

\bigtitle{Pentecost Tuesday.}
\addcontentsline{toc}{section}{Pentecost Tuesday}\phantomsection

\gscore[Ad Magnif.]{Ant. 6. F}{an_pacem_relinquo_vobis_solesmes_1961}
\antiphona{english}{Peace I leave with you, My peace I give unto you; not as the world giveth, give I unto you. Alleluia}

\collectlateng

\texteparacol{\lettrine{A}{d}sit nobis, quǽsumus, Dómine, virtus Spíritus San\-cti:~† quæ et corda nostra cleménter expúrget,~* et ab ómnibus tueátur advérsis. Per Dóminum…in unitáte ejúsdem Spíritus.}{english}{\noindent O Lord, we beseech thee, that thy Holy Spirit may dwell in us in much power, mercifully cleansing our hearts, and shielding us from all things hurtful. Through Jesus Christ…in the unity of the same Holy Ghost.}

\bigtitle{Ember Wednesday.}
\addcontentsline{toc}{section}{Ember Wednesday}\phantomsection

%%Vaticana has cœlo, LA1949 caeli, we will go with cæli 

\gscore[Ad Magnif.]{Ant. 1. f}{an_ego_sum_panis_vivus_solesmes_1961}
\antiphona{english}{I am the Living Bread Which came down from heaven; if any man eat of this Bread, he shall live for ever, and the Bread that I will give is My Flesh, Which I will give for the life of the world. Alleluia}

\collectlateng

\texteparacol{\lettrine{M}{e}ntes nostras quǽsumus Dómine, Paráclitus, qui a te procédit, illúminet:~* et indúcat in omnem, sicut tuus promísit Fílius veritátem. Qui tecum vivit et regnat in unitáte ejúsdem Spíritus.}{english}{\noindent O Lord, we beseech thee, that the Comforter, which proceeded from thee, may enlighten our minds, and lead us into all truth, even as thy Son hath promised unto us.
Who with thee liveth and reigneth in the unity of the same Holy Ghost.}

%%Solesmes does not have † here

\bigtitle{Pentecost Thursday.}
\addcontentsline{toc}{section}{Pentecost Thursday}\phantomsection

\gscore[Ad Magnif.]{Ant. 8. G*}{an_spiritus_qui_solesmes_1961}
\antiphona{english}{The Comforter, Which proceedeth from the Father, alleluia, He shall glorify Me, alleluia, alleluia}

\rubrique{Collect as on the day, \normaltext{\pageref{or_dom_pent}.}}

%%reprinted in antiphonal, but a label with pageref saves space.

\bigtitle{Ember Friday.}
\addcontentsline{toc}{section}{Ember Friday}\phantomsection

\gscore[Ad Magnif.]{Ant. 8. G}{an_paraclitus_autem_solesmes_1961}
\antiphona{english}{\noindent But the Comforter which is the Holy Ghost, whom the Father will send in my name, He shall teach you all things, and bring all things to your remembrance, whatsoever I have said unto you. Alleluia}

\collectlateng

\texteparacol{\lettrine{D}{a}, quǽsumus Ecclésiæ tuæ miséricors Deus:~† ut San\-cto Spíritu congregáta,* hostíli nullátenus incursióne turbétur. Per Dóminum…in unitáte ejúsdem Spíritus.}{english}
{\noindent O most merciful God, we beseech thee to grant unto thy Church that, forasmuch as she is gathered together in the Holy Ghost, no assaults of her enemies may be able to trouble her. Through Jesus Christ…in the unity of the same Holy Ghost.}

\end{document}