% !TEX TS-program = lualatexmk
% !TEX parameter =  --shell-escape

\documentclass[vespers_book_la_eng.tex]{subfiles}

\ifcsname preamble@file\endcsname
  \setcounter{page}{\getpagerefnumber{M-vespers_book_la_eng_sundays_tp}}
\fi

\begin{document}

 \chapter*{SUNDAYS AFTER EASTER.}
\addcontentsline{toc}{chapter}{Sundays after Easter.}\phantomsection
  \header{sundays  after easter.}

\rubrique{Antiphon and psalms of Sundays in Paschal Time, \normaltext{\pageref{M-pss_tp}.}}

\bigtitle{I Sunday after Easter.}
\addcontentsline{toc}{section}{I Sunday after Easter} \phantomsection

\doublemajor

\rubrique{All as in the psalter of Sundays in Paschal Time, \normaltext{\pageref{M-pss_tp},} except as given below.}

\chapterlateng

\scripture{1 Joann. 5, 4.}
\texteparacol{\lettrine{C}{a}ríssimi: Omne quod natum est ex Deo, vincit mundum:~† et hæc est victória, quæ vincit mundum,~* fides nostra.}
{english}{\noindent Beloved: For whatsoever is born of God, overcometh the world: and this is the victory which overcometh the world, our faith.}

\hymnlateng

\label{hy_ad_regias_agni_dapes} \phantomsection
\gscore[]{8.}{hy_ad_regias_agni_dapes_solesmes_1961}

\textes{ad_regias_en}

\anothertone
%\grechangenextscorelinedim{3}{spacebeneathtext}{0.25cm}{scalable}
\grechangenextscorelinedim{3}{spacebeneathtext}{0.15cm}{scalable}
\grechangenextscorelinedim{4,7,10}{spacebeneathtext}{0.05cm}{scalable}
%\grechangenextscorelinedim{4,7,10}{spacebeneathtext}{0.03cm}{scalable}

\gscore[]{4.}{hy_ad_regias_agni_dapes_another_chant_solesmes_1961}

\label{versiculus_tp} \phantomsection
\smallscore{versiculus_tp}

\begin{enpars} \noindent \vv  Stay with us, O Lord, alleluia.

\noindent \rr  Because it is towards evening, alleluia.
\end{enpars}

\gscore[Ad Magnif.]{Ant. 8. c}{an_post_dies_octo_solesmes_1961}
\antiphona{english}{After eight days, the doors being shut, the Lord being entered in, said unto them: Peace be unto you, alleluia}

\collectlateng

\texteparacol{\lettrine{P}{r}æsta, quǽsumus omnípotens Deus:~† ut qui paschália festa perégimus,~* hæc, te largiénte, móribus et vita teneámus. Per Dóminum.}{english}
{\noindent Grant, we beseech thee, almighty God, that we who have celebrated the paschal feast, may, by thy bounty, retain its fruits in our daily habits and behaviour.
Through our Lord.}

\bigtitle{II Sunday after Easter.}
\addcontentsline{toc}{section}{II Sunday after Easter} \phantomsection

\chapterlateng

\scripture{1 Petri 2, 21 – 22.}

\texteparacol{\lettrine{C}{a}ríssimi: Christus passus est pro nobis,~† vobis relínquens exémplum ut sequámini vestígia eius.~* Qui peccátum non fecit, nec invéntus est dolus in ore ejus.}{english}{\noindent Dearly beloved: Christ also suffered for us, leaving you an example that you should follow his steps. Who did no sin, neither was guile found in his mouth.}

\rubrique{Hymnus \normaltext{Ad régias Agni dapes, \pageref{hy_ad_regias_agni_dapes}.} \normaltext{℣. Mane nobíscum, \pageref{versiculus_tp}.}}

\gscore[Ad Magnif.]{Ant. 3. a}{an_ego_sum_pastor_bonus_solesmes_1961}
\antiphona{english}{I am the Good Shepherd, who feed my sheep: and I lay down my life for my sheep. Alleluia}

\collectlateng

\texteparacol{\lettrine{D}{e}us, qui in Fílii tui humilitáte jacéntem mundum erexísti:~† fidélibus tuis perpétuam concéde lætítiam;~* ut, quos perpétuæ mortis eripuísti cásibus, gáudiis fácias pérfrui sempitérnis. Per eúmdem Dóminum.}{english}
{\noindent O God, whose Son hath humbled himself, and who hast through him raised up the whole world, grant to thy faithful people everlasting joy; and as Thou hast delivered them from the bitter pains of eternal death, make them to be glad for ever in thy presence.
Through the same Lord.}

\bigtitle{III Sunday after Easter.}
\addcontentsline{toc}{section}{III Sunday after Easter} \phantomsection

\chapterlateng

\scripture{1 Petri 2, 11.}
\texteparacol{\lettrine{C}{a}ríssimi: Obsecro vos tamquam ádvenas et peregrínos,~† abstinére vos a carnálibus desidériis,~* quæ mílitant advérsus ánimam.}{english}{\noindent Dearly beloved, I beseech you as strangers and pilgrims, to refrain yourselves from carnal desires which war against the soul.}

\rubrique{Hymnus \normaltext{Ad régias Agni dapes, \pageref{hy_ad_regias_agni_dapes}.} \normaltext{℣. Mane nobíscum, \pageref{versiculus_tp}.}}

\gscore[Ad Magnif.]{Ant. 8. G}{an_amen_amen_quia_plorabitis_solesmes}
\antiphona{english}{Amen, amen I say to you that you shall lament and weep, but the world shall rejoice; and you shall be made sorrowful, but your sorrow shall be turned into joy, alleluia}

\collectlateng

\texteparacol{\lettrine{D}{e}us, qui errántibus, ut in viam possint redíre justítiæ, veritátis tuæ lumen osténdis:~† da cun\-ctis qui christiána professióne censéntur, et illa respúere quæ huic inimíca sunt nómini;~* et ea quæ sunt apta sectári.
Per Dóminum nostrum.}{english}
{\noindent Almighty God, who showest to them that be in error the light of thy truth, to the intent that they may return into the way of righteousness grant unto all them that are admitted into the fellowship of Christ's religion, that they may eschew those things that are contrary to their profession, and follow all such things as are agreeable to the same.
Through our Lord.}

\bigtitle{IV Sunday after Easter.}
\addcontentsline{toc}{section}{IV Sunday after Easter} \phantomsection

\chapterlateng

\scripture{Jac. 1, 17.}
\texteparacol{\lettrine{C}{a}ríssimi: Omne datum óptimum, et omne donum perféctum desúrsum est, descéndens a Patre lúminum,~† apud quem non est transmutátio,~* nec vicissitúdinis obumbrátio.}{english}{\noindent Dearly beloved: Every best gift, and every perfect gift, is from above, coming down from the Father of lights, with whom there is no change, nor shadow of alteration.}

\rubrique{Hymnus \normaltext{Ad régias Agni dapes, \pageref{hy_ad_regias_agni_dapes}.} \normaltext{℣. Mane nobíscum, \pageref{versiculus_tp}.}}

\gscore[Ad Magnif.]{Ant. 2. D}{an_vado_ad_eum_sed_quia_solesmes_1961}
\antiphona{english}{I go my way to him That sent Me, but because I have said this to you, sadness has filled your hearts, alleluia}

\collectlateng

\texteparacol{\lettrine{C}{o}ncéde, quǽsumus omnípotens Deus:~† ut, qui ex mérito nostræ actiónis afflígimur,* tuæ grátiæ consolatióne respirémus.
Per Dóminum.}{english}
{\noindent Grant, we beseech thee, Almighty God, that we who for our evil deeds are worthily punished, may, by the comfort of thy grace, mercifully be relieved.
Through our Lord.}

\bigtitle{V Sunday after Easter.}
\addcontentsline{toc}{section}{V Sunday after Easter} \phantomsection

\chapterlateng

\scripture{Jac. 1, 22 – 24.}

\texteparacol{\lettrine{C}{a}ríssimi: Estóte factóres verbi, et non auditóres tantum: falléntes vosmetípsos.~† Quia si quis audítor est verbi, et non factor: hic comparábitur viro consideránti vultum nativitátis suæ in spéculo:~* considerávit enim se, et ábiit, et statim oblítus est qualis fúerit.}{english}{\noindent Dearly beloved: But be ye doers of the word, and not hearers only, deceiving your own selves. For if a man be a hearer of the word, and not a doer, he shall be compared to a man beholding his own countenance in a glass. For he beheld himself, and went his way, and presently forgot what manner of man he was.}

\rubrique{Hymnus \normaltext{Ad régias Agni dapes, \pageref{hy_ad_regias_agni_dapes}.} \normaltext{℣. Mane nobíscum, \pageref{versiculus_tp}.}}

\gscore[Ad Magnif.]{Ant. 8. G*}{an_petite_et_accipietis_solesmes_1961}
\antiphona{english}{Ask, and ye shall receive, that your joy may be full; for the Father himself loveth you, because ye have loved Me, and have believed in Me, alleluia}

\collectlateng

\texteparacol{\lettrine{D}{e}us, a quo bona cuncta procédunt, largíre supplícibus tuis:~†  ut cogitémus te inspiránte quæ recta sunt;~* et, te gubernánte éadem faciámus. Per Dóminum nostrum.}{english}
{\noindent O God, from whom all good things do come, grant to us thy humble servants that by thy holy inspiration we may think those things that be good, and by thy merciful guiding may perform the same.
Through our Lord.}

\end{document}