% !TEX TS-program = lualatexmk
% !TEX parameter =  --shell-escape


\documentclass[12pt]{book}

\input{commonheaders_vespers_people}

\begin{document}
 
 \twosided[pc] %% to allow for the left-right switching of columns needed for bilingual liturgical typesetting (like in missals); it's unclear why this goes BEFORE everything else and not when activating the environment. %% It's OK for paper booklets to not switch, but it's probably best practice to switch.
 
 \thispagestyle{empty}
 
 {\centering{\huge\capspace{{SUNDAYS OF LENT AT VESPERS.}}}\par}
 
 \header{sundays of lent at vespers.}
 
\smalltitle{Before the Office.}

\rubrique{These prayers are said kneeling. End when the priest stands.}

\textebilingue{orationes_ante_do}{orationes_ante_do_eng}

\smalltitle{\prayers}

\rubrique{The celebrant intones the following verse. All sing the response and doxology.}

\begin{enpars}
\rubrique{Make the sign of the cross at the beginning and at the Magnificat, and bow the head when the Trinity is mentioned, when mention is made of the ``Holy Name'' in some way, and at ``Orémus.''}\end{enpars}

\gscore[]{℣.}{Deus_In_Adjutorium_laus_tibi}

\noindent
\begin{enpars}
O God, \cc{} come to my assistance. \rr O Lord, make haste to help me. Glory be to the Father, and to the Son,~* and to the Holy Ghost. As it was in the beginning, is now,~* and ever shall be, world without end. Amen. Praise be unto thee, o Lord, king of eternal glory.\end{enpars}

\smalltitle{First Antiphon. 7. c2.}
\initialscore{1_Dixit_Dominus}

\footnotetext[1]{The words are not repeated, which is always done when the antiphon begins with the first words of the psalm, which continues therefore after the word with which the antiphon ends, whether it was sung in full or simply begun. Regardless, if the words are the same, the antiphon is continued with the psalm.}

\rubrique{\begin{enpars}
Sit at the word ``Sede.'' The cantor sings odd verses; all sing even verses, and this for the hymn and Magnificat as well.
\end{enpars}}

\sdpsalmusbilingue{109}{109_7}{109_en}

\smallscore{an--dixit_dominus_domino--solesmes_1961}

\antiphona{english}{The Lord said to my Lord: Sit thou at my right hand.}

\rubrique{\begin{enpars}
Clerics in descending order, or a cantor, intone the subsequent antiphons. The cantor then continues the psalm.
\end{enpars}}

\gscore[2. Ant.]{3. b}{2_Magna_Opera}

\psalmus{110}{110_3b}{110_3a_b}{110_en}

\smallscore{an--magna_opera_domini--solesmes_1961}

\antiphona{english}{Great are the works of the Lord, sought out according to all his wills.}

\gscore[3. Ant.]{4. g}{3_Qui_timet_Dominum}

\psalmus{111}{111_4g}{111_4g}{111_en}

\smallscore{an--qui_timet_dominum--solesmes_1961}

\antiphona{english}{He who feareth the Lord shall delight exceedingly in his commandments.}

\gscore[4. Ant.]{7. c}{4_Sit_nomen_Domini}

\psalmus{112}{112_7c}{112_7}{112_en}

\smallscore{an--sit_nomen_domini..._in--solesmes_1961}

\antiphona{english}{Blessed be the name of the Lord forever.}

\gscore[5. Ant.]{T. pereg.}{5_Deus_autem_noster}

\psalmus{113}{113_per}{113_per}{113_en}

\smallscore{an--deus_autem_noster--solesmes_1961}

\antiphona{english}{But our God is in heaven, he hath done all things whatsoever he would.}


\rubrique{\begin{enpars}
Having sung the last antiphon in full, stand when the celebrant does.
\end{enpars}}

%\newpage

\chapterlateng

\smalltitle{I Sunday of Lent.}

\scripture{2 Cor. 6, 1 – 2.}
\texteparacol{\lettrine{F}{r}atres: Hortámur vos, ne in vácuum grátiam Dei recipiátis.~† Ait enim: Témpore accépto exaudívi te,* et in die salútis adjúvi te.}{english}{\noindent Brethren, we exhort you, that you receive not the grace of God in vain. For he saith: In an accepted time have I heard thee; and in the day of salvation have I helped thee.}

\smalltitle{II Sunday of Lent.}

\scripture{1 Thess. 4, 1.}

\texteparacol{\lettrine{F}{r}atres: Rogámus vos, et obsecrámus in Dómino Jesu:~† ut, quemádmodum accepístis a nobis, quómodo vos opórteat ambuláre, et placére Deo,* sic et ambulétis, ut abundétis magis.}{english}{\noindent For the rest therefore, brethren, we pray and beseech you in the Lord Jesus, that as you have received from us, how you ought to walk, and to please God, so also you would walk, that you may abound the more.}

\smalltitle{III Sunday of Lent.}

\scripture{Eph. 5, 1 – 2.}
\texteparacol{\lettrine{F}{r}atres: Estóte imitatóres Dei, sicut fílii caríssimi:~† et ambuláte in dilectióne, sicut et Christus diléxit nos, et trádidit semetípsum pro nobis~* oblatiónem et hóstiam Deo in odórem suavitátis.}{english}{\noindent Brethren: Be ye, therefore, followers of God, as most dear children; And walk in love, as Christ also hath loved us, and hath delivered himself for us, an oblation and a sacrifice to God for an odour of sweetness.}

\smalltitle{IV Sunday of Lent.}

\scripture{Gal. 4, 22 – 24.}
\texteparacol{\lettrine{F}{r}atres: Scriptum est quóniam Abraham duos fílios hábuit: unum de ancílla, et unum de líbera:~† sed qui de ancílla, secúndum carnem natus est: qui autem de líbera, per repromissiónem:~* quæ sunt per allegoríam dicta.}{english}{\noindent Brethren: For it is written that Abraham had two sons: the one by a bondwoman, and the other by a free woman. But he who was of the bondwoman, was born according to the flesh: but he of the free woman, was by promise, which things are said by an allegory.}

\smalltitle{Passion Sunday.}

\scripture{Heb. 9, 11 – 12.}

\texteparacol{\lettrine{F}{r}atres: Christus assístens Póntifex futurórum bonórum, per ámplius et perféctius tabernáculum non manu factum, id est, non huius creatiónis:~† neque per sánguinem hircórum aut vitulórum, sed per próprium sánguinem introívit semel in Sancta,~* ætérna redemptióne invénta.}{english}{\noindent Brethren: But Christ, being come an high priest of the good things to come, by a greater and more perfect tabernacle not made with hand, that is, not of this creation: Neither by the blood of goats, or of calves, but by his own blood, entered once into the holies, having obtained eternal redemption.}

\smalltitle{Palm Sunday.}

\scripture{Phil. 2, 5 – 7.}
\texteparacol{\lettrine{F}{r}atres: Hoc enim sentíte in vobis, quod et in Christo Jesu: qui, cum in forma Dei esset, non rapínam arbitrátus est esse se æquálem Deo:~† sed semetípsum exinanívit, formam servi accípiens, in similitúdinem hóminum factus,~* et hábitu invéntus ut homo.}{english}{\noindent Brethren: For let this mind be in you, which was also in Christ Jesus: Who being in the form of God, thought it not robbery to be equal with God: But emptied himself, taking the form of a servant, being made in the likeness of men, and in habit found as a man.}

%% The usual tone and the 3rd tone, used ad libitum, are omitted as we do not use these tones, but they are included in the tex file and the gabc files are available, for the sake of completeness.

\hymnlateng

\rubrique{This hymn is sung at Vespers of Lent, from Saturday before the I Sunday of Lent until the Saturday before Passion Sunday exclusive.}

%\gscore[]{8.}{hy--Lucis_Creator_optime}
%
%\rubrique{\begin{enpars}
%If, however, a commemoration is made of the Blessed Virgin Mary:
%\end{enpars}}
%
%\smallscore{hy--Lucis_Creator_BMV}
%
%\smalltitle{2. Another tone ad lib.}

\gscore[]{2.}{hy_audi_benigne_conditor_textus_antiquus}

\textes{audi_benigne_en}

\textes{cantor_rubric}

\smallscore{versiculus_quadragesimae}

\begin{enpars}\noindent \vv  God hath given his Angels charge over thee.

\noindent \rr To keep thee in all thy ways.\end{enpars}
 
\hymnlateng
 
 \rubrique{This hymn is sung at Vespers of Passiontide, from Saturday before Passion Sunday until Wednesday of Holy Week inclusive.}

\rubrique{\begin{enpars}All kneel for the entire strophe beginning \normaltext{O crux ave Spes única.}\end{enpars}}
 
 \gscore[]{1.}{hy_vexilla_regis_textus_antiquus}
 
\textes{vexilla_regis_en}

 \smallscore{versiculus_passionis}

\begin{enpars}\noindent \vv Deliver me, O Lord, from the evil man.

\noindent \rr Rescue me from the unjust man.\end{enpars}
 
 \rubrique{\begin{enpars}
The celebrant intones the \normaltext{Magnificat} antiphon; this is always proper to the day. The cantor continues the \normaltext{Ma\-gnificat.}\end{enpars}}

\newpage
\bigtitle{Canticle of the Blessed Virgin Mary.}

\smalltitle{I Sunday of Lent.}\label{1_sunday}

\gscore[Ad Magnif.]{Ant. 8. G*}{an_ecce_nunc_tempus_solesmes_1961}
\antiphona{english}{Behold, now is the accepted time behold, now is the day of salvation; in these days therefore let us approve ourselves as the ministers of God, in much patience, in fastings, in watchings, and in love unfeigned}

 \smallscore{Magnificat_8Gstar}
 
\magnificattexte{Magnificat_8}

\smalltitle{II Sunday of Lent.}

\gscore[Ad Magnif.]{Ant. 1. f}{an_visionem_solesmes_1961}
\antiphona{english}{Tell the vision that ye have seen to no man, until the Son of man be risen again from the dead}

 \smallscore{Magnificat_1f}
 
\magnificattexte{Magnificat_1_ordinaire}

\smalltitle{III Sunday of Lent.}

\gscore[Ad Magnif.]{Ant. 8. G}{an_extollens_solesmes_1961}
\antiphona{english}{A certain woman of the company lifted up her voice and said: Blessed is the womb that bare thee, and the paps which Thou hast sucked. But Jesus said unto her: Yea, rather, blessed are they that hear the word of God, and keep it}

 \smallscore{Magnificat_8G}
 
\magnificattexte{Magnificat_8}

\smalltitle{IV Sunday of Lent.}

\gscore[Ad Magnif.]{Ant. 1. g}{an_subiit_ergo_solesmes_1961}
\antiphona{english}{And Jesus went up into a mountain, and there He sat with his disciples}

 \smallscore{Magnificat_1g}
 
\magnificattexte{Magnificat_1_ordinaire}

\smalltitle{Passion Sunday.}

\gscore[Ad Magnif.]{Ant. 2. D}{an_abraham_pater_vester_solesmes_1961} 
\antiphona{english}{Your father Abraham rejoiced to see my day and he saw it, and was glad}

\smallscore{Magnificat_2}

\magnificattexte{Magnificat_2}

\smalltitle{Palm Sunday.}

\gscore[Ad Magnif.]{Ant. 8. G*}{an_scriptum_est_enim__percutiam_solesmes_1961}

\antiphona{english}{It is written: I will smite the Shepherd, and the sheep of the flock shall be scattered abroad, but after I am risen again, I will go before you into Galilee: there shall ye see Me, saith the Lord}

\rubrique{The tone is the same as on the I Sunday of Lent, \normaltext{\pageref{1_sunday}.}}

\textes{magnificat_en_multicols}

 \rubrique{\begin{enpars}
Sit as the antiphon is sung in full. Then the usual dialogue follows. On Sundays, all stand from now until the end of the office, unless one sits for the antiphons sung for the commemorations, according to custom.\end{enpars}}

\textes{dv_la_en}

\myrule

\rubrique{When the officiant is not a deacon:}

\smallscore{domine_exaudi_solemn}

\begin{enpars}
 \vv O Lord hear my prayer.

 \rr And let my cry come unto thee.
\end{enpars}

\myrule

 \rubrique{\begin{enpars}
The celebrant sings the collect of the day, and any commemorations follow. When required, the Suffrage of All Saints is sung in the last place, as below.\end{enpars}}

\bigtitle{Collect.}
\smallskip

\smalltitle{I Sunday of Lent.}

\texteparacol{\lettrine{D}{e}us, qui Ecclésiam tuam ánnua quadragesimáli observatióne puríficas:~† præsta famíliæ tuæ; ut quod a te obtinére abstinéndo nítitur,* hoc bonis opéribus exsequátur. Per Dóminum.}{english}
{\noindent O God, who dost every year purge thy church by the fast of forty days, grant unto this thy family, that what things soever they strive to obtain at thy hand by abstaining from meats, they may ever turn to profit by good works. Through our Lord.}
    
    \smalltitle{II Sunday of Lent.}

\texteparacol{\lettrine{D}{e}us, qui cónspicis omni nos virtúte destítui: intérius exteriúsque custódi;~† ut ab ómnibus adversitátibus muniámur in córpore,* et a právis cogitatiónibus mundémur in mente. Per Dóminum.}{english}
{\noindent O God, Who seest that we have no power of ourselves to help ourselves, keep us both outwardly in our bodies, and inwardly in our souls, that we may be defended from all adversities which may happen to the body, and from all evil thoughts which may assault and hurt the soul. Through our Lord.}
    
    \smalltitle{III Sunday of Lent.}

\texteparacol{ \lettrine[depth=1]{Q}{u}ǽsumus omnípotens Deus, vota humílium réspice:~† atque ad defensiónem nostram,* déxteram tuæ majestátis exténde.
Per Dóminum.}{english}
{\noindent We beeseech thee, almighty God, look upon the hearty desires of thy humble servants, and stretch forth the right hand of thy majesty to be our defense against all our enemies. Through our Lord.}
    
    \smalltitle{IV Sunday of Lent.}

\texteparacol{\lettrine{C}{o}ncéde, quǽsumus omnípotens Deus:~† ut, qui ex mérito nostræ actiónis afflígimur,* tuæ grátiæ consolatióne respirémus.
Per Dóminum.}{english}
{\noindent Grant, we beseech thee, Almighty God, that we who for our evil deeds are worthily punished, may, by the comfort of thy grace, mercifully be relieved. Through our Lord.}
    
    \smalltitle{Passion Sunday.}

\texteparacol{\lettrine[depth=1]{Q}{u}ǽsumus omnípotens Deus, famíliam tuam propítius réspice:~† ut te largiénte, regátur in córpore;~* et te servánte, custodiátur in mente. Per Dóminum.}{english}
{\noindent We beseech thee, Almighty God, mercifully to look upon this thy family, that by thy great goodness they may be governed and preserved evermore, both in body and soul. Through our Lord.}
    
        \smalltitle{Palm Sunday.}

\texteparacol{\lettrine{O}{m}nípotens sempitérne Deus, qui humáno géneri, ad imitándum humilitátis exémplum, Salvatórem nostrum carnem súmere, et crucem subíre fecísti:~† concéde propítius; ut et patiéntiæ ipsíus habére documénta,*  et resurrectiónis consórtia mereámur. Per eúmdem Dóminum nostrum.}{english}
{\noindent Almighty and everlasting God, who of thy tender love towards mankind, hast sent thy Son our Saviour Jesus Christ to take upon him our flesh and to suffer death upon the Cross, that all mankind should follow the example of his great humility; mercifully grant, that we may both follow the example of his patience, and also be made partakers of his resurrection.
Through the same Lord.}

\myrule

\smalltitle{Suffrage of All Saints.}

\gscore[Ant.]{2.}{an_beata_dei_genitrix_virgo_solesmes_1961}
\antiphona{english}{May the Blessed Virgin Mary, Mother of God, and all the Saints, intercede for us with the Lord}

\textebilingue{suffragium}{suffragium_en}

\textes{dv_la_en}

\myrule

\orrubrique

\smallscore{domine_exaudi_solemn}

\myrule

\textes{cantor_rubric}

\gscore[]{6.}{BD_Dominicis_Adventus_Quadragesimae}

\noindent \vv Let us bless the Lord.

\noindent \rr Thanks be to God.

\textebilingue{fidelium_animae}{fidelium_animae_en}

\rubrique{\begin{enpars}
Then the Lord's Prayer is recited silently, and the office is concluded with the following verses, the Marian antiphon, etc., unless Compline follows, or, in parish and other churches not bound to the office in choir, Benediction of the Blessed Sacrament or a sermon, or even some other pious exercises.\end{enpars}}

\rubrique{\begin{enpars}The versicle \normaltext{Dóminus} and response \normaltext{Et vitam } are sung on a lower note.\end{enpars}}

\texteparacol{
\noindent \vv Dóminus det nobis suam pacem.

\noindent \rr Et vitam ætérnam. Amen.}
{english}
{\noindent \vv May the Lord grant us his peace.

\noindent \rr And life everlasting. Amen.}

\rubrique{\begin{enpars}On Sundays, all remain standing.\end{enpars}}

\smalltitle{Ave Regina Cælorum.}

\rubrique{¶ After the Purification, that is, from Compline of February 2 even if the feast is transferred, until Compline of Wednesday in Holy Week.}

%\gscore[]{6.}{an--ave_regina_caelorum--solesmes_1961}
%%\vspace{1ex}
%
%\smalltitle{Simple Tone.}

\gscore[]{6.}{an_ave_regina_caelorum_simple_tone_solesmes}

\textes{ave_regina_la_en}

\rubrique{The versicle and response are sung on a lower note.}
\label{conclusion} \phantomsection

\textes{divinum_auxilium_la_en}
\end{document}