% !TEX TS-program = lualatexmk
% !TEX parameter =  --shell-escape

\documentclass[vespers_book_la_eng.tex]{subfiles}

\ifcsname preamble@file\endcsname
  \setcounter{page}{\getpagerefnumber{M-vespers_book_common_saints}}
\fi

\begin{document}
\addcontentsline{toc}{chapter}{Commons of the Saints}\phantomsection
%\festum{purification of the blessed virgin mary}
\addcontentsline{toc}{section}{Common of Apostles and Evangelists outside of Paschal Time}\phantomsection

\header{commons of the saints.}

\rubrique{\begin{enpars}
In Paschal Time, \normaltext{Allelúia} is added to the versicles and responses given here in the common as at Vespers throughout the year.\end{enpars}}

\atvespers{i}
\firstantiphon{8. c}
\initialscore{an_hoc_est_praeceptum_meum_solesmes_1961}

\antiphona{english}{This is my commandment that you love one another, as I have loved you}


\psalmus{109}{109_8c}{109_8}{109_en}

\rubrique{\begin{enpars}
Clerics in descending order, or a cantor, intone the subsequent antiphons.\end{enpars}}

 \gscore[2. Ant.]{1. g}{an_majorem_caritatem_solesmes_1961}
 \antiphona{english}{Greater love than this no man hath, that a man lay down his life for his friends}
\psalmus{110}{110_1g}{110_1}{110_en}

  \gscore[3. Ant.]{1. a3}{an_vos_amici_mei_estis_solesmes_1961}
  \antiphona{english}{You are my friends if you do the things that I command you, said the Lord}
 \psalmus{111}{111_1a3}{111_1}{111_en}

\gscore[4. Ant.]{1. f}{an_beati_pacifici_solesmes_1961}
\antiphona{english}{You are my friends if you do the things that I command you, said the Lord}
\psalmus{112}{112_1f}{112_1}{112_en}

 \gscore[5. Ant.]{1. g}{an_in_patientia_vestra_solesmes_1961}
\antiphona{english}{In your patience you shall possess your souls}
\psalmus{116}{116_1g}{116_1}{116_en}

\chapterlateng \label{chapter_comm_ap_out_tp} \phantomsection
\scripture{Ephes. 2, 19 – 20.}

\texteparacol{\lettrine{F}{r}atres: Jam non estis hóspites et ádvenæ:~† sed estis cives san\-ctórum et doméstici Dei: superædificáti super fundaméntum Apostolórum et Prophetárum,~* ipso summo angulári lápide Christo Jesu.}{english}{\noindent Brethren: Now therefore you are no more strangers and foreigners; but you are fellow citizens with the saints, and the domestics of God, built upon the foundation of the apostles and prophets, Jesus Christ himself being the chief cornerstone.}

\hymnlateng

\gscore[]{4.}{hy_exsultet_orbis_gaudiis_solesmes_1961}

\textes{exsultet_orbis_lat_eng}

\smallscore{versiculus_1_vesperis_apostolorum_extra_tp}

\begin{enpars} \noindent \vv Their sound hath gone forth into all the earth.

\noindent \rr And their words unto the ends of the world.
 \end{enpars}
 
 \gscore[Ad Magnif.]{1. f}{an_tradent_enim_vos_solesmes_1961}
 
 \antiphona{english}{For they shall deliver you up to councils: and in the synagogues you shall be beaten: and you shall stand before governors and kings for my sake, for a testimony unto them}

\rubrique{The collect is proper to the feast.}

\atvespers{ii}
 
 \addcontentsline{toc}{section}{Common of the Bl. Virgin Mary}\phantomsection

\atvespers{i}

 \firstantiphon{3. a}
\initialscore{an_dum_esset_rex_solesmes_1961}
\antiphona{english}{While the King sitteth at his table, my spikenard sendeth forth the smell thereof}
\psalmus{109}{109_3a}{109_3a_b}{109_en}

\gscore[2. Ant.]{4. A*}{an_laeva_ejus_solesmes_1961}
\antiphona{english}{His left hand is under my head, and his right hand doth embrace me}
\psalmus{112}{112_4A_A_star}{112_4A_A_star}{112_en}

\gscore[3. Ant.]{3. b}{an_nigra_sum_solesmes_1961} %% this is annoyingly stretched
\antiphona{english}{I am black but comely, O ye daughters of Jerusalem; therefore hath the King loved me, and brought me into his chamber}
\psalmus{121}{121_3b}{121_3a_b}{121_en}

\gscore[4. Ant.]{8. G}{an_jam_hiems_transiit_solesmes_1961}
\antiphona{english}{Lo the winter is past, the rain is over and gone. Rise up, my love, and come away}
\psalmus{126}{126_8G}{126_8}{126_en}

\gscore[5. Ant.]{4. A*}{an_speciosa_facta_es_solesmes_1961}
\antiphona{english}{O Holy Mother of God, thou art become beautiful and gentle in thy gladness}
\psalmus{147}{147_4A_A_star}{147_4A_A_star}{147_en}

\chapterlateng

\scripture{Eccli. 24, 14.}
\label{Capitulum_BMV}

\texteparacol{\lettrine{A}{b} inítio et ante sǽcula creáta sum,~† et usque ad futúrum sǽculum non désinam,~* et in habitatióne sancta coram ipso ministrávi.
}{english}{\noindent From the beginning, and before the world, was I created, and unto the world to come I shall not cease to be, and in the holy dwelling place I have ministered before him.}

\hymnlateng

\rubrique{The first verse is sung kneeling.}

\label{hy_ave_maris_stella_mode_1_LA1949}
%%%for now, but cross ref later since there are 3 tones
%\grechangenextscorelinedim{3}{spacebeneathtext}{0.1cm}{scalable}
\gscore[]{1.}{hy_ave_maris_stella_mode_1_LA1949}
\textes{ams_en}

\anothertone

\gscore[]{4.}{hy_ave_maris_stella_mode_4}

\anothertone

\gscore[]{7.}{hy_ave_maris_stella_mode_7}

 \smallscore{versiculus_bmv}

\begin{enpars}
 
\noindent \vv Holy Virgin, my praise by thee accepted be.

\noindent \rr Give me strength against thine enemies.
\end{enpars}

%\admagnificat
%\label{an_sancta_maria_solesmes_1961}
%\phantomsection
\gscore[Ad Magnif.]{Ant. 4. E}{an_sancta_maria_solesmes_1961}
%%The minipage not allowing breaks causes problems: the score will be stretched to fill the page.
\antiphona{english}{O Holy Mary, be thou a help to the helpless, a strength to the fearful, a comfort to the sorrowful; pray for the people, plead for the clergy, make intercession for all women vowed to God: may all that are keeping this thy holy feast day feel the might of thine assistance}

%%we need a better way to handle this! alt text is  too finicky, minipages can't break
%%minipages work IF AND ONLY IF the spread is even-odd, left-right. A page turn WILL NOT DO.

%\newpage
\begin{minipage}[t]{3.3in}%
{\fontsize{9}{11}\selectfont{Our Lady of Mount Carmel.}} \gresetinitiallines{0} \gabcsnippet{(c4) † tu(g)am(g) so(g)lé(gh)mnem(g') com(g)me(g)mo(gf)ra(g)ti(ghg)ó(e.)nem.(e.) (::)}%
\begin{enpars}
\noindent  …thy yearly solemn commemoration…
\end{enpars}
\end{minipage}%
\vspace{\baselineskip}

\begin{minipage}[t]{3.4in}%
{\fontsize{9}{11}\selectfont{Holy Name of Mary.}} \gresetinitiallines{0} \gabcsnippet{(c4) † tu(g)i(g) san(g')cti(f) Nó(gh)mi(g)nis(g') com(g)me(g)mo(gf)ra(g)ti(ghg)ó(e.)nem.(e.) (::)}%
\begin{enpars}
\noindent  …commemoration of thy holy name…
\end{enpars}
\end{minipage}%
\vspace{\baselineskip}

\begin{minipage}[t]{4.9in}% évitez des nombres arbitraires…
{\fontsize{9}{11}\selectfont{Our Lady, Help of Christians.}} \gresetinitiallines{0} \gabcsnippet{(c4) † qui(cd)cúm(fe)que(fg) tu(g)um(g) san(gh)ctum(g') im(g)pló(gf)rant(g) au(ghg)xí(e)li(e)um.(e.) (:) Al(g)le(gf)lú(e.){ia}.(e.) (::)}%
\begin{enpars}
\noindent  …whoever implores thy holy aid, alleluia.
\end{enpars}
\end{minipage}
\vspace{\baselineskip}

\begin{minipage}[t]{2.6in}%
{\fontsize{9}{11}\selectfont{Our Lady, Patroness.}} \gresetinitiallines{0}\gabcsnippet{(c4) †  tu(g)um(g) san(gh)ctum(g) Pa(gf)tro(ghg)cí(e)ni(e)um.(e.) (::)}%
\gresetinitiallines{1}
\begin{enpars}
\noindent … thy holy patronage…
\end{enpars}
\end{minipage}%

\collectlateng
 \label{concede_bmv} \phantomsection
 
\texteparacol{\lettrine{C}{o}ncéde nos famúlos tuos, quǽsumus Domíne Deus, perpétua mentis et corpóris sanitáte gaudére:~† et gloriósa beátæ Mariæ semper Virgínis intercessióne,~* a præsénti liberári tristítia, et æterna perfrúi lætítia. Per Dóminum.}{english}{\noindent Grant, we beseech thee, O Lord God, unto all thy servants, that they may remain continually in the enjoyment of soundness both of mind and body, and by the glorious intercession of the Blessed Mary, always a Virgin, may be delivered from present sadness, and enter into the joy of thine eternal gladness.
Through our Lord.}

\atvespers{ii}

\allexcept

\label{an_beatam_me_dicent_comm_bmv_solesmes_1961}
\phantomsection
\gscore[]{8. G}{an_beatam_me_dicent_comm_bmv_solesmes_1961}
 \antiphona{english}{All generations shall call me blessed, for God hath regarded the lowliness of his hand-maiden}


\addcontentsline{toc}{section}{Common of the Dedication of a Church}\phantomsection

\atvespers{i}

\firstantiphon{7. a}

\initialscore{an_domum_tuam_domine_solesmes_1961}
\antiphona{english}{Holiness becometh thine house, O Lord, in length of days}

\psalmus{109}{109_7a}{109_7}{109_en}

\gscore[2. Ant.]{1. D2}{an_domus_mea_solesmes_1961} 
 \antiphona{english}{My house shall be called the house of prayer}
\psalmus{110}{110_1D2}{110_1D2}{110_en}

\gscore[3. Ant.]{1. f}{an_haec_est_domus_domini_solesmes_1961}
  \antiphona{english}{This is the Lord's house stoutly builded, well founded upon a sure rock}
\psalmus{111}{111_1f}{111_1}{111_en}

\gscore[4. Ant.]{8. c}{an_bene_fundata_est_solesmes_1961}
\antiphona{english}{The Lord's house is well founded upon a sure rock}
\psalmus{112}{112_8c}{112_8}{112_en}

\gscore[5. Ant.]{1. g}{an_lapides_pretiosi_solesmes_1961}
\antiphona{english}{All thy walls are of stones most precious, and the towers of Jerusalem shall be built up with jewels}
\psalmus{147}{147_1g}{147_1}{147_en}

\chapterlateng
 \scripture{Apoc. 21, 2.}

\texteparacol{\lettrine{V}{i}di civitátem sanctam Jerúsalem novam,~† descendéntem de cælo a Deo,~* parátam sicut spónsam ornátam viro suo.}{english} {\noindent I saw the holy city, New Jerusalem, coming down from God out of heaven, prepared as a Bride adorned for her husband.}

\hymnlateng
\gscore[]{1.}{hy_caelestis_urbs_jerusalem_solesmes_1961}

\textes{caelestis_urbs_en}

 \smallscore{versiculus_dedications_ecclesiae_1_vesp}

\begin{otherlanguage}{english} \noindent \vv This is the house of God, stoutly builded.

\noindent \rr Well founded upon a sure rock.
 \end{otherlanguage}
 
 \gscore[Ad Magnif.]{1. g}{an_sanctificavit_dominus_solesmes_1961}
\antiphona{english}{The Most High hath hallowed his tabernacle. For this is the House of God, whereon his Name shall be called, whereof it is written my Name shall be there, saith the Lord}

\collectlateng

\smalltitle{On the day of the dedication itself and on the octave day thereof:}

\texteparacol{\lettrine{D}{e}us, qui invisibíliter ómnia cóntines, et tamen pro salúte géneris humáni signa tuæ poténtiæ visibíliter osténdis:~† templum hoc poténtia tuæ inhabitatiónis illústra, et concéde; ut omnes, qui huc deprecatúri convéniunt,~* ex quacúmque tribulatióne ad te clamáverint, consolatiónis tuæ benefícia consequántur. Per Dóminum.}{english}{\noindent O God, who though unseen, upholdest all things and yet for the salvation of mankind dost visibly show signs of thy power: give glory to this temple by the might of thine indwelling, and grant that all who in their deep distress shall come and call upon thee here may receive thy divine comfort. Through our Lord.}

\smalltitle{On the anniversary and on the anniversary of the dedication of another church.}

\texteparacol{\lettrine{D}{e}us, qui nobis per síngulos annos hujus sancti templi tui consecratiónis réparas diem, et sacris semper mystériis repræséntas incólumes:~† exáudi preces pópuli tui, et præsta; ut quisquis hoc templum benefícia petitúrus ingréditur,~* cun\-cta se impetrásse lætétur. Per Dóminum.}{english}{\noindent O God, who dost every year bring round unto us again the day whereon this thine holy temple was hallowed, and bringest us again in soundness of body and mind to be present at thine holy worship, graciously hear the supplications of thy people, and grant that whosoever shall come into this thine house to ask good at thine hand, may be rejoiced in the obtaining of all his request.
Through our Lord.}

\atvespers{ii}

\allexcept

 \smallscore{versiculus_dedications_ecclesiae_2_vesp}
\begin{otherlanguage}{english} \noindent \vv Holiness becometh thine house, O Lord.

\noindent \rr In length of days.
 \end{otherlanguage}

\gscore[Ad Magnif.]{Ant. 6. F}{an_o_quam_metuendus_solesmes_1961}
\antiphona{english}{How dreadful is this place. Surely this is none other but the house of God, and this is the gate of heaven}

\end{document}