% !TEX TS-program = lualatexmk
% !TEX parameter =  --shell-escape
\documentclass[12pt]{book}

\input{commonheaders_vespers_people} %%% will need to fix spacing problem in psalms at a later date.

\begin{document}
 
 \twosided[pc] %% to allow for the left-right switching of columns needed for bilingual liturgical typesetting (like in missals); it's unclear why this goes BEFORE everything else and not when activating the environment. %% It's OK for paper booklets to not switch, but it's probably best practice to switch.
 
  \thispagestyle{empty}
 
 {\centering{\huge\capspace{{II VESPERS OF THE SACRED HEART}}}\par}\header{ii vespers of the sacred heart and within the octave.}
 
  \vspace{0.5\baselineskip}
 
{\centering{\huge\capspace{{AND DAYS WITHIN THE OCTAVE.}}}\par}
 
\smalltitle{Before the Office.}

\rubrique{These prayers are said kneeling. End when the priest stands.}

\textebilingue{orationes_ante_do}{orationes_ante_do_eng}

\smalltitle{\prayers}

\rubrique{The celebrant intones the following verse. All sing the response and doxology.}

\begin{otherlanguage}{english}
\rubrique{Make the sign of the cross at the beginning and at the Magnificat, and bow the head when the Trinity is mentioned, when mention is made of the ``Holy Name'' in some way, and at ``Orémus.''}\end{otherlanguage}

\rubrique{On the feast itself:}

\gscore[]{℣.}{Deus_in_adjutorium_tonus_solemnior}

\noindent
\begin{otherlanguage}{english}
O God, \cc{} come to my assistance. \rr O Lord, make haste to help me. Glory be to the Father, and to the Son,~* and to the Holy Ghost. As it was in the beginning, is now,~* and ever shall be, world without end. Amen. Alleluia.\end{otherlanguage}

\myrule

\rubrique{On the other days within the octave:}

\gscore[]{℣.}{Deus_In_Adjutorium_Festivus}

\rubrique{On the feast and its octave day, the antiphons are doubled. On all other days, the office is of semidouble rite.}

\smalltitle{First Antiphon. 1. f.}
\initialscore{an_unus_militum_solesmes_1961}

\begin{otherlanguage}{english} \noindent Ant. One of the soldiers with a spear pierced his side, and forthwith came there out Blood and Water.\end{otherlanguage}

\rubrique{\begin{otherlanguage}{english}
Sit at the word ``Sede.'' The cantor sings odd verses; all sing even verses, and this for the hymn and Magnificat, when sung entirely in chant, as well.
\end{otherlanguage}}

\psalmus{109}{109_1f}{109_1f}{109_en}

\rubrique{\begin{otherlanguage}{english}
Clerics in descending order, or a cantor, intone the subsequent antiphons.\end{otherlanguage}}

 \gscore[2. Ant.]{7. c}{an_stans_jesus_solesmes_1961}
% 
 \begin{otherlanguage}{english} \noindent Ant. Jesus stood and cried, saying, If any man thirst, let him come unto me and drink.\end{otherlanguage}
%
\psalmus{110}{110_7c}{110_7c}{110_en}
%
  \gscore[3. Ant.]{3. a2}{an_in_caritate_solesmes_1961}
%  
  \begin{otherlanguage}{english} \noindent Ant. With an everlasting love hath God loved us, and from the day that he was lifted up over the earth, he hath drawn us with loving-kindness unto his Heart. \end{otherlanguage}
%
 \psalmus{115}{115_3a2}{115_3a2_g}{115_en}
%
\gscore[4. Ant.]{4. E}{an_venite_ad_me_solesmes_1961}
%
\begin{otherlanguage}{english} \noindent Ant. Come unto me, all ye that travail and are heavy laden, and I will refresh you.\end{otherlanguage}
%
\psalmus{127}{127_4E}{127_4E}{127_en}

 \gscore[5. Ant.]{5. a}{an_fili_praebe_solesmes_1961}
 
\begin{otherlanguage}{english} \noindent Ant. My son, give me thine heart, and let thine eyes observe my ways.\end{otherlanguage}

\psalmus{147}{147_5}{147_5}{147_en}

\rubrique{\begin{otherlanguage}{english}
Having sung the last antiphon in full, stand when the celebrant does.
\end{otherlanguage}}

\bigtitle{Chapter.}

\smalltitle{On the feast and within the octave:}
 \scripture{Eph. 3, 8 – 9.}

\texteparacol{\lettrine{F}{r}ratres: Mihi ómnium sanctórum mínimo data est grátia hæc,~† in géntibus evangelizáre investigábiles divítias Christi:~* et illumináre omnes, quæ sit dispensátio sacraménti abscónditi a sǽculis in Deo.\par
\rr Deo grátias.}{\noindent Brethren: Unto me, who am less than the least of all saints, is this grace given, that I should preach among the Gentiles the unsearchable riches of Christ; and to make all men see what is the fellowship of the Mystery, which from the beginning of the world hath been hid in God. Thanks be to God.}

\smalltitle{On Sunday within the octave:}
\scripture{1 Petri 5, 6 – 7.}

\texteparacol{\lettrine{C}{a}ríssimi: Humiliámini sub poténti manu Dei, ut vos exáltet in témpore visitatiónis;~† omnem sollicitúdinem vestram proiciéntes in eum,~* quóniam ipsi cura est de vobis.\par
\rr Deo grátias.}{\noindent Beloved:  humble yourselves therefore under the mighty hand of God, that he may exalt you in due time: casting all your care upon him; for he careth for you. Thanks be to God.}

\smalltitle{Hymn.}

\gscore[]{3.}{hy_en_ut_superba_solesmes_1961}

\begin{multicols}{2}
\begin{otherlanguage}{english}
\raggedcolumns
Lo, how the savage crew\\
Of our proud sins hath rent\\
The heart of our all-gracious God,\\
That heart so innocent.

The soldier's quivering lance\\
Our guilt it was that drave,\\
Our wicked deeds that to its point\\
Such cruel sharpness gave.

O wounded heart, whence sprang\\
The Church, the Saviour's bride;\\
Thou door of our salvation's ark\\
Set in its mystic side.

Thou holy fount, whence flows\\
The sacred sevenfold flood,\\
Where we our filthy robes may cleanse\\
In the Lamb's saving blood:

By sorrowful relapse,\\
thee will we rend no more;\\
But like the flames, those types of love,\\
Strive heavenward to soar.

Father and Son supreme\\
And Spirit, hear our cry;\\
Whose is the kingdom, praise and power,\\
Through all eternity.
Amen.

\end{otherlanguage}
\end{multicols}

\rubrique{\begin{otherlanguage}{english}
The cantor sings the verse to which all reply.
\end{otherlanguage}}

\rubrique{On the feast, the solemn tone is used, but on the days within the octave, the ordinary tone is used:}
\label{haurietis}\phantomsection
   \smallscore{versiculus_ii_vesperis_ss_cordis}
   
\begin{otherlanguage}{english}
\noindent \vv With joy shall ye draw water.

\noindent \rr Out of the fountains of the Saviour.

\end{otherlanguage}
\myrule

\rubrique{On Sunday within the octave:}

 \smallscore{versiculus_dom_octave_sc}

\begin{otherlanguage}{english} \noindent \vv The merciful Lord hath instituted a memorial of his wondrous deeds.

\noindent \rr  He hath given Meat unto them that fear him. \end{otherlanguage}
 
 \myrule
 
\smalltitle{Antiphon at the Magnificat.}

\rubrique{On the feast and octave day, sit when the celebrant does, after he sings the first words. Rise at the end. Otherwise, remain standing.}

 \smalltitle{At II Vespers.}
 \label{ad_jesum_autem}\phantomsection
 \gscore[]{1. f}{an_ad_jesum_autem_solesmes_1961}
 
 \begin{otherlanguage}{english} \noindent Ant. But when they came to Jesus and saw that he was dead already, they broke not his legs, but one of the soldiers with a spear pierced his side, and forthwith came there out Blood and Water.\end{otherlanguage}
 
 \smallscore{Magnificat_1_f_Solemnis}

\begin{enumerate}[label=\arabic*.,%
			leftmargin=0pt, %
			itemindent=0mm, %
			labelsep=3pt, %
			labelwidth=0pt, %
			rightmargin=0pt, %
			parsep=0pt, %
			topsep=0pt, %
			itemsep=0pt,%
			start=3]

\item Quia respéxit humilitátem \textit{ancíllæ} \textbf{su}æ:~* ecce enim ex hoc beátam me dicent omnes gene\textit{rati}\textbf{ó}nes.
\item Quia fecit mihi \textit{magna} \textit{qui} \textbf{pot}\textbf{ens} est:~* et sanctum \textit{nomen} \textbf{e}jus.
\item Et misericórdia ejus a progéni\textit{e} \textit{in} \textit{pro}\textbf{gé}\textbf{ni}es~* timén\textit{tibus} \textbf{e}um.
\item Fecit poténtiam in \textit{bráchio} \textbf{su}o:~* dispérsit supérbos mente \textit{cordis} \textbf{su}i.
\item Depósuit po\textit{téntes} \textit{de} \textbf{se}de,~* et exal\textit{távit} \textbf{hú}miles.
\item Esuriéntes \textit{implévit} \textbf{bo}nis:~* et dívites dimí\hspace{0.03em}\textit{sit} \textit{in}\textbf{á}nes.
\item Suscépit Israel \textit{púerum} \textbf{su}um,~* recordátus misericór\textit{diæ} \textbf{su}æ.
\item Sicut locútus est \textit{ad} \textit{pa}\textit{tres} \textbf{no}stros,~* Abraham et sémini e\hspace{0.03em}\textit{jus} \textit{in} \textbf{sǽ}cula.
\item Glória \textit{Pa}\textit{tri,} \textit{et} \textbf{Fí}\textbf{li}o,~* et Spirí\hspace{0.03em}\textit{tui} \textbf{San}cto.
\item Sicut erat in princípio, \textit{et} \textit{nunc,} \textit{et} \textbf{sem}per,~* et in sǽcula sæcu\hspace{0.03em}\textit{lórum.} \textbf{A}men.


\end{enumerate}
    
      \smalltitle{On Sunday within the octave.}
      
          \gscore[Ad Magnif.]{Ant. 6. f}{an_quae_mulier_solesmes_1961}
        \begin{otherlanguage}{english} \noindent Ant. What woman having ten pieces of silver, if she lose one piece, doth not light a candle, and sweep the house, and seek diligently till she find it?
\end{otherlanguage}
      
  
\smallscore{Magnificat_6}

   \smallskip%
    \noindent%
    \begin{enumerate}[%
    		label=\arabic*.,%
			leftmargin=0pt, %
			itemindent=0mm, %
			labelsep=3pt, %
			labelwidth=0pt, %
			rightmargin=0pt, %
			parsep=0pt, %
			topsep=0pt, %
			itemsep=0pt,%
			start=3]%

\item Quia respéxit humilitátem ancíl\hspace{0.03em}\textit{læ} \textbf{su}æ:~* ecce enim ex hoc beátam me dicent omnes gene\textit{rati}\textbf{ó}nes.
\item Quia fecit mihi magna \textit{qui} \textbf{pot}ens est:~* et san\-ctum \textit{nomen} \textbf{e}jus.
\item Et misericórdia ejus a progénie in \textit{pro}\hspace{0.03em}\textbf{gé}nies~* timén\textit{tibus} \textbf{e}um.
\item Fecit poténtiam in bráchi\textit{o} \textbf{su}o:~* dispérsit supérbos mente \textit{cordis} \textbf{su}i.
\item Depósuit poténtes \textit{de} \textbf{se}de,~* et exal\textit{távit} \textbf{hú}miles.
\item Esuriéntes implé\textit{vit} \textbf{bo}nis:~* et dívites dimí\textit{sit} \textit{in}\textbf{á}nes.
\item Suscépit Israel púe\textit{rum} \textbf{su}um,~* recordátus misericór\textit{diæ} \textbf{su}æ.
\item Sicut locútus est ad pa\hspace{0.03em}\textit{tres} \textbf{no}stros,~* Abraham et sémini e\textit{jus} \textit{in} \textbf{sǽ}cula.
\item Glória Patri, \textit{et} \textbf{Fí}lio,~* et Spirí\hspace{0.03em}\textit{tui} \textbf{San}cto.
\item Sicut erat in princípio, et nunc, \textit{et} \textbf{sem}per,~* et in sǽcula sæcu\hspace{0.03em}\textit{lórum.} \textbf{A}men.
\end{enumerate}

\smalltitle{Canticle of the Blessed Virgin Mary.}

\scripture{Luc. 1, 46–55.}

\begin{multicols}{2}
 \begin{otherlanguage}{english}
My soul \cc~* doth magnify the Lord.

And my spirit hath rejoiced~* in God my Saviour.

Because he hath regarded the humility of his handmaid;~* for behold from henceforth all generations shall call me blessed.

Because he that is mighty, hath done great things to me;~* and holy is his name.

And his mercy is from generation unto generations,~* to them that fear him.

He hath shewed might in his arm:~* he hath scattered the proud in the conceit of their heart.

He hath put down the mighty from their seat,~* and hath exalted the humble.

He hath filled the hungry with good things;~* and the rich he hath sent empty away.

He hath received Israel his servant,~* being mindful of his mercy:

As he spoke to our fathers,~* to Abraham and to his seed for ever.

Glory be to the Father, and to the Son,~* and to the Holy Ghost.

As it was in the beginning, is now,~* and ever shall be, world without end. Amen.
\end{otherlanguage}
% \end{paracol}
\end{multicols}

\rubrique{\begin{otherlanguage}{english}
Sit as the antiphon is repeated. Then the usual dialogue follows. All stand from now until the end of the office.\end{otherlanguage}}

\textebilingue{dv}{dv_eng}

   \begin{paracol}{2}
   Orémus.
   
\label{oratio_sc}\phantomsection

\lettrine{D}{e}us, qui nobis in Corde Fílii tui, nostris vulneráto peccátis, infinítos dilectiónis thesáuros misericórditer largíri dignáris:~† concéde quǽsumus; ut illi devótum pietátis nostræ præstántes obséquium,~* dignæ quoque satisfactiónis exhibeámus offícium. Per eúmdem Dóminum.

   ℟. Amen.
   
\switchcolumn
\begin{otherlanguage}{english}
\noindent Let us pray. O God, who hast suffered the Heart of thy Son to be wounded by our sins, and in that very Heart hast bestowed on us the abundant riches of thy love: grant that the devout homage of our hearts, which we render unto him, may of thy mercy be deemed a recompence acceptable in thy sight.
Through the same Jesus Christ, thy Son, Our Lord, Who liveth and reigneth with thee in the unity of the Holy Ghost, God, world without end.
\end{otherlanguage}
    \end{paracol}
    
    \myrule
    
    \smalltitle{On Sunday within the octave:}
       \begin{paracol}{2}
   Orémus.
   
\label{oratio_dom_sc}\phantomsection

\lettrine{P}{r}oté\-ctor in te sperántium, Deus, sine quo nihil est válidum, nihil san\-ctum:~† multíplica super nos misericórdiam tuam; ut, te re\-ctóre, te duce, sic transeámus per bona temporália,~* ut non amittámus ætérna. Per Dóminum.

   ℟. Amen.
   
\switchcolumn
\begin{otherlanguage}{english}
\noindent Let us pray.
O God, the protector of all them that trust in thee, without whom nothing is strong, nothing is holy; increase and multiply upon us thy mercy, that thou being our ruler and guide, we may so pass through things temporal, that we finally lose not the things eternal. Through Jesus Christ, thy Son, Our Lord, Who liveth and reigneth with thee in the unity of the Holy Ghost, God, world without end.
 Amen.
\end{otherlanguage}
    \end{paracol}
    
\rubrique{On Sunday, the following day is commemorated:}

\gscore[Ant.]{1.}{an_ignem_solesmes_1961}

\texteparacol{\noindent \vv Tóllite jugum meum super vos et díscite a me.

\noindent \rr Quia mitis sum et húmilis Corde.
}{\noindent \vv Take my yoke upon you, and learn of me.

\noindent \rr For I am meek and lowly in Heart.}

\rubrique{Oratio \normaltext{Deus, qui nobis, \pageref{oratio_sc}.}}

\rubrique{But if the following day is not of the octave, ant. \normaltext{Ad Jesum autem, \pageref{ad_jesum_autem}. ℣. Hauriétis, \pageref{haurietis},} Oratio \normaltext{Deus, qui nobis, \pageref{oratio_sc}.}} 

\textebilingue{dv}{dv_eng}

\rubrique{\begin{otherlanguage}{english}
The cantor sings the verse to which all reply.
\end{otherlanguage}}

    \smalltitle{At II Vespers.}

\gscore[]{5.}{BD_Solemnis_mode_5}

    \myrule
    
    \smalltitle{On Sunday within the octave:}

\gscore[]{2.}{BD_Semiduplex_Vesp}

\myrule

\noindent \vv Let us bless the Lord.

\noindent \rr Thanks be to God.

\textebilingue{fidelium_animae}{fidelium_animae_en}

\rubrique{The office is as on the feast unless a feast of nine lessons is celebrated instead, and the days within the octave are of semidouble rite.}

\rubrique{On Friday at II Vespers, the office is of double major rite.}
\end{document}