\begin{otherlanguage}{english}
\item 1. When Israel went out of Egypt,~* the house of Jacob from a barbarous people:
\item 2. Judea was made his sanctuary,~* Israel his dominion.
\item 3. The sea saw and fled:~* Jordan was turned back.
\item 4. The mountains skipped like rams,~* and the hills like the lambs of the flock.
\item 5. What ailed thee, O thou sea, that thou didst flee:~* and thou, O Jordan, that thou wast turned back?
\item 6. Ye mountains, that ye skipped like rams,~* and ye hills, like lambs of the flock?
\item 7. At the presence of the Lord the earth was moved,~* at the presence of the God of Jacob:
\item 8. Who turned the rock into pools of water,~* and the stony hill into fountains of waters.
\item 9. Not to us, O Lord, not to us;~* but to thy name give glory.
\item 10. For thy mercy, and for thy truth’s sake:~* lest the Gentiles should say: Where is their God?
\item 11. But our God is in heaven:~* he hath done all things whatsoever he would.
\item 12. The idols of the Gentiles are silver and gold,~* the works of the hands of men.
\item 13. They have mouths and speak not:~* they have eyes and see not.
\item 14. They have ears and hear not:~* they have noses and smell not.
\item 15. They have hands and feel not: they have feet and walk not:~* neither shall they cry out through their throat.
\item 16. Let them that make them become like unto them:~* and all such as trust in them.
\item 17. The house of Israel hath hoped in the Lord:~* he is their helper and their protector.
\item 18. The house of Aaron hath hoped in the Lord:~* he is their helper and their protector.
\item 19. They that fear the Lord have hoped in the Lord:~* he is their helper and their protector.
\item 20. The Lord hath been mindful of us,~* and hath blessed us.
\item 21. He hath blessed the house of Israel:~* he hath blessed the house of Aaron.
\item 22. He hath blessed all that fear the Lord,~* both little and great.
\item 23. May the Lord add blessings upon you:~* upon you, and upon your children.
\item 24. Blessed be you of the Lord,~* who made heaven and earth.
\item 25. The heaven of heaven is the Lord’s:~* but the earth he has given to the children of men.
\item 26. The dead shall not praise thee, O Lord:~* nor any of them that go down to hell.
\item 27. But we that live bless the Lord:~* from this time now and for ever.
\item 28. Glory be to the Father, and to the Son,~* and to the Holy Ghost. 
\item 29. As it was in the beginning, is now,~* and ever shall be, world without end. Amen.
\end{otherlanguage}