\begin{enpars}
\item How have I loved thy law, o Lord!~* it is my meditation all the day.
\item Through thy commandment, thou hast made me wiser than my enemies:~* for it is ever with me.
\item I have understood more than all my teachers:~* because thy testimonies are my meditation.
\item I have had understanding above ancients:~* because I have sought thy commandments.
\item I have restrained my feet from every evil way:~* that I may keep thy words.
\item I have not declined from thy judgments,~* because thou hast set me a law.
\item How sweet are thy words to my palate!~* more than honey to my mouth.
\item By thy commandments I have had understanding:~* therefore have I hated every way of iniquity.
\item Thy word is a lamp to my feet,~* and a light to my paths.
\item I have sworn and am determined~* to keep the judgments of thy justice.
\item I have been humbled, o Lord, exceedingly:~* quicken thou me according to thy word.
\item The free offerings of my mouth make acceptable, o Lord:~* and teach me thy judgments.
\item My soul is continually in my hands:~* and I have not forgotten thy law.
\item Sinners have laid a snare for me:~* but I have not erred from thy precepts.
\item I have purchased thy testimonies for an inheritance for ever:~* because they are the joy of my heart.
\item I have inclined my heart to do thy justifications for ever,~* for the reward.
\item Glory be to the Father, and to the Son,~* and to the Holy Ghost.
\item As it was in the beginning, is now,~* and ever shall be, world without end. Amen.
\end{enpars}