\begin{otherlanguage}{english}
\item I have hated the unjust:~* and I have loved thy law.
\item Thou art my helper and my protector:~* and in thy word I have greatly hoped.
\item Depart from me, ye malignant:~* and I will search the commandments of my God.
\item Uphold me according to thy word, and I shall live:~* and let me not be confounded in my expectation.
\item Help me, and I shall be saved:~* and I will meditate always on thy justifications.
\item Thou hast despised all them that fall off from thy judgments;~* for their thought is unjust.
\item I have accounted all the sinners of the earth prevaricators:~* therefore have I loved thy testimonies.
\item Pierce thou my flesh with thy fear:~* for I am afraid of thy judgments.
\item I have done judgment and justice:~* give me not up to them that slander me.
\item Uphold thy servant unto good:~* let not the proud calumniate me.
\item My eyes have fainted after thy salvation:~* and for the word of thy justice.
\item Deal with thy servant according to thy mercy:~* and teach me thy justifications.
\item I am thy servant:~* give me understanding that I may know thy testimonies.
\item It is time, O Lord, to do:~* they have dissipated thy law.
\item Therefore have I loved thy commandments~* above gold and the topaz.
\item Therefore was I directed to all thy commandments:~* I have hated all wicked ways
\item Glory be to the Father, and to the Son,~* and to the Holy Ghost.
\item As it was in the beginning, is now,~* and ever shall be, world without end. Amen.
\end{otherlanguage}