\begin{enpars}
\item My soul hath fainted after thy salvation:~* and in thy word I have very much hoped.
\item My eyes have failed for thy word,~* saying: When wilt thou comfort me?
\item For I am become like a bottle in the frost:~* I have not forgotten thy justifications.
\item How many are the days of thy servant:~* when wilt thou execute judgment on them that persecute me?
\item The wicked have told me fables:~* but not as thy law.
\item All thy statutes are truth:~* they have persecuted me unjustly, do thou help me.
\item They had almost made an end of me upon earth:~* but I have not forsaken thy commandments.
\item Quicken thou me according to thy mercy:~* and I shall keep the testimonies of thy mouth.
\item For ever, o Lord,~* thy word standeth firm in heaven.
\item Thy truth unto all generations:~* thou hast founded the earth, and it continueth.
\item By thy ordinance the day goeth on:~* for all things serve thee.
\item Unless thy law had been my meditation,~* I had then perhaps perished in my abjection.
\item Thy justifications I will never forget:~* for by them thou hast given me life.
\item I am thine, save thou me:~* for I have sought thy justifications.
\item The wicked have waited for me to destroy me:~* but I have understood thy testimonies.
\item I have seen an end of all perfection:~* thy commandment is exceeding broad.
\item Glory be to the Father, and to the Son,~* and to the Holy Ghost.
\item As it was in the beginning, is now,~* and ever shall be, world without end. Amen.
\end{enpars}