%%the practice that I recall is to rise for Oremus and the prayerr but the Cantus Selecti rubric is before DV!

\rubrique{In fine processionem tantum. At the end of processions only.}

\rubrique{At the end of processions:}

\texteparacol{\rubrique{Surgit Celebrans.}

\vv Dóminus vo\textbf{bís}cum.

\rr Et cum spíritu \textbf{tu}o.

Orémus.

\lettrine{D}{e}us, cui próprium est miseréri semper et párcere:~† súscipe deprecatiónem nostram; ut nos, et omnes fámulos tuos, quos deli\-ctórum caténa constríngit,~* miseratio tuæ pietátis cleménter absólvat.

\lettrine{E}{x}áudi, quǽsumus Dómine, súpplicum preces:~† et confiténtium tibi parce peccátis;~* ut páriter nobis indulgéntiam tríbuas benígnus et pacem.

\lettrine{I}{n}effabilem nobis, Dómine, misericórdiam tuam cleménter osténde:~† ut simul nos et a peccátis ómnibus éxuas,~* et a pœnis quas pro his merémur, erípias.

\lettrine{D}{e}us, qui culpa offénderis, pæniténtia placáris:~† preces pópuli tui supplicántis propítius réspice;~* et flagélla tuæ iracúndiæ, quæ pro peccátis nostris merémur, avérte.

\lettrine{O}{m}nípotens sempitérne Deus, miserere fámulo tuo Pontífici
nostro \nomen:~† et dírige eum secúndum tuarn cleméntiam in
viam salútis ætérnae;~* ut te donánte, tibi plácita cúpiat, et tota
virtúte perfíciat.

\lettrine{D}{e}us, a quo sancta desidéria, re\-cta consília et justa sunt opera:~† da servis tuis illam, quam mundus dare non potest, pacem;~* ut et corda nostra mandátis tuis dédita, et hóstium subláta formídine, témpora sint tua prote\-ctióne tranquílla. 

\lettrine{U}{r}e igne San\-cti Spíritus renes nostros et cor nostrum Dómine:~† ut tibi casto córpore serviámus,~* et mundo corde placeámus.

\lettrine{F}{i}délium Deus ómnium cónditor et redémptor, animábus famulórum famularúmque tuárum remissiónem cunctórum tríbue peccatórum:~† ut indulgéntiam, quam semper optavérunt,~* piis supplicatiónibus consequántur.

\lettrine{A}{c}tiones nostras, quǽsumus Dómine, aspirándo prævéni et adjuvándo proséquere:~† ut cun\-cta orátio et operátio a te semper incípiat~* et per te cœpta finiátur.

\lettrine{O}{m}nípotens sempitérne Deus, qui vivórum domináris simul et mortuórum, omniúmque miseréris, quos tuos fide et ópere futúros esse prænóscis:~† te súpplices exorámus; ut pro quibus effúndere preces decrévimus, quosque vel præsens sǽculum adhuc in carne rétinet vel futúrum jam exútos córpore suscépit,~* intercedéntibus ómnibus San\-ctis tuis, pietátis tuæ cleméntia, ómnium deli\-ctórum suórum véniam consequántur. Per Dóminum nostrum Jesum Christum.

\vv Dóminus vo\textbf{bís}cum.

\rr Et cum spíritu \textbf{tu}o.

\rubrique{Cantores.}

\vv Exáudiat nos omnípotens et miséricors Dóminus.

\rr Amen.

\rubrique{Celebrans, voce recta et depressa.}

\vv Et fidélium ánimæ per misericórdiam Dei requiéscant in pace.

\rr Amen.}
{english}
{\rubrique{The celebrant rises.}

\noindent	\vv The Lord be with you.
	
\noindent \rr And with thy spirit.

Let us pray.

\noindent O God, whose property is always to have mercy and to spare, receive our petition; that we and all thy servants who are bound by the chain of sin may, by the compassion of thy goodness mercifully be absolved.

\noindent Graciously hear, we beseech thee, O Lord, the prayers of thy supplicants and pardon the sins of those who confess to thee: that in thy bounty thou mayest grant us both pardon and peace.

\noindent In thy clemency, O Lord, show unto us thine ineffabile mercy; that thou mayest both free us from sins and deliver us from the punishments which we deserve for them.

\noindent O God, who by sin art offended, and by penance appeased, mercifully regard the prayers of thy people making supplication to thee; and turn away the scourges of thy wrath which we deserve for our sins.

\noindent Almighty, everlasting God, have mercy upon thy servant, \nomen{} our Sovereign Pontiff, and direct him, according to thy clemency, into the way of everlasting salvation; that by thy grace he may desire those things that are pleasing to thee, and perform them with all his strength.

\noindent O God, from whom all holy desires, all good counsels, and all just works do proceed; give unto thy servants that peace which the world cannot give; so that, with our hearts set to obey thy commandments, and freed from the fear of the enemy, we may pass our lives in peace under thy protection.

\noindent Inflame, o Lord, with the fire of the Holy Spirit, our hearts and our desires; that we may serve thee with a chaste body and please thee with a clean heart.

\noindent O God, the creator and redeemer of all the faithful, grant to the souls of thy departed servants the remission of all their sins; that through pious supplications they may obtain the pardon they have always desired.

\noindent Direct, we beseech thee, O Lord, our actions by thy holy inspirations and carry them on by thy gracious assistance; that every prayer and work of ours may always begin with thee and through thee be happily ended.

\noindent Almighty and everlasting God, who hast dominion over the living and the dead, and art merciful to all whom thou foreknowest shall be thine by faith and good works: we humbly beseech thee; that they for whom we intend to pour forth our prayers, whether this present world still detains them in the flesh, or the world to come has already received them out of their bodies, may, through the intercession of all thy saints, and in thy compassionate goodness, obtain the pardon of all their sins. Through our Lord Jesus Christ. Amen.

\noindent	\vv The Lord be with you.
	
\noindent \rr And with thy spirit.

\rubrique{Cantors.}

\noindent	\vv May the almighty and most merciful Lord graciously hear us.

\noindent \rr  Amen.

\rubrique{Celebrant, in a monotone, on a lower note.}

\noindent \vv May the souls of the faithful departed through the mercy of God rest in peace.

\noindent \rr Amen.}