\begin{enpars}\noindent Loving Mother of the Redeemer, 
who remains the accessible Gateway of Heaven,
and Star of the Sea,
Give aid to a falling people 
that strives to rise;
O Thou who begot thy holy Creator,
while all nature marvelled,
Virgin before and after
receiving that ``Ave'' from the mouth of Gabriel,
have mercy on sinners.\end{enpars}

\texteparacol{\rubrique{Ab Adventu usque ad Nativitaem Domini:}

 \vv Angelus Dómini nuntiávit Maríæ.

\rr Et concépit de Spíritu Sancto.

Orémus.

\lettrine{G}{r}átiam tuam, quǽsumus, Dómine, méntibus nostris infúnde:~† ut qui, Angelo nuntiánte, Christi Fílii tui incarnatiónem cognóvimus,~* per passiónem ejus et crucem, ad resurrectiónis glóriam perducámur. Per eumdem Christum Dóminum nóstrum.

\rr  Amen.

\rubrique{A Nativitate Domini usque ad Purificationem B.M.V.:} 

 \vv Post partum Virgo invioláta permansísti.

 \rr Dei Génitrix intercéde pro nobis.

Orémus.

\lettrine{D}{e}us, qui salútis ætérnæ, beátæ Maríæ virginitáte fœcúnda, humáno géneri prǽmia præstitísti:~† tríbue, quǽsumus; ut ipsam pro nobis intercédere sentiámus,~* per quam merúimus auctórem vitæ suscípere, Dóminum nóstrum Jesum Christum Fílium tuum.

\rr  Amen.}{english}{\rubrique{From the first Sunday of Advent until Christmas Eve.}

\noindent \vv The Angel of the Lord declared unto Mary.

\noindent \rr And she conceived by the Holy Ghost.

\noindent Let us pray. Pour forth we beseech thee, O Lord, thy grace into our hearts, that we to whom the Incarnation of Christ, thy Son, was made known by the message of an Angel, may, by his Passion and cross, be brought to the glory of his Resurrection. Through the same Christ our Lord. Amen.

\rubrique{From First Vespers of Christmas until the Purification.}

\noindent \vv After childbirth, O Virgin, thou didst remain inviolate.

\noindent \rr Intercede for us, O Mother of God.

\noindent Let us pray. O God, Who by the fruitful virginity of blessed Mary, hast given to the human race the rewards of eternal salvation: grant, we beseech thee, that we may experience her intercession for us, through whom we deserved to receive the Author of life, our Lord Jesus Christ, thy Son. Amen.}