\begin{multicols}{2}
\raggedcolumns
\begin{enpars}
%%PBC
Begotten of the Father’s love before the world was made, called Alpha and
Omega, he the source and end of all things that are, that were, and that shall
be: forever and ever.

%% 2 Neale/Baker reworked/; 2, 3, 6, 7, 8 missing from PBC

He commanded and it was created,
He said and it was done:
The earth, the heavens, and the depths of the seas
three things in one,
All beneath the height of the sun and the globe of the moon: forever and ever.

%% 3 missing from PBC, reworked Jeff O

He clothes himself with the form of a frail Body,
with members subject to death; lest the human
race, the offspring of Adam, should perish
together with their first Parent, whom the terrible
law had condemned to the depths of hell. 

%%4 PBC
O blessed birth, when the Virgin conceiving of the Holy Spirit brought forth
our salvation, and when the Child, redeemer of the world, lifted his sacred face: forever and ever.

%%5 PBC
Let the height of heaven sing; sing, all ye angels; whatever has any power, sing
unto the praise of God. Let no tongue fall silent, and let every voice sound
forth: forever and ever.

%%6 from Jeff O unsourced translation

Behold him of
whom sang the inspired bards in the ancient
times, and whom the faithful pages of the prophets
had guaranteed in past times 
 the Promised one now suddenly apears.
 Let all creatures praise him: forever and ever.

%%7 Neale/Baker reworked
Blessed Judge of the dead
Blessed King of the living,
The right hand of the Father enthroned in the highest places,
The Just Avenger
coming thenceforth
who purges by thy power
the crimes of all: forever and ever.

%% 8 PBC reworked
Old men and young men, a choir of children, the throng
of mothers and virgins, simple girls, Harmonious in voice, let them resound in pure song: forever and
ever.

%%9/last PBC
To thee, O Christ, with the Father and the Holy Spirit, be hymns, glory,
perennial praise, giving of thanks, honor, virtue, victory, and eternal kingship: forever and ever.


\end{enpars}
\end{multicols}