\texteparacol{\lettrine{N}{ec} quisquam sumit sibi honórem, sed qui vocátur a Deo, tamquam Aaron. 
Sic et Chri\-stus non semetí\-psum clarificávit ut Póntifex fíeret, sed qui locútus est ad eum: Fílius meus es tu, ego hódie génui te. 
Quemádmodum et in álio loco dicit: Tu es sacérdos in ætérnum, secúndum órdinem Melchísedech. 
Qui in diébus carnis suæ preces, supplicationésque ad eum, qui possit illum salvum fácere a morte, cum clamóre válido et lácrimis ófferens, exaudítus est pro sua reveréntia. 
Et quidem cum esset Fílius Dei, dídicit ex iis, quæ passus est, obediéntiam:
Et consummátus, fa\-ctus est ómnibus obtemperántibus sibi causa salútis ætérnæ, 
Appellátus a Deo Póntifex juxta órdinem Melchísedech.}{english}{\noindent Neither doth any man take the honour to himself, but he that is called by God, as Aaron was.
 So Christ also did not glorify himself, that he might be made a high priest: but he that said unto him: Thou art my Son, this day have I begotten thee.
As he saith also in another place: Thou art a priest for ever, according to the order of Melchisedech.
Who in the days of his flesh, with a strong cry and tears, offering up prayers and supplications to him that was able to save him from death, was heard for his reverence.
And whereas indeed he was the Son of God, he learned obedience by the things which he suffered:
And being consummated, he became, to all that obey him, the cause of eternal salvation.
Called by God a high priest according to the order of Melchisedech.}