\texteparacol{\lettrine{N}{o}stis qui convéntus erat malignántium Judæórum, et quæ multitúdo erat operántium iniquitátem. Quam iniquitátem? Quia voluérunt occídere Dóminum Jesum Christum. Tanta ópera bona, inquit, osténdi vobis: pro\-pter quod horum me vultis occídere? Pértulit o\-mnes infírmos eórum, curávit o\-mnes lánguidos eórum, prædicávit regnum cælórum, non tácuit vítia eórum, ut ipsa pótius eis displicérent, non médicus, a quo sanabántur. His ómnibus curatiónibus ejus ingráti, tamquam multa febre phrenétici, insaniéntes in médicum, qui vénerat curáre eos, excogitavérunt consílium perdéndi eum: tamquam ibi voléntes probáre, utrum vere homo sit, qui mori possit, an áliquid super hómines sit, et mori se non permíttat. Verbum ipsórum agnóscimus in Sapiéntia Salomónis: Morte turpíssima, ínquiunt, condemnémus eum. Interrogémus eum: erit enim respéctus in sermónibus illíus. Si enim vere Fílius Dei est, líberet eum.}
{english}
{\noindent We know what secret counsel was that of the wicked Jews, and what insurrection was that of the workers of iniquity. Of what iniquity were they the workers? The murder of our Lord Jesus Christ. Many good works, saith he, have I showed you for which of those works go ye about to kill me? He had borne with all their weaknesses: He had healed all their diseases: He had preached unto them the kingdom of heaven: He had discovered to them their iniquities, that they might rather hate them, than the Physician that came to cure them. And now at last, without gratitude for all the tenderness of his healing love, like men raging in an high delirium, throwing themselves madly on the Physician, who had come to cure them, they took counsel together how they might kill him, as if to see if he were a man and could die, or something more than a man, and that would not let himself die. In the Wisdom of Solomon we recognize their words, Let us condemn him with a shameful death. Let us examine him; for, by his own saying, he shall be respected. If he be the Son of God, let him help him.}