\texteparacol{\lettrine{E}{x}acuérunt tamquam gládium linguas suas. Non dicant Judǽi: Non occídimus Christum. Etenim proptérea eum dedérunt júdici Piláto, ut quasi i\-psi a morte ejus videréntur immúnes. Nam cum dixísset eis Pilátus: Vos eum occídite: respondérunt, Nobis non licet occídere quemquam. Iniquitátem facínoris sui in júdicem hóminem refúndere volébant: sed numquid Deum júdicem fallébant? Quod fecit Pilátus, in eo i\-pso quod fecit, aliquántum párticeps fuit: sed in comparatióne illórum multo i\-pse innocéntior. Institit enim quantum pótuit, ut illum ex eórum mánibus liberáret: nam proptérea flagellátum prodúxit ad eos. Non persequéndo Dóminum flagellávit, sed eórum furóri satisfácere volens: ut vel sic jam mitéscerent, et desínerent velle occídere, cum flagellátum vidérent. Fecit et hoc. At ubi perseveravérunt, nostis illum lavísse manus, et dixísse, quod i\-pse non fecísset, mundum se esse a morte illíus. Fecit tamen. Sed si reus, quia fecit vel invítus: illi innocéntes, qui coëgérunt ut fáceret? Nullo modo. Sed ille dixit in eum senténtiam, et jussit eum crucifígi, et quasi i\-pse occídit: et vos, o Judǽi, occidístis. Unde occidístis? Gládio linguæ: acuístis enim linguas vestras. Et quando percussístis, nisi quando clamástis: Crucifíge, crucifíge?}
{english}{They whetted their tongue like a sword. The Jews cannot say: We did not murder Christ, albeit they gave him over to Pilate his judge, that they themselves might seem free of his death. For when Pilate said unto them, Take ye him: and kill him, they answered, It is not lawful for us to put any man to death. They could throw the blame of their sin upon a human judge: but did they deceive God, the Great Judge? In that which Pilate did, he was their accomplice, but in comparison with them, he had far the lesser sin. Pilate strove as far as he could, to deliver him out of their hands; for the which reason also he scourged him, and brought him forth to them. He scourged not the Lord for cruelty's sake, but in the hope that; he might so slake their wild thirst for blood: that, perchance, even they might be touched with compassion, and cease to lust for his death, when they saw what he was after the flagellation. Even this effort he made! But when Pilate saw that he could not prevail, but that rather a tumult was made, ye know how that he took water, and washed his hands before the multitude, saying: I am innocent of the blood of this Just Person. And yet he delivered him to be crucified! But if he were guilty who did it against his will, were they innocent; who goaded him on to it? No. Pilate gave sentence against him. and commanded him to be crucified. But ye, O ye Jews, ye also are his murderers! Wherewith? With your tongue, whetted like a sword. And when? But when ye cried, Crucify him! Crucify him!}