%\bigtitle{Lectio IV.\\ Ex tractátu sancti Augustíni Epíscopi super Psalmos.}

\scripture{In Psalm 63 ad versum 2.}
\texteparacol{
\lettrine{E}{x} tractátu sancti Augustíni Epíscopi super Psalmos.

\lettrine{P}{r}otexísti me, Deus, a convéntu malignántium, a multitúdine operántium iniquitátem. Jam i\-psum caput nostrum intueámur. Multi Mártyres tália passi sunt, sed nihil sic elúcet, quómodo caput Mártyrum: ibi mélius intuémur, quod illi expérti sunt. Proté\-ctus est a multitúdine malignántium, protegénte se Deo, protegénte carnem suam ipso Fílio, et hómine, quem gerébat: quia fílius hóminis est, et Fílius Dei est. Fílius Dei, pro\-pter formam Dei: fílius hóminis, pro\-pter formam servi, habens in potestáte pónere ánimam suam, et recípere eam. Quid ei potuérunt fácere inimíci? Occidérunt corpus, ánimam non occidérunt. Inténdite. Parum ergo erat, Dóminum hortári Mártyres verbo, nisi firmáret exémplo.}{english}{\noindent From the Treatise of Saint Augustine, bishop, Upon the Psalms.

\noindent Thou hast hidden me from the secret counsel of the wicked, from the insurrection of the workers of iniquity. Now let us fix our eyes upon our head. Many martyrs have. suffered such things as he suffered, but God's hiding of his suffering servants is not so well seen in the martyrs, as it is in the captain of the martyrs. And it is in him that we best see how it fared with them. He was hidden from the secret counsel of the wicked; hidden by God, being himself God; hidden, as touching the manhood, by God the Son, and the very manhood, which is taken into God the Son; because he is the Son of man, and he is the Son of God; Son of God, as being in the form of God; Son of man, as having taken upon him the form of a servant, whose life no man taketh from him, but who layeth it down of himself. He hath power to lay it down, and he hath power to take it again. What then was all that they which hated him could do? They could kill the body, but they were not able to kill the soul. Consider this very earnestly. It had been a small thing for the Lord to preach to the martyrs by his word, if he had not also nerved them by his example.}