\begin{enpars}
\begin{multicols}{2}
\raggedcolumns 
\textit{Priest:}  Lord, have mercy on us.

\textit{Faithful:} Christ, have mercy on us.

\textit{Priest:}  Lord, have mercy on us. Christ, hear us.

\textit{Faithful:} Christ, graciously hear us.

\textit{Priest:}  God, the Father of Heaven,

\textit{Faithful:} have mercy on us.\kern 0.05em\footnote{\raggedright{The faithful respond to each invocation in a similar way.}\par}

God the Son, Redeemer of the world,

God the Holy Spirit,

Holy Trinity, one God,

Holy Mary, pray for us.

Holy Mother of God,

Holy Virgin of virgins,

Mother of Christ,

%Mother of the Church, Paul VI
%%Mother of mercy, Francis
Mother of divine grace,

%%Mother of hope, Francis
Mother most pure,

Mother most chaste,

Mother inviolate,

Mother undefiled,

Mother most amiable,
Mother most admirable,

Mother of good counsel,

Mother of our Creator,
Mother of our Savior,

Virgin most prudent,

Virgin most venerable,

Virgin most renowned,

Virgin most powerful,

Virgin most merciful,

Virgin most faithful,

Mirror of justice,

Seat of wisdom,

Cause of our joy,

Spiritual vessel,

Vessel of honor,

Singular vessel of devotion,

Mystical rose,

Tower of David,

Tower of ivory,

House of gold,

Ark of the covenant,

Gate of heaven,

Morning star,

Health of the sick,

Refuge of sinners,

%%Comfort of Migrants, Francis
Comforter of the afflicted,

Help of Christians,

Queen of Angels,

Queen of Patriarchs,

Queen of Prophets,

Queen of Apostles,

Queen of Martyrs,

Queen of Confessors,

Queen of Virgins,

Queen of all Saints,
Queen conceived without original sin,

Queen assumed into heaven,

Queen of the most holy Rosary,

%Queen of Families, JPII
Queen of Peace,

Lamb of God, who takest away
the sins of the world, 
spare us o Lord.

Lamb of God, who takest away
the sins of the world, 
graciously hear us o Lord.

Lamb of God, who takest away
the sins of the world, 
\end{multicols}

\vv Pray for us, O holy Mother of God, 

\rr That we may be made worthy of the promises of Christ.

Let us pray. \nopagebreak

\lettrine{G}{r}ant, we beseech thee, o Lord God, unto us thy servants, that we may rejoice in continual health of mind and body; and, by the glorious intercession of Blessed Mary, ever Virgin, may we be delivered from present sadness and enter into the joy of thine eternal gladness. Through Christ our Lord.

\rr Amen.

\end{enpars}