\begin{enpars}
\begin{multicols}{2}
\raggedcolumns
\textit{Priest:} Lord, have mercy.

\textit{Faithful:} Christ, have mercy.

\textit{Priest:} Lord, have mercy. Christ hear us.

\textit{Faithful:} Christ, graciously hear us.

\textit{Priest:} God the Father of heaven,

\textit{Faithful:} have mercy on us.\kern 0.05em\footnote{\raggedright{The faithful respond to each invocation in a similar way.}\par}

God the Son, Redeemer of the world,

God the Holy Spirit,

Holy Trinity, one God,

Holy Mary, pray for us.

Saint Joseph,

Illustrious scion of David,

Light of patriarchs,

Spouse of the Mother of God,

Chaste Guardian of the Virgin,

Foster father of the Son of God,

Diligent protector of Christ,

Head of the Holy Family,

Joseph most chaste,

Joseph most just,

Joseph most prudent,

Joseph most strong,

Joseph most obedient,

Joseph most faithful,

Mirror of patience,

Lover of poverty,

Model of workers,

Glory of family life,

Guardian of virgins,

Pillar of families,

Solace of the afflicted,

Hope of the sick,

Patron of the dying,

 Terror of evil demons,
 
 Protector of Holy Church,
  
Lamb of God, who takest away
the sins of the world, spare us o Lord.

Lamb of God, who takest away
the sins of the world, graciously hear us o Lord.

Lamb of God, who takest away the sins of the world, 
have mercy on us.
\end{multicols}

\vv He hath made him lord of his house.

\rr And ruler of all his poseessions.

Let us pray.

\lettrine{O}{} God, who in thine ineffible
providence didst vouchsafe  to choose blessed
Joseph as spouse of thy most holy Mother; grant, we beseech thee, that we be may be made worthy
to have him for our intercession in heaven
whom on earth we venerate as our protector.
Through Christ our Lord. 

\rr Amen.
\end{enpars}