\begin{enpars}
\begin{multicols}{2}
\raggedcolumns 
\textit{Priest:}  Lord, have mercy on us.

\textit{Faithful:} Christ, have mercy on us.

\textit{Priest:}  Lord, have mercy on us. Christ, hear us.

\textit{Faithful:} Christ, graciously hear us.

\textit{Priest:}  God, the Father of Heaven,

\textit{Faithful:} have mercy on us.\kern 0.05em\footnote{\raggedright{The faithful respond to each invocation in a similar way.}\par}

God the Son, Redeemer of the world,

God the Holy Spirit,

Holy Trinity, one God,

Blood of Christ, only Son of the Father, save us.

Blood of Christ, incarnate Word of God,

Blood of Christ, of the new and eternal testament,

Blood of Christ, falling upon the earth in agony,

Blood of Christ, shed profussely in the Scourging,

Blood of Christ, flowing forth in the Crowning of Thorns,

Blood of Christ, poured out on the cross,

Blood of Christ, the price of our salvation,

Blood of Christ, without which there is no forgiveness,

Blood of Christ, Eucharistic drink and refreshment of souls,

%Blood of Christ, in which we are washed

Blood of Christ, stream of mercy,

Blood of Christ, victor over demons,

Blood of Christ, courage of the martyrs

Blood of Christ, strenght of confessors,

Blood of Christ, bringing forth Virgins,

Blood of Christ, help of those in peril,

Blood of Christ, relief of the burdened,

Blood of Christ, solace in sorrow,

Blood of Christ, hope of the penitent,

Blood of Christ, consolation of the dying,

Blood of Christ, peace and tenderness of hearts,

Blood of Christ, pledge of eternal life,

Blood of Christ, freeing souls from purgatory,

Blood of Christ, most worthy of honor of all glory and honor,

Lamb of God, who takest away
the sins of the world, 
spare us o Lord.

Lamb of God, who takest away
the sins of the world, 
graciously hear us o Lord.

Lamb of God, who takest away
the sins of the world, 
\end{multicols}

\vv Thou hast redeemed us, o Lord, with thy blood.

\rr And made for us, o God, a kingdom.

Let us pray.

\lettrine{A}{l}mighty and everlasting God, thou hast appointed thine only Begotten Son
to be the Redeemer of the world and willed to be appeased by his Blood; grant, we beg of thee,
that we may worthily adore this price of our salvation, and through its power be safeguarded from the evils
of this present life, so we may rejoice in its fruits forever in heaven.
Through the same Christ our Lord.

\rr Amen.

\end{enpars}