
\begin{enpars}
\begin{multicols}{2}
\raggedcolumns 
\textit{Priest:}  Lord, have mercy on us.

\textit{Faithful:} Christ, have mercy on us.

\textit{Priest:}  Lord, have mercy on us. Jesus, hear us.

\textit{Faithful:} Jesus, graciously hear us.

\textit{Priest:}  God, the Father of Heaven,

\textit{Faithful:} have mercy on us.\kern 0.05em\footnote{\raggedright{The faithful respond to each invocation in a similar way.}\par}
God the Son, Redeemer of the world,

God the Holy Spirit,

Holy Trinity, one God,

Jesus, Son of the living God,

Jesus, splendor of the Father,

Jesus, brightness of eternal light,

Jesus, king of glory,

Jesus, sun of justice,

Jesus, Son of the Virgin Mary,

Jesus, most amiable,

Jesus, most admirable,

Jesus, father of the world to come,

Jesus, the mighty God,

Jesus, angel of great counsel,

Jesus, most powerful,

Jesus, most patient,

Jesus, most obedient,

Jesus, meek and humble of heart,

Jesus, lover of chastity,

Jesus, lover of us,

Jesus, God of peace,

Jesus, author of life,

Jesus, model of all virtues,

Jesus, zealous lover of souls,

Jesus, our God,

Jesus, our refuge,

Jesus, father of the poor,

Jesus, treasure of the faithful,

Jesus, Good Shepherd,

Jesus, true light,

Jesus, eternal wisdom,

Jesus, infinite goodness,

Jesus, our way and our life,

Jesus, joy of angels,

Jesus, king of patriarchs,

Jesus, master of apostles,

Jesus, teacher of evangelists,

Jesus, strength of Martyrs,

Jesus, light of Confessors,

Jesus, purity of Virgins,

Jesus, crown of all Saints,

Be merciful, spare us, O Jesus.

Be merciful, graciously hear us, O Jesus.

%% is it Jesus, deliver us.?
%%or is it deliver us, O Jesus.
From all evil, Jesus, deliver us.

From all sin,

From thy wrath,

From the snares of the devil,

From the spirit of fornication,

From everlasting death,

From the neglect of thine inspirations,

Through the mystery of thy holy Incarnation,

Through thy Nativity,

Through thine Infancy,

Through thy most divine life,

Through thy Labors,

Through thine agony and Passion,

Through thy Cross and dereliction,

Through thy sufferings,

Through thy death and burial,

Through thy Resurrection,

Through thy Ascension,

Through thine institution of the Most Holy Eucharist,

Through thy joys,

Through thy glory,

Lamb of God, who takest away the sins of the world, spare us, O Jesus,

Lamb of God, who takest away the sins of the world, hear us, O Jesus,

Lamb of God, who takest away the sins of the world, have mercy on us, O Jesus.

\end{multicols}

\vv Jesus, hear us.

\rr Jesus, graciously hear us.

Let us pray. \nopagebreak

\lettrine{O}{} Lord Jesus Christ, who hast said: Ask and ye shall receive, seek and ye shall find, knock and it shall be opened unto you; mercifully attend to our supplications, and grant us the gift of thy divine charity, that we may ever love thee with our whole hearts, and never cease from praising Thee.

\lettrine{G}{i}ve us, O Lord, a perpetual fear and love of thy holy Name, for thou never ceasest to direct and govern by thy grace those whom thou instructest in the solidity of thy love: Who livest and reignest world without end.

\rr Amen.
\end{enpars}