\begin{enpars}
\begin{multicols}{2}
\raggedcolumns 

\textit{Priest:}  Lord, have mercy on us.

\textit{Faithful:} Christ, have mercy on us.

\textit{Priest:}  Lord, have mercy on us. Christ, hear us.

\textit{Faithful:} Christ, graciously hear us.

\textit{Priest:}  God, the Father of Heaven,

\textit{Faithful:} have mercy on us.\kern 0.05em\footnote{\raggedright{The faithful respond to each invocation in a similar way.}\par}

God, the Son, Redeemer of the world,

God, the Holy Ghost,

Holy Trinity, One God,

Heart of Jesus, Son of the Eternal Father,

Heart of Jesus, formed by the Holy Spirit in the womb of the Virgin Mother,

Heart of Jesus, substantially united to the Word of God,

Heart of Jesus, of Infinite Majesty,

Heart of Jesus, sacred temple of God,

Heart of Jesus, tabernacle of the Most High,

Heart of Jesus, House of God and gate of Heaven,

Heart of Jesus, burning furnace of charity,

Heart of Jesus, abode of justice and love,

Heart of Jesus, full of goodness and love,

Heart of Jesus, abyss of all virtues,

Heart of Jesus, most worthy of all praise,

Heart of Jesus, king and center of all hearts,

Heart of Jesus, in whom are all the treasures of wisdom and knowledge,

Heart of Jesus, in whom dwelleth all the fullness of the Divinity.

Heart of Jesus, in whom the Father was well pleased,

Heart of Jesus, of whose fullness we have all received,

Heart of Jesus, desire of the everlasting hills,

Heart of Jesus, patient and most merciful,

Heart of Jesus, enriching all who invoke thee,

Heart of Jesus, fountain of life and holiness,

Heart of Jesus, propitiation for our sins,

Heart of Jesus, loaded down with opprobrium,

Heart of Jesus, bruised for our offenses,

Heart of Jesus, obedient to death,

Heart of Jesus, pierced with a lance,

Heart of Jesus, source of all consolation,

Heart of Jesus, our life and resurrection,

Heart of Jesus, our peace and reconciliation,

Heart of Jesus, victim for our sins,

Heart of Jesus, salvation of those who trust in thee,

Heart of Jesus, hope of those who die in thee,

Heart of Jesus, delight of all the Saints,

Lamb of God, who takest away
the sins of the world,

spare us o Lord.

Lamb of God, who takest away
the sins of the world,

graciously hear us o Lord.

Lamb of God, who takest away
the sins of the world,

have mercy on us.
\end{multicols}

\vv Jesus, meek and humble of heart.

\rr Make our hearts like unto Thine.

Let us pray.

\lettrine{A}{l}mighty and eternal God, look upon the heart of thy dearly beloved Son and upon the praises and satisfaction he offers thee in the name of sinners and for those who seek thy mercy; be thou appeased by worthy homage, and grant us pardon in the name of the same Jesus Christ, thy Son, who liveth and reigneth with thee, world without end. 

\rr Amen.

\textit{Or}

Let us pray.

\lettrine{O}{} God, who hast suffered the heart of thy Son to be wounded by our sins, and in that very heart hast bestowed on us the abundant riches of thy love: grant that the devout homage of our hearts, which we render unto him, may of thy mercy be deemed a recompence acceptable in thy sight. Through the same Christ our Lord.

\rr Amen.

\end{enpars}