%\begin{paracol}{2}
\texteparacol{\lettrine{M}agníficat~* ánima mea Dóminum.\\ Et exsultávit spíritus meus~* in Deo salutári meo.

 Quia respéxit humilitátem ancíllæ suæ:~* ecce enim ex hoc beátam me dicent o\-mnes generatiónes.

 Quia fecit mihi magna qui potens est:~* et san\-ctum nomen ejus.

 Et misericórdia ejus a progénie in progénies~* timéntibus eum.

 Fecit poténtiam in bráchio suo:~* dispérsit supérbos mente cordis sui.

 Depósuit poténtes de sede,~* et exaltávit húmiles.

 Esuriéntes implévit bonis:~* et dívites dimísit inánes.

 Suscépit Israel púerum suum,~* recordátus misericórdiæ suæ.

 Sicut locútus est ad patres no\-stros,~* Abraham, et sémini ejus in sǽcula.

 Glória Patri, et Fílio,~* et Spirítui San\-cto.

 Sicut erat in princípio, et nunc, et semper,~* et in sǽcula sæculórum. Amen.}
 {english}
% \switchcolumn
% \begin{enpars}
%\begin{enpars}
{My soul \cc~* doth magnify the Lord.

And my spirit hath rejoiced~* in God my Saviour.

Because he hath regarded the humility of his handmaid;~* for behold from henceforth all generations shall call me blessed.

Because he that is mighty, hath done great things to me;~* and holy is his name.

And his mercy is from generation unto generations,~* to them that fear him.

He hath shewed might in his arm:~* he hath scattered the proud in the conceit of their heart.

He hath put down the mighty from their seat,~* and hath exalted the humble.

He hath filled the hungry with good things;~* and the rich he hath sent empty away.

He hath received Israel his servant,~* being mindful of his mercy:

As he spoke to our fathers,~* to Abraham and to his seed for ever.

Glory be to the Father, and to the Son,~* and to the Holy Ghost.

As it was in the beginning, is now,~* and ever shall be, world without end. Amen.}
%\end{otherlanguage}
%\end{enpars}
% \end{paracol}