{\centering\Large{\MakeUppercase{\capspace{novena prayer for the gifts of the holy ghost in preparation for the feast of pentecost.}}}\par}

\begin{enpars}
\lettrine{O}{} Lord Jesus Christ, * Who before ascending into heaven did promise to send the Holy Ghost to finish Thy work * in the souls of thy disciples,* deign to grant the same Holy Spirit to me, * that He may perfect in my soul, the work of Thy grace and Thy love.

\lettrine{G}{r}ant me the Spirit of Wisdom, * that I may despise the perishable things of this world, * the Spirit of Understanding to enlighten my mind with the light of Thy divine truth, * the Spirit of Counsel that I may ever choose the surest way of pleasing God and gaining heaven; * the Spirit of Fortitude that I may bear my cross with Thee; * the Spirit of Knowledge that I may know God and know myself * and grow perfect in the science of the Saints; * the Spirit of Piety that I may find the service of God sweet and amiable; * the Spirit of Fear that I may be filled with a loving reverence towards God and may dread in any way to displease Him.

%%original version
%\lettrine{M}{a}rk me, dear Lord, * with the sign of Thy true disciple * and animate me in all things with Thy Spirit. Amen.

%Come, Holy Ghost, fill the hearts of Thy faithful and enkindle in them the fire of Thy love.
%
%\vv Send forth Thy Spirit and they shall be created.
%
%\rr And Thou shalt renew the face of the earth.

\lettrine{M}{a}rk me, dear Lord, * with the sign of Thy true disciple * and animate me in all things with Thy Spirit. * Come, Holy Ghost, 
fill the hearts of Thy faithful and enkindle in them the fire of Thy love. * Send forth Thy Spirit and they shall be created. * And Thou shalt renew the face of the earth. Amen.

\end{enpars}