{\centering\Large{\MakeUppercase{\capspace{novena to saint thérèse of lisieux, the little flower.}}}\par}

\begin{enpars}

\lettrine{O}{} Glorious Saint Thérèse,
whom Almighty God hast raise up to aid
and inspire the human family,
I implore thy miraculous intercession.

\lettrine{T}{h}ou art so powerful in obtaining every need
of body and spirit from the Heart of God.
Holy Mother Church proclaims thee ``Prodigy of Miracles…
the greatest saint of Modern Times.'' Now I fervently beseech thee to answer my petition:

\rubrique{Here mention your intentions. Please also pray for all who are present today. Pray for the intentions of our Holy Father the Pope, and for all the clergy. Pray also for the sick and the dying, and for the suffering souls in purgatory.}

\lettrine{A}{n}d to carry out thy promises of
spending heaven doing good on earth…
of letting fall from Heaven a Shower of Roses.

\lettrine{L}{i}ttle Flower, give me thy childlike faith,
to see the Face of God
in the people and experiences of my life,
and to love God with full confidence.

\lettrine{S}{a}int Thérèse, my Carmelite Sister,
I will fulfill thy plea ``to be made known everywhere''
and I will continue to lead others to Jesus through thee. Amen.

\end{enpars}


%%original has you

%%%O Glorious St. Thérèse,
%%%whom Almighty God has raised up to aid
%%%and inspire the human family,
%%%I implore your Miraculous Intercession.
%%%
%%%You are so powerful in obtaining every need
%%%of body and spirit from the Heart of God.
%%%Holy Mother Church proclaims you “Prodigy of Miracles…
%%%the greatest saint of Modern Times.”
%%%
%%%Now I fervently beseech you to answer my petition:
%%%
%%%and to carry out your promises of
%%%spending heaven doing good on earth…
%%%of letting fall from Heaven a Shower of Roses.
%%%
%%%Little Flower, give me your childlike faith,
%%%to see the Face of God
%%%in the people and experiences of my life,
%%%and to love God with full confidence.
%%%
%%%St. Thérèse, my Carmelite Sister,
%%%I will fulfill your plea “to be made known everywhere”
%%%and I will continue to lead others to Jesus through you.
%%%
%%%Amen.