{\centering\Large{\MakeUppercase{\capspace{novena in preparation for the feast of the immaculate conception of the b.v.m.}}}\par}

\rubrique{\normaltext{Enchiridion Indulgentiarum,} § 34 attaches a partial indulgence to the public novenas before the feasts of the Immaculate Conception, Christmas, and Pentecost, under the usual conditions.}

\begin{enpars}
\lettrine{O}{} Immaculate Virgin Mary, * who wast so pleasing in the sight of God from the first moment of thy conception in the womb of thy mother, Saint Anne. I believe the teaching of our holy Mother the Church, * that in the first instant of thy conception, * by the singular grace and privilege of Almighty God, * in virtue of the merits of Jesus Christ, Savior of the human race and beloved Son, thou wert preserved from all stain of original sin. * I thank God for this wonderful privilege and grace He bestowed upon thee * as I honor thy Immaculate Conception.

\lettrine{O}{} Mary, conceived without sin, pray for us who have recourse to thee! * O Mother of God * intercede for us with thy divine Son, * and obtain for me the intention I ask in this novena:

\rubrique{Here mention your intentions. Please also pray for all who are present today. Pray for the intentions of our Holy Father the Pope, and for all the clergy. Pray also for the sick and the dying, and for the suffering souls in purgatory.}

\lettrine{O}{} Virgin Immaculate, * O lily of purity, * O mother of God and my Mother, * from thy throne in heaven, * turn thine eyes of pity upon me. * Filled with confidence in thy goodness and power, * I beg thee to help me in this journey of life, * which is so full of dangers for my soul. * I entrust myself entirely to thee, * that I may never be the slave of the devil through sin, * but may always life a humble and pure life. * I consecrate myself to thee forever, * for my only desire is to love thy Son Jesus Christ. * O Mary, since none of thy devout servants has ever perished, * may I too be saved. Amen.

\rubrique{Here recite three \normaltext{Hail Marys} and a \normaltext{Glory be} for the intentions of the Holy Father, in hope of obtaining the indulgence attached to this novena.}
\end{enpars}