{\centering\Large{\MakeUppercase{\capspace{tuesday novena to saint anthony.}}}\par}

\begin{enpars}

\rubrique{Said kneeling.}

\lettrine{G}{l}orious Saint Anthony, * safe refuge of the afflicted and distressed, * who by miraculous revelation * have directed all those who seek aid to come to your altar * with the promise * that whoever visits it for nine consecutive Tuesdays * and there piously invokes you, * will feel the power of your intercession. * I, a poor sinner, * encouraged by this promise, * come to you, o powerful Saint, * and with a firm hope I implore your aid, * your protection, * your counsel, * and your blessing.

\lettrine{O}{b}tain for me, I beseech thee, * my request in this necessity (think of your request). * But if it should be opposed to the will of God, * the welfare of my soul, obtain for me such other graces * as shall be conducive to my salvation. * Through Christ our Lord. Amen.

\vv Our Father…

\rr Give us this day…

\vv Hail Mary…

\rr Holy Mary…

\vv Glory be…

\rr As it was…world without end. Amen.

\vv Pray for us, o blessed Anthony.

\rr That we may be made worthy of the promises of Christ.

\rubrique{Priest:}

Let us pray.

\lettrine{A}{l}mighty and eternal God, who glorified thy faithful confessor Anthony with the perpetual gift of working miracles: graciously grant, that what we confidently seek through his merits, we may surely receive by his intercession. Through Christ Our Lord.

\rr Amen.

\rubrique{The priest blesses the faithful with the relic, when the prayers are said outside of the presence of the Most Blessed Sacrament. Then:}

\rubrique{Priest:}

\lettrine{I}{f} then you ask for miracles,
death, error, all calamities,
the leprosy and demons fly,
and health succeeds infirmities.

\rr The sea obeys, and fetters break, and lifeless limbs thou dost restore; whilst treasures lost are found again when young or old thine aid implore.

\rubrique{Priest:}

\lettrine{A}{ll} dangers vanish at your prayer,
and direst need doth quickly flee,
let those who know your power proclaim, 
let Paduans say, these are of thee.

\rr The sea obeys, etc.

\rubrique{Priest:}

To Father, Son may glory be,
and Holy Ghost eternally.

\rr The sea obeys.
\end{enpars}

%%%original has modern English verbs and pronouns
%%whosoever is more archaic but oh well