{\centering\Large{\MakeUppercase{\capspace{novena in preparation for the feast of the annunciation of the b.v.m.}}}\par}

\begin{enpars}

\lettrine{O}{} most holy Virgin Mary to whom God sent the Archangel Gabriel to announce that thou shouldst be the Mother of His only-begotten Son,
pray for us who have recourse to thee.

\lettrine{O}{} most loving Mother, attracted by thy beauty and thy sweetness, and by thy tender compassion, I confidently turn to thee, miserable as I am, and beg of thee to obtain for me from thy dear Son the favor that I request in this novena: 

\rubrique{(Mention your request)}

\lettrine{O}{b}tain for me also, o Queen of heaven, the most lively contrition for my many sins and the grace to imitate closely those virtues which thou hast practiced so faithfully, especially humility, purity, and obedience. Above all, I beg thee to be my Mother and Protectress, to receive me into the number of thy devoted children, and to guide me from thy high throne of glory.

\lettrine{D}{o} not reject my petitions, o Mother of Mercy! Have pity on me, and do not abandon me during life or at the moment of my death. Amen.

\vv Our Father…

\rr Give us this day…

\vv Hail Mary…

\rr Holy Mary…

\vv Glory be…

\rr As it was…world without end. Amen.
\end{enpars}

%original has contemporary English and no additional o's which sounds off

%O most holy Virgin Mary to whom God sent the Archangel Gabriel to announce that you should be the Mother of His only-begotten Son,
%pray for us who have recourse to you.
%
%Most loving Mother, attracted by your beauty and sweetness, and by your tender compassion, I confidently turn to thee, miserable as I am, and beg of you to obtain for me from your dear Son the favor that I request in this novena: 
%
%\rubrique{(Mention your request)}
%
%Obtain for me also, Queen of heaven, the most lively contrition for my many sins and the grace to imitate closely those virtues which you practiced so faithfully, especially humility, purity, and obedience. Above all, I beg you to be my Mother and Protectress, to receive me into the number of your devoted children, and to guide me from your high throne of glory.
%
%Do not reject my petitions, Mother of Mercy! Have ptiy on me, and do not abandon me during life or at the moment of my death. Amen.