{\centering\Large{\MakeUppercase{\capspace{novena to saint joseph.}}}\par}

\begin{enpars}
\lettrine{S}{a}int Joseph, I, your unworthy child, greet you. Thou art the faithful protector and intercessor of all who love and venerate thee. Thou knowest that I have special confidence in thee and that, after Jesus and Mary, I place all my hope of salvation in you, for you arre especially powerful with God and will never abandon your faithful servants. Therefore I humbly invoke you and commend myself, with all who are dear to me and all that belong to me, to your intercession. I beg of you, by your love for Jesus and Mary, not to abandon me during life and to assist me at the hour of my death.

\lettrine{G}{l}orious Saint Joseph, spouse of the Immaculate virgin, obtain for me a pure, humble, charitable mind, and perfect resignation to the divine will. Be my guide, my father, and my model through life that I may merit to die as thou didst in the arms of Jesus and Mary.

\lettrine{L}{o}ving Saint Joseph, faithful follower of Jesus Christ, I raise my heart to you to implore your powerful intercession in obtaining from the divine heart of Jesus all the graces necessary for my spiritual and temporal welfare, particularly the grace of a happy death, and the special grace I now implore:

\rubrique{Here mention your intentions. Please also pray for all who are present today. Pray for the intentions of our Holy Father the Pope, and for all the clergy. Pray also for the sick and the dying, and for the suffering souls in purgatory.}

\lettrine{G}{u}ardian of the Word Incarnate, I feel confident that your prayers in my behalf will be graciously heard before the throne of God. Amen.
\end{enpars}