{\centering\Large{\MakeUppercase{\capspace{novena prayers to the child jesus in preparation for the feast of christmas.}}}\par}

\begin{enpars}

\lettrine{O}{} Divine Infant, *
after the wonders of Your birth in Bethlehem, *
Thou wished to extend Thine infinite mercy to the whole world. *
Thou callest the Wise Men, * by heavenly inspiration, * to Thy crib,
which was a royal throne.
Thou graciously received those holy men
who were obedient to the divine call *
and hastened to Thy feet. *
They recognized and worshipped Thee as Prince of Peace, *
the Redeemer of mankind,
and the very Son of God.

\lettrine{S}{h}ow us also Thy goodness and almighty power. * Enlighten our minds, * strengthen our wills, * and inflame our hearts to know Thee, * to serve thee, and to love Thee in this life, * that we may find our joy in Thee eternally in the life to come.

O Jesus, most powerful Child, * I implore Thee again to help me:

\rubrique{Here mention your intentions. Please also pray for all who are present today. Pray for the intentions of our Holy Father the Pope, and for all the clergy. Pray also for the sick and the dying, and for the suffering souls in purgatory.}

\lettrine{D}{i}vine Child, * through Thy most Holy Mother's powerful intercession and through the boundless mercy of Thine omnipotence as God, * I beseech Thee for a favorable answer to my prayer during this Novena. * Grant me the grace of posessing Thee eternally with Mary and Joseph * and of adoring Thee with Thine holy angels and saints.

\rubrique{Here recite three \normaltext{Hail Marys} and a \normaltext{Glory be} for the intentions of the Holy Father, in hope of obtaining the indulgence attached to this novena.}

\end{enpars}