{\centering\Large{\MakeUppercase{\capspace{novena in preparation for the feast of saint anthony.}}}\par}

\begin{enpars}
\lettrine{I}{} greet thee, Saint Anthony, lily of purity, ornament and glory of Christianity. I greet thee, great saint, filled with divine love. I rejoice at the favors which our Lord has so generously bestowed on thee. I humbly and confidently entreat thee to help me, for I know that God hast given thee charity and pity, as well as power.

\lettrine{G}{o}od Saint Anthony, meek and humble in spite of thy great gifts, and whom we invoke so confidently to help us find that which is lost, pray, I beg of thee, that I may never lose my faith, the gift that I cherish most. Behold my needs:

\rubrique{Here mention your intentions. Please also pray for all who are present today. Pray for the intentions of our Holy Father the Pope, and for all the clergy. Pray also for the sick and the dying, and for the suffering souls in purgatory.}

\lettrine{I}{} beg of thee, by the love thou dhast felt toward the little Jesus when He embraced thee so lovingly, to tell Him now of my wants. Remember how complete thy happiness was when thou didst hold Him in thine arms.

\lettrine{I}{} venerate thee, glorious favorite of God, and bow my guilty head before thee in humble reverence, while I raise my sad heart full of hope toward the throne of divine mercy, and to thee, my intercessor. He who placed Himself in thine arms will now fill thy hands with all I ask of thee.

Obtain for me what I desire, if it be God’s holy Will, and I will make known the wondrous power of thine intercession for the greater glory of God. Amen.
\end{enpars}

%%original has you

%\lettrine{I}{} greet you, Saint Anthony, lily of purity, ornament and glory of Christianity. I greet you, great saint, filled with divine love. I rejoice at the favors our Lord has so generously bestowed on you. I humbly and confidently entreat you to help me, for I know that God has given you charity and pity, as well as power.
%
%\lettrine{G}{o}od Saint Anthony, meek and humble in spite of your great gifts, and whom we invoke so confidently to help us find that which is lost, pray, I beg of you, that I may never lose my faith, the gift that I cherish most.
%
%\lettrine{B}{}hold my needs:
%
%\rubrique{Here mention your intentions. Please also pray for all who are present today. Pray for the intentions of our Holy Father the Pope, and for all the clergy. Pray also for the sick and the dying, and for the suffering souls in purgatory.}
%
%I beg of you, by the love you felt toward the little Jesus when He embraced you so lovingly, to tell Him now of my wants. Remember how complete your happiness was when you held Him in your arms.
%
%\lettrine{I}{} venerate you, glorious favorite of God, and bow my guilty head before you in humble reverence, while I raise my sad heart full of hope toward the throne of divine mercy, and to you, my intercessor. He who placed Himself in your arms will now fill your hands with all I ask of you.
%
%Obtain for me what I desire, if it be God’s holy Will, and I will make known the wondrous power of your intercession for the greater glory of God. Amen.