\begin{enpars} \noindent \vv Let us pray for our Pontiff, \nomen{} \rr The Lord preserve him, and give him life, and make him blessed upon the earth, and deliver him up not up to the will of his enemies.

\texteparacol{\vv  Fiat manus tua super virum déxteræ tuae.

\rr Et super fílium hóminis quem confirmásti tibi.

Orémus.

\lettrine{D}{e}us, ómnium fidélium pastor et re\-ctor, fámulum tuum \nomen, quem pastórem Ecclésiæ tuæ præésse voluísti propítius réspice:~† da ei, quǽsumus, verbo et exémplo, quibus præest, proficere;~* ut ad vitam una cum grege sibi crédito, pervéniat sempitérnam. Per Christum Dominum nostrum. \rr Amen.

\vel

Orémus.

\lettrine{O}{m}nípotens sempitérne Deus, miserere fámulo tuo Pontífici
nostro \nomen:~† et dírige eum secúndum tuarn cleméntiam in
viam salútis ætérnae;~* ut te donánte, tibi plácita cúpiat, et tota
virtúte perfíciat. Per Christum Dominum nostrum.

\rr Amen.}
{english}
{\noindent \vv Let thy hand be upon the man of thy right hand.

\noindent \rr And upon the son of man whom thou madest strong for thyself.

Let us pray. O God, Shepherd and Ruler of all the faithful, look mercifully upon thy servant \nomen, whom thou hast chosen as shepherd to preside over thy church: grant him, we beseech thee, that, by word and example, he may edify those over whom he hath charge, so that together with the flock committed to him, he may attain everlasting life. Through Christ our Lord. Amen.

\orrubrique

Let us pray.  Almighty, everlasting God, have mercy upon thy servant, \nomen{} our Sovereign Pontiff, and direct him, according to thy clemency, into the way of everlasting salvation; that by thy grace he may desire those things that are pleasing to thee, and perform them with all his strength. Amen.}