
\begin{enpars}

\rubrique{Having prayed the Litany of the Sacred Heart, on First Fridays or on other occasions dedicated to the Sacred Heart, the priest recites this prayer.}

\lettrine{M}{o}st sweet Jesus, whose overflowing charity for men is requited by so much
forgetfulness, negligence and contempt, behold us prostrate before thee, eager to
repair by a special act of homage the cruel indifference and injuries to which thy loving
Heart is everywhere subject.

Mindful, alas! that we ourselves have had a share in such great indignities, which we
now deplore from the depths of our hearts, we humbly ask thy pardon and declare
our readiness to atone by voluntary expiation, not only for our own personal offenses,
but also for the sins of those, who, straying far from the path of salvation, refuse in
their obstinate infidelity to follow thee, their shepherd and leader, or, renouncing the
promises of their baptism, have cast off the sweet yoke of thy law.

We are now resolved to expiate each and every deplorable outrage committed
against thee; we are determined to make amends for the manifold offenses against
Christian modesty in unbecoming dress and behavior, for all the foul seductions laid to
ensnare the feet of the innocent, for the frequent violations of Sundays and holydays,
and the shocking blasphemies uttered against thee and thy Saints.
We wish also to make amends for the insults to which thy Vicar on earth and thy
priests are subjected; for the profanation, by conscious neglect or terrible acts of
sacrilege of the very Sacrament of thy divine Love; and lastly for the public crimes
of nations who resist the rights and teaching authority of the Church which thou hast
founded.

Would, O divine Jesus, that we were able to wash away such abominations with our
blood. We now offer, in reparation for these violations of thy divine honor, the
satisfaction thou didst once make to thine Eternal Father on the cross and which thou
continuest to renew daily on our altars; we offer it in union with the acts of atonement of
thy Virgin Mother and all the Saints and of the pious faithful on earth; and we sincerely
promise to make recompense, as far as we can with the help of thy grace, for all neglect
of thy great love and for the sins we and others have committed in the past.
Henceforth, we will live a life of unwavering faith, of purity of conduct, of perfect
observance of the precepts of the Gospel and especially that of charity. We promise
to the best of our power to prevent others from offending thee and to bring as many
as possible to follow thee.

O loving Jesus, through the intercession of the Blessed Virgin Mary, our model in
reparation, deign to receive the voluntary offering we make of this act of expiation;
and by the crowning gift of perseverance keep us faithful unto death in our duty and
the allegiance we owe to thee, so that we may all one day come to that happy home,
where with the Father and the Holy Ghost thou livest and reignest, God, world without
end.

\rr Amen.

\end{enpars}