\texteparacol{\lettrine{Q}{u}o perduxérunt illas scrutatiónes suas, quas perscrutántes defecérunt, ut étiam mórtuo Dómino et sepúlto, custódes pónerent ad sepúlcrum? Dixérunt enim Piláto: Sedúctor ille: hoc appellabátur nómine Dóminus Jesus Christus, ad solátium servórum suórum, quando dicúntur seductóres: ergo illi Piláto: Sedúctor ille, ínquiunt, dixit adhuc vivens: Post tres dies resúrgam. Jube ítaque custodíri sepúlcrum usque in diem tértium, ne forte véniant discípuli ejus, et furéntur eum, et dicant plebi: Surréxit a mórtuis: et erit novíssimus error pejor prióre. Ait illis Pilátus: Habétis custódiam, ite, custodíte sicut scitis. Illi autem abeúntes, muniérunt sepúlcrum, signántes lápidem cum custódibus.}
{english}
{\noindent How far did the accomplishment of their diligent search reach? Even to the setting a watch of soldiers at the sepulchre, to guard the Lord, even after he was dead and buried. For they said unto Pilate: Sir, we remember that that deceiver. This was the term by which they designated the Lord Jesus Christ, and the remembrance that he was so named is a sweet consolation to us his servants, when we are called impostors. So they said unto Pilate, that deceiver said, while he was yet alive: After three days I will rise again. Command therefore that the sepulchre be made sure until the third day, lest his disciples come and steal him away, and say unto the people: He is risen again from the dead: so the last error shall be worse than the first. Pilate said unto them: Ye have a watch; go your way; make it as sure as ye can. So they went and made the sepulchre sure, sealing the stone, and setting a watch.}