\texteparacol{\lettrine{P}{o}suérunt custódes mílites ad sepúlcrum. Concússa terra Dóminus resurréxit: mirácula facta sunt tália circa sepúlcrum, ut et ipsi mílites, qui custódes advénerant, testes fíerent, si vellent vera nuntiáre. Sed avarítia illa, quæ captivávit discípulum cómitem Christi, captivávit et mílitem custódem sepúlcri. Damus, ínquiunt, vobis pecúniam: et dícite, quia vobis dormiéntibus venérunt discípuli ejus, et abstulérunt eum. Vere defecérunt scrutántes scrutatiónes. Quid est quod dixísti, o infélix astútia? Tantúmne déseris lucem consílii pietátis, et in profúnda versútiæ demérgeris, ut hoc dicas: Dícite quia vobis dormiéntibus venérunt discípuli ejus, et abstulérunt eum? Dormiéntes testes ádhibes: vere tu i\-pse obdormísti, qui scrutándo tália defecísti.}
{english}
{\noindent So they went, and made the sepulchre sure, sealing the stone, and setting a watch and anon, behold, there was a great earthquake, and the Lord arose. So great wonders were wrought about the sepulchre that the very soldiers, which were put to guard it, were witnesses thereto, if only they would have told the truth. But the same love of money which had made a slave of that disciple who was a companion of Christ, made slaves also of the soldiers that were put to watch his sepulchre. Some of the watch came into the city, and showed unto the chief priests all the things that were done: and when they were assembled with the elders, and had taken counsel, they gave large money unto the soldiers, saying: Say ye, his disciples came by night and stole him away while we slept. In good sooth, their diligent search had been accomplished and ended before this. What didst thou say, O stupid cunning? Wast thou indeed so utterly void of the light of godly wisdom, and confounded in the bottomless pit of thine own falsehood as to tell them to say: His disciples came by night, and stole him away while we slept? Part of the testimony of thine eye-witnesses was that they were asleep at the time: thou thyself wast asleep not to be able to see that on their own testimony, their testimony must have been worthless.}