\scripture{In Psalmum 63 ad versum 7.}

\texteparacol{\lettrine{E}{x} tractátu sancti Augustíni Epíscopi super Psalmos.

\lettrine{A}{c}cédet homo ad cor altum, et exaltábitur Deus. Illi dixérunt: Quis nos vidébit? Defecérunt scrutántes scrutatiónes, consília mala. Accéssit homo ad i\-psa consília, passus est se tenéri ut homo. Non enim tenerétur nisi homo, aut viderétur nisi homo, aut cæderétur nisi homo, aut crucifigerétur, aut morerétur nisi homo. Accéssit ergo homo ad illas o\-mnes passiónes, quæ in illo nihil valérent, nisi esset homo. Sed si ille non esset homo, non liberarétur homo. Accéssit homo ad cor altum, id est, cor secrétum, obíciens aspéctibus humánis hóminem, servans intus Deum: celans formam Dei, in qua æquális est Patri, et ófferens formam servi, qua minor est Patre.
}{english}
{\noindent We shall attain to thoughts that are very deep: but God shall still be exalted. The enemies of our Lord had communed of laying snares privily; they had said, Who shall see them? They had searched out iniquities; they had accomplished a diligent search. And man attained even unto (the realization of) their counsels, for the Lord, as man, suffered himself to be taken. For he had not been taken at all, unless he had been a man, or seen, unless he had been a man, or smitten, unless he had been a man, or crucified, unless he had been a man, or have died, unless he had been a man. man therefore, he attained unto all those sufferings, which had had nothing in him, unless he had been a man. But if he had not been man, man had not been redeemed. And the Lord as man attained to thoughts that were very deep, yea, secret; showing the manhood to the eyes of men, and keeping the godhead within him; veiling the form of God, as touching which, he is equal to the Father, and manifesting the form of a servant, as touching which, he is inferior to the Father.}