{\centering\Large{\MakeUppercase{\capspace{devotions to saint anthony.}}}\par}

\smalltitle{prayer to the infant jesus in saint anthony's arms.}

O Jesus my Saviour! Who vouchsafed to appear to St Anthony in the form of an infant, I implore thee, through the love thou didst bear to this saint when he dwelt on earth and which thou now bearest to him in heaven, graciously hear my prayer and assist me in my necessities, Who livest and reignest, world without end. Amen.

\smalltitle{prayer to saint anthony for seeking a lost article}

Dear Saint Anthony, thou art the patron of the poor and the helper of all who seek lost articles. Help me to find the object which I have lost so that I will be able to make better use of the time I will gain for God’s greater honor and glory. Grant thy gracious aid to all people who seek that which they have lost, especially those who seek to regain God’s grace. Amen.

\smalltitle{miracle prayer to saint anthony of padua.}

\lettrine{O}{} holy Saint Anthony, the gentlest and
kindest of Saints, thy burning love of
God, thine exalted virtue, and thy great
charity towards thy fellow creatures,
made thee worthy when on earth to
possess miraculous powers such as were
given to no other saint.

\lettrine{M}{i}racle waited on thy word, and that
word thou werest ever ready to speak at the
request of those in trouble.
The anxious prayer of bitter trial was
never addressed to thee in vain.

\lettrine{T}{o} the sick thou gavest back health; thou didst
restore what was lost; the sorrow
stricken were the objects of thy tender
compassion; even the dead thou didst raise to
life when the wounded heart cried out to
thee from the depths of its bitter anguish.

\lettrine{W}{h}en on earth nothing was impossible with thee, except not to have
compassion on those in distress and sorrow.

\lettrine{E}{n}couraged by this thought, and convinced of the efficacy of thy holy
intercession, we kneel before thy holy image, and full of confidence, we
implore you to obtain for us.

\rubrique{Here mention your intention.}

The answer to this our prayer may require a miracle. Even so, are you not
the Saint of Miracles, who, when on earth, had but to speak the mightiest
wonders were wrought!
O gentle and loving St Anthony, you whose heart was ever full of human
sympathy, whispers our prayer into the ears of the Infant Jesus, who loved
to linger in your arms.
One word from you and our prayer will be granted. O, speak but that word
and the gratitude of our heart will ever be yours!

Amen.

Our Father, Hail Mary, and Glory be.



%%original has you

%%O Holy St Anthony, the gentlest and
%%kindest of Saints, your burning love of
%%God, your exalted virtue, and your great
%%charity towards your fellow creatures,
%%made you worthy, when on earth to
%%possess miraculous powers such as were
%%given to no other saint.
%%Miracle waited on your word, and that
%%word you were ever ready to speak at the
%%request of those in trouble.
%%The anxious prayer of bitter trial was
%%never addressed to you in vain.
%%To the sick you gave back health; you
%%restored what was lost; the sorrow
%%stricken were the objects of your tender
%%compassion; even the dead you raised to
%%life when the wounded heart cried out to
%%you from the depths of its bitter anguish.
%%When on earth nothing was impossible with you, except not to have
%%compassion on those in distress and sorrow.
%%Encouraged by this thought, and convinced of the efficacy of your holy
%%intercession, we kneel before your holy image, and full of confidence, we
%%implore you to obtain for us (here mention your request).
%%The answer to this our prayer may require a miracle. Even so, are you not
%%the Saint of Miracles, who, when on earth, had but to speak the mightiest
%%wonders were wrought!
%%O gentle and loving St Anthony, you whose heart was ever full of human
%%sympathy, whispers our prayer into the ears of the Infant Jesus, who loved
%%to linger in your arms.
%%One word from you and our prayer will be granted. O, speak but that word
%%and the gratitude of our heart will ever be yours!
%%Amen.

%%Our Father, Hail Mary, and Glory be.

%%87 87 87
%%we need a tune but there aren't that many in this meter, not GOOD ones anyway

%%If great wonders thou desirest,
%%Hopeful to Saint Anthony pray;
%%Error, Satan, wants the direst,
%%Death and pest his will obey,
%%And the sick, who beg his pity,
%%From their couches haste away.
%%
%%Young and old are ever singing,
%%Praises to Saint Anthony bringing;
%%Stormy ocean calms its passion,
%%Bonds and fetters break in twain,
%%Treasures lost and limbs disabled,
%%These his pow'r restores again.
%%
%%%%Fr Price is missing this verse
%%
%%Padua has been the witness
%%Of these deeds six hundred years;
%%Dangers flee and need must perish,
%%Grief and sorrow disappear,
%%Filling all the world with wonder,
%%While the demons quake with fear.
%%
%%Glory be to God the Father
%%And to His coequal Son,
%%To the Holy Ghost resplendent;
%%One in Three -- Three in One;
%%Praise we Father, Son and Spirit
%%While eternal ages run.