\begin{multicols}{2}
\raggedcolumns
\begin{enpars}
The grieving Mother stood weeping beside the cross where her Son was hanging.

Through her weeping soul, compassionate and grieving, a sword passed.

O how sad and afflicted was that blessed Mother of the Only-begotten!

Who mourned and grieved, the pious Mother, looking at the torment of her glorious Child.

Who is the man that would not weep if he saw the Mother of Christ in such torment?

Who would not be able to feel compassion on beholding Christ’s Mother suffering with her Son?

For the sins of his people she saw Jesus in torment and subjected to the scourge.

She saw her sweet offspring dying, forsaken, while he gave up his spirit.

O Mother, fountain of love, make me feel the power of sorrow, that I may grieve with thee.

Grant that my heart may burn in the love of Christ my Lord, that I may greatly please him.

Holy Mother, grant that the wounds of the Crucified drive deep into my heart.

That of thy wounded Son, who so deigned to suffer for me, I may share the pain.

Let me, pious one, weep with thee, bemoan the Crucified, for as long as I live.

To stand beside the cross with thee, and to join thee in thy weeping, this I desire.

Chosen Virgin of virgins, be not bitter with me, let me weep with thee.

Grant that I may bear the death of Christ, share his Passion, and commemorate his wounds.

Let me be wounded with his wounds, let me be inebriated by the cross and thy Son’s blood.

Lest I burn, set afire by flames, Virgin, may I be defended by thee, on the day of judgement.

Christ, when it is time to pass away, grant that through thy Mother I may come to the palm of victory.

When my body dies, grant that to my soul is given the glory of paradise.
\end{enpars}\end{multicols}